---
title: "05_descriptives"
---

```{r setup}

# 1. Load packages, paths and data configurations
source(here::here("R", "00_setup.R"))

# 2. Set a global theme for plots
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# ESM

```{r ESM_setup}
#| include: false
# load data
esm_data  <- readRDS(file.path(dir_data_ana, "esm_ana.rds"))
# load data configuration file 
load_config("ESM")
```

```{r missingness}
esm_data %>% select(NA_own, PA_own, love, perc_resp, neg_gen, pos_gen,
                    PA_similarity_act, NA_similarity_act, PA_similarity_perc,
                    NA_similarity_perc) %>%
  miss_var_summary()
```

The variables 'neg_gen' (about the occurrence of any (general) negative event since previous beep) and 'pos_gen' (any positive event since previous beep) are very high in missingness (91,4%). This is because these multiple choice questions were only asked on beeps were partner was not present (i.e., in 31,9% of beeps) AND there has not been any contact with partner since previous beep (reducing the proportion of beeps that were asked the neg_gen and pos_gen items to 8,53%). This explains the full missingness.

The moderation by event valence (positive vs negative) that will be analyzed (see pre-registration step 3b of study component 1) will therefore have limited data.

The perceived similarity measures are missing for 18,3% of the beeps. This is the same level as for NA_own, PA_own, etc., and is 1 - compliance . The actual similarity measures are missing for 27,6%, which is more because the calculating depends in both partners answering the underlying items

### Repeated measures correlations

As in @bakdashRepeatedMeasuresCorrelation2017 we look at some repeated measure correlations. This is the common within person or within dyad correlation.

#### Intra-individual Correlation: PA vs NA ($r_{rm} = -0.60, p < .001$)

```{r}
#| warning: false

rmcorr(person, PA_own, NA_own, esm_data)
plot(rmcorr(person, PA_own, NA_own, esm_data))
rmcorr(person, cPA_own, cNA_own, esm_data)      # showing the same but on already centered vars
plot(rmcorr(person, cPA_own, cNA_own, esm_data))  # one line because already centered
```

The correlation between an individual's own PA and NA is strong and negative. The \`rmcorr\` plot shows consistent negative slopes across participants, suggesting that as a person's positive affect increases relative to their mean, their negative affect decreases.

#### Dyadic PA Convergence: Own PA vs Partner PA ($r_{rm} = 0.23, p < .001$)

```{r}
#| warning: false

rmcorr(dyad, cPA_own, cPA_part_act, esm_data)
plot(rmcorr(dyad, cPA_own, cPA_part_act, esm_data))
```

#### Dyadic NA Convergence: Own NA vs Partner NA ($r_{rm} = 0.22, p < .001$)

```{r}
#| warning: false

rmcorr(dyad, cNA_own, cNA_part_act, esm_data)
plot(rmcorr(dyad, cNA_own, cNA_part_act, esm_data))
```

### Intraclass Correlation Coefficient (ICC)

How much of the variance in the data is due to differences between people or dyads rather than within people or dyads (like moment to moment fluctuations). We calculated ICCs using a model-based approach rather than the `psych` package to properly account for the 3-level nested structure of our ESM data (Moments within Persons within Dyads) and to handle the unbalanced nature of our repeated measurements.

```{r}

# Variables to analyze
target_vars <- c("PA_own", "NA_own", "NA_part_perc", "PA_part_perc", "love", "perc_resp", "PA_similarity_perc", "NA_similarity_perc")

# Run the loop
icc_table <- target_vars %>%
  map_dfr(function(v) {
    # Construct the 3-level formula: Moments < Person < Dyad
    formula_str <- as.formula(paste0(v, " ~ 1 + (1 | dyad / person)"))
    # Fit the model
    model <- lmer(formula_str, data = esm_data)
    # Extract ICCs and format into a single row
    icc_data <- icc(model, by_group = TRUE)
    # Create a clean data frame for this variable
    data.frame(
      Variable = v,
      ICC_Dyad = icc_data$ICC[icc_data$Group == "dyad"],
      ICC_Person = icc_data$ICC[icc_data$Group == "person:dyad"]
    )
  })

# View the results
print(icc_table)

# Calculate the Residual (Momentary) variance
plot_data <- icc_table %>%
  mutate(Momentary = 1 - (ICC_Dyad + ICC_Person)) %>%
  pivot_longer(cols = c(ICC_Dyad, ICC_Person, Momentary), 
               names_to = "Level", values_to = "Proportion")

plot <- ggplot(plot_data, aes(x = Variable, y = Proportion, fill = Level)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_discrete(type = hcl.colors(4, palette = "Grays", alpha = 0.8)) +
  labs(title = "Variance Decomposition by Variable",
       x = "", y = "Percentage of Total Variance") +
  theme_minimal()

# show plot
print(plot)

# save plot
plot_counter <- 1
prefix <- "ICC"
plot_name <- sprintf("%s_%d_icc_variance_decomposition_1.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1

```

The ICC analysis for the three-level variables indicates that a substantial portion of the variance in affect and relationship perceptions is attributable to momentary, within-person fluctuations, as both dyad-level and person-level ICCs remain relatively low to moderate. Perceived responsiveness stands out with the highest dyad-level stability (ICC=0.35), suggesting it is more strongly characterized by shared couple-level baselines than the other variables. In contrast, the similarity perception variables (`PA_similarity_perc` and `NA_similarity_perc`) show very low ICCs at both the dyad and person levels, indicating that how similar partners perceive themselves to be is a highly dynamic state that changes significantly from moment to moment rather than being a stable trait of the individual or the dyad.

Technical note:\
Y∼1+(1∣dyad/person), decomposes the total variance into three independent sources:

1.  σdyad2$​: Variance of the dyad-level random intercept (between-dyad differences).

2.  σperson2​: Variance of the person-level random intercept nested within dyads (stable differences between partners).

3.  σerror2​: The residual variance (moment-to-moment or within-person fluctuations).

4.  The ICC is the proportion of variance of the total variance.

The **total variance** is the sum: σtotal2​=σdyad2​+σperson2​+σerror2​.

Since the `_act` similarity scores are identical for both partners, they are mathematically "dyad-level" constants at each time point, so that their data only has two levels of meaningful variation.

```{r}

dyad_level_vars <- c("PA_similarity_act", "NA_similarity_act")

# Simplified 2-level loop (Moments < Dyad)
icc_table_dyad_only <- dyad_level_vars %>%
  map_dfr(function(v) {
    formula_str <- as.formula(paste0(v, " ~ 1 + (1 | dyad)"))
    model <- lmer(formula_str, data = esm_data)
    icc_data <- icc(model) # Standard ICC for 2-level
    data.frame(
      Variable = v,
      ICC_Dyad = icc_data$ICC_adjusted,
      ICC_Person = "-" # Explicitly NA as it's not applicable
    )
  })

print(icc_table_dyad_only)

plot_data <- icc_table_dyad_only %>%
  mutate(Momentary = 1 - (ICC_Dyad)) %>%
  pivot_longer(cols = c(ICC_Dyad, Momentary), 
               names_to = "Level", values_to = "Proportion")

plot <- ggplot(plot_data, aes(x = Variable, y = Proportion, fill = Level)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_discrete(type = hcl.colors(4, palette = "Grays", alpha = 0.8)) +
  labs(title = "Variance Decomposition by Variable",
       x = "", y = "Percentage of Total Variance") +
  theme_minimal()

print(plot)

plot_counter <- 1
prefix <- "ICC"
plot_name <- sprintf("%s_%d_icc_variance_decomposition_2.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1
```

For the variables that are identical for both partners within a dyad, such as `PA_similarity_act`, the variance was decomposed into dyad-level stability versus momentary fluctuations. The results show a low dyad-level ICC (ICC=0.04), which reveals that actual emotional similarity is not a particularly stable characteristic of the couples in this sample. Instead, the vast majority of the variance in actual similarity is driven by transient, time-specific co-fluctuations, suggesting that partners move in and out of emotional alignment throughout the sampling period.

# BG

```{r BG_setup}
# load data
bg_data   <- readRDS(file.path(dir_data_ana, "bg_ana.rds"))
demographics <- readRDS(file.path(dir_data_ana, "demographics.rds"))
# load data configuration file
load_config("BG")
```

## Descriptive statistics (demographics)

```{r demographics}

plot_counter <- 1
prefix <- "Demo"

# 1. AGE (numeric)
header("Age")
print(summary(demographics$age))
cat(sprintf("Missing age: %d\n", sum(is.na(demographics$age))))
plot <- demographics %>%
  filter(!is.na(age)) %>%
  ggplot(aes(x = age)) +
  geom_histogram(binwidth = 2) +
  labs(title = "Age distribution", x = "Age (years)", y = "Count") +
  theme_minimal()
plot
plot_name <- sprintf("%s_%d_age_hist.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1

# 2. GENDER
header("Gender")
labels <- get_label(demographics, "gender")
print(freq_table(labels))
plot <- plot_bar_categorical(labels, "Gender Distribution", "Gender")
plot
plot_name <- sprintf("%s_%d_gender_bar.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1

# 3. NATIONALITY
header("Nationality")
labels <- get_label(demographics, "nat")
print(freq_table(labels))
plot <- plot_bar_categorical(labels, "Nationality Distribution", "Nationality")
plot
plot_name <- sprintf("%s_%d_nationality_bar.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1

# 4. ETHNICITY
header("Ethnicity")
labels <- get_label(demographics, "etn")
print(freq_table(labels))
plot <- plot_bar_categorical(labels, "Ethnicity Distribution", "Ethnicity")
plot
plot_name <- sprintf("%s_%d_ethnicity_bar.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1

# 5. EDUCATION
header("Education")
labels <- get_label(demographics, "edu")
print(freq_table(labels))
plot <- plot_bar_categorical(labels, "Eduction Level Distribution", "Education")
plot
plot_name <- sprintf("%s_%d_education_bar.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1

# 6. CHILDREN
header("Children")
labels <- get_label(demographics, "child")
print(freq_table(labels))
plot <- plot_bar_categorical(labels, "Children (Yes/No)", "Children?")
plot
plot_name <- sprintf("%s_%d_chile_bar.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1

# 7. LIVING TOGETHER
header("Living together")
labels <- get_label(demographics, "liv_tog")
print(freq_table(labels))
plot <- plot_bar_categorical(labels, "Living Together? (Yes/No)", "Living Together?")
plot
plot_name <- sprintf("%s_%d_liv_tog_bar.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1

# 8. RELATIONSHIP DURATION (Numeric - months)
header("Relationship duration (months)")
print(summary(demographics$rel_dur))
cat(sprintf("Missing relationship duration: %d\n", sum(is.na(demographics$rel_dur))))
plot <- demographics %>%
  filter(!is.na(rel_dur)) %>%
  ggplot(aes(x = rel_dur)) +
  geom_histogram(binwidth = 2) +
  labs(title = "Relationship Duration Distribution", x = "Duration (Months)", y = "Count") +
  theme_minimal()
plot
plot_name <- sprintf("%s_%d_rel_dur_hist.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1

# 9. MEDICATION
header("Medication")
labels <- get_label(demographics, "med")
print(freq_table(labels))
plot <- plot_bar_categorical(labels, "Medication Use? (Yes/No)", "Living Together?")
plot
plot_name <- sprintf("%s_%d_med_bar.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1
```

## DCI descriptive statistics

### DCI scores histogram

```{r dci_scores_histogram}

plot_counter <- 1
prefix <- "DCI"

plot_dci_total_hist <- bg_data %>%
  filter(!is.na(dci_total)) %>%
  ggplot(aes(x = dci_total)) +
  geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
  labs(title = "Distribution of Total DCI Scores", 
       subtitle = "Sum of 30 items (Negative items reversed)",
       x = "Total Score", y = "Count") +
  theme_minimal()
plot_dci_total_hist
plot_name <- sprintf("%s_%d_dci_score_hist.png", prefix, plot_counter)
save_plot(plot_dci_total_hist, plot_name)
plot_counter <- plot_counter + 1
```

### Reliability of DCI scale

```{r dci_reliability}

# A. Cronbach's Alpha
alpha_res <- bg_data %>%
    select(all_of(scale_DCI)) %>%
    alpha()
cat(sprintf("Cronbach's alpha for DCI scale = %f", alpha_res$total$raw_alpha))

# B. McDonald's Omega
omega_res <- bg_data %>%
    select(all_of(scale_DCI)) %>%
    omega(nfactors = 3, plot = TRUE)
omega_res <- suppressMessages(omega(bg_data %>% select(all_of(scale_DCI)), 
                                             nfactors = 3, plot = FALSE))
cat(sprintf("\nMcDonald's omega for DCI scale = %f", omega_res$omega.tot))
```

### DCI dyadic correlation

```{r dci_dyadic_correlation}

# Create dataframe with dci_total and we_ness of partners in same row
dci_dyadic <- bg_data %>%
  mutate(role = if_else(person < 700, "P1", "P2")) %>%
  select(dyad, role, dci_total, we_ness) %>%
  group_by(dyad) %>%
  pivot_wider(
    names_from = role, 
    values_from = c(dci_total, we_ness)
    )

# Calculate Correlation
cor_total <- cor(dci_dyadic$dci_total_P1, dci_dyadic$dci_total_P2, use = "complete.obs")
header(sprintf("Dyadic Correlation (Total DCI): r = %.3f", cor_total))

# Plot Scatter (Visualizing the correlation)
plot_dci_total_corr <- dci_dyadic %>%
  ggplot(aes(x = dci_total_P1, y = dci_total_P2)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", color = "blue", se = FALSE) +
  labs(title = "Dyadic Correlation: Total DCI",
       subtitle = paste0("Pearson r = ", round(cor_total, 3)),
       x = "Partner 1 Score", y = "Partner 2 Score") +
  theme_minimal()
plot_dci_total_corr

# Save plot
plot_name <- sprintf("%s_%d_dci_total_scatter.png", prefix, plot_counter)
save_plot(plot_dci_total_corr, plot_name)
plot_counter <- plot_counter + 1

```

## We-ness descriptive statistics

### We-ness scores histogram and differences between partners

```{r weness_scores_histogram}

plot_counter <-1
prefix <- "WENESS"

# Histogram of We-ness
plot_weness_hist <- bg_data %>%
  filter(!is.na(we_ness)) %>%
  ggplot(aes(x = we_ness)) +
  geom_histogram(binwidth = 0.1, fill = "forestgreen", color = "white") +
  labs(title = "Distribution of 'We-ness' Scores", 
       subtitle = "Average of 22 Positive items (No negative items)",
       x = "We-ness Score", y = "Count") +
  theme_minimal()
plot_weness_hist

# Save Histogram
plot_name <- sprintf("%s_%d_weness_hist.png", prefix, plot_counter)
save_plot(plot_weness_hist, plot_name)
plot_counter <- plot_counter + 1

# Calculate Absolute Difference in We-ness
dci_dyadic <- dci_dyadic %>%
  mutate(
    weness_diff = abs(we_ness_P1 - we_ness_P2)
  )

cat("Summary of Absolute Differences in We-ness between Partners:\n")
summary(dci_dyadic$weness_diff) %>% print()

# Plot Histogram of Differences
plot_weness_diff <- dci_dyadic %>%
  filter(!is.na(weness_diff)) %>%
  ggplot(aes(x = weness_diff)) +
  geom_histogram(binwidth = 0.1, fill = "firebrick", color = "white") +
  labs(title = "Distribution of We-ness Differences", 
       subtitle = "Absolute difference between Partner 1 and Partner 2",
       x = "Difference Score (|P1 - P2|)", y = "Count of Couples") +
  theme_minimal()
plot_weness_diff

# Save plot
plot_name <- sprintf("%s_%d_weness_diff_hist.png", prefix, plot_counter)
save_plot(plot_weness_diff, plot_name)
plot_counter <- plot_counter + 1
```

### Reliability of we-ness construct

```{r weness_reliability}

# A. Cronbach's Alpha
alpha_res <- bg_data %>%
    select(all_of(DCI_weness_items)) %>%
    alpha()
cat(sprintf("Cronbach's alpha for we-nesse = %f\n", alpha_res$total$raw_alpha))

# B. McDonald's Omega
omega_res <- bg_data %>%
    select(all_of(DCI_weness_items)) %>%
    omega(nfactors = 3, plot = FALSE)
cat(sprintf("\nMcDonald's omega for we-ness = %f\n\n", omega_res$omega.tot))
print(omega_res)
omega.diagram(omega_res, main = "We-ness Factor Structure")
print(diagram)

# Save diagram
plot_name <- sprintf("%s_%d_weness_factor_structure.png", prefix, plot_counter)
save_base_plot(omega.diagram(omega_res, main = "We-ness Factor Structure"), plot_name)
plot_counter <- plot_counter + 1

```

A hierarchical factor analysis (Schmid-Leiman bifactor solution) was conducted to evaluate the internal structure of the We-Ness construct. The analysis supports a hierarchical structure consisting of one General Factor (*g*) and three distinct specific group factors, confirming that "We-Ness" is an overarching construct that encompasses both giving and receiving support.

**1. Reliability and Global Fit** The scale demonstrates excellent total reliability (ωt​=0.87), indicating that the items form a cohesive total scale. The Hierarchical Omega (ωh​=0.42) indicates that the General Factor accounts for a substantial portion (42%) of the common variance.

**2. Factor Interpretation** The variance is partitioned into a general "We-Ness" trait and three specific sub-domains corresponding to the provider of the support and the focus of communication:

-   **General Factor (*g*) — "Global We-Ness":** This factor represents the fundamental dyadic cohesion. It captures the shared tendency of the couple to function as a unit, regardless of the specific coping strategy employed.

-   **Factor 1 — Supportive Dyadic Coping (Perceived Partner Support):** *(Items: DCI 5, 6, 8, 9, 13, 14)* This factor captures the variance associated with **receiving support**. These items describe the individual's perception of their partner's active efforts to help them (e.g., "My partner shows empathy and understanding to me" ). It represents the "You help me" dimension.

-   **Factor 2 — Stress Communication Balance:** *(Items: Positive loading on 1–4, Negative loading on 16–19)* This bipolar factor captures the **communicative dynamic** of the dyad. It differentiates between focusing on one's own stress (Self-Communication, items 1–4) versus focusing on the partner's stress (Partner-Communication, items 16–19). This represents the "Who is talking?" dimension.

-   **Factor 3 — Supportive Dyadic Coping (Own Provided Support):** *(Items: DCI 20, 21, 24, 28, 29, 30)* This factor captures the variance associated with **providing support**. Unlike Factor 1, these items reflect the individual's own active efforts to support their partner (e.g., "I show empathy and understanding to my partner" ). It represents the "I help you" dimension.

### We-ness dyadic correlation

```{r weness_dyadic_correlation}

# Dyadic Correlation We-ness
cor_weness <- cor(dci_dyadic$we_ness_P1, dci_dyadic$we_ness_P2, use = "complete.obs")
header(sprintf("Dyadic Correlation (We-ness): r = %.3f", cor_weness))

# Plot Scatter We-ness
plot_weness_corr <- dci_dyadic %>%
  ggplot(aes(x = we_ness_P1, y = we_ness_P2)) +
  geom_point(alpha = 0.6, color = "darkgreen") +
  geom_smooth(method = "lm", color = "black", se = FALSE) +
  labs(title = "Dyadic Correlation: We-ness",
       subtitle = paste0("Pearson r = ", round(cor_weness, 3)),
       x = "Partner 1 We-ness", y = "Partner 2 We-ness") +
  theme_minimal()
plot_weness_corr

# Save plot
plot_name <- sprintf("%s_%d_weness_scatter.png", prefix, plot_counter)
save_plot(plot_weness_corr, plot_name)
plot_counter <- plot_counter + 1

```

# VMR

```{r VMR_setup}
# load data
vmr_data  <- readRDS(file.path(dir_data_ana, "vmr_ana.rds"))
# load data configuration file 
load_config("VMR")
```

# POST

```{r POST_setup}
# load data
post_data  <- readRDS(file.path(dir_data_ana, "post_ana.rds"))
# load data configuration file 
load_config("POST")
```
