---
title: "Data Descriptives"
---

```{r setup}
#| include: false

# 1. Load packages, paths and data configurations
source(here::here("R", "00_setup.R"))

# 2. Set a global theme for plots
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Demographics

```{r demographics setup}
# load data
demographics <- readRDS(file.path(dir_data_ana, "demographics.rds"))
# plot initialisation
plot_counter <- 1
prefix <- "Demo"
```

## Age

```{r age}

# AGE = numeric
print(summary(demographics$age))
cat(sprintf("Missing age: %d\n", sum(is.na(demographics$age))))
plot <- demographics %>%
  filter(!is.na(age)) %>%
  ggplot(aes(x = age)) +
  geom_histogram(binwidth = 2) +
  labs(title = "Age distribution", x = "Age (years)", y = "Count") +
  theme_minimal()
plot
plot_name <- sprintf("%s_%d_age_hist.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1

```

## Gender

```{r gender}
# gender = factor
labels <- get_label(demographics, "gender")
print(freq_table(labels))
plot <- plot_bar_categorical(labels, "Gender Distribution", "Gender")
plot
plot_name <- sprintf("%s_%d_gender_bar.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1
```

## Nationality

```{r nationality}
# nationality = factor
labels <- get_label(demographics, "nat")
print(freq_table(labels))
plot <- plot_bar_categorical(labels, "Nationality Distribution", "Nationality")
plot
plot_name <- sprintf("%s_%d_nationality_bar.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1

```

## Ethnicity

```{r ethnicity}
# ethnicity = factor
labels <- get_label(demographics, "etn")
print(freq_table(labels))
plot <- plot_bar_categorical(labels, "Ethnicity Distribution", "Ethnicity")
plot
plot_name <- sprintf("%s_%d_ethnicity_bar.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1
```

## Education

```{r education}
# education = factor
labels <- get_label(demographics, "edu")
print(freq_table(labels))
plot <- plot_bar_categorical(labels, "Eduction Level Distribution", "Education")
plot
plot_name <- sprintf("%s_%d_education_bar.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1
```

## Children

```{r children}
# children = factor
labels <- get_label(demographics, "child")
print(freq_table(labels))
plot <- plot_bar_categorical(labels, "Children (Yes/No)", "Children?")
plot
plot_name <- sprintf("%s_%d_chile_bar.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1
```

## Living Together

```{r living together}
# living together = factor
labels <- get_label(demographics, "liv_tog")
print(freq_table(labels))
plot <- plot_bar_categorical(labels, "Living Together? (Yes/No)", "Living Together?")
plot
plot_name <- sprintf("%s_%d_liv_tog_bar.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1
```

## Relationship Duration

```{r relationship duration}
# relationship duration = numeric
print(summary(demographics$rel_dur))
cat(sprintf("Missing relationship duration: %d\n", sum(is.na(demographics$rel_dur))))
plot <- demographics %>%
  filter(!is.na(rel_dur)) %>%
  ggplot(aes(x = rel_dur)) +
  geom_histogram(binwidth = 2) +
  labs(title = "Relationship Duration Distribution", x = "Duration (Months)", y = "Count") +
  theme_minimal()
plot
plot_name <- sprintf("%s_%d_rel_dur_hist.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1
```

## Medication

```{r medication}
# medication = factor
labels <- get_label(demographics, "med")
print(freq_table(labels))
plot <- plot_bar_categorical(labels, "Medication Use? (Yes/No)", "Living Together?")
plot
plot_name <- sprintf("%s_%d_med_bar.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1
```

------------------------------------------------------------------------

# ESM

```{r ESM_setup}
#| include: false
# load data
esm_data  <- readRDS(file.path(dir_data_ana, "esm_ana.rds"))
# load data configuration file 
load_config("ESM")
# plot initialisation
plot_counter <- 1
prefix <- "ESM"
```

## Missingness

```{r missingness}
esm_data %>% select(NA_own, PA_own, love, perc_resp, neg_gen, pos_gen,
                    PA_similarity_act, NA_similarity_act, PA_similarity_perc,
                    NA_similarity_perc) %>%
             miss_var_summary()
plot <- gg_miss_var(esm_data %>% select(NA_own, PA_own, love, perc_resp, neg_gen, pos_gen), show_pct = TRUE) + labs(title = "ESM Missingness per Variable")
# save plot
plot_name <- sprintf("%s_%d_Missingness.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1
# print plot
plot
```

The variables 'neg_gen' (about the occurrence of any (general) negative event since previous beep) and 'pos_gen' (any positive event since previous beep) are very high in missingness (91,4%). This is because these multiple choice questions were only asked on beeps were partner was not present (i.e., in 31,9% of beeps) AND there has not been any contact with partner since previous beep (reducing the proportion of beeps that were asked the neg_gen and pos_gen items to 8,53%). This explains the full missingness.

The moderation by event valence (positive vs negative) that will be analyzed (see pre-registration step 3b of study component 1) will therefore have more limited data.

The perceived similarity measures are missing for 18,3% of the beeps. This is the same level as for NA_own, PA_own, etc., and is ≈ 1 - compliance . The actual similarity measures are missing for 27,6%, which is higher because the calculating depends on both partners answering the underlying items.

## Repeated measures correlations

As in @bakdashRepeatedMeasuresCorrelation2017 we look at repeated measure correlations as descriptive measures but refrain from using these for inference (see below). The questions in this section is **"When one person in the couple deviates from their average, does the other person deviate from their average in the same direction?"**

We will use the function `rmcorr::rmcorr()`, which calculates the overall within-individual relationship among paired measures assessed on two or more occasions. The first argument of `rmcorr()` specifies the grouping over which commonality is sought:

-   When using using 'person' as the first argument in `rmcorr::rmcorr()`, this is the correlation that is common across persons between two variables' deviations from their baseline (so for example own PA and own NA, what is the 'shared' or 'pooled' correlation of the within-person deviations from baseline?).

-   When using 'dyad' as the first argument in `rmcorr::rmcorr()`, this is the correlation between two partner's deviations from baseline that is common across dyads. This treats the correlation as characteristic of the relationship, not of the individuals.

As mentioned, we will refrain from using the results from `rmcorr()` for inference. When grouping dyadic data by person in `rmcorr()`, the model treats partners as independent entities, ignoring the shared variance between them (because they belong to the same dyad). This violation of independence artificially inflates the degrees of freedom (df) and deflates p-values, making significance tests overly optimistic. However, the descriptive correlation coefficient (r) accurately represents the common within-person association across the sample.

### Personal: Intrapersonal dynamics (individual level)

#### Own PA vs own NA

($r_{rm} = -0.60, p < .001$)

```{r ESM_rmcorr_PA_own_NA_own}
#| warning: false
rmc <- rmcorr(person, PA_own, NA_own, esm_data)
# results
create_rmc_plot_title(rmc, "ESM - ")
rmc
plot(rmc, main = plot_title)
# save plot
plot_name <- sprintf("%s_%d_rmcorr_pa_own_na_own.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1

# for illustration: the same but on already centered vars, so one regression line
rmc <- rmcorr(person, cPA_own, cNA_own, esm_data)
# results
create_rmc_plot_title(rmc, "ESM - ")
rmc
plot(rmc, main = plot_title)
# save plot
plot_name <- sprintf("%s_%d_rmcorr_cpa_own_cna_own.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1

```

The common within-person correlation between an individual's own PA and NA is strong and negative. The \`rmcorr\` plot shows consistent negative slopes across participants, suggesting that as a person's positive affect increases relative to their mean, their negative affect decreases.

#### Own PA vs perceived partner's PA

($r_{rm} = 0.36, p < .001$)

```{r ESM_rmcorr_PA_own_PA_perceived_partner}
#| warning: false
rmc <- rmcorr(person, cPA_own, cPA_part_perc, esm_data)
create_rmc_plot_title(rmc, "ESM - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_pa_own_pa_partner.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1
```

While there was significant correlation between partners' actual Positive Affect ($r_{rm}​=.23$), participants' perceptions of their partners' PA were more strongly correlated with their own internal PA ($r_{rm}​=.36$), suggesting a degree of perceptual projection and/or shared context effects.

#### Own NA vs perceived partner's NA

($r_{rm} = 0.36, p < .001$)

```{r ESM_rmcorr_NA_own_NA_perceived_partner}
#| warning: false
rmc <- rmcorr(person, cNA_own, cNA_part_perc, esm_data)
create_rmc_plot_title(rmc, "ESM - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_na_own_na_partner_perc.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1
```

#### Actual PA similarity vs perceived PA similarity

($r_{rm} = 0.62, p < .001$)

How well tuned is a person to the "state of the union"? Even though 'Actual Similarity' is a dyad-level variable repeated for both partners, the *perception* of that reality is an individual cognitive process. Grouping by person allows the model to capture how well each individual tracks the shared reality.

```{r ESM_rmcorr_actual_perceived_sim_PA}
#| warning: false
rmc <- rmcorr(person, PA_similarity_act, PA_similarity_perc, esm_data)
create_rmc_plot_title(rmc, "ESM - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_pa_sim_act_pa_sim_perc.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1
```

High, but probably can be because of the common element of own PA in both similarity measures.

#### Actual NA similarity vs perceived NA similarity

($r_{rm} = 0.49, p < .001$)

```{r ESM_rmcorr_actual_perceived_sim_NA}
#| warning: false
rmc <- rmcorr(person, NA_similarity_act, NA_similarity_perc, esm_data)
create_rmc_plot_title(rmc, "ESM - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_na_sim_act_na_sim_perc.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1
```

High, but part of the association may be mechanical because both indices share the actor’s own affect component (and both combine elevation and mismatch).

### Dyadic: Interpersonal covariation

#### Own PA vs Partner PA

($r_{rm} = 0.23, p < .001$)

When looking at two variables that come from different partners (e.g., own PA and patner's PA) we will use grouping by dyad.

Dyad members are treated as indistinguishable. The correlation between Partner A and Partner B is mathematically identical to the correlation between Partner B and Partner A, including both partners' rows for a dyad-level analysis results in data duplication. This redundancy artificially doubles the degrees of freedom and inflates statistical significance without adding unique information about the couple's synchrony. Limiting the dataset to one row per dyad ensures that the analysis reflects the true number of independent observations. When grouping by 'person' (as above), all rows (both partners) are retained because the analysis targets individuals; this introduces statistical non-independence but removing half the subjects would bias the estimate of the average person's experience.

```{r ESM_rmcorr_PA_own_PA_partner}
#| warning: false

# filter to have only one line per dyad.
esm_data_dyad <- esm_data %>% filter(person < 700)

rmc <- rmcorr(dyad, cPA_own, cPA_part_act, esm_data_dyad)
create_rmc_plot_title(rmc, "ESM - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_pa_own_pa_partner.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1
```

This result suggests a modest syncing of actual PA.

#### Own NA vs Partner NA

($r_{rm} = 0.22, p < .001$)

```{r ESM_rmcorr_NA_own_NA_partner}
#| warning: false
rmc <- rmcorr(dyad, cNA_own, cNA_part_act, esm_data_dyad)
create_rmc_plot_title(rmc, "ESM - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_pa_own_na_partner.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1
```

This result suggests a modest syncing of actual NA.

## Intraclass Correlation Coefficient (ICC)

How much of the variance in the data is due to differences between people or dyads rather than within people or dyads (like moment to moment fluctuations). We calculated ICCs using a model-based approach rather than the `psych` package to properly account for the 3-level nested structure of our ESM data (Moments within Persons within Dyads).

```{r ESM_icc}

# Variables to analyze
target_vars <- c("PA_own", "NA_own", "NA_part_perc", "PA_part_perc", "love", "perc_resp", "PA_similarity_perc", "NA_similarity_perc")

# Run the loop
icc_table <- target_vars %>%
  map_dfr(function(v) {
    # Construct the 3-level formula: Moments < Person < Dyad
    formula_str <- as.formula(paste0(v, " ~ 1 + (1 | dyad / person)"))
    # Fit the model
    model <- lmer(formula_str, data = esm_data)
    # Extract ICCs and format into a single row, icc from performance package
    icc_data <- icc(model, by_group = TRUE)
    # Create a clean data frame for this variable
    d_icc <- icc_data$ICC[icc_data$Group == "dyad"]
    p_icc <- icc_data$ICC[icc_data$Group == "person:dyad"] 
    data.frame(
      Variable = v,
      ICC_Dyad = d_icc,
      ICC_Person = p_icc,
      ICC_Moment = 1- d_icc - p_icc
    )
  })

# View the results
print(icc_table)

# Calculate the Residual (Momentary) variance
plot_data <- icc_table %>%
  mutate(Momentary = 1 - (ICC_Dyad + ICC_Person)) %>%
  pivot_longer(cols = c(ICC_Dyad, ICC_Person, Momentary), 
               names_to = "Level", values_to = "Proportion")

plot <- ggplot(plot_data, aes(x = Variable, y = Proportion, fill = Level)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_discrete(type = hcl.colors(4, palette = "Grays", alpha = 0.8)) +
  labs(title = "ESM - Variance Decomposition by Variable",
       x = "", y = "Percentage of Total Variance") +
  theme_minimal()

# show plot
print(plot)

# save plot
plot_name <- sprintf("%s_%d_icc_variance_decomposition_1.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1

```

The ICC analysis for the three-level variables indicates that a substantial portion of the variance in affect and relationship perceptions is attributable to momentary, within-person fluctuations, as both dyad-level and person-level ICCs remain relatively low to moderate. Perceived responsiveness stands out with the highest dyad-level stability (ICC=0.35), suggesting it is more strongly characterized by shared couple-level baselines than the other variables. In contrast, the similarity perception variables (`PA_similarity_perc` and `NA_similarity_perc`) show very low ICCs at both the dyad and person levels, indicating that how similar partners perceive themselves to be is a highly dynamic state that changes significantly from moment to moment rather than being a stable trait of the individual or the dyad. Note that the Momentary ICC also contains measurement error.

Technical note:\
$Y∼1+(1∣dyad/person)$ , decomposes the total variance into three independent sources:

1.  $\sigma_{dyad}^2$​: Variance of the dyad-level random intercept (between-dyad differences).

2.  $\sigma_{person}^2$​: Variance of the person-level random intercept nested within dyads (stable differences between partners).

3.  $\sigma_{error}^2$​: The residual variance (moment-to-moment or within-person fluctuations).

4.  The ICC is the proportion of variance of the total variance:

    -   **ICC_Dyad:** Stable differences between couples

    -   **ICC_Person:** Stable differences between individuals within those couples

    -   **ICC_Moment:** The "within-person" variance. This captures how much a person deviates from their own average throughout the day (includes error).

The **total variance** is the sum: $\sigma_{total}^2 = \sigma_{dyad}^2 + \sigma_{person}^2 + \sigma_{error}^2$.

Since the `_act` similarity scores are identical for both partners, they are mathematically "dyad-level" constants at each time point, so their data only has two levels of meaningful variation.

```{r ESM_icc_actual_similarity}

dyad_level_vars <- c("PA_similarity_act", "NA_similarity_act")

# Simplified 2-level loop (Moments < Dyad)
icc_table_dyad_only <- dyad_level_vars %>%
  map_dfr(function(v) {
    formula_str <- as.formula(paste0(v, " ~ 1 + (1 | dyad)"))
    model <- lmer(formula_str, data = esm_data)
    icc_data <- icc(model) # Standard ICC for 2-level
    data.frame(
      Variable = v,
      ICC_Dyad = icc_data$ICC_adjusted,
      ICC_Person = "-" # Explicitly NA as it's not applicable
    )
  })

print(icc_table_dyad_only)

plot_data <- icc_table_dyad_only %>%
  mutate(Momentary = 1 - (ICC_Dyad)) %>%
  pivot_longer(cols = c(ICC_Dyad, Momentary), 
               names_to = "Level", values_to = "Proportion")

plot <- ggplot(plot_data, aes(x = Variable, y = Proportion, fill = Level)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_discrete(type = hcl.colors(4, palette = "Grays", alpha = 0.8)) +
  labs(title = "Variance Decomposition by Variable",
       x = "", y = "Percentage of Total Variance") +
  theme_minimal()

print(plot)

plot_name <- sprintf("%s_%d_icc_variance_decomposition_2.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1
```

For the variables that are identical for both partners within a dyad, such as `PA_similarity_act`, the variance was decomposed into dyad-level stability versus momentary fluctuations. The results show a low dyad-level ICC (ICC=0.04), which reveals that actual emotional similarity is not a particularly stable characteristic of the couples in this sample. Instead, the vast majority of the variance in actual similarity is driven by transient, time-specific co-fluctuations, suggesting that partners move in and out of emotional alignment throughout the sampling period.

## Trajectory Plots

Change the variables 'participant' and 'variable' in the code below to choose which trajectory for which participants is plotted.

```{r trajectory plots}

participants <- c(154, 854) # change this vector to participants to be plotted
variable <- "love"     # change this variable to the variable to be plottee
plot_trajectory(esm_data, variable, participants)
 
```

## Similarity indices

Theoretically, the similarity indices can range from -200 to +100:

-   The -200 would be when one partner has baseline of +100 and rates at 0 (so -100 deviation) while the other partner has baseline of 0 and rates at +100 (so deviation of +100), giving a mean elevation of 0, while their distance is at 200.

-   The +100 would be when both partners' baseline would be at 0 and they both rate at +100 (so deviation of +100), bringing the mean elevation to 100 and the distance to 0.

```{r base_theoretical_similarity_surface}

# Theoretical ranges -100 to 100 because of centering
x <- seq(-100, 100, length.out = 200) # 'resolution' of 200
y <- seq(-100, 100, length.out = 200)

# function Z = (X + Y)/2 - abs(X - Y)
similarity <- function(X, Y) {
  elevation <- (X + Y) / 2
  mismatch <- abs(X - Y)
  return(elevation - mismatch)
}

# create z matrix, 'outer' applies similarity function to all combinations.
z <- outer(x, y, similarity)

# dynamic plot using plotly::plot_ly()
base_surface <- plot_ly(x = ~x, y = ~y, z = ~z) %>%
  add_surface(
    colorbar = list(title = "Similarity"),
    colorscale = "Viridis",
    opacity = 0.8
  ) 
```

### Theoretical similarity surface

```{r theoretical_surface}
base_surface %>%
  layout(
    title = list(text = "Similarity Surface: S = (X+Y)/2 - |X-Y|", y = 0.95),
    scene = list(
      xaxis = list(title = "Affect 1 (X)"),
      yaxis = list(title = "Affect 2 (Y)"),
      zaxis = list(title = "Similarity Index (S)"),
      # initial camera angle
      camera = list(eye = list(x = 1.5, y = 1.5, z = 1.2))
    )
  )
```

In practice, these edge cases will not appear. In the data, the similarity indices ranges are between approximately -150 and +70, and the bulk of the data has similarity between -75 and +50.

### Similarity in the sample

#### Ranges and distribution

```{r practical_ranges}

sim_cols <- c("PA_similarity_act", "NA_similarity_act", 
              "PA_similarity_perc", "NA_similarity_perc")

draw_violin_plot <- function() {
  vioplot(esm_data[, sim_cols],
            main = "Range and Distribution of Affect Similarity",
            col = c("#A6CEE3", "#FB9A99", "#A6CEE3", "#FB9A99"),
            names = c("PA\nActual", "NA\nActual", "PA\nPerceived", "NA\nPerceived"),
            areaEqual = FALSE,
            rectCol = "gray30", lineCol = "gray30", border = "gray30"
    )
    title(ylab = "Similarity (S = M - D)", line = 2)
    abline(h = 0, col = "black", lty = 2, lwd = 1.5)
}
draw_violin_plot()

plot_name <- sprintf("%s_%d_violin.png", prefix, plot_counter)
save_base_plot(draw_violin_plot(), plot_name)
plot_counter <- plot_counter + 1

# numerical ranges
cat("\nRanges (Min/Max):\n")
print(sapply(esm_data[, sim_cols], range, na.rm = TRUE))
```

#### Actual PA similarity

```{r PA_act_obs_vs_theo}
#| warning: false

base_surface %>%
  add_layer(
    data = esm_data, # %>% filter(person == 703),
    x = ~cPA_own,
    y = ~cPA_part_act,    
    z = ~PA_similarity_act,     
    title = "PA Actual Similarity: Observed vs. Theoretical",
    x_lab = "Actor PA (Cent.)",
    y_lab = "Partner PA (Cent.)",
    z_lab = "Actual PA Similarity Index"
  )
```

#### Actual NA similarity

```{r NA_act_obs_vs_theo}
#| warning: false

base_surface %>%
  add_layer(
    data = esm_data,
    x = ~cNA_own,
    y = ~cNA_part_act,    
    z = ~NA_similarity_act,   
    title = "NA Actual Similarity: Observed vs. Theoretical",
    x_lab = "Actor NA (Cent.)",
    y_lab = "Partner NA (Cent.)",
    z_lab = "Actual NA Similarity Index"
  )
```

#### Perceived PA similarity

```{r PA_perc_obs_vs_theo}
#| warning: false

base_surface %>%
  add_layer(
    data = esm_data, #%>% filter(dyad == 3),
    x = ~cPA_own,
    y = ~cPA_part_perc,    
    z = ~PA_similarity_perc,   
    title = "PA Perceived Similarity: Observed vs. Theoretical",
    x_lab = "Actor PA (Cent.)",
    y_lab = "Perceived Partner PA (Cent.)",
    z_lab = "Perceived PA Similarity Index"
  )
```

#### Perceived NA similarity

```{r NA_perc_obs_vs_theo}
#| warning: false

base_surface %>%
  add_layer(
    data = esm_data,
    x = ~cNA_own,
    y = ~cNA_part_perc,    
    z = ~NA_similarity_perc,   
    title = "NA Perceived Similarity: Observed vs. Theoretical",
    x_lab = "Actor NA (Cent.)",
    y_lab = "Perceived Partner NA (Cent.)",
    z_lab = "Perceived NA Similarity Index"
  )
```

## Time trend sensitivity analysis

### Time trend in momentary love

From the preregistration: *"Rather than including time covariates in all primary models, we will test for systematic time trends in momentary love across the ESM period (e.g., habituation or drift across the two-week period), for example by fitting* $𝐿_{𝑚𝑝𝑑} = 𝑏_{0,𝑝𝑑} + 𝛽_𝑇𝑇_𝑚 + 𝜀_{𝑚𝑝𝑑}$*, with* $𝑇_𝑚$ *the chronological order of beeps, centered at the study midpoint."* In this model, $𝐿_{𝑚𝑝𝑑}$ denotes the momentary love felt for the partner by person 𝑝 in dyad 𝑑 at moment 𝑚, and $b_{0,pd} = \beta_0 + v_{0,d}+u_{0, pd}$, the sum of the fixed intercept, the dyad-level random intercept and the person-level random intercept.

The preregistration continues : *"When indicated, we will re-estimate models including a linear time covariate (and, where feasible, day-in-study and time-of-day) as an additional robustness check. As an additional diagnostic, we will also inspect whether the similarity indices show systematic time trends."*

```{r create_centered_time_trend}
# Create chronological order of beeps centered at study midpoint (T_m)
# --> the Intercept then represents the average love at the middle of the study
data_w_time <- esm_data %>%
  mutate(
    time_centered = beep - mean(beep, na.rm = TRUE),
  )
```

```{r love_time_trend}
# preregistered model : Love_mpd = b0_pd + beta*Time + error
# We nest Person within Dyad for the random intercept (b0_pd).

model_love_time_trend <- lmer(love ~ time_centered + (1 | dyad/person), data = data_w_time)
summary(model_love_time_trend)
```

The fixed effect for time_centered is slightly, but significantly positive, indicating a minor drift upwards (love increases over time). Because of the large dataset, power is large and tiny fluctuations can be detected. A trend of 0.039 for every next beep, amounts to 0.2 (on a 0-100 slider scale) per weekday and 0.4 per weekend day. Over the full study duration, this is an increase of 2.7 points. This is indistinguishable from measurement noise (e.g., trying to hit 75 and sliding to 77). **The trend is thus statistically significant but practically meaningless.**

### Time trend in similarity

In the preregistration, we also write: "*As an additional diagnostic, we will also inspect whether the similarity indices show systematic time trends."*

```{r data_prep_2}
# only retain one row per dyad x beep to avoiud double counting
data_similarity <- data_w_time %>% filter(person < 700)
```

When modelling the actual similarity indices (which are symmetric within a couple) in function of time, we use one line in the dataset for every dyad, since using both rows would treat both partners as independent observations and artificially inflate degrees of freedom and significance.

#### Actual PA similarity

```{r actual_PA_similarity_time_trend}
# "Do partners become more/less similar in PA over the 2 weeks?"
model_act_PA_sim_time_trend <- lmer(PA_similarity_act ~ time_centered + (1 | dyad), data = data_similarity)
summary(model_act_PA_sim_time_trend)
```

Dyads are slightly increasing in actual PA similarity over the course of the two weeks $\beta = 0.039, p < .001$. Again, this is a very small effects, given the theoretical (-200, +100) and practical (-153, +62) range of this index. One beep further, actual PA similarity increases by 0.039, which would amount to an increase of 3.51 over the ESM period.

Note the high residual variance (454.21), compared to the smaller variances for the dyad random intercept. This is because, as we've seen when calculating ICC, actual PA similarity is a highly fluctuating state. This residual variance also means that the slight upward time trend that was detected is a 'slow tide' beneath 'huge waves'.

#### Actual NA similarity

```{r actual_NA_similarity_time_trend}
# "Do partners become more/less similar in NA over the 2 weeks?"
model_act_NA_sim_time_trend <- lmer(NA_similarity_act ~ time_centered + (1 | dyad), data = data_similarity)
summary(model_act_NA_sim_time_trend)
```

Dyads are slightly increasing in actual NA similarity over the course of the two weeks $\beta = 0.024, p < .001$. Again, this is a very small effects, given the theoretical (-200, +100) and practical (-106, +60) range of this index. One beep further, actual NA similarity increases by 0.024, which would amount to an increase of 2.16 over the ESM period.

Again, note the high residual variance (168.70), compared to the smaller variances for the dyad random intercept. This is lower than for actual PA similarity but alsi indicates that actual NA similarity is a fluctuating state. This residual variance also means that the slight upward time trend that was detected is a 'slow tide' beneath 'huge waves'.

#### Perceived PA similarity

For the perceived similarity indices, we use the data from all persons, since the perceived indices are person-specific.

```{r perceived_PA_similarity_time_trend}
# "Does perceived PA similarity increase/decrease over the 2 weeks?"
model_perc_PA_sim_time_trend <- lmer(PA_similarity_perc ~ time_centered + (1 | dyad/person), data = data_w_time)
summary(model_perc_PA_sim_time_trend)
```

Persons within dyads are slightly increasing ($\beta = 0.05, p < .001$) in perceived PA similarity over the course of the two weeks. Again, this is a very small effects, given the theoretical (-200, +100) and practical (-146, +50) range of this index. One beep further, perceived PA similarity increases by 0.05, which would amount to an increase of 4.48 over the ESM period.

Note tagain he high residual variance (408.67), compared to the smaller variances for the person (10) and dyad (6.92) random intercepts. This is because, as we've seen when calculating ICC, perceived PA similarity is a highly fluctuating state. The variance proportions found here largely correspond to what was found in the ICC calculations. This residual variance also means that the slight upward time trend that was detected is a 'slow tide' beneath 'huge waves'.

#### Perceived NA similarity

```{r perceived_NA_similarity_time_trend}
# "Does perceived NA similarity increase/decrease over the 2 weeks?"
model_perc_NA_sim_time_trend <- lmer(NA_similarity_perc ~ time_centered + (1 | dyad/person), data = data_w_time)
summary(model_perc_NA_sim_time_trend)
```

Persons within dyads are showing a marginally significant, very slight increase in perceived NA similarity over the course of the two weeks ($\beta = 0.008, p=.048$). However, this effect is negligible, even compared to the perceived PA similarity findings. One beep further, perceived NA similarity increases by only 0.008, which would amount to a total increase of just 0.74 points over the entire ESM period (90 beeps). Given the theoretical range of the index (-200 to +100) and the practical range of (-105, +71), this trend is practically invisible.

Note again the high residual variance (\~174.6) relative to the person (9.7) and dyad (13.6) random intercept variances. As with PA, perceived NA similarity is a highly fluctuating state rather than a stable trait, although the overall volatility is lower here than for PA (Residual variance of \~175 vs \~409). The 'slow tide' metaphor applies here as well, but in this case, the tide is barely moving at all beneath the waves.

## Autocorrelation analysis

```{r esm_autocorrelation}
#| warning: false

target_vars <- c("cPA_own", "cNA_own", "cPA_part_act", "cNA_part_act", 
                 "cPA_part_perc", "cNA_part_perc", "clove")

# calculate Lag-1 ac per person
# We group by dyad/person to ensure we don't lag across different people
acf_results <- esm_data %>%
  group_by(dyad, person) %>% 
  arrange(dyad, person, beep) %>% # order by time!
  summarise(
    across(all_of(target_vars), 
          ~ cor(., dplyr::lag(.), use = "pairwise.complete.obs"),
           .names = "{.col}"
    ),
    .groups = "drop"
  )

# reshape for easy checking/plotting
acf_long <- acf_results %>%
  pivot_longer(cols = all_of(target_vars), 
               names_to = "Variable", 
               values_to = "Autocorrelation")

# summary
cat("--- Average Lag-1 Autocorrelation (Temporal Dependence) ---\n")
acf_summary <- acf_long %>%
  group_by(Variable) %>%
  summarise(
    Mean_ACF = mean(Autocorrelation, na.rm = TRUE),
    Median_ACF = median(Autocorrelation, na.rm = TRUE),
    SD_ACF = sd(Autocorrelation, na.rm = TRUE)
  )
print(acf_summary)

# Plot
ggplot(acf_long, aes(x = Variable, y = Autocorrelation, fill = Variable)) +
  geom_boxplot(alpha = 0.6, outlier.shape = NA) +
  geom_jitter(width = 0.2, alpha = 0.3, size = 1) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_minimal() +
  labs(
    title = "Distribution of Lag-1 Autocorrelation per Person",
    subtitle = "Higher values indicate stronger temporal dependence (validating AR(1))",
    y = "Lag-1 Correlation (r)",
    x = NULL
  ) +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))
```

The autocorrelation values range from 0.30 (love) to 0.38 (perceived PA similarity), which is a moderate temporal dependence. The SDs are \~ .20–.25, so there is substantial heterogeneity across persons. If this would be 0 then AR(1) would be unnecessary, and if this would be close to 1 there would be more like a random walk.\
Without doing statistical testing, at first sight perception of similarity seems to be 'stickier' (more temporal inert or persistent) than actual similarity.

**Important:** we calculate lag across days, so correlate end-of-day with next-day which might create a discontinuity. Raw ACF, as calculated here, can be influenced by these discontinuities or driven by time-of-day, etc. The real target is the residual autocorrelation after fitting the moddel. Furthermore, this analyses presumes an equal spacing between beeps.

Visual inspection of the lag-1 autocorrelation for the centered variables revealed moderate temporal dependence (​≈.30−.38), indicating that any assumption of independent within-person errors would be violated. Consistent with the preregistration and the recommendations for intensive longitudinal data (@bolgerIntensiveLongitudinalMethods2013 and @snijderstomMultilevelAnalysisIntroduction2012), we will use a first-order autoregressive \[AR(1)\] covariance structure to account for this serial correlation. Given the caveats in the analysis here, we will fir models with and without the AR(1) structure.

```{r ESM unload}
#| include: false
clean_config("ESM")
```

```{r BG_setup}
# load data
bg_data   <- readRDS(file.path(dir_data_ana, "bg_ana.rds"))
# load data configuration file
load_config("BG")
# plot initialisation
plot_counter <- 1
prefix <- "BG"
```

## Correlation analyses

We preregistered: *" We will additionally fit models including the individual (actual or perceived) affect levels as covariates to check whether the effect of similarity changes meaningfully. Specifically, when feasible (i.e., converging), we will add the person-mean centered actor and partner affects and the corresponding average actor and partner affect levels across the ESM period (grand-mean centered) as covariates. In other words, we test whether ‘we feel the same’ predicts love beyond simply ‘I feel good’. **We will also inspect correlations between momentary love and the candidate covariates.** If these correlations are small (e.g., \|r\| \< .30; following the reasoning in Sels et al., 2020), this provides some reassurance even if the larger sensitivity models do not converge."*

The idea is that in the models prediciting momentary love from similarity, the similarity indices can pick up effects of the underlying individual affect levels. See @selsActualPerceivedEmotional2020. This can arguably even be more the case when using our similarity constructs because these contains an elevation term (the mean). This is why we will also inspect the individual component-outcome associations (via correlations) as a heuristic for whether omitting the individual affect levels is likely to matter. If these correlations are low (and we, as Sels, proposed \|r\| \< .30 as a heuristic criterion), this reduces the concern that a similarity effect is merely a disguised level effect.

```{r predictors_correlation_w_love }

# See Sels et al., 2020

corr_df <- esm_data %>%
  select(
   clove,
   PA_similarity_act, NA_similarity_act,
   PA_similarity_perc, NA_similarity_perc,
   cPA_own, cNA_own,
   cPA_part_act, cNA_part_act,
   cPA_part_perc, cNA_part_perc
  )

corr_outcome <- corr_df %>%
  summarise(across(-clove, ~ cor(clove, .x, use = "pairwise.complete.obs"))) %>%
  pivot_longer(cols = everything(), names_to = "predictor", values_to = "r") %>%
  mutate(abs_r = abs(r)) %>%
  arrange(desc(abs_r))

corr_outcome

cor_matrix <- cor(corr_df, use = "pairwise.complete.obs")

plot <- plot_corr_heatmap(
  cor_matrix,
  "Outcome–predictor correlations (person-mean centered variables)"
)

print(plot) 

plot_name <- sprintf("%s_%d_outcome_predictors_correlations.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1
```

To evaluate the potential role of individual affect levels, we examined the preregistered diagnostic correlations between the person-mean centered affect components and the centered outcome (`clove`). We focus on centered variables here because the fixed effect for the time-varying similarity predictor $S^{within}$ in our preregistered mixed models estimates associations within persons. We do not report parallel between-person diagnostics here, as the present checks are intended solely to probe momentary (within-person) confounding relevant to $S^{within}$. Between-person effects are addressed directly in the multilevel models through the inclusion of the decomposed $S^{between}$ component and corresponding between-level covariates in the sensitivity analyses.

**1. Correlation with Outcome (`clove`).** Consistent with standard dyadic research findings (e.g., @kennyDyadicDataAnalysis2020), actor effects are generally stronger than partner effects. The person-mean centered affect components (`cPA_own`, `cNA_own`) show moderate bivariate correlations with `clove` ($r≈.34$ for PA, $r≈−.32$ for NA). Partner effects were smaller ($r≈.17$ for PA, $r≈−.18$ for NA).

**2. Correlation with Similarity Predictors.** The similarity indices themselves showed substantial bivariate overlap with the affect components, as expected given their construction (calculated as Mean − \|Difference\|). This is **a definition-induced overlap**. From the data:

-   **For Positive Affect (PA):** The actual similarity index (`PA_similarity_act)` correlates strongly with individual affect levels ($r≈.63$ for both actor and partner).

-   **For Negative Affect (NA):** The overlap was moderate ($r≈.37$).

Consequently, models that include affect levels as covariates shift the interpretation of the similarity coefficient toward similarity beyond levels.

**3. Sensitivity Rationale.** @selsActualPerceivedEmotional2020 used low correlations as a heuristic to omit component controls, but our checks here reveal that it necessitates a more rigorous approach. Because affect levels correlate small-to-moderately with the outcome ($r≈.17–.34$) **and** strongly with the similarity predictor ($r≈.37–.63$), excluding them risks an ambiguous interpretation in which the variance of feeling good (or bad) is attributed to 'similarity'.

Therefore, the preregistered sensitivity analyses are important for interpretability and for evaluating whether the similarity construct contributes beyond level. In these models, we will add the person-mean centered momentary affect levels (actor and partner) and the grand-mean centered average affect levels as covariates. This matches the decomposition of our primary similarity predictor (into $S^{within}$ and $S^{between}$). If the hypothesized similarity effect attenuates substantially after controlling for levels, this would suggest that the index's predictive power is largely driven by level.

## Distribution analyses

### love

```{r love_skewness_analysis}

# 1. set thresholds
high <- 95
cutoff_perc <- 95  # % cutoff
cutoff_no <- 5   # cutoff in number of beeps NOT at ceiling of 100  

esm_data <- esm_data %>%
  mutate(
    is_high = love >= high,
    is_ceiling = love == 100
    )

# 2. distribution of love
plot_hist_numeric(esm_data, "love", "Ditribution of 'love'", "love (0-100)",2)

# 3. Summary
cat("Total beeps:", nrow(esm_data), "\n")
cat("Answered beeps with love >=", high, ":", sum(esm_data$is_high, na.rm = TRUE), "\n")
cat("Percentage with love >=", high, ":", round(mean(esm_data$is_high, na.rm = TRUE) * 100, 1), "% of answered beeps\n")
cat("Answered beeps at with love = 100:", sum(esm_data$is_ceiling, na.rm = TRUE), "\n")
cat("Percentage with love = 100:", round(mean(esm_data$is_ceiling, na.rm = TRUE) * 100, 1), "% of answered beeps\n\n")

# 4. Person-level summary
person_summary <- esm_data %>%
  group_by(person, dyad) %>%
  summarise(
    n_obs          = n(),
    n_answ         = sum(!is.na(love)),
    n_ceiling      = sum(is_ceiling, na.rm = TRUE),
    not_ceiling    = sum(!is.na(love)) - sum(is_ceiling, na.rm = TRUE),
    pct_ceiling    = mean(is_ceiling, na.rm = TRUE) * 100,
    .groups = "drop"
  ) %>%
  arrange(desc(pct_ceiling), desc(n_answ))

print(person_summary, n = 15)

heavy_persons_perc <- person_summary %>% filter(pct_ceiling >= cutoff_perc)
heavy_persons_no <- person_summary %>% filter(not_ceiling <= cutoff_no)
cat(sprintf("\nPersons with >= %d%% ceiling responses: %d\n\n", cutoff_perc, nrow(heavy_persons_perc)))
cat(sprintf("\nPersons with <= %d responses not at 100%%: %d\n\n", cutoff_no, nrow(heavy_persons_no)))

# 5. Trajectory plots
variable <- "love" 
participants <- c(26, 726) 
plot_trajectory(esm_data, variable, participants)
# plot_trajectory(esm_data, "PA_own", participants)
participants <- c(179, 879) 
plot_trajectory(esm_data, variable, participants)
participants <- c(154, 854) 
plot_trajectory(esm_data, variable, participants)
participants <- c(172, 872) 
plot_trajectory(esm_data, variable, participants)
participants <- c(80, 780) 
plot_trajectory(esm_data, variable, participants)
participants <- c(35, 735) 
plot_trajectory(esm_data, variable, participants)
participants <- c(151, 851) 
plot_trajectory(esm_data, variable, participants)
participants <- c(154, 854) 
plot_trajectory(esm_data, variable, participants)
participants <- c(108, 808) 
plot_trajectory(esm_data, variable, participants)
participants <- c(47, 747) 
plot_trajectory(esm_data, variable, participants)
participants <- c(143, 843) 
plot_trajectory(esm_data, variable, participants)

# 6. Create sensitivity analysis datasets
dyads_to_exclude <- heavy_persons_no$dyad

esm_data_ceiling_love <- esm_data %>%
  filter(!dyad %in% dyads_to_exclude)

cat("\nN after excluding dyads with heavy-ceiling persons:", nrow(esm_data_ceiling_love), "\n")

plot_hist_numeric(esm_data_ceiling_love, "love", "Ditribution of 'love'", "love (0-100)",2)

# 7. Summary
cat("Total beeps:", nrow(esm_data_ceiling_love), "\n")
cat("Answered beeps with love >=", high, ":", sum(esm_data_ceiling_love$is_high, na.rm = TRUE), "\n")
cat("Percentage with love >=", high, ":", round(mean(esm_data_ceiling_love$is_high, na.rm = TRUE) * 100, 1), "% of answered beeps\n")
cat("Answered beeps at with love = 100:", sum(esm_data_ceiling_love$is_ceiling, na.rm = TRUE), "\n")
cat("Percentage with love = 100:", round(mean(esm_data_ceiling_love$is_ceiling, na.rm = TRUE) * 100, 1), "% of answered beeps\n\n")
```

The outcome variable `love` is measured on a 0–100 scale. Inspection of its distribution reveals a pronounced ceiling effect: approximately $17.2\%$ of answered beeps received a score of exactly 100, and an additional $4\%$ scored 95 or above. This ceiling is expected to produce non-normality and heteroscedasticity in the residuals of the linear mixed-effects models: the assumption of normally distributed residuals is violated at the upper tail, and residual variance is likely smaller for observations near the ceiling than for those in the middle of the distribution. While LMMs are reasonably robust to moderate departures from normality, particularly with large samples, the severity of the ceiling here warrants explicit examination of whether conclusions from the primary model hold when these observations are reduced or removed.

To assess robustness, we will run two sensitivity analyses based on the degree to which individual participants show ceiling-dominated response patterns. Participants with very few love-ratings at lower than 100 contribute almost no within-person variability in the outcome, providing little information for estimating the within-person associations that are central to our research questions. We therefore identified participants whose non-ceiling beep count fell at or below a threshold of 5 and 10 , and excluded the corresponding dyads entirely from each sensitivity dataset. Exclusion at the dyad level is appropriate given the dyadic structure of the data: the key predictors (including within-dyad similarity) are computed from both partners' scores simultaneously, so removing one partner renders the dyadic predictors undefined for all shared moments, making dyad-level exclusion more coherent than person-level exclusion.

The primary model was fitted on the full dataset. The two sensitivity models were fitted on the datasets excluding dyads with at least one partner falling below the non-ceiling threshold of 5 and 10 beeps, respectively. Consistency of the fixed-effect estimates across all three models would indicate that the ceiling responses do not materially distort the primary conclusions.

------------------------------------------------------------------------

# BG

## DCI descriptive statistics

### DCI scores histogram

```{r dci_scores_histogram}

plot_dci_total_hist <- bg_data %>%
  filter(!is.na(dci_total)) %>%
  ggplot(aes(x = dci_total)) +
  geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
  labs(title = "Distribution of Total DCI Scores", 
       subtitle = "Sum of 30 items (Negative items reversed)",
       x = "Total Score", y = "Count") +
  theme_minimal()
plot_dci_total_hist
plot_name <- sprintf("%s_%d_dci_score_hist.png", prefix, plot_counter)
save_plot(plot_dci_total_hist, plot_name)
plot_counter <- plot_counter + 1
```

Because the 30 DCI items used are Likert 1-5, the total DCI score is between 30 and 150.

## Reliability of DCI scale

### Alpha and Omega ($\alpha$ and $\omega$)

psych::omega() estimates reliability using a factor model and assumes that items may refelect:

-   One general factor (what items have in common)

-   Several group factors (clusters or subscales)

In the function, we therefore have to specify the number of group factors we expect. The DCI has the following 9 subscales (@bodenmannDyadicCopingInventory2018), with items 36 and 37 for evaluation (Evaluation of dyadic coping; (EDC):

1.  Stress communicated by oneself (SCO): items 1, 2, 3, and 4

2.  Supportive dyadic coping by oneself (SDC): items 20, 21, 23, 24, and 29

3.  Delegated dyadic coping by oneself (DDCO): items 28, and 30

4.  Negative dyadic coping by oneself (NDCO): items 22, 25, 26, and 27

5.  Stress communication of the partner (SCP): items 16, 17, 18, and 19

6.  Supportive dyadic coping of the partner (SDCP): items 5, 6, 8, 9, and 13

7.  Delegated dyadic coping of the partner (DDCP): items 12 and 14

8.  Negative dyadic coping by partner (NDCP): items 7, 10, 11, and 15

9.  Common dyadic coping (CDC): items 31, 32, 33, 34, and 35

Since items 31-37 are missing in the data, 8 subscales would be expected, and nfactors is set to 8.

```{r dci_reliability}

# A. Cronbach's Alpha
alpha_res <- bg_data %>%
    select(all_of(scale_DCI)) %>%
    alpha()
cat(sprintf("Cronbach's alpha for DCI scale = %f", alpha_res$total$raw_alpha))

# B. McDonald's Omega
omega_res <- bg_data %>%
    select(all_of(scale_DCI)) %>%
    omega(nfactors = 8, plot = FALSE)
cat(sprintf("\nMcDonald's omega for DCI scale = %f\n\n", omega_res$omega.tot))
```

### Hierarchical EFA with 8 factors

The Schmid-Leiman transformation produces a bifactor representation (one general factor that all items can load on + orthogonal group factors capturing leftover shared variance). This. matches how questionnaires are actually built. It also produces a hierarchical omega, being the reliability of the general factor as measured by the total score (how defensible is it to treat the total score as measuring one common trait). We can look at the factor structure in more detail.

Running the exploratory hierarchical factor analysis (@schmidDevelopmentHierarchicalFactor1957) with nfactors=8, the question is whether we can find the original 8 subscales on our data.\
We use Pearson correlations (thus without the poly = TRUE argument for polychoric correlations), which should be a good approximation for 5 Likert items. As a sensitivity, we will recalculate with polychoric correlations.

```{r dci_efa_8}
omega_res <- bg_data %>%
    select(all_of(scale_DCI)) %>%
    omega(nfactors = 8, plot = FALSE)
omega.diagram(omega_res) # by default factor loadings of > 0.2 are printed
omega.diagram(omega_res, cut = .3)
print(omega_res)
print(omega_res, cut = .3)
plot_name <- sprintf("%s_%d_dci_omega_factor_analysis_fact8_cut2.png", prefix, plot_counter)
save_base_plot(omega.diagram(omega_res), plot_name)
plot_counter <- plot_counter + 1
plot_name <- sprintf("%s_%d_dci_omega_factor_analysis_fact8_cut3.png", prefix, plot_counter)
save_base_plot(omega.diagram(omega_res, cut = .3), plot_name)
plot_counter <- plot_counter + 1
```

The DCI has very high overall reliability ($\alpha = 0.85,\ \omega = 0.92$) and the 8-factor structure fits the data closely (RMSEA/Root Mean Square Error of Approximation = 0.037 \< 0.05, BIC = -900.29).

\[Note: h2=communality (proportion of variance explained by common factors, u2= uniqueness (proportion of variance not explained by common factors, 1-h2), p2=proportion of variance attributable to g, com=complexity (how cross-loaded is thjis item).\]

Omega Hierarchical = 0.58, meaning that about 58% of the variance is due to "general coping", while the remaining variance is driven by the 8 factors (using cutoff for loadings of \> .2, in bold italics when \> .3):

|  |  |  |  |  |
|---------------|---------------|---------------|:--------------|---------------|
| **Factor** | **Items Loading...** | ... and items missing... | **...for interpretation** | **Other items loading** |
| **F1\*** | ***5*****, *6***, 8, ***9*** | 13 | Supportive Partner (SDCP)  | 24 |
|  | 7, ***10*****, *11***, 15 | \- | part 1 of Negative Partner (NDCP)  |  |
| **F2\*** | ***20, 21, 24*** | 23, 29 | part 1 of Supportive Self (SDC)  | 1, 26, 27 |
| **F3\*** | ***16, 17, 18, 19*** | \- | Stress Comm. Partner (SCP)  | 7 (-), ***8, 13***, 15, 21 |
| **F4\*** | ***1, 2, 3, 4*** | \- | Stress Comm. Self (SCO)  | 8, 13, 14, 21, 27 |
| **F5\*** | ***22, 25***, 27 | 26 | Negative Self (NDCO)  | 2 (-), 13 |
|  | 7, ***10, 11*** | 15 | part 2 of Negative Partner (NDCP)  |  |
| **F6\*** | ***28, 30*** | \- | Delegated Self (DDCO)  | 17, 18, 24, 27 (-), ***29*** |
| **F7\*** | ***23, 29*** | 20, 21, 24 | part 2 of Supportive Self (SDC) | ***26***, 27 |
| **F8\*** | ***12, 14*** | \- | Delegated Partner (DDCP)  | 2 |

Note: use fa.parallel to determine how many factors are clearly above noise.

```{r dci_horn}
bg_data %>%
  select(all_of(scale_DCI)) %>%
  fa.parallel()
```

### Polychoric sensitivities

```{r dci_efa_8_polychoric}
omega_res <- bg_data %>%
    select(all_of(scale_DCI)) %>%
    omega(nfactors = 8, plot = FALSE, poly = TRUE)
omega.diagram(omega_res) # by default factor loadings of > 0.2 are printed
print(omega_res, cut = .3)
plot_name <- sprintf("%s_%d_dci_omega_factor_analysis_fact8_cut2_poly.png", prefix, plot_counter)
save_base_plot(omega.diagram(omega_res), plot_name)
plot_counter <- plot_counter + 1
```

**Polychoric correlations** assume each item is an **ordinal discretization** of an underlying continuous latent response, and try to recover the correlation between those latent responses. The results produced here, indicate some instability or overfactoring:

**DCI12 has h2 = 1.00 and u2 = 0.00** and a very high group loading (F7 = .94). This could indicate a **Heywood** situation or a correlation structure that is near-deterministic for that item. With polychorics this often happens when an item has **very sparse categories** (e.g., almost everyone answers 1 or 5), making polychoric estimation unstable.

### Hierarchical EFA with 7 factors

```{r dci_efa_7}
omega_res <- bg_data %>%
    select(all_of(scale_DCI)) %>%
    omega(nfactors = 7, plot = FALSE)
omega.diagram(omega_res) # by default factor loadings of > 0.2 are printed
omega.diagram(omega_res, cut = .3)
print(omega_res)
print(omega_res, cut = .3)
plot_name <- sprintf("%s_%d_dci_omega_factor_analysis_fact7_cut2.png", prefix, plot_counter)
save_base_plot(omega.diagram(omega_res), plot_name)
plot_counter <- plot_counter + 1
plot_name <- sprintf("%s_%d_dci_omega_factor_analysis_fact7_cut3.png", prefix, plot_counter)
save_base_plot(omega.diagram(omega_res, cut = .3), plot_name)
plot_counter <- plot_counter + 1
```

|  |  |  |  |  |
|---------------|---------------|---------------|:--------------|---------------|
| **Factor** | **Items Loading...** | ... and items missing... | **...for interpretation** | **Other items loading** |
| **F1\*** | ***5*****, *6***, 8, ***9, 13*** | \- | Supportive Partner (SDCP)  | ***16, 19,*** 26, 28(-) |
|  | 7, ***10*****, *11***, 15 | \- | part 1 of Negative Partner (NDCP)  |  |
| **F2\*** | ***20, 21, 24*** | 23, 29 | part 1 of Supportive Self (SDC)  | 1, 6, 26, 27 |
| **F3\*** | ***16, 17, 18, 19*** | \- | Stress Comm. Partner (SCP)  | 7 (-), 8, 13, 15, 26(-), ***28,*** 30 |
| **F4\*** | ***1, 2, 3, 4*** | \- | Stress Comm. Self (SCO)  | 13, 23(-), 30 |
| **F5\*** | ***23, 29*** | 20, 21, 24 | part 2 of Supportive Self (SDC)  | 26 |
|  | ***28***, 30 | \- | Delegated Self (DDCO)  |  |
| **F6\*** | ***7,*** 10, 11 | 15 | part 2 of Negative Partner (NDCP) |  |
|  |  |  | mix? | ***22, 23(-), 25,*** 30 |
| **F7\*** | ***12, 14*** | \- | Delegated Partner (DDCP)  | 2 |

### DCI ICC

```{r dci_icc}

model_dci <- lmer(dci_total ~ 1 + (1 | dyad), data = bg_data)
icc_res <- performance::icc(model_dci)
icc_val <- icc_res$ICC_adjusted

icc_table <- data.frame(
  Level = c("ICC_Dyad", "ICC_Person (incl. error)"),
  Proportion = c(icc_val, 1 - icc_val)
)

print(icc_table)

plot <- ggplot(icc_table, aes(x = "", y = Proportion, fill = Level)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = scales::percent(Proportion, accuracy = 0.1)), 
            position = position_stack(vjust = 0.5), color = "white") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent) +
 scale_fill_discrete(type = hcl.colors(4, palette = "Grays", alpha = 0.8)) +
  labs(title = "Variance in DCI Total",
       subtitle = "How much of the score depends on the couple?",
       x = "", y = "Proportion of Variance") +
  theme_minimal()

print(plot)

plot_name <- sprintf("%s_%d_dci_total_icc.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1

```

### DCI pairwise dyadic correlation

Another valid way to look at the dyadic correlation is to use every dyad twice but with partners inversed.

```{r dci_dyadic_correlation}

# Create dataframe with dci_total and we_ness of partners in same row
dci_dyadic <- bg_data %>%
  mutate(role = if_else(person < 700, "P1", "P2")) %>%
  select(dyad, role, dci_total, we_ness) %>% # include we_ness for later chunk
  group_by(dyad) %>%
  pivot_wider(
    names_from = role, 
    values_from = c(dci_total, we_ness)
    ) %>%
  ungroup()

# create double-entry data/symmetrical because partners are indistinguishable
dci_double <- bind_rows(
  dci_dyadic %>% select(val_x = dci_total_P1, val_y = dci_total_P2),
  dci_dyadic %>% select(val_x = dci_total_P2, val_y = dci_total_P1)
)

# Calculate Correlation
cor_total <- cor(dci_double$val_x, dci_double$val_y, use = "complete.obs")
header(sprintf("Dyadic Correlation (Total DCI): r = %.3f", cor_total))

# Plot Scatter 
plot_dci_total_corr <- dci_double %>%
  ggplot(aes(x = val_x, y = val_y)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", color = "lightblue", se = FALSE) +
  geom_abline(intercept = 0, slope = 1, color = "lightgrey") +
  labs(title = "Dyadic Correlation: Total DCI",
       subtitle = paste0("Pearson r = ", round(cor_total, 3)),
       x = "Partner A Score", y = "Partner B Score") +
  theme_minimal() +
  coord_fixed()
plot_dci_total_corr

# Save plot
plot_name <- sprintf("%s_%d_dci_total_scatter.png", prefix, plot_counter)
save_plot(plot_dci_total_corr, plot_name)
plot_counter <- plot_counter + 1

```

## We-ness descriptive statistics

### We-ness scores histogram and differences between partners

```{r weness_scores_histogram}

# Histogram of We-ness
plot_weness_hist <- bg_data %>%
  filter(!is.na(we_ness)) %>%
  ggplot(aes(x = we_ness)) +
  geom_histogram(binwidth = 0.1, fill = "lightgreen", color = "white") +
  labs(title = "Distribution of 'We-ness' Scores", 
       subtitle = "Average of 22 Positive items (No negative items)",
       x = "We-ness Score", y = "Count") +
  theme_minimal()
plot_weness_hist

# Save Histogram
plot_name <- sprintf("%s_%d_weness_hist.png", prefix, plot_counter)
save_plot(plot_weness_hist, plot_name)
plot_counter <- plot_counter + 1

# Calculate Absolute Difference in We-ness
dci_dyadic <- dci_dyadic %>%
  mutate(
    weness_diff = abs(we_ness_P1 - we_ness_P2)
  )

cat("Summary of Absolute Differences in We-ness between Partners:\n")
summary(dci_dyadic$weness_diff) %>% print()

# Plot Histogram of Differences
plot_weness_diff <- dci_dyadic %>%
  filter(!is.na(weness_diff)) %>%
  ggplot(aes(x = weness_diff)) +
  geom_histogram(binwidth = 0.1, fill = "tomato3", color = "white") +
  labs(title = "Distribution of We-ness Differences", 
       subtitle = "Absolute difference between Partner 1 and Partner 2",
       x = "Difference Score (|P1 - P2|)", y = "Count of Couples") +
  theme_minimal()
plot_weness_diff

# Save plot
plot_name <- sprintf("%s_%d_weness_diff_hist.png", prefix, plot_counter)
save_plot(plot_weness_diff, plot_name)
plot_counter <- plot_counter + 1
```

## Reliability of we-ness construct

### Alpha and Omega

```{r weness_reliability}
# A. Cronbach's Alpha
alpha_res <- bg_data %>%
    select(all_of(DCI_weness_items)) %>%
    alpha()
cat(sprintf("Cronbach's alpha for we-nesse = %f\n", alpha_res$total$raw_alpha))

# B. McDonald's Omega
omega_res <- bg_data %>%
    select(all_of(DCI_weness_items)) %>%
    omega(nfactors = 6, plot = FALSE)
cat(sprintf("\nMcDonald's omega for we-ness = %f\n\n", omega_res$omega.tot))
```

### Hierarchical EFA with 6 factors

Since we exclude the negative scales (NDCO and NDCP, with items 7, 10, 11, 15, 22, 25, 26, and 27), we expect 6 subscales in our we-ness construct.

A hierarchical factor analysis (Schmid-Leiman bifactor solution) was conducted to evaluate the internal structure of the We-Ness construct. The analysis supports a hierarchical structure consisting of one General Factor (*g*) and three distinct specific group factors, confirming that "We-Ness" is an overarching construct that encompasses both giving and receiving support.

```{r weness_efa_6}

# A. Cronbach's Alpha
alpha_res <- bg_data %>%
    select(all_of(DCI_weness_items)) %>%
    alpha()
cat(sprintf("Cronbach's alpha for we-nesse = %f\n", alpha_res$total$raw_alpha))

# B. McDonald's Omega
omega_res <- bg_data %>%
    select(all_of(DCI_weness_items)) %>%
    omega(nfactors = 6, plot = FALSE)
cat(sprintf("\nMcDonald's omega for we-ness = %f\n\n", omega_res$omega.tot))
print(omega_res)
omega.diagram(omega_res, main = "We-ness Factor Structure")

# Save diagram
plot_name <- sprintf("%s_%d_weness_factor_structure.png", prefix, plot_counter)
save_base_plot(omega.diagram(omega_res, main = "We-ness Factor Structure"), plot_name)
plot_counter <- plot_counter + 1

```

|  |  |  |  |  |
|---------------|---------------|---------------|:--------------|---------------|
| **Factor** | **Items Loading...** | ... and items missing... | **...for interpretation** | **Other items loading** |
| **F1\*** | ***5*****, *6***, ***8***, ***9, 13*** | \- | Supportive Partner (SDCP)  | 14, ***16***, 19, 28(-) |
| **F2\*** | ***20, 21, 24*** | 23, 29 | Part 1 of Supportive Self (SDC)  | ***1***, 30 |
| **F3\*** | ***16, 17, 18, 19*** | \- | Stress Comm. Partner (SCP)  | ***8, 13,*** ***28,*** 30 |
| **F4\*** | ***1, 2, 3, 4*** | \- | Stress Comm. Self (SCO)  | 13, 14, 23(-) |
| F5\* | ***23, 29*** | 20, 21, 24 | Part 2 of Supportive Self (SDC) | \- |
|  | ***28***, 30 | \- | Delegated Self (DDCO)  | \- |
| F6\* | ***12, 14*** | \- | Delegated Partner (DDCP) | 2 |

### We-ness ICC

```{r weness_icc}

model_weness <- lmer(we_ness ~ 1 + (1 | dyad), data = bg_data)
icc_res <- performance::icc(model_weness)
icc_val <- icc_res$ICC_adjusted

icc_table <- data.frame(
  Level = c("ICC_Dyad", "ICC_Person (incl. error)"),
  Proportion = c(icc_val, 1 - icc_val)
)

print(icc_table)

plot <- ggplot(icc_table, aes(x = "", y = Proportion, fill = Level)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = scales::percent(Proportion, accuracy = 0.1)), 
            position = position_stack(vjust = 0.5), color = "white") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent) +
 scale_fill_discrete(type = hcl.colors(4, palette = "Grays", alpha = 0.8)) +
  labs(title = "Variance in We-ness",
       subtitle = "How much of the score depends on the couple?",
       x = "", y = "Proportion of Variance") +
  theme_minimal()

print(plot)

plot_name <- sprintf("%s_%d_weness_icc.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1


```

### We-ness pairwise dyadic correlation

```{r weness_dyadic_correlation}

# create double-entry data/symmetrical because partners are indistinguishable
dci_double <- bind_rows(
  dci_dyadic %>% select(val_x = we_ness_P1, val_y = we_ness_P2),
  dci_dyadic %>% select(val_x = we_ness_P2, val_y = we_ness_P1)
)

# Calculate Correlation
cor_total <- cor(dci_double$val_x, dci_double$val_y, use = "complete.obs")
header(sprintf("Dyadic Correlation (We-ness): r = %.3f", cor_total))

# Plot Scatter
plot <- dci_double %>%
  ggplot(aes(x = val_x, y = val_y)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", color = "blue", se = FALSE) +
  geom_abline(intercept = 0, slope = 1, color = "lightgrey") +
  labs(title = "Dyadic Correlation: We-ness",
       subtitle = paste0("Pearson r = ", round(cor_total, 3)),
       x = "Partner A Score", y = "Partner B Score") +
  theme_minimal() +
  coord_fixed()
plot

# Save plot
plot_name <- sprintf("%s_%d_weenss_total_scatter.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1
```

```{r BG unload}
#| include: false
clean_config("BG")
```

------------------------------------------------------------------------

# VMR

```{r VMR_setup}
# load data
vmr_data  <- readRDS(file.path(dir_data_ana, "vmr_ana.rds"))
# load data configuration file 
load_config("VMR")
# plot initialisation
plot_counter <- 1
prefix <- "VMR"
```

## Repeated measures correlations

We will use `rmcorr:rmcorr()` on the VMR data in a similar way as on the ESM data. We will calculate these correlations separately for the positive and the negative interaction, next to pooled across interactions.

```{r create_positive_and_negative}
vmr_data_pos <- vmr_data %>% filter(topic == "positive")
vmr_data_neg <- vmr_data %>% filter(topic == "negative")
```

### Personal: Intrapersonal dynamics (individual level)

#### Own PA vs own NA

($r_{rm} = -0.72, p < .001$)

```{r VMR_rmcorr_PA_own_NA_own}
#| warning: false

# All interactions combined
rmc <- rmcorr(person, PA_own, NA_own, vmr_data)
create_rmc_plot_title(rmc, "VMR All - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_pa_own_na_own_all.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1

# Positive Interaction
rmc <- rmcorr(person, PA_own, NA_own, vmr_data_pos)
create_rmc_plot_title(rmc, "VMR Positive - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_pa_own_na_own_pos.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1

# Negative Interaction
rmc <- rmcorr(person, PA_own, NA_own, vmr_data_neg)
create_rmc_plot_title(rmc, "VMR Negative - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_pa_own_na_own_neg.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1

```

#### Own PA vs perceived partner's PA

($r_{rm} = 0.7, p < .001$)

```{r VMR_rmcorr_PA_own_PA_perceived_partner}
#| warning: false

# All interactions combined
rmc <- rmcorr(person, cPA_own, cPA_part_perc, vmr_data)
create_rmc_plot_title(rmc, "VMR All - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_pa_own_pa_partner_all.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1

# Positive Interaction
rmc <- rmcorr(person, cPA_own, cPA_part_perc, vmr_data_pos)
create_rmc_plot_title(rmc, "VMR Positive - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_pa_own_pa_partner_pos.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1

# Negative Interaction
rmc <- rmcorr(person, cPA_own, cPA_part_perc, vmr_data_neg)
create_rmc_plot_title(rmc, "VMR Positive - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_pa_own_pa_partner.png_neg", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1
```

#### Own NA vs perceived partner's NA

($r_{rm} = 0.6, p < .001$)

```{r VMR_rmcorr_NA_own_NA_perceived_partner}
#| warning: false

# All interactions combined
rmc <- rmcorr(person, cNA_own, cNA_part_perc, vmr_data)
create_rmc_plot_title(rmc, "VMR All - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_na_own_na_partner_perc_all.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1

# Positive Interaction
rmc <- rmcorr(person, cNA_own, cNA_part_perc, vmr_data_pos)
create_rmc_plot_title(rmc, "VMR Positive - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_na_own_na_partner_perc_pos.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1

# Negative Interaction
rmc <- rmcorr(person, cNA_own, cNA_part_perc, vmr_data_neg)
create_rmc_plot_title(rmc, "VMR Positive - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_na_own_na_partner_perc_neg.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1
```

#### Actual PA similarity vs perceived PA similarity

($r_{rm} = 0.67, p < .001$)

How well tuned is a person to the "state of the union"?

```{r VMR_rmcorr_actual_perceived_sim_PA}
#| warning: false

# All interactions combined
rmc <- rmcorr(person, PA_similarity_act, PA_similarity_perc, vmr_data)
create_rmc_plot_title(rmc, "VMR All - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_pa_sim_act_pa_sim_perc_all.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1

# Positive Interaction
rmc <- rmcorr(person, PA_similarity_act, PA_similarity_perc, vmr_data_pos)
create_rmc_plot_title(rmc, "VMR Positive - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_pa_sim_act_pa_sim_perc_pos.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1

# Negative Interaction
rmc <- rmcorr(person, PA_similarity_act, PA_similarity_perc, vmr_data_neg)
create_rmc_plot_title(rmc)
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_pa_sim_act_pa_sim_perc_neg.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1
```

#### Actual NA similarity vs perceived NA similarity

($r_{rm} = 0.59, p < .001$)

How well tuned is a person to the "state of the union"?

```{r VMR_rmcorr_actual_perceived_sim_NA}
#| warning: false

# All interactions combined
rmc <- rmcorr(person, NA_similarity_act, NA_similarity_perc, vmr_data)
create_rmc_plot_title(rmc, "VMR All - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_na_sim_act_na_sim_perc_all.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1

# Positive Interaction
rmc <- rmcorr(person, NA_similarity_act, NA_similarity_perc, vmr_data_pos)
create_rmc_plot_title(rmc, "VMR Positive - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_na_sim_act_na_sim_perc_pos.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1

# Negative Interaction
rmc <- rmcorr(person, NA_similarity_act, NA_similarity_perc, vmr_data_neg)
create_rmc_plot_title(rmc)
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_na_sim_act_na_sim_perc_neg.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1
```

### Dyadic: Interpersonal covariation

```{r create_wide_vmr}
vmr_data_dyad <- vmr_data %>% filter(person < 700)
vmr_data_dyad_pos <- vmr_data_dyad %>% filter(topic == "positive")
vmr_data_dyad_neg <- vmr_data_dyad %>% filter(topic == "negative")
```

#### Own PA vs Partner PA

($r_{rm} = 0.45, p < .001$)

```{r VMR_rmcorr_PA_own_PA_partner}
#| warning: false

# All interactions combined
rmc <- rmcorr(dyad, cPA_own, cPA_part_act, vmr_data_dyad)
create_rmc_plot_title(rmc, "VMR All - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_pa_own_pa_partner_all.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1

# Positive Interaction
rmc <- rmcorr(dyad, cPA_own, cPA_part_act, vmr_data_dyad_pos)
create_rmc_plot_title(rmc, "VMR Positive - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_pa_own_pa_partner_pos.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1

# Negative Interaction
rmc <- rmcorr(dyad, cPA_own, cPA_part_act, vmr_data_dyad_neg)
create_rmc_plot_title(rmc, "VMR Negative - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_pa_own_pa_partner_neg.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1

```

#### Own NA vs Partner NA

($r_{rm} = 0.42, p < .001$)

```{r VMR_rmcorr_NA_own_NA_partner}
#| warning: false

# All interactions combined
rmc <- rmcorr(dyad, cNA_own, cNA_part_act, vmr_data_dyad)
create_rmc_plot_title(rmc, "VMR All - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_na_own_na_partner_all.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1

# Positive Interaction
rmc <- rmcorr(dyad, cNA_own, cNA_part_act, vmr_data_dyad_pos)
create_rmc_plot_title(rmc, "VMR Positive - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_na_own_na_partner_pos.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1

# Negative Interaction
rmc <- rmcorr(dyad, cNA_own, cNA_part_act, vmr_data_dyad_neg)
create_rmc_plot_title(rmc, "VMR Negative - ")
rmc
plot(rmc, main = plot_title)
plot_name <- sprintf("%s_%d_rmcorr_na_own_na_partner_neg.png", prefix, plot_counter)
save_base_plot(plot(rmc, main = plot_title), plot_name)
plot_counter <- plot_counter + 1
```

## Intraclass Correlation Coefficient (ICC)

### All interactions combined

```{r vmr_icc_all}

# Variables to analyze
target_vars_vmr <- c(
  "PA_own", "NA_own",
  "NA_part_perc", "PA_part_perc",
  "PA_similarity_perc", "NA_similarity_perc"
)

# Run the loop
icc_table_all <- target_vars_vmr %>%
  map_dfr(function(v) {
    # Construct the 3-level formula: Moments < Person < Dyad
    formula_str <- as.formula(paste0(v, " ~ 1 + (1 | dyad / person)"))
    # Fit the model
    model <- lmer(formula_str, data = vmr_data)
    # Extract ICCs and format into a single row, icc from performance package
    icc_data <- icc(model, by_group = TRUE)
    # Create a clean data frame for this variable
    d_icc <- icc_data$ICC[icc_data$Group == "dyad"]
    p_icc <- icc_data$ICC[icc_data$Group == "person:dyad"] 
    data.frame(
      Variable = v,
      ICC_Dyad = d_icc,
      ICC_Person = p_icc,
      ICC_Moment = 1- d_icc - p_icc
    )
  })

# View the results
print(icc_table_all)

# Calculate the Residual (Momentary) variance
plot_data <- icc_table_all %>%
  mutate(Momentary = 1 - (ICC_Dyad + ICC_Person)) %>%
  pivot_longer(cols = c(ICC_Dyad, ICC_Person, Momentary), 
               names_to = "Level", values_to = "Proportion")

plot <- ggplot(plot_data, aes(x = Variable, y = Proportion, fill = Level)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_discrete(type = hcl.colors(4, palette = "Grays", alpha = 0.8)) +
  labs(title = "VMR All - Variance Decomposition by Variable",
       x = "", y = "Percentage of Total Variance") +
  theme_minimal()

# show plot
print(plot)

# save plot
plot_name <- sprintf("%s_%d_icc_variance_decomposition_all_1.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1
```

```{r vmr_icc_act_all}
# Variables to analyze (VMR)
dyad_level_vars <- c("PA_similarity_act", "NA_similarity_act")

# Simplified 2-level loop (Segments < Dyad)
icc_table_dyad_only_all <- dyad_level_vars %>%
  map_dfr(function(v) {
    formula_str <- as.formula(paste0(v, " ~ 1 + (1 | dyad)"))
    model <- lmer(formula_str, data = vmr_data)
    icc_data <- icc(model) # Standard ICC for 2-level
    data.frame(
      Variable = v,
      ICC_Dyad = icc_data$ICC_adjusted,
      ICC_Person = "-" # not applicable in 2-level model
    )
  })

print(icc_table_dyad_only_all)

plot_data <- icc_table_dyad_only_all %>%
  mutate(Momentary = 1 - (ICC_Dyad)) %>%
  pivot_longer(
    cols = c(ICC_Dyad, Momentary),
    names_to = "Level",
    values_to = "Proportion"
  )

plot <- ggplot(plot_data, aes(x = Variable, y = Proportion, fill = Level)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_discrete(type = hcl.colors(4, palette = "Grays", alpha = 0.8)) +
  labs(title = "VMR All - Variance Decomposition by Variable",
       x = "", y = "Percentage of Total Variance") +
  theme_minimal()

print(plot)

plot_name <- sprintf("%s_%d_vmr_icc_variance_decomposition_all_2.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1

```

### Positive interaction

```{r vmr_icc_pos}

# Variables to analyze
target_vars_vmr <- c(
  "PA_own", "NA_own",
  "NA_part_perc", "PA_part_perc",
  "PA_similarity_perc", "NA_similarity_perc"
)

# Run the loop
icc_table_pos <- target_vars_vmr %>%
  map_dfr(function(v) {
    # Construct the 3-level formula: Moments < Person < Dyad
    formula_str <- as.formula(paste0(v, " ~ 1 + (1 | dyad / person)"))
    # Fit the model
    model <- lmer(formula_str, data = vmr_data_pos)
    # Extract ICCs and format into a single row, icc from performance package
    icc_data <- icc(model, by_group = TRUE)
    # Create a clean data frame for this variable
    d_icc <- icc_data$ICC[icc_data$Group == "dyad"]
    p_icc <- icc_data$ICC[icc_data$Group == "person:dyad"] 
    data.frame(
      Variable = v,
      ICC_Dyad = d_icc,
      ICC_Person = p_icc,
      ICC_Moment = 1- d_icc - p_icc
    )
  })

# View the results
print(icc_table_pos)

# Calculate the Residual (Momentary) variance
plot_data <- icc_table_pos %>%
  mutate(Momentary = 1 - (ICC_Dyad + ICC_Person)) %>%
  pivot_longer(cols = c(ICC_Dyad, ICC_Person, Momentary), 
               names_to = "Level", values_to = "Proportion")

plot <- ggplot(plot_data, aes(x = Variable, y = Proportion, fill = Level)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_discrete(type = hcl.colors(4, palette = "Grays", alpha = 0.8)) +
  labs(title = "VMR Positive - Variance Decomposition by Variable",
       x = "", y = "Percentage of Total Variance") +
  theme_minimal()

# show plot
print(plot)

# save plot
plot_name <- sprintf("%s_%d_icc_variance_decomposition_pos_1.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1
```

```{r vmr_icc_act_pos}
# Variables to analyze (VMR)
dyad_level_vars <- c("PA_similarity_act", "NA_similarity_act")

# Simplified 2-level loop (Segments < Dyad)
icc_table_dyad_only_pos <- dyad_level_vars %>%
  map_dfr(function(v) {
    formula_str <- as.formula(paste0(v, " ~ 1 + (1 | dyad)"))
    model <- lmer(formula_str, data = vmr_data_pos)
    icc_data <- icc(model) # Standard ICC for 2-level
    data.frame(
      Variable = v,
      ICC_Dyad = icc_data$ICC_adjusted,
      ICC_Person = "-" # not applicable in 2-level model
    )
  })

print(icc_table_dyad_only_pos)

plot_data <- icc_table_dyad_only_pos %>%
  mutate(Momentary = 1 - (ICC_Dyad)) %>%
  pivot_longer(
    cols = c(ICC_Dyad, Momentary),
    names_to = "Level",
    values_to = "Proportion"
  )

plot <- ggplot(plot_data, aes(x = Variable, y = Proportion, fill = Level)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_discrete(type = hcl.colors(4, palette = "Grays", alpha = 0.8)) +
  labs(title = "VMR Positive - Variance Decomposition by Variable",
       x = "", y = "Percentage of Total Variance") +
  theme_minimal()

print(plot)

plot_name <- sprintf("%s_%d_vmr_icc_variance_decomposition_pos_2.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1

```

### Negative interaction

```{r vmr_icc_neg}

# Variables to analyze
target_vars_vmr <- c(
  "PA_own", "NA_own",
  "NA_part_perc", "PA_part_perc",
  "PA_similarity_perc", "NA_similarity_perc"
)

# Run the loop
icc_table_neg <- target_vars_vmr %>%
  map_dfr(function(v) {
    # Construct the 3-level formula: Moments < Person < Dyad
    formula_str <- as.formula(paste0(v, " ~ 1 + (1 | dyad / person)"))
    # Fit the model
    model <- lmer(formula_str, data = vmr_data_neg)
    # Extract ICCs and format into a single row, icc from performance package
    icc_data <- icc(model, by_group = TRUE)
    # Create a clean data frame for this variable
    d_icc <- icc_data$ICC[icc_data$Group == "dyad"]
    p_icc <- icc_data$ICC[icc_data$Group == "person:dyad"] 
    data.frame(
      Variable = v,
      ICC_Dyad = d_icc,
      ICC_Person = p_icc,
      ICC_Moment = 1- d_icc - p_icc
    )
  })

# View the results
print(icc_table_neg)

# Calculate the Residual (Momentary) variance
plot_data <- icc_table_neg %>%
  mutate(Momentary = 1 - (ICC_Dyad + ICC_Person)) %>%
  pivot_longer(cols = c(ICC_Dyad, ICC_Person, Momentary), 
               names_to = "Level", values_to = "Proportion")

plot <- ggplot(plot_data, aes(x = Variable, y = Proportion, fill = Level)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_discrete(type = hcl.colors(4, palette = "Grays", alpha = 0.8)) +
  labs(title = "VMR Negative - Variance Decomposition by Variable",
       x = "", y = "Percentage of Total Variance") +
  theme_minimal()

# show plot
print(plot)

# save plot
plot_name <- sprintf("%s_%d_icc_variance_decomposition_neg_1.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1
```

```{r vmr_icc_act_neg}
# Variables to analyze (VMR)
dyad_level_vars <- c("PA_similarity_act", "NA_similarity_act")

# Simplified 2-level loop (Segments < Dyad)
icc_table_dyad_only_neg <- dyad_level_vars %>%
  map_dfr(function(v) {
    formula_str <- as.formula(paste0(v, " ~ 1 + (1 | dyad)"))
    model <- lmer(formula_str, data = vmr_data_neg)
    icc_data <- icc(model) # Standard ICC for 2-level
    data.frame(
      Variable = v,
      ICC_Dyad = icc_data$ICC_adjusted,
      ICC_Person = "-" # not applicable in 2-level model
    )
  })

print(icc_table_dyad_only_neg)

plot_data <- icc_table_dyad_only_neg %>%
  mutate(Momentary = 1 - (ICC_Dyad)) %>%
  pivot_longer(
    cols = c(ICC_Dyad, Momentary),
    names_to = "Level",
    values_to = "Proportion"
  )

plot <- ggplot(plot_data, aes(x = Variable, y = Proportion, fill = Level)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_discrete(type = hcl.colors(4, palette = "Grays", alpha = 0.8)) +
  labs(title = "VMR Negative - Variance Decomposition by Variable",
       x = "", y = "Percentage of Total Variance") +
  theme_minimal()

print(plot)

plot_name <- sprintf("%s_%d_vmr_icc_variance_decomposition_neg_2.png", prefix, plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1

```

### Comparison

```{r VMR_icc_comparison}
# 1) Bind + reshape to wide with a stable naming scheme: <Interaction>__<Metric>
icc_table_wide <- bind_rows(
  "All"      = icc_table_all,
  "Pos" = icc_table_pos,
  "Neg" = icc_table_neg,
  .id = "Interaction"
) %>%
  pivot_longer(
    cols      = c(ICC_Dyad, ICC_Person, ICC_Moment),
    names_to  = "Metric",
    values_to = "Value"
  ) %>%
  pivot_wider(
    names_from  = c(Interaction, Metric),
    values_from = Value,
    names_sep   = "_"
  ) %>%
  select(
    Variable,
    All_ICC_Dyad,      All_ICC_Person,      All_ICC_Moment,
    Pos_ICC_Dyad, Pos_ICC_Person, Pos_ICC_Moment,
    Neg_ICC_Dyad, Neg_ICC_Person, Neg_ICC_Moment
  )

# 2) gt table with spanners + repeated subheaders + vertical separators
icc_table_wide %>%
  gt() %>%
  tab_spanner(
    label   = "All interactions",
    columns = c(All_ICC_Dyad, All_ICC_Person, All_ICC_Moment)
  ) %>%
  tab_spanner(
    label   = "Positive interaction",
    columns = c(Pos_ICC_Dyad, Pos_ICC_Person, Pos_ICC_Moment)
  ) %>%
  tab_spanner(
    label   = "Negative interaction",
    columns = c(Neg_ICC_Dyad, Neg_ICC_Person, Neg_ICC_Moment)
  ) %>%
  cols_label(
    All_ICC_Dyad = "Dyad", All_ICC_Person = "Person", All_ICC_Moment = "Moment",
    Pos_ICC_Dyad = "Dyad", Pos_ICC_Person = "Person", Pos_ICC_Moment = "Moment",
    Neg_ICC_Dyad = "Dyad", Neg_ICC_Person = "Person", Neg_ICC_Moment = "Moment"
  ) %>%
  fmt_number(
    columns = -Variable,
    decimals = 2
  ) %>%
  cols_align("left",   columns = Variable) %>%
  cols_align("center", columns = -Variable) %>%
  tab_options(table.font.size = px(12),
    table.background.color = "white",
    heading.background.color = "white",
    column_labels.background.color = "white",
    row_group.background.color = "white"
  )
```

Some observations:

-   Within positive interactions, perceived NA similarity fluctuates less moment-to-moment and is more attributable to stable person/dyad differences. In a positive interaction, noticing negative cues may be more person/couple-specific, hence more between-person/dyad. Also perceived PA similarity becomes more dyad/person-linked in negative interactions than in positive interactions.

    Note: not analysed, but floor effects in NA during positive interactions can cause this I think.

-   Dyad-level variance is much larger in negative interactions for NA_own: .10 (All) → .04 (Positive) → .26 (Negative). This suggests couples differ much more in their overall NA during negative interactions.

-   The perceived partner affects show more stable components in the valenced subsets. So perceptions of the partner are less “moment-only” once you condition on interaction valence.

## Similarity indices

### Theoretical similarity surface

The similarity indices are constructed in the same way as for the ESM data, so the `base_surface` is the same.

### Similarity in the sample

#### Ranges and distribution

```{r VMR_similarity_pracical_ranges}

sim_cols <- c("PA_similarity_act", "NA_similarity_act", 
              "PA_similarity_perc", "NA_similarity_perc")

draw_violin_plot <- function() {
  vioplot(vmr_data[, sim_cols],
        main = "Range and Distribution of Affect Similarity",
        col = c("#A6CEE3", "#FB9A99", "#A6CEE3", "#FB9A99"),
        names = c("PA\nActual", "NA\nActual", "PA\nPerceived", "NA\nPerceived"),
        areaEqual = FALSE,   # FALSE = width is proportional to data density
        rectCol = "gray30",  # Color of the inner boxplot
        lineCol = "gray30",  # Color of the median line
        border = "gray30"    # Color of the violin border
)
title(ylab = "Similarity (S = M - D)", line = 2)
abline(h = 0, col = "black", lty = 2, lwd = 1.5) # Zero line
}

draw_violin_plot()

plot_name <- sprintf("%s_%d_violin.png", prefix, plot_counter)
save_base_plot(draw_violin_plot(), plot_name)
plot_counter <- plot_counter + 1

# numerical ranges
cat("Ranges (Min/Max):\n")
print(sapply(vmr_data[, sim_cols], range, na.rm = TRUE))
```

#### Actual PA similarity

```{r VMR_PA_act_obs_vs_theo}
#| warning: false

base_surface %>%
  add_layer(
    data = vmr_data,
    x = ~cPA_own,
    y = ~cPA_part_act,    
    z = ~PA_similarity_act,     
    title = "PA Actual Similarity: Observed vs. Theoretical",
    x_lab = "Actor PA (Cent.)",
    y_lab = "Partner PA (Cent.)",
    z_lab = "Actual PA Similarity Index"
  )
```

#### Actual NA similarity

```{r VMR_NA_act_obs_vs_theo}
#| warning: false

base_surface %>%
  add_layer(
    data = vmr_data,
    x = ~cNA_own,
    y = ~cNA_part_act,    
    z = ~NA_similarity_act,   
    title = "NA Actual Similarity: Observed vs. Theoretical",
    x_lab = "Actor NA (Cent.)",
    y_lab = "Partner NA (Cent.)",
    z_lab = "Actual NA Similarity Index"
  )
```

#### Perceived PA similarity

```{r VMR_PA_perc_obs_vs_theo}
#| warning: false

base_surface %>%
  add_layer(
    data = vmr_data,
    x = ~cPA_own,
    y = ~cPA_part_perc,    
    z = ~PA_similarity_perc,   
    title = "PA Perceived Similarity: Observed vs. Theoretical",
    x_lab = "Actor PA (Cent.)",
    y_lab = "Perceived Partner PA (Cent.)",
    z_lab = "Perceived PA Similarity Index"
  )
```

#### Perceived NA similarity

```{r VMR_NA_perc_obs_vs_theo}
#| warning: false

base_surface %>%
  add_layer(
    data = vmr_data,
    x = ~cNA_own,
    y = ~cNA_part_perc,    
    z = ~NA_similarity_perc,   
    title = "NA Perceived Similarity: Observed vs. Theoretical",
    x_lab = "Actor NA (Cent.)",
    y_lab = "Perceived Partner NA (Cent.)",
    z_lab = "Perceived NA Similarity Index"
  )
```

## Autocorrelation analysis

```{r vmr_autocorrelation}
#| warning: false

target_vars <- c("cPA_own", "cNA_own", "cPA_part_act", "cNA_part_act", 
                 "cPA_part_perc", "cNA_part_perc")

# calculate Lag-1 ac per person
# We group by dyad/person to ensure we don't lag across different people
acf_results <- vmr_data %>%
  group_by(dyad, person, topic) %>% # also group by topic here
  arrange(dyad, person, topic, segment) %>% # order by topic and time!
  summarise(
    across(all_of(target_vars), 
          ~ cor(., dplyr::lag(.), use = "pairwise.complete.obs"),
           .names = "{.col}"
    ),
    .groups = "drop"
  )

# reshape for easy checking/plotting
acf_long <- acf_results %>%
  pivot_longer(cols = all_of(target_vars), 
               names_to = "Variable", 
               values_to = "Autocorrelation")

# summary
cat("--- Average Lag-1 Autocorrelation (Temporal Dependence) ---\n")
acf_summary <- acf_long %>%
  group_by(topic, Variable) %>%
  summarise(
    Mean_ACF = mean(Autocorrelation, na.rm = TRUE),
    Median_ACF = median(Autocorrelation, na.rm = TRUE),
    SD_ACF = sd(Autocorrelation, na.rm = TRUE)
  )
print(acf_summary)

# Plot
ggplot(acf_long, aes(x = Variable, y = Autocorrelation, fill = Variable)) +
  geom_boxplot(alpha = 0.6, outlier.shape = NA) +
  geom_jitter(width = 0.2, alpha = 0.3, size = 1) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_wrap(~topic) +
  theme_minimal() +
  labs(
    title = "Distribution of Lag-1 Autocorrelation per Person",
    subtitle = "Faceted by interaction type",
    y = "Lag-1 Correlation (r)",
    x = NULL
  ) +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))

```

The mean lag-1 autocorrelations in the VMR data are lower than in the ESM data and have large between-person variability (SDs). Because each person only has T=16 segments (per interaction topic), each correlation r is estimated from only n=T-1=15 adjacent pairs (or fewer when missingness is present). This can lead to

-   substantial downward bias in estimates of the correlations, known as the Hurwicz or Kendall bias (see Table 4 in @solanasLagoneAutocorrelationShort2010)

-   substantial sampling error. Using Fisher's z-transformation, the transformed correlation is approximately normal with $SE(z)≈1/\sqrt{(n−3)}$​, see @fisherProbableErrorCoefficient1921. Via back-transormation this implies $SE(r)≈(1−r^2)/\sqrt{n−3}$​. However, this SE is a rough heuristic and can even be too small since it assumes independent pairs, while the data are serially dependent.

The lower lag-1 correlations and higher SDs in VMR may therefore come from these bias and noise effects due to short series length in addition to potentially faster, more reactive affect dynamics at the 30 seconds timescale.

```{r VMR unload}
#| include: false
clean_config("VMR")
```

------------------------------------------------------------------------

# POST

```{r POST_setup}
# load data
post_data  <- readRDS(file.path(dir_data_ana, "post_ana.rds"))
# load data configuration file 
load_config("POST")
# plot intialisation
plot_counter <- 1
prefix <- "POST"
```

### Relation between closeness and love

```{r data_prep_1}
# from wide to long format, with two rows per person, one for pos, one for neg
post_data_long <- post_data %>%
  pivot_longer(
    cols = c(love_neg, love_pos, close_neg, close_pos),
    names_to = c(".value", "context"), # Splitting names into variable + context
    names_sep = "_"                    # Separator is the underscore
  )

data_neg <- post_data_long %>% filter(context == "neg")
data_pos <- post_data_long %>% filter(context == "pos")
```

```{r pos_neg_pooled}
#| warning: false

# Q: "Is Love associated with Closeness across all contexts?"

# rmcorr
rmcorr(person, love, close, dataset = post_data_long)

# multilevel model
model_overall <- lmer(close ~ love + (1 | dyad/person), data = post_data_long)
summary(model_overall)
```

Random intercepts absorb person and dyad baselines (stable dyad and person differences) such that fixed effects are more reliable. Random effects absorb the stable, baseline differences between dyads (e.g., Dyad A is always closer than Dyad B). This prevents these baseline differences from skewing the results and ensures the Fixed Effect (the relationship between Love and Closeness) is estimated with the correct level of certainty (standard error), avoiding false positives due to non-independence.

Strong, positive, and statistically significant association between the love felt after pooled interaction and the closeness felt. Slope = .69, p \< .001.

```{r neg}
model_neg <- lmer(close ~ love + (1 | dyad), data = data_neg)
summary(model_neg)

# data_neg_scaled <- data_neg %>% 
#   mutate(love = scale(love), close = scale(close))

# r_neg_model <- lmer(close ~ love + (1 | dyad), data = data_neg_scaled)
# summary(r_neg_model) # The estimate for 'love' is your correlation r.
```

Strong, positive, and statistically significant association between the love felt after a negative interaction and the closeness felt after that same interaction. Slope = .66, p \< .001. Dyad variance = 75.31 and residual variance = 154.59. ICC = 75.31/(75.31 + 154.59) = 0.33. About 33% of variation in closeness is due to the dyad the person belongs to.

```{r pos}
model_pos <- lmer(close ~ love + (1 | dyad), data = data_pos)
summary(model_pos)
```

Even stronger association with slope of .81 (p \< .001).

Remarkable: the variance for the random dyad intercept is 0. This should be interpreted carefully but suggests that the model found no stable differences between couples in how much closeness people feel. So, in the context of positive interactions, knowing which couple someone belongs to provides no additional information about their closeness **beyond what is already explained by their love**. After the negative scenario, couples had different baselines (variance = 75.31). **Some couples stayed closer than others even when love was low after the negative interaction.**

The association between love and closeness was significantly stronger following positive interactions (β=0.81) compared to negative interactions (β=0.66). This suggests that while love acts as a buffer during conflict—maintaining connection despite the friction of the argument—it acts as an enhancer during positive moments, where the absence of defensive barriers allows feelings of affection to translate more directly into feelings of intimacy.

```{r POST unload}
#| include: false
clean_config("POST")
```
