---
title: "05_descriptives"
---

```{r setup}

# 1. Load packages, paths and data configurations
source(here::here("R", "00_setup.R"))

# 2. Set a global theme for plots
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# ESM

```{r ESM_setup}
# load data
esm_data  <- readRDS(file.path(dir_data_ana, "esm_ana.rds"))
# load data configuration file 
load_config("ESM")
```

### Repeated measures correlations

As in @bakdashRepeatedMeasuresCorrelation2017 we look at some repeated measure correlations. This is the common within person or within dyad correlation.

```{r rmcorr analysis}
#| warning: false

# (Bakdash & Marusich, 2017)
rmcorr(person, PA_own, NA_own, esm_data)
plot(rmcorr(person, PA_own, NA_own, esm_data))
rmcorr(dyad, cPA_own, cNA_part_act, esm_data)
plot(rmcorr(dyad, cPA_own, cNA_part_act, esm_data))


```

# BG

```{r BG_setup}
# load data
bg_data   <- readRDS(file.path(dir_data_ana, "bg_ana.rds"))
demographics <- readRDS(file.path(dir_data_ana, "demographics.rds"))
# load data configuration file
load_config("BG")
```

## Descriptive statistics (demographics)

```{r demographics}

plot_counter <- 1

# 1. AGE (numeric)
header("Age")
print(summary(demographics$age))
cat(sprintf("Missing age: %d\n", sum(is.na(demographics$age))))
plot <- demographics %>%
  filter(!is.na(age)) %>%
  ggplot(aes(x = age)) +
  geom_histogram(binwidth = 2) +
  labs(title = "Age distribution", x = "Age (years)", y = "Count") +
  theme_minimal()
plot
plot_name <- sprintf("A-%d_age_hist.png", plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1

# 2. GENDER
header("Gender")
labels <- get_label(demographics, "gender")
print(freq_table(labels))
plot <- plot_bar_categorical(labels, "Gender Distribution", "Gender")
plot
plot_name <- sprintf("A-%d_gender_bar.png", plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1

# 3. NATIONALITY
header("Nationality")
labels <- get_label(demographics, "nat")
print(freq_table(labels))
plot <- plot_bar_categorical(labels, "Nationality Distribution", "Nationality")
plot
plot_name <- sprintf("A-%d_nationality_bar.png", plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1

# 4. ETHNICITY
header("Ethnicity")
labels <- get_label(demographics, "etn")
print(freq_table(labels))
plot <- plot_bar_categorical(labels, "Ethnicity Distribution", "Ethnicity")
plot
plot_name <- sprintf("A-%d_ethnicity_bar.png", plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1

# 5. EDUCATION
header("Education")
labels <- get_label(demographics, "edu")
print(freq_table(labels))
plot <- plot_bar_categorical(labels, "Eduction Level Distribution", "Education")
plot
plot_name <- sprintf("A-%d_education_bar.png", plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1

# 6. CHILDREN
header("Children")
labels <- get_label(demographics, "child")
print(freq_table(labels))
plot <- plot_bar_categorical(labels, "Children (Yes/No)", "Children?")
plot
plot_name <- sprintf("A-%d_chile_bar.png", plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1

# 7. LIVING TOGETHER
header("Living together")
labels <- get_label(demographics, "liv_tog")
print(freq_table(labels))
plot <- plot_bar_categorical(labels, "Living Together? (Yes/No)", "Living Together?")
plot
plot_name <- sprintf("A-%d_liv_tog_bar.png", plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1

# 8. RELATIONSHIP DURATION (Numeric - months)
header("Relationship duration (months)")
print(summary(demographics$rel_dur))
cat(sprintf("Missing relationship duration: %d\n", sum(is.na(demographics$rel_dur))))
plot <- demographics %>%
  filter(!is.na(rel_dur)) %>%
  ggplot(aes(x = rel_dur)) +
  geom_histogram(binwidth = 2) +
  labs(title = "Relationship Duration Distribution", x = "Duration (Months)", y = "Count") +
  theme_minimal()
plot
plot_name <- sprintf("A-%d_rel_dur_hist.png", plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1

# 9. MEDICATION
header("Medication")
labels <- get_label(demographics, "med")
print(freq_table(labels))
plot <- plot_bar_categorical(labels, "Medication Use? (Yes/No)", "Living Together?")
plot
plot_name <- sprintf("A-%d_med_bar.png", plot_counter)
save_plot(plot, plot_name)
plot_counter <- plot_counter + 1
```

## DCI descriptive statistics

### DCI scores histogram

```{r dci_scores_histogram}

plot_dci_total_hist <- bg_data %>%
  filter(!is.na(dci_total)) %>%
  ggplot(aes(x = dci_total)) +
  geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
  labs(title = "Distribution of Total DCI Scores", 
       subtitle = "Sum of 30 items (Negative items reversed)",
       x = "Total Score", y = "Count") +
  theme_minimal()
plot_dci_total_hist
plot_name <- sprintf("A-%d_dci_score_hist.png", plot_counter)
save_plot(plot_dci_total_hist, plot_name)
plot_counter <- plot_counter + 1
```

### Reliability of DCI scale

```{r dci_reliability}

# A. Cronbach's Alpha
alpha_res <- bg_data %>%
    select(all_of(scale_DCI)) %>%
    alpha()
cat(sprintf("Cronbach's alpha for DCI scale = %f", alpha_res$total$raw_alpha))

# B. McDonald's Omega
omega_res <- bg_data %>%
    select(all_of(scale_DCI)) %>%
    omega(nfactors = 3, plot = TRUE)
omega_res <- suppressMessages(omega(bg_data %>% select(all_of(scale_DCI)), 
                                             nfactors = 3, plot = FALSE))
cat(sprintf("\nMcDonald's omega for DCI scale = %f", omega_res$omega.tot))
```

### DCI dyadic correlation

```{r dci_dyadic_correlation}

# Create dataframe with dci_total and we_ness of partners in same row
dci_dyadic <- bg_data %>%
  mutate(role = if_else(person < 700, "P1", "P2")) %>%
  select(dyad, role, dci_total, we_ness) %>%
  group_by(dyad) %>%
  pivot_wider(
    names_from = role, 
    values_from = c(dci_total, we_ness)
    )

# Calculate Correlation
cor_total <- cor(dci_dyadic$dci_total_P1, dci_dyadic$dci_total_P2, use = "complete.obs")
header(sprintf("Dyadic Correlation (Total DCI): r = %.3f", cor_total))

# Plot Scatter (Visualizing the correlation)
plot_dci_total_corr <- dci_dyadic %>%
  ggplot(aes(x = dci_total_P1, y = dci_total_P2)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", color = "blue", se = FALSE) +
  labs(title = "Dyadic Correlation: Total DCI",
       subtitle = paste0("Pearson r = ", round(cor_total, 3)),
       x = "Partner 1 Score", y = "Partner 2 Score") +
  theme_minimal()
plot_dci_total_corr

# Save plot
plot_name <- sprintf("A-%d_dci_total_scatter.png", plot_counter)
save_plot(plot_dci_total_corr, plot_name)
plot_counter <- plot_counter + 1

```

## We-ness descriptive statistics

### We-ness scores histogram and differences between partners

```{r weness_scores_histogram}

# Histogram of We-ness
plot_weness_hist <- bg_data %>%
  filter(!is.na(we_ness)) %>%
  ggplot(aes(x = we_ness)) +
  geom_histogram(binwidth = 0.1, fill = "forestgreen", color = "white") +
  labs(title = "Distribution of 'We-ness' Scores", 
       subtitle = "Average of 22 Positive items (No negative items)",
       x = "We-ness Score", y = "Count") +
  theme_minimal()
plot_weness_hist

# Save Histogram
plot_name <- sprintf("A-%d_weness_hist.png", plot_counter)
save_plot(plot_weness_hist, plot_name)
plot_counter <- plot_counter + 1

# Calculate Absolute Difference in We-ness
dci_dyadic <- dci_dyadic %>%
  mutate(
    weness_diff = abs(we_ness_P1 - we_ness_P2)
  )

cat("Summary of Absolute Differences in We-ness between Partners:\n")
summary(dci_dyadic$weness_diff) %>% print()

# Plot Histogram of Differences
plot_weness_diff <- dci_dyadic %>%
  filter(!is.na(weness_diff)) %>%
  ggplot(aes(x = weness_diff)) +
  geom_histogram(binwidth = 0.1, fill = "firebrick", color = "white") +
  labs(title = "Distribution of We-ness Differences", 
       subtitle = "Absolute difference between Partner 1 and Partner 2",
       x = "Difference Score (|P1 - P2|)", y = "Count of Couples") +
  theme_minimal()
plot_weness_diff

# Save plot
plot_name <- sprintf("A-%d_weness_diff_hist.png", plot_counter)
save_plot(plot_weness_diff, plot_name)
plot_counter <- plot_counter + 1
```

### Reliability of we-ness construct

```{r weness_reliability}

# A. Cronbach's Alpha
alpha_res <- bg_data %>%
    select(all_of(DCI_weness_items)) %>%
    alpha()
cat(sprintf("Cronbach's alpha for we-nesse = %f\n", alpha_res$total$raw_alpha))

# B. McDonald's Omega
omega_res <- bg_data %>%
    select(all_of(DCI_weness_items)) %>%
    omega(nfactors = 3, plot = FALSE)
cat(sprintf("\nMcDonald's omega for we-ness = %f\n\n", omega_res$omega.tot))
print(omega_res)
diagram <- omega.diagram(omega_res, main = "We-ness Factor Structure")
print(diagram)

# Save diagram
plot_name <- sprintf("A-%d_weness_factor_structure.png", plot_counter)
save_base_plot(diagram, plot_name)
plot_counter <- plot_counter + 1

```

A hierarchical factor analysis (Schmid-Leiman bifactor solution) was conducted to evaluate the internal structure of the We-Ness construct. The analysis supports a hierarchical structure consisting of one General Factor (*g*) and three distinct specific group factors, confirming that "We-Ness" is an overarching construct that encompasses both giving and receiving support.

**1. Reliability and Global Fit** The scale demonstrates excellent total reliability (ωt​=0.87), indicating that the items form a cohesive total scale. The Hierarchical Omega (ωh​=0.42) indicates that the General Factor accounts for a substantial portion (42%) of the common variance.

**2. Factor Interpretation** The variance is partitioned into a general "We-Ness" trait and three specific sub-domains corresponding to the provider of the support and the focus of communication:

-   **General Factor (*g*) — "Global We-Ness":** This factor represents the fundamental dyadic cohesion. It captures the shared tendency of the couple to function as a unit, regardless of the specific coping strategy employed.

-   **Factor 1 — Supportive Dyadic Coping (Perceived Partner Support):** *(Items: DCI 5, 6, 8, 9, 13, 14)* This factor captures the variance associated with **receiving support**. These items describe the individual's perception of their partner's active efforts to help them (e.g., "My partner shows empathy and understanding to me" ). It represents the "You help me" dimension.

-   **Factor 2 — Stress Communication Balance:** *(Items: Positive loading on 1–4, Negative loading on 16–19)* This bipolar factor captures the **communicative dynamic** of the dyad. It differentiates between focusing on one's own stress (Self-Communication, items 1–4) versus focusing on the partner's stress (Partner-Communication, items 16–19). This represents the "Who is talking?" dimension.

-   **Factor 3 — Supportive Dyadic Coping (Own Provided Support):** *(Items: DCI 20, 21, 24, 28, 29, 30)* This factor captures the variance associated with **providing support**. Unlike Factor 1, these items reflect the individual's own active efforts to support their partner (e.g., "I show empathy and understanding to my partner" ). It represents the "I help you" dimension.

### We-ness dyadic correlation

```{r weness_dyadic_correlation}

# Dyadic Correlation We-ness
cor_weness <- cor(dci_dyadic$we_ness_P1, dci_dyadic$we_ness_P2, use = "complete.obs")
header(sprintf("Dyadic Correlation (We-ness): r = %.3f", cor_weness))

# Plot Scatter We-ness
plot_weness_corr <- dci_dyadic %>%
  ggplot(aes(x = we_ness_P1, y = we_ness_P2)) +
  geom_point(alpha = 0.6, color = "darkgreen") +
  geom_smooth(method = "lm", color = "black", se = FALSE) +
  labs(title = "Dyadic Correlation: We-ness",
       subtitle = paste0("Pearson r = ", round(cor_weness, 3)),
       x = "Partner 1 We-ness", y = "Partner 2 We-ness") +
  theme_minimal()
plot_weness_corr

# Save plot
plot_name <- sprintf("A-%d_weness_scatter.png", plot_counter)
save_plot(plot_weness_corr, plot_name)
plot_counter <- plot_counter + 1

```

# VMR

```{r VMR_setup}
# load data
vmr_data  <- readRDS(file.path(dir_data_ana, "vmr_ana.rds"))
# load data configuration file 
load_config("VMR")
```

# POST

```{r POST_setup}
# load data
post_data  <- readRDS(file.path(dir_data_ana, "post_ana.rds"))
# load data configuration file 
load_config("POST")
```
