---
title: "Study Component 1: Affect Similarity in Daily Life (ESM)"
---

```{r setup}
#| include: false

# 1. Load packages, paths and data configurations
source(here::here("R", "00_setup.R"))

# 2. Set a global theme for plots
theme_set(theme_minimal() + theme(legend.position = "bottom"))

```

```{r ESM_setup}
#| include: false
# load data
esm_data  <- readRDS(file.path(dir_data_ana, "esm_ana.rds"))
# load data configuration file 
load_config("ESM")
# plot initialisation
plot_counter <- 1
prefix <- "ESM"
```

# Introduction

See also [OSF preregistration](https://doi.org/10.17605/OSF.IO/V7T9F){.uri}.

## Overview

Study Component 1 examines how **momentary affect similarity** in daily life relates to **momentary love toward the partner**, using experience sampling data from romantic couples.

We distinguish:

-   **Positive Affect (PA)** and **Negative Affect (NA)** (analyzed separately)
-   **Actual similarity** (self-reported affect compared to partner's self-reported affect)
-   **Perceived similarity** (self-reported affect compared to actor's perception of partner's affect)

Similarity is conceptualized as **shared, co-experienced elevated affect**, operationalized as:

-   Dyadic elevation (mean level)
-   Penalized by mismatch (absolute difference)

The main research questions are:

1.  Does momentary affect similarity predict momentary love? (H1)
2.  Is this association mediated by perceived similarity? (H2)
3.  Are these associations moderated by perceived responsiveness and event valence? (H3)

All analyses use multilevel models with moments (m) nested within persons (p) nested within dyads (d).

## Conceptualization and Operationalization

### Notation and Multilevel Structure

Let:

-   $L_{mpd}$ = momentary love
-   $A_{mpd}$ = actor affect
-   $P^{act}_{mpd}$ = partner's actual affect
-   $P^{perc}_{mpd}$ = partner affect as perceived by actor

Three-level structure:

-   Moments ($m$)
-   Persons ($p$)
-   Dyads ($d$)

### Person-Mean Centering

We person-mean center affect variables:

$$
X_{mpd} = A_{mpd} - \bar{A}_{pd}
$$

$$
Y^{act}_{mpd} = P^{act}_{mpd} - \bar{P}^{act}_{pd}
$$

$$
Y^{perc}_{mpd} = P^{perc}_{mpd} - \bar{P}^{perc}_{pd}
$$

### Similarity indices

#### Actual Similarity

Mean component:

$$
M^{act}_{mpd} = \frac{X_{mpd} + Y^{act}_{mpd}}{2}
$$

Mismatch component:

$$
D^{act}_{mpd} = |X_{mpd} - Y^{act}_{mpd}|
$$

Similarity index:

$$
S^{act}_{mpd} = M^{act}_{mpd} - D^{act}_{mpd}
$$

Actual affect similarity is symmetric within a couple at a given moment (it compares the two partners‚Äô self-reported affect), such that $S^{act}_{mpd}$ is conceptually a dyad‚Äìmoment construct that is identical for both partners.

#### Perceived Similarity

$$
M^{perc}_{mpd} = \frac{X_{mpd} + Y^{perc}_{mpd}}{2}
$$

$$
D^{perc}_{mpd} = |X_{mpd} - Y^{perc}_{mpd}|
$$

$$
S^{perc}_{mpd} = M^{perc}_{mpd} - D^{perc}_{mpd}
$$

Perceived similarity is actor-specific because it compares the actor‚Äôs affect to the actor‚Äôs perception of the partner‚Äôs affect, so $S^{perc}_{mpd}$ can differ between partners within the same dyad.

### Decomposition into Within and Between Components

We split between a between-dyad component and a within-dyad (momentary) component, to estimate and interpret separately the within-dyad time-varying association and the stable (between-dyad) association.

#### Actual Similarity

Dyad mean, and grand-mean:

$$
\bar{S}^{act}_d, \bar{\bar{S}}^{act}
$$

Within-dyad:

$$
S^{act,within}_{mpd} = S^{act}_{mpd} - \bar{S}^{act}_d
$$

Between-dyad:

$$
\bar{S}^{act,between}_d = \bar{S}^{act}_d - \bar{\bar{S}}^{act}
$$

The decomposition can be understood by specifying a model in different steps. For example, for the step 1a model (see below):

$$
L_{mpd} = b_{0,pd} + 
\beta_1 S^{act,within}_{mpd} +
\beta_2 \bar{S}^{act,between}_d +
\varepsilon_{mpd},\;\text{with}\; b_{0,pd} = \beta_0 + v_{0,d} + u_{0,pd}
$$

we could equally write

1.  Level 1 (modelling the momentary level): $L_{mpd} = \pi_{0,pd} + \beta_1 S^{act,within}_{mpd} + \varepsilon_{mpd},$

    with $\pi_{0,pd}$ the total person intercept (baseline love for person p in dyad d, averaged across moments). This is what happens within a person.

2.  Level 2 (modelling the person level, within a dyad): $œÄ_{0,pd}‚Äã=Œ¥_{0,d}‚Äã\;+\;u_{0,pd},$

    with $\delta_{0,d}$‚Äã the total dyad intercept (baseline love for dyad, averaged across partners) and $u_{0,pd}$ the person-specific random effect, capturing how much a person deviates from the dyad average.

3.  Level 3 (modelling the dyad level): $Œ¥_{0,d}‚Äã=Œ≤_0‚Äã+Œ≤_2‚Äã(\bar{S}_d^{act,between}‚Äã)+v_{0,d}‚Äã$

    with $v_{0,d}$ the dyad-specific random effect, capturing the couple effect (how much couples deviate from each other).

#### Perceived Similarity

Person mean, and grand-mean:

$$
\bar{S}^{perc}_{pd}, \bar{\bar{S}}^{perc}
$$

Within-person:

$$
S^{perc,within}_{mpd} = S^{perc}_{mpd} - \bar{S}^{perc}_{pd}
$$

Between-person:

$$
\bar{S}^{perc,between}_{pd} = \bar{S}^{perc}_{pd} - \bar{\bar{S}}^{perc}
$$

## Assumptions and Robustness Checks

### Random Effects

-   Random intercepts for dyads
-   Random intercepts for persons nested within dyads
-   No random slopes in primary models

### Temporal Dependence

#### CAR(1)

We use a continuous AR(1) residual covariance structure within persons. The preregistration specifies: "*Because beeps are unequally spaced, we will implement AR(1) using the actual time stamps (continuous-time AR(1))*". Beeps are unequally space for example because of missed beeps, no beeps during the night or during working hours, next to slight variations because of the protocol.

This means that we assume $\text{Corr}(\varepsilon_{mpd},\varepsilon_{{m'}pd})=\phi^{|\Delta t|}$ with $\Delta t$ the time difference between beeps at moment m and m'. An equivalent expression is $\text{Corr}(\varepsilon_{mpd},\varepsilon_{{m'}pd}) = \exp(-\lambda \Delta t)$ with $\lambda=-\log(\phi)$, $\lambda$ being a 'decay' parameter.

The unit/scale of the continuous time variable can be freely chosen and we have opted for **1 hour units**; the $\phi\in(0,1)$ will be estimated from the data and will reflect the correlation between residuals from observations that are one time unit apart. The choice of the time unit thus only determines the interpretation of $\phi$.

Note 1: If the time variable is chosen too small (e.g., seconds), the true $\phi$ might be effectively 0 or 1. If the likelihood surface is flat near the starting value of 0.2, the optimizer may simply return 0.2 as the final estimate without moving, leading to spurious results. If $\phi$ is estimated as 0.2, this can be a red flag. It is best to use the average time between beeps as time unit.

Note 2: Using lagged love in the equation instead of a CAR(1) residuals would model love inertia explicitly and answer a different question: does similarity predict love above prior love?

#### Time differences between beeps

Based on the columns 'time' (expressed in hours since the first beep for every person, increasing over the ESM period) and 'day_index', we can analyse how equal beeps are separated.

```{r within_dat_time_differences_between_beeps}
# time differences per person
esm_data_calc <- esm_data %>%
  group_by(person, day_index) %>%
  mutate(within_day_time_diff = time - lag(time)) %>%
  ungroup()

#  average difference per person within days
person_means <- esm_data_calc %>%
  group_by(person) %>%
  summarise(
    avg_diff_person = mean(within_day_time_diff, na.rm = TRUE)
  )

# overall grand-mean difference
grand_mean <- mean(esm_data_calc$within_day_time_diff, na.rm = TRUE)
sprintf("Overall Grand-Mean Difference: %f hours", grand_mean)

# histogram of person-average differences
plot_hist_numeric(
  data = person_means, 
  var = "avg_diff_person", 
  title = "Distribution of Person-Average Within-Day Time Intervals", 
  xlab = "Average Time Between Beeps (Hours)",
  binwidth = 1/60 # minutes
)
```

```{r overnight_time_differences_between_beeps}

# overnight differences
overnight_calc <- esm_data %>%
  group_by(person) %>%
  mutate(
    is_day_change = day_index > lag(day_index)
  ) %>%
  filter(is_day_change == TRUE) %>%
  mutate(overnight_time_diff = time - lag(time)) %>%
  ungroup()

# average overnight difference per person
person_overnight_means <- overnight_calc %>%
  group_by(person) %>%
  summarise(
    avg_overnight_diff = mean(overnight_time_diff, na.rm = TRUE)
  )

# overall grand-mean
grand_mean_overnight <- mean(overnight_calc$overnight_time_diff, na.rm = TRUE)
sprintf("Overall Grand-Mean Overnight Difference: %f hours", grand_mean_overnight)

# histogram
plot_hist_numeric(
  data = person_overnight_means, 
  var = "avg_overnight_diff", 
  title = "Distribution of Person-Average Overnight Intervals", 
  xlab = "Average Overnight Gap (Hours)",
  binwidth = 0.25 
)
```

### Sensitivity Analyses

1.  **Elevation sensitivity** (actor/partner affect covariates)

    We will in run all models with the individual affect levels as covariates to analyse what our similarity indices add above and beyond the individual affect levels. This is conceptually important because our similarity operationalization includes a level component (the mean of the individual affect scores), creating inherent overlap between similarity and individual levels.

    As an initial check, we inspected bivariate correlations on the within-person/dyad scale (i.e., between centered momentary love as(outcome) and candidate covariates, as in @selsActualPerceivedEmotional2020. These analyses (see`/scripts/05_descriptives.qmd`) indicated that momentary affect levels were small-to-moderately associated with momentary love (largest $r‚âà.34$) and showed substantial construct-induced overlap with the similarity indices.

    Given this overlap, models that omit affect levels would leave the similarity coefficient open to an ambiguous interpretation. Including affect levels therefore allows us to evaluate whether similarity predicts love beyond level effects.

2.  **Time-trend sensitivity**

    We tested for systematic time trends in momentary love across the ESM period by fitting $ùêø_{ùëöùëùùëë} = ùëè_{0,ùëùùëë} + ùõΩ_ùëáùëá_ùëö + ùúÄ_{ùëöùëùùëë}$, with $ùëá_ùëö$ the chronological order of beeps, centered at the study midpoint (see `/scripts/05_descriptives.qmd`), and found that there was a small, statistically significant but practically meaningless drift upwards (love increases over time). It is therefore **not indicated to rerun models with a time covariate**.

3.  **Random-slope robustness checks**

    We will attempt models that add random slopes for key level 1 predictors like within-dyad actual similarity or within-person perceived similarity.

## Model specification in R

We use the function `lme()` from the `nlme` package, designed for¬Ýlinear mixed-effects models with nested random effects, and it also allows¬Ýcorrelated within-group residuals (see <https://rdrr.io/cran/nlme/man/lme.html>). Contrary to `lme4`, `nlme` allows easy specification of residual correlation structures. See also @pinheiroNlmeLinearNonlinear1999 and @batesFittingLinearMixedEffects2015.

The lme() formula has the core signature:

> `lme(fixed, data, random, correlation, weights, subset, method, na.action, control, ...)`

with the arguments

-   `fixed` specifying the fixed-effects regression

-   `random` specifying the random effects and how the data is nested. \~ 1 indicates random intercepts, \~ x random slopes for variable x, and dyad/person applies this for dyad and person-within-dyad. x/y means that y is nested in x.

-   `correlation` adds a residual correlation structure. We use `correlation = corCAR1(form = ~ time | dyad/person)` to define a continuous-time AR(1) structure. Because we specify the grouping dyad/person, the correlarion only applies within the same group; different groups are treated as uncorrelated. The time variable does not need to be integer-valued. By default, $\phi$ is esrtimated starting from a value of $0.2$ and then allowed to vary to be estimated from the data. See also <https://rdrr.io/cran/nlme/man/corCAR1.html>.

-   `na.action` specifies \<hat to do with rows that have missing values. The default is `na.fail`, meaning that `lme()` will stop with an error. To allow missing beeps, we will specify `na.action = na.exclude`. Both na.omit and na.exclude drop incomplete rows for estimation

-   `control` adjusts optimisation behaviour and convergence (optimiser choice, iteration limits, tolerances,...), passing a list from `lmeControl()`. The optimiser choice is passed with `opt = c("nlminb","optim")`, defaulting to `"nlminb"` (nonlinear minimization subject to box constraints; generally robust. `optim` is a general-purpose optimiser offering several algorithms, like BFGS (Broyden-Fletcher-Goldfarb-Shanno) which sometimes navigates complex surfaces better (`optimMethod = "BFGS"`, the default when choosing `opt = "optim"`).

    We will use default values and switch only in case of convergence issues. In that case, we will increase the iteration limits and/or try `opt = "optim"`.

-   `method` chooses whether the model is fitted using restricted maximum likelihood ("REML", default) or full maximum likelihood ("ML"). We will use the default method = "REML". Note however that results are then no comparable across differenr fixed-effects model specifications (see guidance for `lme()` on <https://stat.ethz.ch/pipermail/r-help/2006-October/115739.html>). Comparing different fixed effects should use `method = "ML"`.

    Note that lme() does not scale internally, so in case of convergence problems, data needs to be pre-scaled.

------------------------------------------------------------------------

# Step 1a: Main Effects of Actual Similarity (H1.1-H1.2)

Primary multilevel model:

$$
L_{mpd} = b_{0,pd} + 
\beta_1 S^{act,within}_{mpd} +
\beta_2 \bar{S}^{act,between}_d +
\varepsilon_{mpd}
$$

Random intercept structure:

$$
b_{0,pd} = \beta_0 + v_{0,d} + u_{0,pd}
$$

-   $v_{0,d} \sim N(0, \tau_d^2)$
-   . .
-   Continuous time AR(1) residual structure within persons

Interpretation:

-   $\beta_1$: within-dyad association. When I (we) am (are) more similar **than usual** (moment-to-moment fluctuation), do I feel more love than usual?

-   $\beta_2$: between-dyad association. Do dyads who are generally/on average more similar report generally/on average higher levels of love?

```{r normality_test_love}
# histogram
plot_hist_numeric(esm_data, "love", "Ditribution of Outcome Variable 'love'", "love (0-100)",2)
# QQ-plot
esm_data %>%
  ggplot(aes(sample = love)) +
  stat_qq() +
  stat_qq_line(color = "red") +
  labs(title = "Normal Q-Q Plot",
       subtitle = "Points should follow the red line for normality") +
  theme_minimal()
# shapiro-wilk too sensitive for n > 5000
shapiro_results <- esm_data %>%
  slice_sample(n=3000) %>% # sample because shapiro.test works for n <5000
  pull(love) %>% # pull gives vector as needed by shapiro.test(), select gives df
  shapiro.test()

print(shapiro_results)
```

## H1.1 ‚Äì Actual PA Similarity

### Results

```{r h11_model}

esm_dat <- esm_data %>% arrange(dyad, person, time) # should already be the case

# --- model in nlme ---
h11 <- lme(
  fixed = love ~ PA_similarity_act_w + PA_similarity_act_b,
  random = ~ 1 | dyad/person,                 # v0,d and u0,pd
  correlation = corCAR1(form = ~ time | dyad/person),  # continuous AR(1) 
  data = esm_data,
  method = "REML",   # the default but specified for transparancy
  na.action = na.exclude,
  control = lmeControl(opt = "nlminb")   # the default but specified for transparancy
)

summary(h11)
intervals(h11, which = "fixed")  # CIs for fixed effects + variance components + AR(1) rho
VarCorr(h11)

# Optional: extract AR(1) parameter (rho)
coef(h11$modelStruct$corStruct, unconstrained = FALSE)

```

1.  **Within-person effect is significant:** $Œ≤_1=0.113$, $p<.001$, 95% CI:¬Ý$[0.104,¬Ý0.122]$

    When a person has higher-than-usual actual PA similarity with their partner at a given moment (relative to the typical similarity), they also report higher love at that moment.

    Within persons, a 1 SD increase in actual PA similarity was associated with a 2.9-point increase in momentary love (0‚Äì100).

2.  **Between-person/dyad effect is not significant**: $\beta_2=0.214$, $p = .43$ , 95% CI $[‚àí0.314,0.742]$.\
    only 102 dyads informing between estimate.

3.  $\phi=0.434$ is the estimated correlation between residuals for 1-hour time differences.

### Robustness/Sensitivity

There are no estimation changes when re-estimating the model with

-   `control = lmeControl(opt = "optim")`

-   `method = "ML"`

-   with `correlation = corCAR1(form = ~ time | dyad/person/day_index)`, which makes that the correlation between residuals from the last beep on a day and the first beep of the next day are explicitly set to 0.

### Fit diagnostics

```{r h11_fit_diagnostics}

diag_df <- data.frame(
  fitted = fitted(h11), # fitted values
  res_raw = resid(h11, type = "response")    # raw residuals
)

# residuals vs fitted (raw residuals)
plot(diag_df$fitted, diag_df$res_raw,
     xlab = "Fitted values",
     ylab = "Raw residuals (y - fitted)",
     main = "Raw residuals vs Fitted")
abline(h = 0, lty = 2)
lines(lowess(diag_df$fitted, diag_df$res_raw, delta = 1e-6), lwd = 2)

# QQ-plot of residuals
qqnorm(diag_df$res_raw, main = "Normal QQ-plot (raw residuals)")
qqline(diag_df$res_raw, lwd = 2)

# Scale-location plot (variance vs fitted)
plot(diag_df$fitted, sqrt(abs(diag_df$res_raw)),
     xlab = "Fitted values",
     ylab = "sqrt(|raw residual|)",
     main = "Scale-Location (heteroscedasticity check)")
lines(lowess(diag_df$fitted, sqrt(abs(diag_df$res_raw)), delta = 1e-6), lwd = 2)

# do fitted values exceed 100?
# Conditional fitted values (includes random effects)
pred_cond <- diag_df$fitted

# fixed-effects-only predictions for nlme objects
pred_fixed <- predict(h11, level = 0)

boundary_summary <- function(pred, upper = 100, lower = 0, name = "pred") {
  over  <- pred > upper
  under <- pred < lower
  data.frame(
    which = name,
    n = length(pred),
    n_over = sum(over, na.rm = TRUE),
    pct_over = 100 * mean(over, na.rm = TRUE),
    max_over = if (any(over, na.rm = TRUE)) max(pred[over], na.rm = TRUE) else NA_real_,
    n_under = sum(under, na.rm = TRUE),
    pct_under = 100 * mean(under, na.rm = TRUE),
    min_under = if (any(under, na.rm = TRUE)) min(pred[under], na.rm = TRUE) else NA_real_
  )
}

bound_cond  <- boundary_summary(pred_cond,  name = "conditional (includes RE)")
bound_fixed <- boundary_summary(pred_fixed, name = "fixed only (level=0)")

rbind(bound_cond, bound_fixed)

## visualize where violations occur
plot(pred_cond, pch = 16,
     ylab = "Conditional fitted love",
     xlab = "Observation index",
     main = "Conditional fitted values with 0/100 bounds")
abline(h = c(0, 100), lty = 2)

## ---- 6) Quick check: residuals near the ceiling (optional) ----
# If you have the observed y aligned, check residual patterns at high observed values

y_obs <- esm_data$love
ok <- !is.na(y_obs) & !is.na(pred_cond) & !is.na(res_pearson)

plot(y_obs[ok], res_pearson[ok],
     xlab = "Observed love (0-100)",
     ylab = "Pearson residuals",
     main = "Residuals vs Observed (ceiling check)")
abline(h = 0, lty = 2)
lines(lowess(y_obs[ok], res_pearson[ok], delta = 1e-6), lwd = 2)

```

## H1.2 ‚Äì Actual NA Similarity

```{r}

```

------------------------------------------------------------------------

# Step 1b: Additive Model (H1.3)

$$
L_{mpd} =
b_{0,pd} +
\beta_1 S^{act,within}_{mpd} +
\beta_2 \bar{S}^{act,between}_d +
\beta_3 S^{perc,within}_{mpd} +
\beta_4 \bar{S}^{perc,between}_{pd} +
\varepsilon_{mpd}
$$

Goals:

-   Test whether perceived similarity predicts love above actual similarity

-   Examine attenuation of actual similarity coefficients

Interpretation:

-   $ùõΩ_3$ indicates whether at moments when an actor perceives more similarity than they typically do (i.e., higher $S^{perc, within}_{mpd}$), they report higher/lower love over and above what can be explained by actual similarity at that moment.

-   $ùõΩ_4$ indicates whether persons who tend to perceive more similarity on average report higher/lower love overall, controlling for the dyad‚Äôs average actual similarity. As noted in the preregistration, this between-person component combines differences between dyads and stable differences between partners within dyads.

If $ùõΩ_3$ and $ùõΩ_4$ are comparable in direction to the corresponding actual-similarity effects in Step 1a, and the actual-similarity association with love are being reduced in magnitude when perceived similarity is included, this is consistent with H1.3.

## H1.3 ‚Äì Adding Perceived Similarity

```{r}

```

------------------------------------------------------------------------

# Step 1c: Actual √ó Perceived Interaction

## Exploratory analysis

$$
L_{mpd} =
b_{0,pd} +
\beta_1 S^{act,within}_{mpd} +
\beta_2 \bar{S}^{act,between}_d +
\beta_3 S^{perc,within}_{mpd} +
\beta_4 \bar{S}^{perc,between}_{pd}
$$$$
+ \beta_5 (S^{act,within}_{mpd} \times S^{perc,within}_{mpd})
+ \beta_6 (\bar{S}^{act,between}_d \times \bar{S}^{perc,between}_{pd})
+ \varepsilon_{mpd}
$$

Predicted values will be probed at ‚àí1 SD, mean, and +1 SD.

```{r}

```

------------------------------------------------------------------------

# Step 2: Multilevel mediation (H2.1-H2.2)

We test whether perceived similarity mediates the association between actual similarity and love.

**Mediator Model**

$$
S^{perc}_{mpd} =
t_{0,pd} +
\theta_1 S^{act,within}_{mpd} +
\theta_2 \bar{S}^{act,between}_d +
\varepsilon^M_{mpd}
$$

**Outcome Model**

$$
L_{mpd} =
f_{0,pd} +
\phi_1 S^{perc,within}_{mpd} +
\phi_2 \bar{S}^{perc,between}_{pd} +
\phi_3 S^{act,within}_{mpd} +
\phi_4 \bar{S}^{act,between}_d +
\varepsilon^L_{mpd}
$$

Indirect effects:

-   Within-person: $\theta_1 \phi_1$
-   Between-dyad: $\theta_2 \phi_2$

## H2.1 - PA mediation

```{r}

```

## H2.2 - NA mediation

```{r}

```

------------------------------------------------------------------------

# Step 3a: Moderation by responsiveness (H3.1-H3.2)

General model structure:

$$
L_{mpd} =
... +
\beta (Similarity \times Responsiveness) +
\varepsilon_{mpd}
$$

Interpretation: - Responsiveness as amplifier (PA) - Responsiveness as buffer (NA)

## H3.1 ‚Äì PA Similarity √ó Responsiveness

```{r}

```

## H3.2 ‚Äì NA Similarity √ó Responsiveness

```{r}

```

------------------------------------------------------------------------

# Step 3b: Moderation by Event Valence (H3.3-H3.4)

Event valence derived from:

-   Positive event
-   Negative event

Mismatch hypotheses:

-   PA similarity strongest in negative moments
-   NA similarity most harmful in positive moments

## H3.3 ‚Äì PA Similarity √ó Event Valence

```{r}

```

## H3.4 ‚Äì NA Similarity √ó Event Valence

```{r}

```
