---
title: "Study Component 1: Affect Similarity in Daily Life (ESM)"
---

```{r setup}
#| include: false

# 1. Load packages, paths and data configurations
source(here::here("R", "00_setup.R"))

# 2. Set a global theme for plots
theme_set(theme_minimal() + theme(legend.position = "bottom"))

```

```{r ESM_setup}
#| include: false
# load data
esm_data  <- readRDS(file.path(dir_data_ana, "esm_ana.rds"))
# load data configuration file 
load_config("ESM")
# plot initialisation
plot_counter <- 1
prefix <- "ESM"
```

# Introduction

See also [OSF preregistration](https://doi.org/10.17605/OSF.IO/V7T9F){.uri}.

## Overview

Study Component 1 examines how **momentary affect similarity** in daily life relates to **momentary love toward the partner**, using experience sampling data from romantic couples.

We distinguish:

-   **Positive Affect (PA)** and **Negative Affect (NA)** (analyzed separately)
-   **Actual similarity** (self-reported affect compared to partner's self-reported affect)
-   **Perceived similarity** (self-reported affect compared to actor's perception of partner's affect)

Similarity is conceptualized as **shared, co-experienced elevated affect**, operationalized as:

-   Dyadic elevation (mean level)
-   Penalized by mismatch (absolute difference)

The main research questions are:

1.  Does momentary affect similarity predict momentary love? (H1)
2.  Is this association mediated by perceived similarity? (H2)
3.  Are these associations moderated by perceived responsiveness and event valence? (H3)

All analyses use multilevel models with moments (m) nested within persons (p) nested within dyads (d).

## Conceptualization and Operationalization

### Notation and Multilevel Structure

Let:

-   $L_{mpd}$ = momentary love
-   $A_{mpd}$ = actor affect
-   $P^{act}_{mpd}$ = partner's actual affect
-   $P^{perc}_{mpd}$ = partner affect as perceived by actor

Three-level structure:

-   Moments ($m$)
-   Persons ($p$)
-   Dyads ($d$)

### Person-Mean Centering

We person-mean center affect variables:

$$
X_{mpd} = A_{mpd} - \bar{A}_{pd}
$$

$$
Y^{act}_{mpd} = P^{act}_{mpd} - \bar{P}^{act}_{pd}
$$

$$
Y^{perc}_{mpd} = P^{perc}_{mpd} - \bar{P}^{perc}_{pd}
$$

### Similarity indices

#### Actual Similarity

Mean component:

$$
M^{act}_{mpd} = \frac{X_{mpd} + Y^{act}_{mpd}}{2}
$$

Mismatch component:

$$
D^{act}_{mpd} = |X_{mpd} - Y^{act}_{mpd}|
$$

Similarity index:

$$
S^{act}_{mpd} = M^{act}_{mpd} - D^{act}_{mpd}
$$

Actual affect similarity is symmetric within a couple at a given moment (it compares the two partners’ self-reported affect), such that $S^{act}_{mpd}$ is conceptually a dyad–moment construct that is identical for both partners.

#### Perceived Similarity

$$
M^{perc}_{mpd} = \frac{X_{mpd} + Y^{perc}_{mpd}}{2}
$$

$$
D^{perc}_{mpd} = |X_{mpd} - Y^{perc}_{mpd}|
$$

$$
S^{perc}_{mpd} = M^{perc}_{mpd} - D^{perc}_{mpd}
$$

Perceived similarity is actor-specific because it compares the actor’s affect to the actor’s perception of the partner’s affect, so $S^{perc}_{mpd}$ can differ between partners within the same dyad.

### Decomposition into Within and Between Components

We split between a between-dyad component and a within-dyad (momentary) component, to estimate and interpret separately the within-dyad time-varying association and the stable (between-dyad) association.

#### Actual Similarity

Dyad mean, and grand-mean:

$$
\bar{S}^{act}_d, \bar{\bar{S}}^{act}
$$

Within-dyad:

$$
S^{act,within}_{mpd} = S^{act}_{mpd} - \bar{S}^{act}_d
$$

Between-dyad:

$$
\bar{S}^{act,between}_d = \bar{S}^{act}_d - \bar{\bar{S}}^{act}
$$

The decomposition can be understood by specifying a model in different steps. For example, for the step 1a model (see below):

$$
L_{mpd} = b_{0,pd} + 
\beta_1 S^{act,within}_{mpd} +
\beta_2 \bar{S}^{act,between}_d +
\varepsilon_{mpd},\;\text{with}\; b_{0,pd} = \beta_0 + v_{0,d} + u_{0,pd}
$$

we could equally write

1.  Level 1 (modelling the momentary level): $L_{mpd} = \pi_{0,pd} + \beta_1 S^{act,within}_{mpd} + \varepsilon_{mpd},$

    with $\pi_{0,pd}$ the total person intercept (baseline love for person p in dyad d, averaged across moments). This is what happens within a person.

2.  Level 2 (modelling the person level, within a dyad): $π_{0,pd}​=δ_{0,d}​\;+\;u_{0,pd},$

    with $\delta_{0,d}$​ the total dyad intercept (baseline love for dyad, averaged across partners) and $u_{0,pd}$ the person-specific random effect, capturing how much a person deviates from the dyad average.

3.  Level 3 (modelling the dyad level): $δ_{0,d}​=β_0​+β_2​(\bar{S}_d^{act,between}​)+v_{0,d}​$

    with $v_{0,d}$ the dyad-specific random effect, capturing the couple effect (how much couples deviate from each other).

#### Perceived Similarity

Person mean, and grand-mean:

$$
\bar{S}^{perc}_{pd}, \bar{\bar{S}}^{perc}
$$

Within-person:

$$
S^{perc,within}_{mpd} = S^{perc}_{mpd} - \bar{S}^{perc}_{pd}
$$

Between-person:

$$
\bar{S}^{perc,between}_{pd} = \bar{S}^{perc}_{pd} - \bar{\bar{S}}^{perc}
$$

## Assumptions and Robustness Checks

### Random Effects

-   Random intercepts for dyads
-   Random intercepts for persons nested within dyads
-   No random slopes in primary models

### Temporal Dependence

#### CAR(1)

We use a continuous AR(1) residual covariance structure within persons. The preregistration specifies: "*Because beeps are unequally spaced, we will implement AR(1) using the actual time stamps (continuous-time AR(1))*". Beeps are unequally space for example because of missed beeps, no beeps during the night or during working hours, next to slight variations because of the protocol.

This means that we assume $\text{Corr}(\varepsilon_{mpd},\varepsilon_{{m'}pd})=\phi^{|\Delta t|}$ with $\Delta t$ the time difference between beeps at moment m and m'. An equivalent expression is $\text{Corr}(\varepsilon_{mpd},\varepsilon_{{m'}pd}) = \exp(-\lambda \Delta t)$ with $\lambda=-\log(\phi)$, $\lambda$ being a 'decay' parameter.

The unit/scale of the continuous time variable can be freely chosen and we have opted for **1 hour units**; the $\phi\in(0,1)$ will be estimated from the data and will reflect the correlation between residuals from observations that are one time unit apart. The choice of the time unit thus only determines the interpretation of $\phi$.

Note 1: If the time variable is chosen too small (e.g., seconds), the true $\phi$ might be effectively 0 or 1. If the likelihood surface is flat near the starting value of 0.2, the optimizer may simply return 0.2 as the final estimate without moving, leading to spurious results. If $\phi$ is estimated as 0.2, this can be a red flag. It is best to use the average time between beeps as time unit.

Note 2: Using lagged love in the equation instead of a CAR(1) residuals would model love inertia explicitly and answer a different question: does similarity predict love above prior love?

#### Time differences between beeps

Based on the columns 'time' (expressed in hours since the first beep for every person, increasing over the ESM period) and 'day_index', we can analyse how equal beeps are separated.

```{r within_dat_time_differences_between_beeps}
# time differences per person
esm_data_calc <- esm_data %>%
  group_by(person, day_index) %>%
  mutate(within_day_time_diff = time - lag(time)) %>%
  ungroup()

#  average difference per person within days
person_means <- esm_data_calc %>%
  group_by(person) %>%
  summarise(
    avg_diff_person = mean(within_day_time_diff, na.rm = TRUE)
  )

# overall grand-mean difference
grand_mean <- mean(esm_data_calc$within_day_time_diff, na.rm = TRUE)
sprintf("Overall Grand-Mean Difference: %f hours", grand_mean)

# histogram of person-average differences
plot_hist_numeric(
  data = person_means, 
  var = "avg_diff_person", 
  title = "Distribution of Person-Average Within-Day Time Intervals", 
  xlab = "Average Time Between Beeps (Hours)",
  binwidth = 1/60 # minutes
)
```

```{r overnight_time_differences_between_beeps}

# overnight differences
overnight_calc <- esm_data %>%
  group_by(person) %>%
  mutate(
    is_day_change = day_index > lag(day_index)
  ) %>%
  filter(is_day_change == TRUE) %>%
  mutate(overnight_time_diff = time - lag(time)) %>%
  ungroup()

# average overnight difference per person
person_overnight_means <- overnight_calc %>%
  group_by(person) %>%
  summarise(
    avg_overnight_diff = mean(overnight_time_diff, na.rm = TRUE)
  )

# overall grand-mean
grand_mean_overnight <- mean(overnight_calc$overnight_time_diff, na.rm = TRUE)
sprintf("Overall Grand-Mean Overnight Difference: %f hours", grand_mean_overnight)

# histogram
plot_hist_numeric(
  data = person_overnight_means, 
  var = "avg_overnight_diff", 
  title = "Distribution of Person-Average Overnight Intervals", 
  xlab = "Average Overnight Gap (Hours)",
  binwidth = 0.25 
)
```

### Sensitivity Analyses

1.  **Elevation sensitivity** (actor/partner affect covariates)

    We will in run all models with the individual affect levels as covariates to analyse what our similarity indices add above and beyond the individual affect levels. This is conceptually important because our similarity operationalization includes a level component (the mean of the individual affect scores), creating inherent overlap between similarity and individual levels.

    As an initial check, we inspected bivariate correlations on the within-person/dyad scale (i.e., between centered momentary love as(outcome) and candidate covariates, as in @selsActualPerceivedEmotional2020. These analyses (see`/scripts/05_descriptives.qmd`) indicated that momentary affect levels were small-to-moderately associated with momentary love (largest $r≈.34$) and showed substantial construct-induced overlap with the similarity indices.

    Given this overlap, models that omit affect levels would leave the similarity coefficient open to an ambiguous interpretation. Including affect levels therefore allows us to evaluate whether similarity predicts love beyond level effects.

2.  **Time-trend sensitivity**

    We tested for systematic time trends in momentary love across the ESM period by fitting $𝐿_{𝑚𝑝𝑑} = 𝑏_{0,𝑝𝑑} + 𝛽_𝑇𝑇_𝑚 + 𝜀_{𝑚𝑝𝑑}$, with $𝑇_𝑚$ the chronological order of beeps, centered at the study midpoint (see `/scripts/05_descriptives.qmd`), and found that there was a small, statistically significant but practically meaningless drift upwards (love increases over time). It is therefore **not indicated to rerun models with a time covariate**.

3.  **Random-slope robustness checks**

    We will attempt models that add random slopes for key level 1 predictors like within-dyad actual similarity or within-person perceived similarity.

## Model specification in R

We use the function `lme()` from the `nlme` package, designed for linear mixed-effects models with nested random effects, and it also allows correlated within-group residuals (see <https://rdrr.io/cran/nlme/man/lme.html>). Contrary to `lme4`, `nlme` allows easy specification of residual correlation structures. See also @pinheiroNlmeLinearNonlinear1999 and @batesFittingLinearMixedEffects2015.

The lme() formula has the core signature:

> `lme(fixed, data, random, correlation, weights, subset, method, na.action, control, ...)`

with the arguments

-   `fixed` specifying the fixed-effects regression

-   `random` specifying the random effects and how the data is nested. \~ 1 indicates random intercepts, \~ x random slopes for variable x, and dyad/person applies this for dyad and person-within-dyad. x/y means that y is nested in x.

-   `correlation` adds a residual correlation structure. We use `correlation = corCAR1(form = ~ time | dyad/person)` to define a continuous-time AR(1) structure. Because we specify the grouping dyad/person, the correlarion only applies within the same group; different groups are treated as uncorrelated. The time variable does not need to be integer-valued. By default, $\phi$ is esrtimated starting from a value of $0.2$ and then allowed to vary to be estimated from the data. See also <https://rdrr.io/cran/nlme/man/corCAR1.html>.

-   `na.action` specifies \<hat to do with rows that have missing values. The default is `na.fail`, meaning that `lme()` will stop with an error. To allow missing beeps, we will specify `na.action = na.exclude`. Both na.omit and na.exclude drop incomplete rows for estimation

-   `control` adjusts optimisation behaviour and convergence (optimiser choice, iteration limits, tolerances,...), passing a list from `lmeControl()`. The optimiser choice is passed with `opt = c("nlminb","optim")`, defaulting to `"nlminb"` (nonlinear minimization subject to box constraints; generally robust. `optim` is a general-purpose optimiser offering several algorithms, like BFGS (Broyden-Fletcher-Goldfarb-Shanno) which sometimes navigates complex surfaces better (`optimMethod = "BFGS"`, the default when choosing `opt = "optim"`).

    We will use default values and switch only in case of convergence issues. In that case, we will increase the iteration limits and/or try `opt = "optim"`.

-   `method` chooses whether the model is fitted using restricted maximum likelihood ("REML", default) or full maximum likelihood ("ML"). We will use the default method = "REML". Note however that results are then no comparable across differenr fixed-effects model specifications (see guidance for `lme()` on <https://stat.ethz.ch/pipermail/r-help/2006-October/115739.html>). Comparing different fixed effects should use `method = "ML"`.

    Note that lme() does not scale internally, so in case of convergence problems, data needs to be pre-scaled.

------------------------------------------------------------------------

# Step 1a: Main Effects of Actual Similarity (H1.1-H1.2)

Primary multilevel model:

$$
L_{mpd} = b_{0,pd} + 
\beta_1 S^{act,within}_{mpd} +
\beta_2 \bar{S}^{act,between}_d +
\varepsilon_{mpd}
$$

Random intercept structure:

$$
b_{0,pd} = \beta_0 + v_{0,d} + u_{0,pd}
$$

-   $v_{0,d} \sim N(0, \tau_d^2)$
-   $𝑢_{0,𝑝𝑑}∼𝑁(0, 𝜏_𝑝^2)$

With continuous time AR(1) residual structure within persons.

Interpretation:

-   $\beta_1$: within-dyad association. When I (we) am (are) more similar **than usual** (moment-to-moment fluctuation), do I feel more love than usual?

-   $\beta_2$: between-dyad association. Do dyads who are generally/on average more similar report generally/on average higher levels of love?

## H1.1 – Actual PA Similarity

### Results

```{r h11_model}

esm_data <- esm_data %>% arrange(dyad, person, time) # should already be the case

# model in nlme
h11 <- lme(
   fixed = love ~ PA_similarity_act_w + PA_similarity_act_b,
   random = ~ 1 | dyad/person,                 # v0,d and u0,pd
   correlation = corCAR1(form = ~ time | dyad/person),  # continuous AR(1) 
   data = esm_data,
   method = "REML",   # the default but specified for transparancy
   na.action = na.exclude,
   control = lmeControl(opt = "nlminb")   # default but specified for transparancy
)

summary(h11)
intervals(h11, which = "fixed")  # 95% CIs for fixed effects
VarCorr(h11)

```

1.  $β_0=75.94$, $p<.001$, 95% CI: $[73.48,78.38]$ is the reported love when all predictors are zero, thus for the average dyad at an average moment (given the centering).

2.  **Within-person effect is significant:** $β_1=0.113$, $p<.001$, 95% CI: $[0.104, 0.122]$

    When a person has higher-than-usual actual PA similarity with their partner at a given moment (relative to the typical similarity), they also report higher love at that moment.

    Within persons, a 10 point increase in actual PA similarity was associated with a 1.13-point (CI \[1.04, 1.22\])increase in momentary love (0–100).

3.  **Between-person/dyad effect is not significant**: $\beta_2=0.214$, $p = .43$ , 95% CI $[−0.314,0.742]$.\
    only 102 dyads informing between estimate.

    So there is no reliable evidence that dyads who are (on average) more similar also have (on average) higher love.

4.  The dyad random intercept $SD = 9.82$ (26% of the variance) and person-within-dyad random intercept $SD = 11.09$ (34% of the variance). Residual $SD = 12.05$ (40% of the variance), the remaining moment-to-moment unexplained variability.

5.  $\phi=0.434$ is the estimated correlation between residuals for 1-hour time differences.

### Fit diagnostics

#### Heteroscedasticity and Linearity

Heteroscadasticity: normalized residuals should look like independent, uncorrelated noise (homoscedastic and without patterns).

Linearity: We add the loess curve (lowess = LOcally WEighted Scatterplot Smoother; local regression lines in a small smoothing window forming a smooth continuous line), which should be relatively flat and close to the horizontal zero line .

Note: there are three types of residuals:

-   raw (`type = "response"`**):** observed value minus predicted value; generally not useful for diagnostics in complexer multilevel model.

-   standardized or Pearson (`type = "pearson"`): raw residuals divided by the estimated standard deviation (standard error) of the errors, putting the residuals on a standard normal scale (mean =0, SD = 1). These residuals will still exhibit the autocorrelation structure that was explicitly modeled.

-   normalized (`type = "normalized"`). these are the standardized residuals but 'uncorrelated' by the specified correlation structure ("standardized residuals pre-multiplied by the inverse of the error correlation matrix").

```{r h11_residuals}
#| warning: false

# Use augment() from broom package
diag_data <- getData(h11) %>%
  as_tibble() %>%
  mutate(
    .fitted     = fitted(h11),
    .resid_norm = resid(h11, type = "normalized")
  )

# Plot residuals vs fitted
ggplot(diag_data, aes(x = .fitted, y = .resid_norm)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  geom_smooth(method = "loess", se = FALSE, color = "blue") +
  labs(title = "Normalized Residuals vs. Fitted Values",
       subtitle = "Check for linearity (loess) and homoscedasticity",
       x = "Fitted Values",
       y = "Normalized Residuals") +
  theme_minimal()

plot_hist_numeric(esm_data, "love", "Ditribution of Outcome Variable 'love'", "love (0-100)",2)

# residuals by dyad
ggplot(diag_data, aes(x = factor(dyad), y = .resid_norm)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.5) +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(
    title = "Normalized Residuals by Dyad",
    subtitle = "Homoscedasticity across dyads",
    x = "Dyad",
    y = "Normalized Residuals"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 6)) 

# residuals vs time
ggplot(diag_data, aes(x = time, y = .resid_norm)) +
  geom_point(alpha = 0.3) +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  geom_smooth(method = "loess", se = FALSE, color = "blue") +
  labs(
    title = "Normalized Residuals vs. Time",
    subtitle = "Check for: Constant spread (Stationarity) and Flat Line (No missing time trends)",
    x = "Time (Beep/Day)",
    y = "Normalized Residuals"
  ) +
  theme_minimal()
```

There is **clear heteroscedasticity**, violating the constant variance assumption. This is likely because the **outcome scale is bounded to a maximum of 100** and thus compressing the variance near the ceiling. See also the histogram of outcome variable `love`.

Other observations:

-   There are some strong outliers

-   The LOESS line is smooth and close to flat

-   There seems to be a small subgroup with low predicted love

Looking at the normalized residuals per dyad:

-   No strong evidence that certain dyads are systematically over- or under-predicted. The random dyad intercept seems to be doing its job in capturing between-dyad mean differences.

-   Some variability in residual spread across dyads, but not extreme or structurally patterned. No strong dyad-specific heteroscedasticity

Looking at the residuals in function of time, we again see the small time trend, which we previosuly tested and found to be significant but immaterial.

#### Normality of residuals (QQ-plot)

```{r}
# QQ-plot of residuals
qqnorm(diag_data$.resid_norm, main = "Normal QQ-plot (normalized residuals)")
qqline(diag_data$.resid_norm, lwd = 2)
```

In the central part of the distribution, between quantiles -1.5 and +1.5, the points are close to the qq-line, and thus approximate normality. There are however strong deviations in the tails, especially the left tail. Again, it is likely that this is because of the bounded scale, where near the +100, the positive residuals are compressed and the residual distribution becomes naturally left-skewed (heavy negative tail).

#### Remaining autocorrelation

```{r h11_remaining_autocorrelation}
plot(ACF(h11, resType = "normalized", maxLag = 10), 
     alpha = 0.05, 
     main = "ACF of Normalized Residuals")
```

This plot suggests there is **some remaining short-term serial dependence** in the residuals that the `corCAR1(~ time | dyad/person)` does not fully absorb.

#### Prediction vs data (boundary check)

```{r h11_fit_vs_data}
#| warning: false

# conditional on all random intercepts
pred_cond <- diag_data$.fitted
# fixed-effects-only
pred_fixed <- predict(h11, level = 0)

bound_cond  <- boundary_summary(pred_cond,  name = "conditional (includes RE)")
bound_fixed <- boundary_summary(pred_fixed, name = "fixed only")

rbind(bound_cond, bound_fixed)

# visualize where violations occur - conditional
plot(pred_cond, pch = 16,
     ylab = "Mixed effect (fixed + random effects) predicitions of love",
     xlab = "Beeps",
     main = "Mixed effect fitted values with 0/100 bounds")
abline(h = c(0, 100), lty = 2)

plot(pred_fixed, pch = 16,
     ylab = "Fixed effect predictions of love",
     xlab = "Beeps",
     main = "Fixed effect fitted values with 0/100 bounds")
abline(h = c(0, 100), lty = 2)

# --- Observed vs predicted (more direct "fit vs data" check) ---
diag_plot <- diag_data %>%
  mutate(
    pred_cond  = pred_cond,
    pred_fixed = pred_fixed
  ) %>%
  pivot_longer(
    cols = c(pred_cond, pred_fixed),
    names_to = "pred_type",
    values_to = "pred"
  ) %>%
  mutate(
    pred_type = recode(
      pred_type,
      pred_cond = "mixed effect (all RE)",
      pred_fixed = "fixed only (level = 0)"
    )
  )

ggplot(diag_plot, aes(x = pred, y = love)) +
  geom_point(alpha = 0.25) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  coord_cartesian(xlim = c(0, 100), ylim = c(0, 100)) +
  facet_wrap(~ pred_type) +
  labs(
    title = "Observed vs predicted love",
    subtitle = "Dashed line = perfect prediction; bounds shown via axis limits (0–100)",
    x = "Predicted love",
    y = "Observed love"
  ) +
  theme_minimal()

```

-   **Fixed-effect predictions:** Because only the fixed effects are used, predicted love varies little (horizontally). The large vertical scatter around these predictions mainly reflects moment-to-moment residual variation (Residual $variance = 145.17, SD = 12.05$).

<!-- -->

-   **Mixed-effect predictions**: Adding dyad/person random intercepts shifts predictions across units, producing extra horizontal spread in predicted love due to between-unit baseline differences (Dyad $variance = 96.47, SD = 9.82$; Person-within-dyad $variance = 123.02, SD = 11.09$), while the remaining vertical scatter is still driven by the residual ($variance = 145.17, SD = 12.05$).

-   Clear boundary effects

### Robustness/Sensitivity

There are no estimation changes when re-estimating the model with

-   `control = lmeControl(opt = "optim")`

-   `method = "ML"`

-   with `correlation = corCAR1(form = ~ time | dyad/person/day_index)`, which makes that the correlation between residuals from the last beep on a day and the first beep of the next day are explicitly set to 0.

#### Adding random slopes

```{r h11_model_random_slopes}

# esm_data <- esm_data %>%
#   mutate(dyad_person = interaction(dyad, person, drop = TRUE))
# 
# # model in nlme
# h11_rs_dyad <- lme(
#   fixed = love ~ PA_similarity_act_w + PA_similarity_act_b,
#   random = list(
#     dyad        = pdDiag(~ 1 + PA_similarity_act_w),  # dyad-specific intercept + slope
#     dyad_person = ~ 1                                      # person-specific intercept
#   ),
#   correlation = corCAR1(form = ~ time | dyad_person),
#   data = esm_data,
#   method = "REML",
#   na.action = na.exclude,
#   control = lmeControl(opt = "nlminb")
# )
# 
# summary(h11)
# intervals(h11, which = "fixed")  # 95% CIs for fixed effects
# VarCorr(h11)
```

## H1.2 – Actual NA Similarity

### Results

```{r h12_model}

h12 <- lme(
   fixed = love ~ NA_similarity_act_w + NA_similarity_act_b,
   random = ~ 1 | dyad/person,                 # v0,d and u0,pd
   correlation = corCAR1(form = ~ time | dyad/person),  # continuous AR(1) 
   data = esm_data,
   method = "REML",   # the default but specified for transparancy
   na.action = na.exclude,
   control = lmeControl(opt = "nlminb")   # default but specified for transparancy
)

summary(h12)
intervals(h12, which = "fixed")  # 95% CIs for fixed effects
VarCorr(h12)

```

------------------------------------------------------------------------

1.  $β_0=75.91$, $p<.001$, 95% CI: $[73.54,78.27]$ is the reported love when all predictors are zero, thus for the average dyad at an average moment (given the centering).

2.  **Within-person effect is significant:** $β_1=-0.152$, $p<.001$, 95% CI: $[-0.167, -0.136]$

    When a dyad has higher-than-usual actual NA similarity at a given moment (relative to the typical similarity), partners report lower love at that moment.

    Within persons, a 10 point increase in actual NA similarity was associated with a $-1.52$-point (CI $[-1.67, -1.36]$)increase in momentary love (0–100).

3.  **Between-person/dyad effect is significant**: $\beta_2=0.621$, $p < .01$ , 95% CI $[0.197,1.045]$.

    Dyad that are higher on average in actual NA similarity, report **higher** love at that moment. A 10-point increase in actual similarity leads to a 6.2 point increase in love.

4.  The dyad random intercept $SD = 9.22$ and person-within-dyad random intercept $SD = 11.09$. Residual $SD = 12.18$, the remaining moment-to-moment unexplained variability.

5.  $\phi=0.447$ is the estimated correlation between residuals for 1-hour time differences.

Because our similarity index reflects co-elevated negative affect (both partners high together), the negative within-dyad effect likely captures acute distress episodes during which love temporarily decreases, whereas the positive between-dyad effect may reflect stable couple differences (e.g., assortative mating, convergence, shared-stressor context and/or stronger communal coping among dyads that frequently experience and coordinate around shared distress).

# Step 1b: Additive Model (H1.3)

$$
L_{mpd} =
b_{0,pd} +
\beta_1 S^{act,within}_{mpd} +
\beta_2 \bar{S}^{act,between}_d +
\beta_3 S^{perc,within}_{mpd} +
\beta_4 \bar{S}^{perc,between}_{pd} +
\varepsilon_{mpd}
$$

Goals:

-   Test whether perceived similarity predicts love above actual similarity

-   Examine attenuation of actual similarity coefficients

Interpretation:

-   $𝛽_3$ indicates whether at moments when an actor perceives more similarity than they typically do (i.e., higher $S^{perc, within}_{mpd}$), they report higher/lower love over and above what can be explained by actual similarity at that moment.

-   $𝛽_4$ indicates whether persons who tend to perceive more similarity on average report higher/lower love overall, controlling for the dyad’s average actual similarity. As noted in the preregistration, this between-person component combines differences between dyads and stable differences between partners within dyads.

If $𝛽_3$ and $𝛽_4$ are comparable in direction to the corresponding actual-similarity effects in Step 1a, and the actual-similarity association with love are being reduced in magnitude when perceived similarity is included, this is consistent with H1.3.

## H1.3 – Adding Perceived Similarity

### H1.3 - PA

```{r h13_model_PA}

h13_PA <- lme(
   fixed = love ~ PA_similarity_act_w + PA_similarity_act_b + PA_similarity_perc_w + PA_similarity_perc_b,
   random = ~ 1 | dyad/person,                 # v0,d and u0,pd
   correlation = corCAR1(form = ~ time | dyad/person),  # continuous AR(1) 
   data = esm_data,
   method = "REML",   # the default but specified for transparancy
   na.action = na.exclude,
   control = lmeControl(opt = "nlminb")   # default but specified for transparancy
)

summary(h13_PA)
intervals(h13_PA, which = "fixed")  # 95% CIs for fixed effects
VarCorr(h13_PA)
```

1.  $\beta_0 =75.92$, $p<.001$, 95% CI: $[73.52,78.31]$

2.  **Within-person effects (momentary deviations from one’s/one dyad’s typical level)**

    -   **Actual PA similarity:** statistically significant, $β_1 = 0.032$, 95% CI $[0.020, 0.043]$.

        When partners are more actually similar in PA than usual at a given moment, love is slightly higher at that moment (net of perceived similarity).

        A 10-point increase in momentary actual PA similarity $\rightarrow$ $+0.32$ points in love (0–100).

    -   **Perceived PA similarity:** statistically significant, $β_3 = 0.149$, 95% CI $[0.137, 0.161]$.

        When someone perceives higher-than-usual PA similarity with the partner, love is higher at that moment (net of actual similarity).

        A 10-point increase in momentary perceived PA similarity $\rightarrow$ $+1.49$ points in love.

3.  **Between-person/dyad effects (stable differences in typical levels across the study period)**

    -   **Actual PA similarity**: not statistically significant, $β_2 = −0.355$, 95% CI $[-0.962, 0.252]$.

        People in dyads who are typically higher in actual PA similarity do not show reliably different overall love, as in the model without perceived similarity.

    -   **Perceived PA similarity:** statistically significant, $β_4 = 0.859$, 95% CI $[0.370, 1.349]$.

        Persons who typically perceive more PA similarity report higher overall love across moments.

        A 10-point higher typical perceived PA similarity $\rightarrow$ $+8.59$ points higher love on average.

This is consistent with our hypothesis: $𝛽_3=0.149$ has the same direction as $β_1=0.113$ from H1.1 and $𝛽_4=0.859$ has the same direction as $\beta_2=0.214$ from H1.1 (although that wasn't significant) and the $\beta_1=0.032$ and $\beta_2=−0.355$ (not significant) here are reduced in magnitude.

### H1.3 - NA

```{r h13_model_NA}

h13_NA <- lme(
   fixed = love ~ NA_similarity_act_w + NA_similarity_act_b + NA_similarity_perc_w + NA_similarity_perc_b,
   random = ~ 1 | dyad/person,                 # v0,d and u0,pd
   correlation = corCAR1(form = ~ time | dyad/person),  # continuous AR(1) 
   data = esm_data,
   method = "REML",   # the default but specified for transparancy
   na.action = na.exclude,
   control = lmeControl(opt = "nlminb")   # default but specified for transparancy
)

summary(h13_NA)
intervals(h13_NA, which = "fixed")  # 95% CIs for fixed effects
VarCorr(h13_NA)

```

1.  $\beta_0 = 75.92$, $p<.001$, 95% CI: $[73.58,78.27]$

2.  Within-person effects (momentary deviations from one’s/one dyad’s typical level)

    -   **Actual NA similarity**: statistically significant, $\beta_1 = -0.094$, 95% CI $[-0.112, -0.077]$.

        When partners are more actually similar in NA than usual at a given moment, love is lower at that moment (net of perceived similarity).

        A 10-point increase in momentary actual NA similarity $\rightarrow$ $-0.94$ points in love (0–100).

    -   **Perceived NA similarity**: statistically significant, $\beta_3 = -0.127$, 95% CI $[-0.144, -0.110]$.

        When someone perceives higher-than-usual NA similarity with the partner, love is lower at that moment (net of actual similarity).

        A 10-point increase in momentary perceived NA similarity $\rightarrow$ $-1.27$ points in love.

3.  Between-person/dyad effects (stable differences in typical levels across the study period)

    -   **Actual NA similarity**: not statistically significant, $\beta_2 = 0.010$, 95% CI $[-0.547, 0.567]$.

        People in dyads who are typically higher in actual NA similarity do not show reliably different overall love once perceived similarity is included.

    -   **Perceived NA similarity**: statistically significant, $\beta_4 = 0.893$, 95% CI $[0.358, 1.428]$.

        Persons who typically perceive more NA similarity report higher overall love across moments.

        A 10-point higher typical perceived NA similarity $\rightarrow$ $+8.93$ points higher love on average.

------------------------------------------------------------------------

# Step 1c: Actual × Perceived Interaction

## Exploratory analysis

$$
L_{mpd} =
b_{0,pd} +
\beta_1 S^{act,within}_{mpd} +
\beta_2 \bar{S}^{act,between}_d +
\beta_3 S^{perc,within}_{mpd} +
\beta_4 \bar{S}^{perc,between}_{pd}
$$$$
+ \beta_5 (S^{act,within}_{mpd} \times S^{perc,within}_{mpd})
+ \beta_6 (\bar{S}^{act,between}_d \times \bar{S}^{perc,between}_{pd})
+ \varepsilon_{mpd}
$$

Predicted values will be probed at −1 SD, mean, and +1 SD.

### PA

```{r interaction_model_PA}

im_PA <- lme(
   fixed = love ~ 
     PA_similarity_act_w + PA_similarity_act_b +
     PA_similarity_perc_w + PA_similarity_perc_b+
     PA_similarity_act_w:PA_similarity_perc_w +
     PA_similarity_act_b:PA_similarity_perc_b,
   random = ~ 1 | dyad/person,                 # v0,d and u0,pd
   correlation = corCAR1(form = ~ time | dyad/person),  # continuous AR(1) 
   data = esm_data,
   method = "REML",   # the default but specified for transparancy
   na.action = na.exclude,
   control = lmeControl(opt = "nlminb")   # default but specified for transparancy
)

summary(im_PA)
intervals(im_PA, which = "fixed")  # 95% CIs for fixed effects
VarCorr(im_PA)
```

\$\\beta_0 = 74.19\$

Within-person effects (momentary deviations from one’s/one dyad’s typical level)

Actual PA similarity: statistically significant, \$\\beta_1 = 0.045\$, 95% CI \$\[0.033, 0.057\]\$.

When partners are more actually similar in PA than usual at a given moment, love is higher at that moment (net of perceived similarity and their within-person interaction).

A 10-point increase in momentary actual PA similarity \$\\rightarrow\$ \$+0.45\$ points in love (0–100), evaluated at the typical (i.e., zero-deviation) level of perceived PA similarity.

Perceived PA similarity: statistically significant, \$\\beta_3 = 0.173\$, 95% CI \$\[0.160, 0.186\]\$.

When someone perceives higher-than-usual PA similarity with the partner, love is higher at that moment (net of actual similarity and their within-person interaction).

A 10-point increase in momentary perceived PA similarity \$\\rightarrow\$ \$+1.73\$ points in love, evaluated at the typical (i.e., zero-deviation) level of actual PA similarity.

Within-person interaction (actual \$\\times\$ perceived PA similarity): statistically significant, \$\\beta_5 = 0.00146\$, 95% CI \$\[0.00121, 0.00171\]\$.

The positive association between actual PA similarity and love is stronger at moments when perceived PA similarity is also higher than usual (and vice versa). Equivalently, the effect of perceived PA similarity on love becomes more positive when actual PA similarity is higher than usual.

Because the interaction is small in raw units, an intuitive scaling is: a 10-point higher-than-usual perceived PA similarity increases the momentary effect of actual PA similarity by \$10 \\times 0.00146 = 0.0146\$ love-points per 1-point of actual similarity (i.e., by \$0.146\$ love-points per 10-point change in actual similarity).

Between-person/dyad effects (stable differences in typical levels across the study period)

Actual PA similarity: not statistically significant, \$\\beta_2 = -0.310\$, 95% CI \$\[-0.899, 0.278\]\$.

People in dyads who are typically higher in actual PA similarity do not show reliably different overall love once perceived similarity and the between-person interaction are included.

Perceived PA similarity: statistically significant, \$\\beta_4 = 1.209\$, 95% CI \$\[0.670, 1.748\]\$.

Persons who typically perceive more PA similarity report higher overall love across moments.

A 10-point higher typical perceived PA similarity \$\\rightarrow\$ \$+12.09\$ points higher love on average, evaluated at the typical (centered) level of typical actual PA similarity.

Between-person interaction (actual \$\\times\$ perceived PA similarity): statistically significant, \$\\beta_6 = 0.0929\$, 95% CI \$\[0.0239, 0.1619\]\$.

The between-person association between perceived PA similarity and average love is stronger in dyads/persons who are typically higher in actual PA similarity (and vice versa). That is, when couples are on average higher in actual PA similarity, the link between typically perceived PA similarity and overall love is more positive.

A 10-point higher typical perceived PA similarity increases the between-person effect of typical actual PA similarity by \$10 \\times 0.0929 = 0.929\$ love-points per 1-point of actual similarity (or \$9.29\$ love-points per 10-point change in actual similarity).

### NA

```{r interaction_model_NA}

im_NA <- lme(
   fixed = love ~ 
     NA_similarity_act_w + NA_similarity_act_b +
     NA_similarity_perc_w + NA_similarity_perc_b+
     NA_similarity_act_w:NA_similarity_perc_w +
     NA_similarity_act_b:NA_similarity_perc_b,
   random = ~ 1 | dyad/person,                 # v0,d and u0,pd
   correlation = corCAR1(form = ~ time | dyad/person),  # continuous AR(1) 
   data = esm_data,
   method = "REML",   # the default but specified for transparancy
   na.action = na.exclude,
   control = lmeControl(opt = "nlminb")   # default but specified for transparancy
)

summary(im_NA)
intervals(im_NA, which = "fixed")  # 95% CIs for fixed effects
VarCorr(im_NA)

```

------------------------------------------------------------------------

# Step 2: Multilevel mediation (H2.1-H2.2)

We test whether perceived similarity mediates the association between actual similarity and love.

**Mediator Model**

$$
S^{perc}_{mpd} =
t_{0,pd} +
\theta_1 S^{act,within}_{mpd} +
\theta_2 \bar{S}^{act,between}_d +
\varepsilon^M_{mpd}
$$

**Outcome Model**

$$
L_{mpd} =
f_{0,pd} +
\phi_1 S^{perc,within}_{mpd} +
\phi_2 \bar{S}^{perc,between}_{pd} +
\phi_3 S^{act,within}_{mpd} +
\phi_4 \bar{S}^{act,between}_d +
\varepsilon^L_{mpd}
$$

Indirect effects:

-   Within-person: $\theta_1 \phi_1$
-   Between-dyad: $\theta_2 \phi_2$

## H2.1 - PA mediation

```{r}

```

## H2.2 - NA mediation

```{r}

```

------------------------------------------------------------------------

# Step 3a: Moderation by responsiveness (H3.1-H3.2)

General model structure:

$$
L_{mpd} =
... +
\beta (Similarity \times Responsiveness) +
\varepsilon_{mpd}
$$

Interpretation: - Responsiveness as amplifier (PA) - Responsiveness as buffer (NA)

## H3.1 – PA Similarity × Responsiveness

```{r}

```

## H3.2 – NA Similarity × Responsiveness

```{r}

```

------------------------------------------------------------------------

# Step 3b: Moderation by Event Valence (H3.3-H3.4)

Event valence derived from:

-   Positive event
-   Negative event

Mismatch hypotheses:

-   PA similarity strongest in negative moments
-   NA similarity most harmful in positive moments

## H3.3 – PA Similarity × Event Valence

```{r}

```

## H3.4 – NA Similarity × Event Valence

```{r}

```
