---
title: "Study Component 2: Affect Similarity in Laboratory Interactions (VMR)"
---

```{r setup}
#| include: false

# 1. Load packages, paths and data configurations
source(here::here("R", "00_setup.R"))

# 2. Set a global theme for plots
theme_set(theme_minimal() + theme(legend.position = "bottom"))

```

```{r SC2_setup}
#| include: false
# load analysis datasets
vmr_data  <- readRDS(file.path(dir_data_ana, "vmr_ana.rds"))
post_data <- readRDS(file.path(dir_data_ana, "post_ana.rds"))
bg_data   <- readRDS(file.path(dir_data_ana, "bg_ana.rds"))

# plot initialisation
plot_counter <- 1
prefix <- "SC2"
```

# Introduction

See also [OSF preregistration](https://doi.org/10.17605/OSF.IO/V7T9F){.uri}.

## Overview

Study Component 2 examines how **interaction-averaged affect similarity** during structured positive and negative laboratory interactions relates to **post-interaction love and closeness**, using video-mediated recall (VMR) data from romantic couples.

As in Study Component 1, we distinguish:

-   **Positive Affect (PA)** and **Negative Affect (NA)** (analyzed separately)
-   **Actual similarity** (self-reported affect compared to partner's self-reported affect)
-   **Perceived similarity** (self-reported affect compared to actor's perception of partner's affect)

Similarity is again conceptualized as **shared, co-experienced elevated affect**, operationalized as dyadic elevation (mean level) penalized by mismatch (absolute difference). For Study Component 2, we average the segment-level similarity indices across the interaction.

The main research questions are:

1.  Does interaction-averaged affect similarity predict post-interaction love and closeness? (H1)
2.  Is this association mediated by perceived similarity? (H2)
3.  Are these associations moderated by trait we-ness and interaction context? (H3)

All analyses use multilevel models with persons ($p$) nested within dyads ($d$), estimated separately for the positive and negative interaction contexts ($c$). For context moderation (H3.3–H3.4), a pooled model includes both contexts.

There are two post-interaction outcomes: love and closeness. Models are estimated for both outcomes.

## Conceptualization and Operationalization

### Notation and Multilevel Structure

Let:

-   $O_{pdc}$ = post-interaction outcome (love or closeness) for person $p$ in dyad $d$ after interaction $c$
-   $A_{mpdc}$ = actor affect at VMR segment $m$
-   $P^{act}_{mpdc}$ = partner's actual affect
-   $P^{perc}_{mpdc}$ = partner affect as perceived by actor

Interaction context: $c = 0$ (negative interaction), $c = 1$ (positive interaction).

### Person-Mean Centering

Affect variables are person-mean centered across all VMR segments and across both interactions for each person $p$ in dyad $d$, as implemented in `04_data_construct.R`:

$$
X_{mpdc} = A_{mpdc} - \frac{1}{N_{pd}} \sum_{m,c} A_{mpdc}
$$

and analogously for $Y^{act}_{mpdc}$ and $Y^{perc}_{mpdc}$.

### Similarity Indices

For each VMR segment $m$ in interaction $c$ within dyad $d$:

$$
S^{act}_{mpdc} = \frac{X_{mpdc} + Y^{act}_{mpdc}}{2} - |X_{mpdc} - Y^{act}_{mpdc}|
$$

$$
S^{perc}_{mpdc} = \frac{X_{mpdc} + Y^{perc}_{mpdc}}{2} - |X_{mpdc} - Y^{perc}_{mpdc}|
$$

These are computed at the segment level in `04_data_construct.R`. For the interaction-level analyses, we average across segments within each interaction $c$:

$$
\bar{S}^{act}_{pdc} = \frac{1}{N_{pdc}} \sum_m S^{act}_{mpdc}, \quad \bar{S}^{perc}_{pdc} = \frac{1}{N_{pdc}} \sum_m S^{perc}_{mpdc}
$$

### Random Intercepts and Estimation

For models estimated separately within each interaction context (Steps 1a–1c, 2, 3a), there is one observation per person, so only a **dyad-level random intercept** is included.

For the context moderation model (Step 3b), data from both interactions are pooled (two observations per person), and random intercepts for both **dyads** and **persons nested within dyads** are included.

Unlike Study Component 1, we do **not** specify AR(1) residual structures because the outcomes are at the interaction level.

### Trait We-ness

Trait we-ness ($W^T_{pd}$) is constructed from the Dyadic Coping Inventory (DCI) as the mean of selected items reflecting positive shared dyadic coping (see `04_data_construct.R`). It is grand-mean centered so that $0$ reflects the average level of trait we-ness in the sample.

## Hypotheses Summary

-   **H1.1** (PA): Higher average actual PA similarity → higher post-interaction love/closeness ($\beta_1 > 0$).
-   **H1.2** (NA): Higher average actual NA similarity → lower post-interaction love/closeness ($\beta_1 < 0$).
-   **H1.3**: Perceived similarity added alongside actual similarity shows comparable direction; actual similarity coefficients attenuate.
-   **H2.1** (PA mediation): Actual PA similarity → perceived PA similarity → higher love/closeness (positive indirect effect).
-   **H2.2** (NA mediation): Actual NA similarity → perceived NA similarity → lower love/closeness (negative indirect effect).
-   **H3.1** (PA × we-ness): Higher trait we-ness amplifies the positive PA similarity–evaluation association ($\beta_3 > 0$).
-   **H3.2** (NA × we-ness): Higher trait we-ness attenuates the negative NA similarity–evaluation association ($\beta_3 > 0$).
-   **H3.3** (PA × context): PA similarity more beneficial in the negative than positive interaction ($\beta_3 < 0$).
-   **H3.4** (NA × context): NA similarity more detrimental in the positive than negative interaction ($\beta_3 < 0$).

# Data Preparation

## Aggregate VMR Similarity to Interaction Level

```{r aggregate_vmr}

# Average similarity indices across VMR segments within each interaction (topic)
vmr_agg <- vmr_data %>%
  group_by(dyad, person, topic) %>%
  summarise(
    PA_sim_act  = mean(PA_similarity_act,  na.rm = TRUE),
    PA_sim_perc = mean(PA_similarity_perc, na.rm = TRUE),
    NA_sim_act  = mean(NA_similarity_act,  na.rm = TRUE),
    NA_sim_perc = mean(NA_similarity_perc, na.rm = TRUE),
    n_segments  = n(),
    .groups = "drop"
  )

cat(sprintf("VMR aggregated: %d rows (persons × interactions)\n", nrow(vmr_agg)))
cat(sprintf("Unique persons: %d\n", n_distinct(vmr_agg$person)))
cat(sprintf("Unique dyads:   %d\n", n_distinct(vmr_agg$dyad)))
```

## Reshape Post-Interaction Outcomes to Long Format

```{r reshape_post}

# Reshape post-interaction data to long format with context indicator
post_long <- post_data %>%
  pivot_longer(
    cols = c(love_neg, close_neg, love_pos, close_pos),
    names_to = c(".value", "topic"),
    names_pattern = "(love|close)_(neg|pos)"
  ) %>%
  mutate(topic = ifelse(topic == "neg", "negative", "positive"))

cat(sprintf("POST reshaped: %d rows\n", nrow(post_long)))
```

## Merge Datasets

```{r merge_sc2}

# Merge VMR aggregates with post-interaction outcomes
sc2 <- vmr_agg %>%
  inner_join(post_long, by = c("dyad", "person", "topic"))

# Merge with trait we-ness from background questionnaire
sc2 <- sc2 %>%
  left_join(bg_data %>% select(person, we_ness), by = "person")

# Create context dummy (0 = negative, 1 = positive)
sc2 <- sc2 %>%
  mutate(context = ifelse(topic == "positive", 1, 0))

# Grand-mean center trait we-ness
sc2 <- sc2 %>%
  mutate(cwe_ness = we_ness - mean(we_ness, na.rm = TRUE))

cat(sprintf("SC2 merged dataset: %d rows, %d persons, %d dyads\n",
            nrow(sc2), n_distinct(sc2$person), n_distinct(sc2$dyad)))

# Grand-mean centered similarity indices for Step 1c interaction models
# Centering is done within each context as specified in the preregistration
sc2 <- sc2 %>%
  group_by(topic) %>%
  mutate(
    PA_sim_act_gmc  = PA_sim_act  - mean(PA_sim_act,  na.rm = TRUE),
    PA_sim_perc_gmc = PA_sim_perc - mean(PA_sim_perc, na.rm = TRUE),
    NA_sim_act_gmc  = NA_sim_act  - mean(NA_sim_act,  na.rm = TRUE),
    NA_sim_perc_gmc = NA_sim_perc - mean(NA_sim_perc, na.rm = TRUE)
  ) %>%
  ungroup()

# Split by context for separate analyses
sc2_neg <- sc2 %>% filter(topic == "negative")
sc2_pos <- sc2 %>% filter(topic == "positive")

cat(sprintf("Negative context: %d obs\n", nrow(sc2_neg)))
cat(sprintf("Positive context: %d obs\n", nrow(sc2_pos)))
```

## Model Specification in R

For models estimated within a single context (Steps 1a–1c, 2, 3a), one observation per person per interaction, only a dyad-level random intercept is estimable:

> `lme(fixed = outcome ~ predictor(s), random = ~ 1 | dyad, data = ..., method = "REML", na.action = na.exclude)`

For the pooled context moderation model (Step 3b), two observations per person allow a nested random intercept structure:

> `lme(fixed = outcome ~ predictor(s), random = ~ 1 | dyad/person, data = ..., method = "REML", na.action = na.exclude)`

# Step 1a: Main Effects of Actual Similarity (H1.1–H1.2)

Primary model, estimated separately within each interaction context $c$:

$$
O_{pdc} = b_{0,d} + \beta_1 \bar{S}^{act}_{pdc} + \varepsilon_{pdc}
$$

Random intercept structure:

$$
b_{0,d} = \beta_0 + v_{0,d}
$$

-   $v_{0,d} \sim N(0, \tau_d^2)$: dyad-level random intercept

Interpretation:

-   $\beta_1$: association between interaction-averaged actual similarity and the post-interaction outcome. For PA (H1.1), expected $\beta_1 > 0$; for NA (H1.2), expected $\beta_1 < 0$.

## H1.1 – Actual PA Similarity

### Love

#### Negative interaction

```{r h11_PA_love_neg}
#| output-fold: true
#| output-summary: "H1.1 PA → Love (negative interaction)"

h11_PA_love_neg <- lme(
  fixed = love ~ PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h11_PA_love_neg)
```

**Intercept** $\beta_0 = 89.05$, 95% CI $[79.88,98.23]$, $p < .001$

This is the expected post-interaction love when $\bar{S}^{act}_{pdc} = 0$, i.e., when a person's interaction-averaged actual PA similarity is zero. As we discussed, that means the elevation and mismatch exactly cancel out across the interaction. On average, the partners were at their affect baselines or any shared elevation was exactly offset by divergence.

The value of 89 is high on a 0–100 scale, suggesting that even after a **negative** interaction (disagreement topic), couples report feeling quite loving toward their partner. This might reflect that the lab setting is relatively benign, or that love is a more stable construct that does not drop sharply from a single 8-minute disagreement.

**Slope** $\beta_1 = 0.379$, 95% CI $[0.060,0.697]$, $p=.020$

For every one-unit increase in interaction-averaged actual PA similarity, post-interaction love increases by 0.38 points. Since PA similarity is on a scale derived from 0–100 affect ratings, a 10-unit increase in average similarity corresponds to +3.8 love points. The effect is significant but modest, and the CI is wide, reflecting substantial uncertainty with only 98 dyad-level data points informing this estimate. The direction is consistent with H1.1: more shared PA during the negative interaction predicts more love afterward.

**Random effects and variance decomposition**

**Dyad intercept** $SD=8.84$, variance $= 78.16$

**Residual** $SD=18.35$, variance $= 336.86$

The total variance is $78.16+336.86=415.02$

The **ICC** (proportion of variance at the dyad level) is $78.16/415.02=0.188$, or about $19\%$. This means that roughly one fifth of the variability in post-interaction love is attributable to stable differences between couples, while the remaining 81% is within-dyad, differences between the two partners within the same couple. The residual is quite large ($SD=18.35$) relative to the dyad intercept ($SD=8.84$). This means partners within the same dyad can differ substantially in how much love they report after the same interaction. That makes sense: two people can experience the same disagreement quite differently. An ICC of 0.19 is modest but non-negligible, confirming that the dyad random intercept is justified: ignoring the nesting would underestimate standard errors.

**The model provides initial support for H1.1** in the negative interaction context: couples who share more PA on average during the disagreement interaction report more love afterward. But the effect is modest, estimated with considerable uncertainty

#### Positive interaction

```{r h11_PA_love_pos}
#| output-fold: true
#| output-summary: "H1.1 PA → Love (positive interaction)"

h11_PA_love_pos <- lme( fixed = love ~ PA_sim_act, random = ~ 1 | dyad, data = sc2_pos, method = "REML", na.action = na.exclude )

model_output(h11_PA_love_pos)
```

**Intercept** $\beta_0 = 85.97$, 95% CI $[82.79, 89.15]$, $p < .001$.

Expected post-interaction love when interaction-averaged actual PA similarity is zero. The intercept is lower than in the negative interaction model ($89.05$), which might seem counterintuitive since one might expect higher love after a positive interaction. However, this comparison is not straightforward because the intercept is evaluated at $\bar{S}^{act} = 0$, and the distribution of PA similarity likely differs between contexts (the positive interaction may have a higher mean PA similarity such that $\bar{S}^{act} = 0$ indicates a 'below average couple' , while hte negative interaction may have lower mean PA similarity so the intercept is looking at a 'higher than average couple'. The intercept being lower in the positive context ($85.97$) than the negative context ($89.05$) may simply mean: a below-average-similarity couple in a positive interaction reports less love than an above-average-similarity couple in a negative interaction.

**Slope** $\beta_1 = 0.459$, 95% CI $[0.127, 0.792]$, $p = .007$.

For every one-unit increase in interaction-averaged actual PA similarity during the positive interaction, post-interaction love increases by $0.46$ points. A 10-unit increase in average similarity corresponds to $+4.6$ love points. The effect is somewhat stronger and more precisely estimated than in the negative interaction ($\beta_1 = 0.379$, $p = .020$), and the evidence is clearer ($p = .007$ vs. $p = .020$). Like the negative interaction, **the direction is consistent with H1.1**.

**Random effects and variance decomposition**

The dyad random intercept is $SD = 10.47$ (variance $= 109.64$) and the residual is $SD = 13.28$ (variance $= 176.33$), yielding a total variance of $285.97$. The ICC is $109.64 / 285.97 = 0.383$, notably higher than in the negative interaction ($0.188$). This means that **after the positive interaction, nearly 40% of the variability in love is between couples rather than between partners within couples. Partners agree more about their love after a positive interaction than after a negative one, which makes sense: a positive discussion may produce a more uniform relational experience within couples, while a disagreement is more likely to be experienced asymmetrically.**

The residual $SD = 13.28$ is considerably smaller than in the negative interaction ($SD = 18.35$). Combined with the higher ICC, this indicates that post-interaction love after a positive interaction is both more homogeneous within couples and less variable overall.

The standardized residuals range from $-3.63$ to $+1.97$, showing the same left-skewed pattern observed across all love models (ceiling compression).

------------------------------------------------------------------------

### Closeness

#### Negative interaction

```{r h11_PA_close_neg}
#| output-fold: true
#| output-summary: "H1.1 PA → Closeness (negative interaction)"

h11_PA_close_neg <- lme(
  fixed = close ~ PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h11_PA_close_neg)
```

**Intercept** $\beta_0 = 92.07$, 95% CI $[82.10, 102.04]$, $p < .001$.

Expected post-interaction closeness when interaction-averaged actual PA similarity is zero. The value is high on the 0–100 scale and even higher than the corresponding love intercept in the same context ($89.05$), suggesting that couples report feeling very close after a negative interaction. There is a wide CI ($\approx 20$ points).

**Slope** $\beta_1 = 0.432$, 95% CI $[0.086, 0.778]$, $p = .015$.

For every one-unit increase in interaction-averaged actual PA similarity during the negative interaction, post-interaction closeness increases by $0.43$ points. A 10-unit increase corresponds to $+4.3$ closeness points. The effect is slightly larger than for love in the same context ($\beta_1 = 0.379$, $p = .020$) and has a comparable level of evidence. T**he direction is consistent with H1.1: sharing more PA during the disagreement predicts feeling closer afterward.**

**Random effects and variance decomposition**

The dyad random intercept is $SD = 12.75$ (variance $= 162.65$) and the residual is $SD = 16.06$ (variance $= 257.79$), yielding a total variance of $420.44$. The ICC is $162.65 / 420.44 = 0.387$. This is substantially higher than the ICC for love in the same context ($0.188$), meaning that partners within the same couple agree much more about their closeness after a negative interaction than about their love. Nearly 40% of the variability in post-interaction closeness is between couples, with the remaining 60% reflecting differences between partners within couples.

#### Positive interaction

```{r h11_PA_close_pos}
#| output-fold: true
#| output-summary: "H1.1 PA → Closeness (positive interaction)"

h11_PA_close_pos <- lme(
  fixed = close ~ PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h11_PA_close_pos)
```

**Intercept** $\beta_0 = 87.27$, 95% CI $[84.36, 90.17]$, $p < .001$.

Expected post-interaction closeness when interaction-averaged actual PA similarity is zero. As with love, the intercept is lower than in the negative interaction ($92.07$), likely reflecting the same issue: the distribution of PA similarity differs between contexts, so the intercept evaluates at a different relative position in the data rather than reflecting a genuine difference in closeness across contexts.

**Slope** $\beta_1 = 0.376$, 95% CI $[0.072, 0.679]$, $p = .016$.

For every one-unit increase in interaction-averaged actual PA similarity during the positive interaction, post-interaction closeness increases by $0.38$ points. A 10-unit increase corresponds to $+3.8$ closeness points. The effect is comparable to the love outcome in the positive interaction ($\beta_1 = 0.459$, $p = .007$) and to closeness in the negative interaction ($\beta_1 = 0.432$, $p = .015$), though slightly smaller in magnitude. T**he direction is again consistent with H1.1.**

**Random effects and variance decomposition**

The dyad random intercept is $SD = 9.37$ (variance $= 87.81$) and the residual is $SD = 12.42$ (variance $= 154.31$), yielding a total variance of $242.12$. The ICC is $87.81 / 242.12 = 0.363$. This is similar to the love ICC in the positive interaction ($0.383$) and to the closeness ICC in the negative interaction ($0.387$), suggesting that roughly a third to two fifths of the variability in post-interaction evaluations is between couples across both outcomes in both contexts, with the exception of love after the negative interaction, where the ICC is notably lower ($0.188$). L**ove after a disagreement is the outcome where partners diverge the most.**

The total variance ($242.12$) is the lowest of the four Step 1a models, indicating that closeness after a positive interaction is the least variable outcome overall. This is consistent with a positive discussion producing both strong agreement within couples and modest differentiation across couples.

### Summary of results

|   | Negative Interaction | Positive Interaction |
|:-----------------------|:-----------------------|:-----------------------|
| **Love** | $\beta_1 = 0.379$, $p = .020$, 95% CI $[0.060, 0.697]$; ICC $= 18.8\%$ | $\beta_1 = 0.459$, $p = .007$, 95% CI $[0.127, 0.792]$; ICC $= 38.3\%$ |
| **Closeness** | $\beta_1 = 0.432$, $p = .015$, 95% CI $[0.086, 0.778]$; ICC $= 38.7\%$ | $\beta_1 = 0.376$, $p = .016$, 95% CI $[0.072, 0.679]$; ICC $= 36.3\%$ |

: Actual PA similarity predicting post-interaction evaluations (H1.1). All models include a dyad-level random intercept. ICC reflects the proportion of variance between dyads (i.e., the expected within-dyad partner correlation).

### Diagnostics

```{r h11_PA_love_neg_diagnostics}
#| warning: false
#| output-fold: true
#| output-summary: "Diagnostics H1.1 PA love (negative interaction)"

lme_diagnostics(h11_PA_love_neg)
```

#### **Diagnostics - H1.1 Actual PA → Love (Negative Interaction)**

**Residuals vs. fitted values.** The LOESS curve deviates from the zero line, dipping below zero at lower fitted values (around 55–65) and running slightly above zero for higher fitted values (75–90). This suggests mild nonlinearity: the model somewhat overpredicts love for couples with low predicted values and slightly underpredicts at the high end. There is also visible heteroscedasticity — the residual spread is larger at lower fitted values and compresses toward higher fitted values, consistent with a ceiling effect where most observations cluster near the upper bound of the 0–100 scale. The bulk of the data concentrates at fitted values above 75, with relatively few observations below 70 informing the left side of the plot.

**QQ-plot.** The central range (theoretical quantiles between approximately $-1.5$ and $+1.5$) follows the reference line reasonably well, indicating approximate normality in the bulk of the distribution. However, the left tail deviates substantially — observed residuals are more extreme than expected under normality, with several points falling well below the line. The right tail is compressed (points fall below the line), meaning the model produces fewer large positive residuals than a normal distribution would predict. This asymmetry is the same ceiling-driven pattern observed in the SC1 models: love ratings near 100 leave no room for positive residuals, producing a left-skewed residual distribution.

These patterns are not ideal but are expected given the bounded outcome scale and the high baseline love observed in this sample. The fixed-effect estimates remain interpretable, but standard errors and $p$-values should be treated with some caution.

```{r h11_PA_love_pos_diagnostics}
#| warning: false
#| output-fold: true
#| output-summary: "Diagnostics H1.1 PA love (positive interaction)"

lme_diagnostics(h11_PA_love_pos)
```

#### Diagnostics - H1.1 Actual PA → Love (Positive Interaction)

**Residuals vs. fitted values.** The LOESS curve runs below zero for lower fitted values (50–75) and approaches zero for higher fitted values (80–95). The pattern is similar to the negative interaction but more pronounced: there are a few isolated observations with very low fitted values (around 50–60) and large negative residuals, pulling the LOESS curve down on the left. Heteroscedasticity is again present — residual spread is wider at lower fitted values and compresses toward the ceiling. The data concentration is shifted further right compared to the negative interaction, with most observations above fitted value 80.

**QQ-plot.** The central range follows the reference line well. The left tail shows the familiar deviation — observed residuals more extreme than expected — with a few notable outliers around theoretical quantile $-3$. The right tail is compressed. Overall comparable to the negative interaction diagnostics.

```{r h11_PA_close_neg_diagnostics}
#| warning: false
#| output-fold: true
#| output-summary: "Diagnostics H1.1 PA closeness (negative interaction)"

lme_diagnostics(h11_PA_love_neg)
```

#### Diagnostics - H1.1 Actual PA → Closeness (Negative Interaction)

**Residuals vs. fitted values.** The LOESS curve shows a similar pattern to the love models: dipping below zero at lower fitted values (55–70) and levelling off near zero at higher fitted values (75–90). The heteroscedasticity pattern is consistent — wider residual spread at lower fitted values, compression at higher values. The data distribution is very similar to the love model in the same context.

**QQ-plot.** Slightly better behaved than the love models. The central range tracks the reference line closely. The left tail deviates but somewhat less dramatically — fewer extreme outliers pulling away from the line. The right tail shows moderate compression. Overall, the normality assumption is approximately met in the central range with the expected ceiling-driven departures in the tails.

```{r h11_PA_close_pos_diagnostics}
#| warning: false
#| output-fold: true
#| output-summary: "Diagnostics H1.1 PA closeness (positive[] interaction)"

lme_diagnostics(h11_PA_love_neg)

```

#### Diagnostics - H1.1 Actual PA → Closeness (Positive Interaction)

**Residuals vs. fitted values.** Very similar to the other three models: LOESS below zero at lower fitted values, converging toward zero at higher values. The residual spread again compresses toward higher fitted values. The bulk of the data sits above fitted value 75, consistent with the positive interaction generally producing high closeness ratings.

**QQ-plot.** The best-behaved of the four models. The central range follows the reference line closely, and the left-tail deviations are present but somewhat milder. There is one notable outlier below theoretical quantile $-2.5$. The right tail is compressed as in the other models.

#### Diagnostics - Summary

All four diagnostic panels show the same constellation: mild nonlinearity in the LOESS curve driven by sparse low-fitted-value observations, heteroscedasticity from ceiling compression, and left-skewed residuals in the QQ-plot. These patterns are structural consequences of the bounded 0–100 outcome scale combined with high baseline love and closeness in this sample. The closeness models show slightly better residual behaviour than the love models, and the positive interaction models show slightly denser concentration at high fitted values. None of the models show patterns that would invalidate the fixed-effect conclusions, but the ceiling-driven violations warrant caution when interpreting standard errors and $p$-values.

## H1.2 – Actual NA Similarity

### Love

```{r h12_NA_love_neg}
#| output-fold: true
#| output-summary: "H1.2 NA → Love (negative interaction)"

h12_NA_love_neg <- lme(
  fixed = love ~ NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h12_NA_love_neg)
```

```{r h12_NA_love_pos}
#| output-fold: true
#| output-summary: "H1.2 NA → Love (positive interaction)"

h12_NA_love_pos <- lme(
  fixed = love ~ NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h12_NA_love_pos)
```

### Closeness

```{r h12_NA_close_neg}
#| output-fold: true
#| output-summary: "H1.2 NA → Closeness (negative interaction)"

h12_NA_close_neg <- lme(
  fixed = close ~ NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h12_NA_close_neg)
```

```{r h12_NA_close_pos}
#| output-fold: true
#| output-summary: "H1.2 NA → Closeness (positive interaction)"

h12_NA_close_pos <- lme(
  fixed = close ~ NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h12_NA_close_pos)
```

# Step 1b: Adding Perceived Similarity (H1.3)

$$
O_{pdc} = b_{0,d} + \beta_1 \bar{S}^{act}_{pdc} + \beta_2 \bar{S}^{perc}_{pdc} + \varepsilon_{pdc}
$$

with the same dyad-level random intercept as in Step 1a.

Evidence consistent with H1.3: $\beta_2$ in the expected direction, and attenuation of $\beta_1$ relative to Step 1a.

## PA Similarity

### Love

```{r h13_PA_love_neg}
#| output-fold: true
#| output-summary: "H1.3 PA → Love (negative interaction)"

h13_PA_love_neg <- lme(
  fixed = love ~ PA_sim_act + PA_sim_perc,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h13_PA_love_neg)
```

```{r h13_PA_love_pos}
#| output-fold: true
#| output-summary: "H1.3 PA → Love (positive interaction)"

h13_PA_love_pos <- lme(
  fixed = love ~ PA_sim_act + PA_sim_perc,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h13_PA_love_pos)
```

### Closeness

```{r h13_PA_close_neg}
#| output-fold: true
#| output-summary: "H1.3 PA → Closeness (negative interaction)"

h13_PA_close_neg <- lme(
  fixed = close ~ PA_sim_act + PA_sim_perc,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h13_PA_close_neg)
```

```{r h13_PA_close_pos}
#| output-fold: true
#| output-summary: "H1.3 PA → Closeness (positive interaction)"

h13_PA_close_pos <- lme(
  fixed = close ~ PA_sim_act + PA_sim_perc,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h13_PA_close_pos)
```

## NA Similarity

### Love

```{r h13_NA_love_neg}
#| output-fold: true
#| output-summary: "H1.3 NA → Love (negative interaction)"

h13_NA_love_neg <- lme(
  fixed = love ~ NA_sim_act + NA_sim_perc,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h13_NA_love_neg)
```

```{r h13_NA_love_pos}
#| output-fold: true
#| output-summary: "H1.3 NA → Love (positive interaction)"

h13_NA_love_pos <- lme(
  fixed = love ~ NA_sim_act + NA_sim_perc,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h13_NA_love_pos)
```

### Closeness

```{r h13_NA_close_neg}
#| output-fold: true
#| output-summary: "H1.3 NA → Closeness (negative interaction)"

h13_NA_close_neg <- lme(
  fixed = close ~ NA_sim_act + NA_sim_perc,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h13_NA_close_neg)
```

```{r h13_NA_close_pos}
#| output-fold: true
#| output-summary: "H1.3 NA → Closeness (positive interaction)"

h13_NA_close_pos <- lme(
  fixed = close ~ NA_sim_act + NA_sim_perc,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h13_NA_close_pos)
```

# Step 1c (Exploratory): Actual × Perceived Interaction

To examine whether the association of one form of similarity with post-interaction evaluations depends on the level of the other, we include interaction terms. Grand-mean centered (within each interaction context) similarity indices are used so that main effects remain interpretable at the average level of the moderator.

$$
O_{pdc} = b_{0,d} + \beta_1 \tilde{S}^{act}_{pdc} + \beta_2 \tilde{S}^{perc}_{pdc} + \beta_3 (\tilde{S}^{act}_{pdc} \times \tilde{S}^{perc}_{pdc}) + \varepsilon_{pdc}
$$

with the same dyad-level random intercept as before. $\tilde{S}$ denotes the grand-mean centered (within each context) interaction-averaged similarity indices.

If the interaction term $\beta_3$ is significant, we probe simple slopes at low ($-1$ SD), average (0), and high ($+1$ SD) levels of the moderator.

## PA Similarity

### Love

```{r im_PA_love_neg}
#| output-fold: true
#| output-summary: "Interaction PA → Love (negative)"

im_PA_love_neg <- lme(
  fixed = love ~ PA_sim_act_gmc + PA_sim_perc_gmc + PA_sim_act_gmc:PA_sim_perc_gmc,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(im_PA_love_neg)
```

```{r im_PA_love_pos}
#| output-fold: true
#| output-summary: "Interaction PA → Love (positive)"

im_PA_love_pos <- lme(
  fixed = love ~ PA_sim_act_gmc + PA_sim_perc_gmc + PA_sim_act_gmc:PA_sim_perc_gmc,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(im_PA_love_pos)
```

### Probing (negative interaction, love)

```{r im_PA_love_neg_probe}
#| output-fold: true
#| output-summary: "Simple slopes — PA love (negative)"

# Probe points (−1 SD, 0, +1 SD)
act_sd_neg  <- sd(sc2_neg$PA_sim_act_gmc,  na.rm = TRUE)
perc_sd_neg <- sd(sc2_neg$PA_sim_perc_gmc, na.rm = TRUE)

# Simple slopes of actual PA similarity at levels of perceived PA similarity
slopes_act <- emtrends(
  im_PA_love_neg,
  specs = ~ PA_sim_perc_gmc,
  var   = "PA_sim_act_gmc",
  at    = list(PA_sim_perc_gmc = c(-perc_sd_neg, 0, perc_sd_neg))
)

summary(slopes_act, infer = TRUE)

# Simple slopes of perceived PA similarity at levels of actual PA similarity
slopes_perc <- emtrends(
  im_PA_love_neg,
  specs = ~ PA_sim_act_gmc,
  var   = "PA_sim_perc_gmc",
  at    = list(PA_sim_act_gmc = c(-act_sd_neg, 0, act_sd_neg))
)

summary(slopes_perc, infer = TRUE)
```

### Closeness

```{r im_PA_close_neg}
#| output-fold: true
#| output-summary: "Interaction PA → Closeness (negative)"

im_PA_close_neg <- lme(
  fixed = close ~ PA_sim_act_gmc + PA_sim_perc_gmc + PA_sim_act_gmc:PA_sim_perc_gmc,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(im_PA_close_neg)
```

```{r im_PA_close_pos}
#| output-fold: true
#| output-summary: "Interaction PA → Closeness (positive)"

im_PA_close_pos <- lme(
  fixed = close ~ PA_sim_act_gmc + PA_sim_perc_gmc + PA_sim_act_gmc:PA_sim_perc_gmc,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(im_PA_close_pos)
```

## NA Similarity

### Love

```{r im_NA_love_neg}
#| output-fold: true
#| output-summary: "Interaction NA → Love (negative)"

im_NA_love_neg <- lme(
  fixed = love ~ NA_sim_act_gmc + NA_sim_perc_gmc + NA_sim_act_gmc:NA_sim_perc_gmc,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(im_NA_love_neg)
```

```{r im_NA_love_pos}
#| output-fold: true
#| output-summary: "Interaction NA → Love (positive)"

im_NA_love_pos <- lme(
  fixed = love ~ NA_sim_act_gmc + NA_sim_perc_gmc + NA_sim_act_gmc:NA_sim_perc_gmc,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(im_NA_love_pos)
```

### Closeness

```{r im_NA_close_neg}
#| output-fold: true
#| output-summary: "Interaction NA → Closeness (negative)"

im_NA_close_neg <- lme(
  fixed = close ~ NA_sim_act_gmc + NA_sim_perc_gmc + NA_sim_act_gmc:NA_sim_perc_gmc,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(im_NA_close_neg)
```

```{r im_NA_close_pos}
#| output-fold: true
#| output-summary: "Interaction NA → Closeness (positive)"

im_NA_close_pos <- lme(
  fixed = close ~ NA_sim_act_gmc + NA_sim_perc_gmc + NA_sim_act_gmc:NA_sim_perc_gmc,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(im_NA_close_pos)
```

# Step 2: Mediation by Perceived Similarity (H2.1–H2.2)

We test whether interaction-averaged perceived similarity mediates the association between interaction-averaged actual similarity and post-interaction evaluations. Mediation is estimated separately within each interaction context.

**Mediator Model:**

$$
\bar{S}^{perc}_{pdc} = t_{0,d} + \theta_1 \bar{S}^{act}_{pdc} + \varepsilon^M_{pdc}
$$

**Outcome Model:**

$$
O_{pdc} = f_{0,d} + \varphi_1 \bar{S}^{perc}_{pdc} + \varphi_2 \bar{S}^{act}_{pdc} + \varepsilon^O_{pdc}
$$

Both with dyad-level random intercepts.

Indirect effect: $\theta_1 \varphi_1$. Confidence intervals via Monte Carlo simulation.

## H2.1 – PA Mediation

### Negative Interaction

#### c-path (total effect)

The c-paths are the Step 1a models: `h11_PA_love_neg` and `h11_PA_close_neg`.

#### a-path (mediator model)

```{r h21_PA_mediator_neg}
#| output-fold: true
#| output-summary: "Mediator model PA (negative)"

h21_PA_med_neg <- lme(
  fixed = PA_sim_perc ~ PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h21_PA_med_neg)
```

#### b- and c'-paths (outcome model — love)

```{r h21_PA_out_love_neg}
#| output-fold: true
#| output-summary: "Outcome model PA → Love (negative)"

h21_PA_out_love_neg <- lme(
  fixed = love ~ PA_sim_perc + PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h21_PA_out_love_neg)
```

#### b- and c'-paths (outcome model — closeness)

```{r h21_PA_out_close_neg}
#| output-fold: true
#| output-summary: "Outcome model PA → Closeness (negative)"

h21_PA_out_close_neg <- lme(
  fixed = close ~ PA_sim_perc + PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h21_PA_out_close_neg)
```

#### Monte Carlo CI – Love (negative interaction)

```{r h21_PA_MC_love_neg}
#| output-fold: true
#| output-summary: "Monte Carlo indirect effect — PA love (negative)"

set.seed(42)
n_sim <- 20000

# Extract coefficients and SEs from mediator and outcome models
a_est     <- fixef(h21_PA_med_neg)["PA_sim_act"]
se_a      <- sqrt(vcov(h21_PA_med_neg)["PA_sim_act", "PA_sim_act"])
b_est     <- fixef(h21_PA_out_love_neg)["PA_sim_perc"]
se_b      <- sqrt(vcov(h21_PA_out_love_neg)["PA_sim_perc", "PA_sim_perc"])
cprime    <- fixef(h21_PA_out_love_neg)["PA_sim_act"]
c_total   <- fixef(h11_PA_love_neg)["PA_sim_act"]

# Monte Carlo draws
a_sim <- rnorm(n_sim, mean = a_est, sd = se_a)
b_sim <- rnorm(n_sim, mean = b_est, sd = se_b)
indirect_sim <- a_sim * b_sim

result <- mc_ci(indirect_sim, a_est, b_est, "PA love neg")

cat(sprintf("Indirect effect: %.4f, 95%% MC CI [%.4f, %.4f], p = %s\n",
            result$point, result$ci_low, result$ci_hi,
            ifelse(result$p < .001, "< .001", sprintf("%.3f", result$p))))
cat(sprintf("Total effect (c): %.4f\n", c_total))
cat(sprintf("Direct effect (c'): %.4f\n", cprime))
```

#### Monte Carlo CI – Closeness (negative interaction)

```{r h21_PA_MC_close_neg}
#| output-fold: true
#| output-summary: "Monte Carlo indirect effect — PA closeness (negative)"

b_est_c   <- fixef(h21_PA_out_close_neg)["PA_sim_perc"]
se_b_c    <- sqrt(vcov(h21_PA_out_close_neg)["PA_sim_perc", "PA_sim_perc"])
cprime_c  <- fixef(h21_PA_out_close_neg)["PA_sim_act"]
c_total_c <- fixef(h11_PA_close_neg)["PA_sim_act"]

b_sim_c <- rnorm(n_sim, mean = b_est_c, sd = se_b_c)
indirect_sim_c <- a_sim * b_sim_c

result_c <- mc_ci(indirect_sim_c, a_est, b_est_c, "PA close neg")

cat(sprintf("Indirect effect: %.4f, 95%% MC CI [%.4f, %.4f], p = %s\n",
            result_c$point, result_c$ci_low, result_c$ci_hi,
            ifelse(result_c$p < .001, "< .001", sprintf("%.3f", result_c$p))))
cat(sprintf("Total effect (c): %.4f\n", c_total_c))
cat(sprintf("Direct effect (c'): %.4f\n", cprime_c))
```

### Positive Interaction

#### a-path (mediator model)

```{r h21_PA_mediator_pos}
#| output-fold: true
#| output-summary: "Mediator model PA (positive)"

h21_PA_med_pos <- lme(
  fixed = PA_sim_perc ~ PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h21_PA_med_pos)
```

#### b- and c'-paths (outcome models)

```{r h21_PA_out_love_pos}
#| output-fold: true
#| output-summary: "Outcome model PA → Love (positive)"

h21_PA_out_love_pos <- lme(
  fixed = love ~ PA_sim_perc + PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h21_PA_out_love_pos)
```

```{r h21_PA_out_close_pos}
#| output-fold: true
#| output-summary: "Outcome model PA → Closeness (positive)"

h21_PA_out_close_pos <- lme(
  fixed = close ~ PA_sim_perc + PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h21_PA_out_close_pos)
```

#### Monte Carlo CI – Love (positive interaction)

```{r h21_PA_MC_love_pos}
#| output-fold: true
#| output-summary: "Monte Carlo indirect effect — PA love (positive)"

a_est_p   <- fixef(h21_PA_med_pos)["PA_sim_act"]
se_a_p    <- sqrt(vcov(h21_PA_med_pos)["PA_sim_act", "PA_sim_act"])
b_est_p   <- fixef(h21_PA_out_love_pos)["PA_sim_perc"]
se_b_p    <- sqrt(vcov(h21_PA_out_love_pos)["PA_sim_perc", "PA_sim_perc"])
cprime_p  <- fixef(h21_PA_out_love_pos)["PA_sim_act"]
c_total_p <- fixef(h11_PA_love_pos)["PA_sim_act"]

a_sim_p <- rnorm(n_sim, mean = a_est_p, sd = se_a_p)
b_sim_p <- rnorm(n_sim, mean = b_est_p, sd = se_b_p)
indirect_sim_p <- a_sim_p * b_sim_p

result_p <- mc_ci(indirect_sim_p, a_est_p, b_est_p, "PA love pos")

cat(sprintf("Indirect effect: %.4f, 95%% MC CI [%.4f, %.4f], p = %s\n",
            result_p$point, result_p$ci_low, result_p$ci_hi,
            ifelse(result_p$p < .001, "< .001", sprintf("%.3f", result_p$p))))
cat(sprintf("Total effect (c): %.4f\n", c_total_p))
cat(sprintf("Direct effect (c'): %.4f\n", cprime_p))
```

#### Monte Carlo CI – Closeness (positive interaction)

```{r h21_PA_MC_close_pos}
#| output-fold: true
#| output-summary: "Monte Carlo indirect effect — PA closeness (positive)"

b_est_pc   <- fixef(h21_PA_out_close_pos)["PA_sim_perc"]
se_b_pc    <- sqrt(vcov(h21_PA_out_close_pos)["PA_sim_perc", "PA_sim_perc"])
cprime_pc  <- fixef(h21_PA_out_close_pos)["PA_sim_act"]
c_total_pc <- fixef(h11_PA_close_pos)["PA_sim_act"]

b_sim_pc <- rnorm(n_sim, mean = b_est_pc, sd = se_b_pc)
indirect_sim_pc <- a_sim_p * b_sim_pc

result_pc <- mc_ci(indirect_sim_pc, a_est_p, b_est_pc, "PA close pos")

cat(sprintf("Indirect effect: %.4f, 95%% MC CI [%.4f, %.4f], p = %s\n",
            result_pc$point, result_pc$ci_low, result_pc$ci_hi,
            ifelse(result_pc$p < .001, "< .001", sprintf("%.3f", result_pc$p))))
cat(sprintf("Total effect (c): %.4f\n", c_total_pc))
cat(sprintf("Direct effect (c'): %.4f\n", cprime_pc))
```

## H2.2 – NA Mediation

### Negative Interaction

#### a-path

```{r h22_NA_mediator_neg}
#| output-fold: true
#| output-summary: "Mediator model NA (negative)"

h22_NA_med_neg <- lme(
  fixed = NA_sim_perc ~ NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h22_NA_med_neg)
```

#### b- and c'-paths

```{r h22_NA_out_love_neg}
#| output-fold: true
#| output-summary: "Outcome model NA → Love (negative)"

h22_NA_out_love_neg <- lme(
  fixed = love ~ NA_sim_perc + NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h22_NA_out_love_neg)
```

```{r h22_NA_out_close_neg}
#| output-fold: true
#| output-summary: "Outcome model NA → Closeness (negative)"

h22_NA_out_close_neg <- lme(
  fixed = close ~ NA_sim_perc + NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h22_NA_out_close_neg)
```

#### Monte Carlo CI (negative interaction)

```{r h22_NA_MC_neg}
#| output-fold: true
#| output-summary: "Monte Carlo indirect effects — NA (negative)"

# Love
a_NA_neg     <- fixef(h22_NA_med_neg)["NA_sim_act"]
se_a_NA_neg  <- sqrt(vcov(h22_NA_med_neg)["NA_sim_act", "NA_sim_act"])
b_NA_love_neg <- fixef(h22_NA_out_love_neg)["NA_sim_perc"]
se_b_NA_love_neg <- sqrt(vcov(h22_NA_out_love_neg)["NA_sim_perc", "NA_sim_perc"])

a_NA_sim_neg <- rnorm(n_sim, mean = a_NA_neg, sd = se_a_NA_neg)
b_NA_sim_love_neg <- rnorm(n_sim, mean = b_NA_love_neg, sd = se_b_NA_love_neg)
ind_NA_love_neg <- a_NA_sim_neg * b_NA_sim_love_neg

res_NA_love_neg <- mc_ci(ind_NA_love_neg, a_NA_neg, b_NA_love_neg, "NA love neg")

cat("--- Love (negative) ---\n")
cat(sprintf("Indirect: %.4f, 95%% MC CI [%.4f, %.4f], p = %s\n",
            res_NA_love_neg$point, res_NA_love_neg$ci_low, res_NA_love_neg$ci_hi,
            ifelse(res_NA_love_neg$p < .001, "< .001", sprintf("%.3f", res_NA_love_neg$p))))

# Closeness
b_NA_close_neg <- fixef(h22_NA_out_close_neg)["NA_sim_perc"]
se_b_NA_close_neg <- sqrt(vcov(h22_NA_out_close_neg)["NA_sim_perc", "NA_sim_perc"])

b_NA_sim_close_neg <- rnorm(n_sim, mean = b_NA_close_neg, sd = se_b_NA_close_neg)
ind_NA_close_neg <- a_NA_sim_neg * b_NA_sim_close_neg

res_NA_close_neg <- mc_ci(ind_NA_close_neg, a_NA_neg, b_NA_close_neg, "NA close neg")

cat("\n--- Closeness (negative) ---\n")
cat(sprintf("Indirect: %.4f, 95%% MC CI [%.4f, %.4f], p = %s\n",
            res_NA_close_neg$point, res_NA_close_neg$ci_low, res_NA_close_neg$ci_hi,
            ifelse(res_NA_close_neg$p < .001, "< .001", sprintf("%.3f", res_NA_close_neg$p))))
```

### Positive Interaction

```{r h22_NA_mediator_pos}
#| output-fold: true
#| output-summary: "Mediator model NA (positive)"

h22_NA_med_pos <- lme(
  fixed = NA_sim_perc ~ NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h22_NA_med_pos)
```

```{r h22_NA_out_love_pos}
#| output-fold: true
#| output-summary: "Outcome model NA → Love (positive)"

h22_NA_out_love_pos <- lme(
  fixed = love ~ NA_sim_perc + NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h22_NA_out_love_pos)
```

```{r h22_NA_out_close_pos}
#| output-fold: true
#| output-summary: "Outcome model NA → Closeness (positive)"

h22_NA_out_close_pos <- lme(
  fixed = close ~ NA_sim_perc + NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h22_NA_out_close_pos)
```

#### Monte Carlo CI (positive interaction)

```{r h22_NA_MC_pos}
#| output-fold: true
#| output-summary: "Monte Carlo indirect effects — NA (positive)"

# Love
a_NA_pos     <- fixef(h22_NA_med_pos)["NA_sim_act"]
se_a_NA_pos  <- sqrt(vcov(h22_NA_med_pos)["NA_sim_act", "NA_sim_act"])
b_NA_love_pos <- fixef(h22_NA_out_love_pos)["NA_sim_perc"]
se_b_NA_love_pos <- sqrt(vcov(h22_NA_out_love_pos)["NA_sim_perc", "NA_sim_perc"])

a_NA_sim_pos <- rnorm(n_sim, mean = a_NA_pos, sd = se_a_NA_pos)
b_NA_sim_love_pos <- rnorm(n_sim, mean = b_NA_love_pos, sd = se_b_NA_love_pos)
ind_NA_love_pos <- a_NA_sim_pos * b_NA_sim_love_pos

res_NA_love_pos <- mc_ci(ind_NA_love_pos, a_NA_pos, b_NA_love_pos, "NA love pos")

cat("--- Love (positive) ---\n")
cat(sprintf("Indirect: %.4f, 95%% MC CI [%.4f, %.4f], p = %s\n",
            res_NA_love_pos$point, res_NA_love_pos$ci_low, res_NA_love_pos$ci_hi,
            ifelse(res_NA_love_pos$p < .001, "< .001", sprintf("%.3f", res_NA_love_pos$p))))

# Closeness
b_NA_close_pos <- fixef(h22_NA_out_close_pos)["NA_sim_perc"]
se_b_NA_close_pos <- sqrt(vcov(h22_NA_out_close_pos)["NA_sim_perc", "NA_sim_perc"])

b_NA_sim_close_pos <- rnorm(n_sim, mean = b_NA_close_pos, sd = se_b_NA_close_pos)
ind_NA_close_pos <- a_NA_sim_pos * b_NA_sim_close_pos

res_NA_close_pos <- mc_ci(ind_NA_close_pos, a_NA_pos, b_NA_close_pos, "NA close pos")

cat("\n--- Closeness (positive) ---\n")
cat(sprintf("Indirect: %.4f, 95%% MC CI [%.4f, %.4f], p = %s\n",
            res_NA_close_pos$point, res_NA_close_pos$ci_low, res_NA_close_pos$ci_hi,
            ifelse(res_NA_close_pos$p < .001, "< .001", sprintf("%.3f", res_NA_close_pos$p))))
```

# Step 3a: Moderation by Trait We-ness (H3.1–H3.2)

We examine whether the similarity–evaluation association depends on trait we-ness ($W^T_{pd}$, grand-mean centered). Models are estimated separately within each interaction context.

For actual similarity:

$$
O_{pdc} = b_{0,d} + \beta_1 \bar{S}^{act}_{pdc} + \beta_2 W^T_{pd} + \beta_3 (\bar{S}^{act}_{pdc} \times W^T_{pd}) + \varepsilon_{pdc}
$$

For perceived similarity:

$$
O_{pdc} = b_{0,d} + \beta_1 \bar{S}^{perc}_{pdc} + \beta_2 W^T_{pd} + \beta_3 (\bar{S}^{perc}_{pdc} \times W^T_{pd}) + \varepsilon_{pdc}
$$

Both with dyad-level random intercept. The conditional similarity–outcome association is $\beta_1 + \beta_3 W^T_{pd}$.

For PA (H3.1): expected $\beta_3 > 0$ (we-ness amplifies positive effects). For NA (H3.2): expected $\beta_3 > 0$ (we-ness buffers negative effects, making the negative slope less negative).

## H3.1 – PA × We-ness

### Actual PA Similarity

#### Love

```{r h31_PA_act_love_neg}
#| output-fold: true
#| output-summary: "H3.1 actual PA × we-ness → Love (negative)"

h31_PA_act_love_neg <- lme(
  fixed = love ~ PA_sim_act + cwe_ness + PA_sim_act:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h31_PA_act_love_neg)
```

```{r h31_PA_act_love_pos}
#| output-fold: true
#| output-summary: "H3.1 actual PA × we-ness → Love (positive)"

h31_PA_act_love_pos <- lme(
  fixed = love ~ PA_sim_act + cwe_ness + PA_sim_act:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h31_PA_act_love_pos)
```

#### Probing (if interaction significant)

```{r h31_PA_act_probe_neg}
#| output-fold: true
#| output-summary: "Simple slopes — actual PA × we-ness → Love (negative)"

we_sd <- sd(sc2_neg$cwe_ness, na.rm = TRUE)

slopes_h31 <- emtrends(
  h31_PA_act_love_neg,
  specs = ~ cwe_ness,
  var   = "PA_sim_act",
  at    = list(cwe_ness = c(-we_sd, 0, we_sd))
)

slopes_h31_df <- as.data.frame(summary(slopes_h31, infer = TRUE))

level_labels <- c("Low (−1 SD)", "Average", "High (+1 SD)")

tbl_h31 <- data.frame(
  `We-ness level` = level_labels,
  `We-ness value` = sprintf("%.2f", slopes_h31_df$cwe_ness),
  `Simple slope`  = sprintf("%.4f", slopes_h31_df$PA_sim_act.trend),
  `SE`            = sprintf("%.4f", slopes_h31_df$SE),
  `95% CI`        = sprintf("[%.4f, %.4f]", slopes_h31_df$lower.CL, slopes_h31_df$upper.CL),
  `t`             = sprintf("%.3f", slopes_h31_df$t.ratio),
  `p`             = ifelse(slopes_h31_df$p.value < .001, "< .001",
                           sprintf("%.3f", slopes_h31_df$p.value)),
  check.names = FALSE
)

kable(tbl_h31,
  align    = c("l", "c", "c", "c", "c", "c", "c"),
  booktabs = TRUE,
  caption  = "Simple slopes of actual PA similarity on love at low, average, and high trait we-ness (negative interaction)"
) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    position   = "left"
  ) %>%
  footnote(
    general = "Simple slopes estimated at −1 SD, mean, and +1 SD of trait we-ness (grand-mean centered).",
    general_title = "Note.",
    footnote_as_chunk = TRUE
  )
```

#### Closeness

```{r h31_PA_act_close_neg}
#| output-fold: true
#| output-summary: "H3.1 actual PA × we-ness → Closeness (negative)"

h31_PA_act_close_neg <- lme(
  fixed = close ~ PA_sim_act + cwe_ness + PA_sim_act:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h31_PA_act_close_neg)
```

```{r h31_PA_act_close_pos}
#| output-fold: true
#| output-summary: "H3.1 actual PA × we-ness → Closeness (positive)"

h31_PA_act_close_pos <- lme(
  fixed = close ~ PA_sim_act + cwe_ness + PA_sim_act:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h31_PA_act_close_pos)
```

### Perceived PA Similarity

#### Love

```{r h31_PA_perc_love_neg}
#| output-fold: true
#| output-summary: "H3.1 perceived PA × we-ness → Love (negative)"

h31_PA_perc_love_neg <- lme(
  fixed = love ~ PA_sim_perc + cwe_ness + PA_sim_perc:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h31_PA_perc_love_neg)
```

```{r h31_PA_perc_love_pos}
#| output-fold: true
#| output-summary: "H3.1 perceived PA × we-ness → Love (positive)"

h31_PA_perc_love_pos <- lme(
  fixed = love ~ PA_sim_perc + cwe_ness + PA_sim_perc:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h31_PA_perc_love_pos)
```

#### Closeness

```{r h31_PA_perc_close_neg}
#| output-fold: true
#| output-summary: "H3.1 perceived PA × we-ness → Closeness (negative)"

h31_PA_perc_close_neg <- lme(
  fixed = close ~ PA_sim_perc + cwe_ness + PA_sim_perc:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h31_PA_perc_close_neg)
```

```{r h31_PA_perc_close_pos}
#| output-fold: true
#| output-summary: "H3.1 perceived PA × we-ness → Closeness (positive)"

h31_PA_perc_close_pos <- lme(
  fixed = close ~ PA_sim_perc + cwe_ness + PA_sim_perc:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h31_PA_perc_close_pos)
```

## H3.2 – NA × We-ness

### Actual NA Similarity

#### Love

```{r h32_NA_act_love_neg}
#| output-fold: true
#| output-summary: "H3.2 actual NA × we-ness → Love (negative)"

h32_NA_act_love_neg <- lme(
  fixed = love ~ NA_sim_act + cwe_ness + NA_sim_act:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h32_NA_act_love_neg)
```

```{r h32_NA_act_love_pos}
#| output-fold: true
#| output-summary: "H3.2 actual NA × we-ness → Love (positive)"

h32_NA_act_love_pos <- lme(
  fixed = love ~ NA_sim_act + cwe_ness + NA_sim_act:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h32_NA_act_love_pos)
```

#### Closeness

```{r h32_NA_act_close_neg}
#| output-fold: true
#| output-summary: "H3.2 actual NA × we-ness → Closeness (negative)"

h32_NA_act_close_neg <- lme(
  fixed = close ~ NA_sim_act + cwe_ness + NA_sim_act:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h32_NA_act_close_neg)
```

```{r h32_NA_act_close_pos}
#| output-fold: true
#| output-summary: "H3.2 actual NA × we-ness → Closeness (positive)"

h32_NA_act_close_pos <- lme(
  fixed = close ~ NA_sim_act + cwe_ness + NA_sim_act:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h32_NA_act_close_pos)
```

### Perceived NA Similarity

#### Love

```{r h32_NA_perc_love_neg}
#| output-fold: true
#| output-summary: "H3.2 perceived NA × we-ness → Love (negative)"

h32_NA_perc_love_neg <- lme(
  fixed = love ~ NA_sim_perc + cwe_ness + NA_sim_perc:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h32_NA_perc_love_neg)
```

```{r h32_NA_perc_love_pos}
#| output-fold: true
#| output-summary: "H3.2 perceived NA × we-ness → Love (positive)"

h32_NA_perc_love_pos <- lme(
  fixed = love ~ NA_sim_perc + cwe_ness + NA_sim_perc:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h32_NA_perc_love_pos)
```

#### Closeness

```{r h32_NA_perc_close_neg}
#| output-fold: true
#| output-summary: "H3.2 perceived NA × we-ness → Closeness (negative)"

h32_NA_perc_close_neg <- lme(
  fixed = close ~ NA_sim_perc + cwe_ness + NA_sim_perc:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h32_NA_perc_close_neg)
```

```{r h32_NA_perc_close_pos}
#| output-fold: true
#| output-summary: "H3.2 perceived NA × we-ness → Closeness (positive)"

h32_NA_perc_close_pos <- lme(
  fixed = close ~ NA_sim_perc + cwe_ness + NA_sim_perc:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h32_NA_perc_close_pos)
```

# Step 3b: Context Moderation — Negative vs. Positive Interaction (H3.3–H3.4)

We examine whether the similarity–evaluation association differs between the positive and negative interaction contexts. A dummy variable $C_{pdc} = c$ ($c = 0$ negative, $c = 1$ positive) is included in pooled models. Since these models contain two observations per person, both dyad-level and person-level random intercepts are included.

For actual similarity:

$$
O_{pdc} = b_{0,pd} + \beta_1 \bar{S}^{act}_{pdc} + \beta_2 C_{pdc} + \beta_3 (\bar{S}^{act}_{pdc} \times C_{pdc}) + \varepsilon_{pdc}
$$

$$
b_{0,pd} = \beta_0 + v_{0,d} + u_{0,pd}
$$

For perceived similarity: analogous model with $\bar{S}^{perc}_{pdc}$.

$\beta_1$ is the similarity–evaluation slope for the negative interaction ($C_{pdc} = 0$), and $\beta_1 + \beta_3$ is the slope for the positive interaction.

For PA (H3.3): expected $\beta_3 < 0$ (PA similarity more beneficial in negative interaction). For NA (H3.4): expected $\beta_3 < 0$ (NA similarity more harmful in positive interaction).

## H3.3 – PA × Context

### Actual PA Similarity

#### Love

```{r h33_PA_act_love}
#| output-fold: true
#| output-summary: "H3.3 actual PA × context → Love"

h33_PA_act_love <- lme(
  fixed = love ~ PA_sim_act + context + PA_sim_act:context,
  random = ~ 1 | dyad/person,
  data = sc2,
  method = "REML",
  na.action = na.exclude
)

model_output(h33_PA_act_love)
```

```{r h33_PA_act_love_slopes}
#| output-fold: true
#| output-summary: "Simple slopes — actual PA × context → Love"

# Simple slopes at negative (context = 0) and positive (context = 1) interaction
slopes_h33 <- emtrends(
  h33_PA_act_love,
  specs = ~ context,
  var   = "PA_sim_act",
  at    = list(context = c(0, 1))
)

slopes_h33_df <- as.data.frame(summary(slopes_h33, infer = TRUE))

tbl_h33 <- data.frame(
  `Interaction context` = c("Negative (0)", "Positive (1)"),
  `Simple slope`        = sprintf("%.4f", slopes_h33_df$PA_sim_act.trend),
  `SE`                  = sprintf("%.4f", slopes_h33_df$SE),
  `95% CI`              = sprintf("[%.4f, %.4f]", slopes_h33_df$lower.CL, slopes_h33_df$upper.CL),
  `t`                   = sprintf("%.3f", slopes_h33_df$t.ratio),
  `p`                   = ifelse(slopes_h33_df$p.value < .001, "< .001",
                                 sprintf("%.3f", slopes_h33_df$p.value)),
  check.names = FALSE
)

kable(tbl_h33,
  align    = c("l", "c", "c", "c", "c", "c"),
  booktabs = TRUE,
  caption  = "Simple slopes of actual PA similarity on love in negative and positive interactions (H3.3)"
) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    position   = "left"
  )
```

#### Closeness

```{r h33_PA_act_close}
#| output-fold: true
#| output-summary: "H3.3 actual PA × context → Closeness"

h33_PA_act_close <- lme(
  fixed = close ~ PA_sim_act + context + PA_sim_act:context,
  random = ~ 1 | dyad/person,
  data = sc2,
  method = "REML",
  na.action = na.exclude
)

model_output(h33_PA_act_close)
```

### Perceived PA Similarity

#### Love

```{r h33_PA_perc_love}
#| output-fold: true
#| output-summary: "H3.3 perceived PA × context → Love"

h33_PA_perc_love <- lme(
  fixed = love ~ PA_sim_perc + context + PA_sim_perc:context,
  random = ~ 1 | dyad/person,
  data = sc2,
  method = "REML",
  na.action = na.exclude
)

model_output(h33_PA_perc_love)
```

#### Closeness

```{r h33_PA_perc_close}
#| output-fold: true
#| output-summary: "H3.3 perceived PA × context → Closeness"

h33_PA_perc_close <- lme(
  fixed = close ~ PA_sim_perc + context + PA_sim_perc:context,
  random = ~ 1 | dyad/person,
  data = sc2,
  method = "REML",
  na.action = na.exclude
)

model_output(h33_PA_perc_close)
```

## H3.4 – NA × Context

### Actual NA Similarity

#### Love

```{r h34_NA_act_love}
#| output-fold: true
#| output-summary: "H3.4 actual NA × context → Love"

h34_NA_act_love <- lme(
  fixed = love ~ NA_sim_act + context + NA_sim_act:context,
  random = ~ 1 | dyad/person,
  data = sc2,
  method = "REML",
  na.action = na.exclude
)

model_output(h34_NA_act_love)
```

```{r h34_NA_act_love_slopes}
#| output-fold: true
#| output-summary: "Simple slopes — actual NA × context → Love"

slopes_h34 <- emtrends(
  h34_NA_act_love,
  specs = ~ context,
  var   = "NA_sim_act",
  at    = list(context = c(0, 1))
)

slopes_h34_df <- as.data.frame(summary(slopes_h34, infer = TRUE))

tbl_h34 <- data.frame(
  `Interaction context` = c("Negative (0)", "Positive (1)"),
  `Simple slope`        = sprintf("%.4f", slopes_h34_df$NA_sim_act.trend),
  `SE`                  = sprintf("%.4f", slopes_h34_df$SE),
  `95% CI`              = sprintf("[%.4f, %.4f]", slopes_h34_df$lower.CL, slopes_h34_df$upper.CL),
  `t`                   = sprintf("%.3f", slopes_h34_df$t.ratio),
  `p`                   = ifelse(slopes_h34_df$p.value < .001, "< .001",
                                 sprintf("%.3f", slopes_h34_df$p.value)),
  check.names = FALSE
)

kable(tbl_h34,
  align    = c("l", "c", "c", "c", "c", "c"),
  booktabs = TRUE,
  caption  = "Simple slopes of actual NA similarity on love in negative and positive interactions (H3.4)"
) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    position   = "left"
  )
```

#### Closeness

```{r h34_NA_act_close}
#| output-fold: true
#| output-summary: "H3.4 actual NA × context → Closeness"

h34_NA_act_close <- lme(
  fixed = close ~ NA_sim_act + context + NA_sim_act:context,
  random = ~ 1 | dyad/person,
  data = sc2,
  method = "REML",
  na.action = na.exclude
)

model_output(h34_NA_act_close)
```

### Perceived NA Similarity

#### Love

```{r h34_NA_perc_love}
#| output-fold: true
#| output-summary: "H3.4 perceived NA × context → Love"

h34_NA_perc_love <- lme(
  fixed = love ~ NA_sim_perc + context + NA_sim_perc:context,
  random = ~ 1 | dyad/person,
  data = sc2,
  method = "REML",
  na.action = na.exclude
)

model_output(h34_NA_perc_love)
```

#### Closeness

```{r h34_NA_perc_close}
#| output-fold: true
#| output-summary: "H3.4 perceived NA × context → Closeness"

h34_NA_perc_close <- lme(
  fixed = close ~ NA_sim_perc + context + NA_sim_perc:context,
  random = ~ 1 | dyad/person,
  data = sc2,
  method = "REML",
  na.action = na.exclude
)

model_output(h34_NA_perc_close)
```
