---
title: "Study Component 2: Affect Similarity in Laboratory Interactions (VMR)"
---

```{r setup}
#| include: false

# 1. Load packages, paths and data configurations
source(here::here("R", "00_setup.R"))

# 2. Set a global theme for plots
theme_set(theme_minimal() + theme(legend.position = "bottom"))

```

```{r SC2_setup}
#| include: false
# load analysis datasets
vmr_data  <- readRDS(file.path(dir_data_ana, "vmr_ana.rds"))
post_data <- readRDS(file.path(dir_data_ana, "post_ana.rds"))
bg_data   <- readRDS(file.path(dir_data_ana, "bg_ana.rds"))

# plot initialisation
plot_counter <- 1
prefix <- "SC2"
```

# Introduction

See also [OSF preregistration](https://doi.org/10.17605/OSF.IO/V7T9F){.uri}.

## Overview

Study Component 2 examines how **interaction-averaged affect similarity** during structured positive and negative laboratory interactions relates to **post-interaction love and closeness**, using video-mediated recall (VMR) data from romantic couples.

As in Study Component 1, we distinguish:

-   **Positive Affect (PA)** and **Negative Affect (NA)** (analyzed separately)
-   **Actual similarity** (self-reported affect compared to partner's self-reported affect)
-   **Perceived similarity** (self-reported affect compared to actor's perception of partner's affect)

Similarity is again conceptualized as **shared, co-experienced elevated affect**, operationalized as dyadic elevation (mean level) penalized by mismatch (absolute difference). For Study Component 2, we average the segment-level similarity indices across the interaction.

The main research questions are:

1.  Does interaction-averaged affect similarity predict post-interaction love and closeness? (H1)
2.  Is this association mediated by perceived similarity? (H2)
3.  Are these associations moderated by trait we-ness and interaction context? (H3)

All analyses use multilevel models with persons ($p$) nested within dyads ($d$), estimated separately for the positive and negative interaction contexts ($c$). For context moderation (H3.3–H3.4), a pooled model includes both contexts.

There are two post-interaction outcomes: love and closeness. Models are estimated for both outcomes.

## Conceptualization and Operationalization

### Notation and Multilevel Structure

Let:

-   $O_{pdc}$ = post-interaction outcome (love or closeness) for person $p$ in dyad $d$ after interaction $c$
-   $A_{mpdc}$ = actor affect at VMR segment $m$
-   $P^{act}_{mpdc}$ = partner's actual affect
-   $P^{perc}_{mpdc}$ = partner affect as perceived by actor

Interaction context: $c = 0$ (negative interaction), $c = 1$ (positive interaction).

### Person-Mean Centering

Affect variables are person-mean centered across all VMR segments and across both interactions for each person $p$ in dyad $d$, as implemented in `04_data_construct.R`:

$$
X_{mpdc} = A_{mpdc} - \frac{1}{N_{pd}} \sum_{m,c} A_{mpdc}
$$

and analogously for $Y^{act}_{mpdc}$ and $Y^{perc}_{mpdc}$.

### Similarity Indices

For each VMR segment $m$ in interaction $c$ within dyad $d$:

$$
S^{act}_{mpdc} = \frac{X_{mpdc} + Y^{act}_{mpdc}}{2} - |X_{mpdc} - Y^{act}_{mpdc}|
$$

$$
S^{perc}_{mpdc} = \frac{X_{mpdc} + Y^{perc}_{mpdc}}{2} - |X_{mpdc} - Y^{perc}_{mpdc}|
$$

These are computed at the segment level in `04_data_construct.R`. For the interaction-level analyses, we average across segments within each interaction $c$:

$$
\bar{S}^{act}_{pdc} = \frac{1}{N_{pdc}} \sum_m S^{act}_{mpdc}, \quad \bar{S}^{perc}_{pdc} = \frac{1}{N_{pdc}} \sum_m S^{perc}_{mpdc}
$$

### Random Intercepts and Estimation

For models estimated separately within each interaction context (Steps 1a–1c, 2, 3a), there is one observation per person, so only a **dyad-level random intercept** is included.

For the context moderation model (Step 3b), data from both interactions are pooled (two observations per person), and random intercepts for both **dyads** and **persons nested within dyads** are included.

Unlike Study Component 1, we do **not** specify AR(1) residual structures because the outcomes are at the interaction level.

### Trait We-ness

Trait we-ness ($W^T_{pd}$) is constructed from the Dyadic Coping Inventory (DCI) as the mean of selected items reflecting positive shared dyadic coping (see `04_data_construct.R`). It is grand-mean centered so that $0$ reflects the average level of trait we-ness in the sample.

## Hypotheses Summary

-   **H1.1** (PA): Higher average actual PA similarity → higher post-interaction love/closeness ($\beta_1 > 0$).
-   **H1.2** (NA): Higher average actual NA similarity → lower post-interaction love/closeness ($\beta_1 < 0$).
-   **H1.3**: Perceived similarity added alongside actual similarity shows comparable direction; actual similarity coefficients attenuate.
-   **H2.1** (PA mediation): Actual PA similarity → perceived PA similarity → higher love/closeness (positive indirect effect).
-   **H2.2** (NA mediation): Actual NA similarity → perceived NA similarity → lower love/closeness (negative indirect effect).
-   **H3.1** (PA × we-ness): Higher trait we-ness amplifies the positive PA similarity–evaluation association ($\beta_3 > 0$).
-   **H3.2** (NA × we-ness): Higher trait we-ness attenuates the negative NA similarity–evaluation association ($\beta_3 > 0$).
-   **H3.3** (PA × context): PA similarity more beneficial in the negative than positive interaction ($\beta_3 < 0$).
-   **H3.4** (NA × context): NA similarity more detrimental in the positive than negative interaction ($\beta_3 < 0$).

# Data Preparation

## Aggregate VMR Similarity to Interaction Level

```{r aggregate_vmr}

# Average similarity indices across VMR segments within each interaction (topic)
vmr_agg <- vmr_data %>%
  group_by(dyad, person, topic) %>%
  summarise(
    PA_sim_act  = mean(PA_similarity_act,  na.rm = TRUE),
    PA_sim_perc = mean(PA_similarity_perc, na.rm = TRUE),
    NA_sim_act  = mean(NA_similarity_act,  na.rm = TRUE),
    NA_sim_perc = mean(NA_similarity_perc, na.rm = TRUE),
    n_segments  = n(),
    .groups = "drop"
  )

cat(sprintf("VMR aggregated: %d rows (persons × interactions)\n", nrow(vmr_agg)))
cat(sprintf("Unique persons: %d\n", n_distinct(vmr_agg$person)))
cat(sprintf("Unique dyads:   %d\n", n_distinct(vmr_agg$dyad)))
```

## Reshape Post-Interaction Outcomes to Long Format

```{r reshape_post}

# Reshape post-interaction data to long format with context indicator
post_long <- post_data %>%
  pivot_longer(
    cols = c(love_neg, close_neg, love_pos, close_pos),
    names_to = c(".value", "topic"),
    names_pattern = "(love|close)_(neg|pos)"
  ) %>%
  mutate(topic = ifelse(topic == "neg", "negative", "positive"))

cat(sprintf("POST reshaped: %d rows\n", nrow(post_long)))
```

## Merge Datasets

```{r merge_sc2}

# Merge VMR aggregates with post-interaction outcomes
sc2 <- vmr_agg %>%
  inner_join(post_long, by = c("dyad", "person", "topic"))

# Merge with trait we-ness from background questionnaire
sc2 <- sc2 %>%
  left_join(bg_data %>% select(person, we_ness), by = "person")

# Create context dummy (0 = negative, 1 = positive)
sc2 <- sc2 %>%
  mutate(context = ifelse(topic == "positive", 1, 0))

# Grand-mean center trait we-ness
sc2 <- sc2 %>%
  mutate(cwe_ness = we_ness - mean(we_ness, na.rm = TRUE))

cat(sprintf("SC2 merged dataset: %d rows, %d persons, %d dyads\n",
            nrow(sc2), n_distinct(sc2$person), n_distinct(sc2$dyad)))

# Grand-mean centered similarity indices for Step 1c interaction models
# Centering is done within each context as specified in the preregistration
sc2 <- sc2 %>%
  group_by(topic) %>%
  mutate(
    PA_sim_act_gmc  = PA_sim_act  - mean(PA_sim_act,  na.rm = TRUE),
    PA_sim_perc_gmc = PA_sim_perc - mean(PA_sim_perc, na.rm = TRUE),
    NA_sim_act_gmc  = NA_sim_act  - mean(NA_sim_act,  na.rm = TRUE),
    NA_sim_perc_gmc = NA_sim_perc - mean(NA_sim_perc, na.rm = TRUE)
  ) %>%
  ungroup()

# Split by context for separate analyses
sc2_neg <- sc2 %>% filter(topic == "negative")
sc2_pos <- sc2 %>% filter(topic == "positive")

cat(sprintf("Negative context: %d obs\n", nrow(sc2_neg)))
cat(sprintf("Positive context: %d obs\n", nrow(sc2_pos)))
```

## Descriptive Overview

```{r sc2_descriptives}

# Descriptive statistics per context
sc2 %>%
  group_by(topic) %>%
  summarise(
    n = n(),
    across(c(love, close, PA_sim_act, PA_sim_perc, NA_sim_act, NA_sim_perc, we_ness),
           list(M = ~mean(.x, na.rm = TRUE), SD = ~sd(.x, na.rm = TRUE)),
           .names = "{.col}_{.fn}")
  ) %>%
  pivot_longer(-topic) %>%
  pivot_wider(names_from = topic, values_from = value)

# Number of VMR segments per person-interaction
summary(vmr_agg$n_segments)
```

## Model Specification in R

For models estimated within a single context (Steps 1a–1c, 2, 3a), one observation per person per interaction, only a dyad-level random intercept is estimable:

> `lme(fixed = outcome ~ predictor(s), random = ~ 1 | dyad, data = ..., method = "REML", na.action = na.exclude)`

For the pooled context moderation model (Step 3b), two observations per person allow a nested random intercept structure:

> `lme(fixed = outcome ~ predictor(s), random = ~ 1 | dyad/person, data = ..., method = "REML", na.action = na.exclude)`

# Step 1a: Main Effects of Actual Similarity (H1.1–H1.2)

Primary model, estimated separately within each interaction context $c$:

$$
O_{pdc} = b_{0,d} + \beta_1 \bar{S}^{act}_{pdc} + \varepsilon_{pdc}
$$

Random intercept structure:

$$
b_{0,d} = \beta_0 + v_{0,d}
$$

-   $v_{0,d} \sim N(0, \tau_d^2)$: dyad-level random intercept

Interpretation:

-   $\beta_1$: association between interaction-averaged actual similarity and the post-interaction outcome. For PA (H1.1), expected $\beta_1 > 0$; for NA (H1.2), expected $\beta_1 < 0$.

## H1.1 – Actual PA Similarity

### Love

```{r h11_PA_love_neg}
#| output-fold: true
#| output-summary: "H1.1 PA → Love (negative interaction)"

h11_PA_love_neg <- lme(
  fixed = love ~ PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h11_PA_love_neg)
```

```{r h11_PA_love_pos}
#| output-fold: true
#| output-summary: "H1.1 PA → Love (positive interaction)"

h11_PA_love_pos <- lme(
  fixed = love ~ PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h11_PA_love_pos)
```

### Closeness

```{r h11_PA_close_neg}
#| output-fold: true
#| output-summary: "H1.1 PA → Closeness (negative interaction)"

h11_PA_close_neg <- lme(
  fixed = close ~ PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h11_PA_close_neg)
```

```{r h11_PA_close_pos}
#| output-fold: true
#| output-summary: "H1.1 PA → Closeness (positive interaction)"

h11_PA_close_pos <- lme(
  fixed = close ~ PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h11_PA_close_pos)
```

### Diagnostics

```{r h11_PA_diagnostics}
#| warning: false
#| output-fold: true
#| output-summary: "Diagnostics H1.1 PA love (negative interaction)"

lme_diagnostics(h11_PA_love_neg)
```

## H1.2 – Actual NA Similarity

### Love

```{r h12_NA_love_neg}
#| output-fold: true
#| output-summary: "H1.2 NA → Love (negative interaction)"

h12_NA_love_neg <- lme(
  fixed = love ~ NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h12_NA_love_neg)
```

```{r h12_NA_love_pos}
#| output-fold: true
#| output-summary: "H1.2 NA → Love (positive interaction)"

h12_NA_love_pos <- lme(
  fixed = love ~ NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h12_NA_love_pos)
```

### Closeness

```{r h12_NA_close_neg}
#| output-fold: true
#| output-summary: "H1.2 NA → Closeness (negative interaction)"

h12_NA_close_neg <- lme(
  fixed = close ~ NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h12_NA_close_neg)
```

```{r h12_NA_close_pos}
#| output-fold: true
#| output-summary: "H1.2 NA → Closeness (positive interaction)"

h12_NA_close_pos <- lme(
  fixed = close ~ NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h12_NA_close_pos)
```

# Step 1b: Adding Perceived Similarity (H1.3)

$$
O_{pdc} = b_{0,d} + \beta_1 \bar{S}^{act}_{pdc} + \beta_2 \bar{S}^{perc}_{pdc} + \varepsilon_{pdc}
$$

with the same dyad-level random intercept as in Step 1a.

Evidence consistent with H1.3: $\beta_2$ in the expected direction, and attenuation of $\beta_1$ relative to Step 1a.

## PA Similarity

### Love

```{r h13_PA_love_neg}
#| output-fold: true
#| output-summary: "H1.3 PA → Love (negative interaction)"

h13_PA_love_neg <- lme(
  fixed = love ~ PA_sim_act + PA_sim_perc,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h13_PA_love_neg)
```

```{r h13_PA_love_pos}
#| output-fold: true
#| output-summary: "H1.3 PA → Love (positive interaction)"

h13_PA_love_pos <- lme(
  fixed = love ~ PA_sim_act + PA_sim_perc,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h13_PA_love_pos)
```

### Closeness

```{r h13_PA_close_neg}
#| output-fold: true
#| output-summary: "H1.3 PA → Closeness (negative interaction)"

h13_PA_close_neg <- lme(
  fixed = close ~ PA_sim_act + PA_sim_perc,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h13_PA_close_neg)
```

```{r h13_PA_close_pos}
#| output-fold: true
#| output-summary: "H1.3 PA → Closeness (positive interaction)"

h13_PA_close_pos <- lme(
  fixed = close ~ PA_sim_act + PA_sim_perc,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h13_PA_close_pos)
```

## NA Similarity

### Love

```{r h13_NA_love_neg}
#| output-fold: true
#| output-summary: "H1.3 NA → Love (negative interaction)"

h13_NA_love_neg <- lme(
  fixed = love ~ NA_sim_act + NA_sim_perc,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h13_NA_love_neg)
```

```{r h13_NA_love_pos}
#| output-fold: true
#| output-summary: "H1.3 NA → Love (positive interaction)"

h13_NA_love_pos <- lme(
  fixed = love ~ NA_sim_act + NA_sim_perc,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h13_NA_love_pos)
```

### Closeness

```{r h13_NA_close_neg}
#| output-fold: true
#| output-summary: "H1.3 NA → Closeness (negative interaction)"

h13_NA_close_neg <- lme(
  fixed = close ~ NA_sim_act + NA_sim_perc,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h13_NA_close_neg)
```

```{r h13_NA_close_pos}
#| output-fold: true
#| output-summary: "H1.3 NA → Closeness (positive interaction)"

h13_NA_close_pos <- lme(
  fixed = close ~ NA_sim_act + NA_sim_perc,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h13_NA_close_pos)
```

# Step 1c (Exploratory): Actual × Perceived Interaction

To examine whether the association of one form of similarity with post-interaction evaluations depends on the level of the other, we include interaction terms. Grand-mean centered (within each interaction context) similarity indices are used so that main effects remain interpretable at the average level of the moderator.

$$
O_{pdc} = b_{0,d} + \beta_1 \tilde{S}^{act}_{pdc} + \beta_2 \tilde{S}^{perc}_{pdc} + \beta_3 (\tilde{S}^{act}_{pdc} \times \tilde{S}^{perc}_{pdc}) + \varepsilon_{pdc}
$$

with the same dyad-level random intercept as before. $\tilde{S}$ denotes the grand-mean centered (within each context) interaction-averaged similarity indices.

If the interaction term $\beta_3$ is significant, we probe simple slopes at low ($-1$ SD), average (0), and high ($+1$ SD) levels of the moderator.

## PA Similarity

### Love

```{r im_PA_love_neg}
#| output-fold: true
#| output-summary: "Interaction PA → Love (negative)"

im_PA_love_neg <- lme(
  fixed = love ~ PA_sim_act_gmc + PA_sim_perc_gmc + PA_sim_act_gmc:PA_sim_perc_gmc,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(im_PA_love_neg)
```

```{r im_PA_love_pos}
#| output-fold: true
#| output-summary: "Interaction PA → Love (positive)"

im_PA_love_pos <- lme(
  fixed = love ~ PA_sim_act_gmc + PA_sim_perc_gmc + PA_sim_act_gmc:PA_sim_perc_gmc,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(im_PA_love_pos)
```

### Probing (negative interaction, love)

```{r im_PA_love_neg_probe}
#| output-fold: true
#| output-summary: "Simple slopes — PA love (negative)"

# Probe points (−1 SD, 0, +1 SD)
act_sd_neg  <- sd(sc2_neg$PA_sim_act_gmc,  na.rm = TRUE)
perc_sd_neg <- sd(sc2_neg$PA_sim_perc_gmc, na.rm = TRUE)

# Simple slopes of actual PA similarity at levels of perceived PA similarity
slopes_act <- emtrends(
  im_PA_love_neg,
  specs = ~ PA_sim_perc_gmc,
  var   = "PA_sim_act_gmc",
  at    = list(PA_sim_perc_gmc = c(-perc_sd_neg, 0, perc_sd_neg))
)

summary(slopes_act, infer = TRUE)

# Simple slopes of perceived PA similarity at levels of actual PA similarity
slopes_perc <- emtrends(
  im_PA_love_neg,
  specs = ~ PA_sim_act_gmc,
  var   = "PA_sim_perc_gmc",
  at    = list(PA_sim_act_gmc = c(-act_sd_neg, 0, act_sd_neg))
)

summary(slopes_perc, infer = TRUE)
```

### Closeness

```{r im_PA_close_neg}
#| output-fold: true
#| output-summary: "Interaction PA → Closeness (negative)"

im_PA_close_neg <- lme(
  fixed = close ~ PA_sim_act_gmc + PA_sim_perc_gmc + PA_sim_act_gmc:PA_sim_perc_gmc,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(im_PA_close_neg)
```

```{r im_PA_close_pos}
#| output-fold: true
#| output-summary: "Interaction PA → Closeness (positive)"

im_PA_close_pos <- lme(
  fixed = close ~ PA_sim_act_gmc + PA_sim_perc_gmc + PA_sim_act_gmc:PA_sim_perc_gmc,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(im_PA_close_pos)
```

## NA Similarity

### Love

```{r im_NA_love_neg}
#| output-fold: true
#| output-summary: "Interaction NA → Love (negative)"

im_NA_love_neg <- lme(
  fixed = love ~ NA_sim_act_gmc + NA_sim_perc_gmc + NA_sim_act_gmc:NA_sim_perc_gmc,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(im_NA_love_neg)
```

```{r im_NA_love_pos}
#| output-fold: true
#| output-summary: "Interaction NA → Love (positive)"

im_NA_love_pos <- lme(
  fixed = love ~ NA_sim_act_gmc + NA_sim_perc_gmc + NA_sim_act_gmc:NA_sim_perc_gmc,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(im_NA_love_pos)
```

### Closeness

```{r im_NA_close_neg}
#| output-fold: true
#| output-summary: "Interaction NA → Closeness (negative)"

im_NA_close_neg <- lme(
  fixed = close ~ NA_sim_act_gmc + NA_sim_perc_gmc + NA_sim_act_gmc:NA_sim_perc_gmc,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(im_NA_close_neg)
```

```{r im_NA_close_pos}
#| output-fold: true
#| output-summary: "Interaction NA → Closeness (positive)"

im_NA_close_pos <- lme(
  fixed = close ~ NA_sim_act_gmc + NA_sim_perc_gmc + NA_sim_act_gmc:NA_sim_perc_gmc,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(im_NA_close_pos)
```

# Step 2: Mediation by Perceived Similarity (H2.1–H2.2)

We test whether interaction-averaged perceived similarity mediates the association between interaction-averaged actual similarity and post-interaction evaluations. Mediation is estimated separately within each interaction context.

**Mediator Model:**

$$
\bar{S}^{perc}_{pdc} = t_{0,d} + \theta_1 \bar{S}^{act}_{pdc} + \varepsilon^M_{pdc}
$$

**Outcome Model:**

$$
O_{pdc} = f_{0,d} + \varphi_1 \bar{S}^{perc}_{pdc} + \varphi_2 \bar{S}^{act}_{pdc} + \varepsilon^O_{pdc}
$$

Both with dyad-level random intercepts.

Indirect effect: $\theta_1 \varphi_1$. Confidence intervals via Monte Carlo simulation.

## H2.1 – PA Mediation

### Negative Interaction

#### c-path (total effect)

The c-paths are the Step 1a models: `h11_PA_love_neg` and `h11_PA_close_neg`.

#### a-path (mediator model)

```{r h21_PA_mediator_neg}
#| output-fold: true
#| output-summary: "Mediator model PA (negative)"

h21_PA_med_neg <- lme(
  fixed = PA_sim_perc ~ PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h21_PA_med_neg)
```

#### b- and c'-paths (outcome model — love)

```{r h21_PA_out_love_neg}
#| output-fold: true
#| output-summary: "Outcome model PA → Love (negative)"

h21_PA_out_love_neg <- lme(
  fixed = love ~ PA_sim_perc + PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h21_PA_out_love_neg)
```

#### b- and c'-paths (outcome model — closeness)

```{r h21_PA_out_close_neg}
#| output-fold: true
#| output-summary: "Outcome model PA → Closeness (negative)"

h21_PA_out_close_neg <- lme(
  fixed = close ~ PA_sim_perc + PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h21_PA_out_close_neg)
```

#### Monte Carlo CI – Love (negative interaction)

```{r h21_PA_MC_love_neg}
#| output-fold: true
#| output-summary: "Monte Carlo indirect effect — PA love (negative)"

set.seed(42)
n_sim <- 20000

# Extract coefficients and SEs from mediator and outcome models
a_est     <- fixef(h21_PA_med_neg)["PA_sim_act"]
se_a      <- sqrt(vcov(h21_PA_med_neg)["PA_sim_act", "PA_sim_act"])
b_est     <- fixef(h21_PA_out_love_neg)["PA_sim_perc"]
se_b      <- sqrt(vcov(h21_PA_out_love_neg)["PA_sim_perc", "PA_sim_perc"])
cprime    <- fixef(h21_PA_out_love_neg)["PA_sim_act"]
c_total   <- fixef(h11_PA_love_neg)["PA_sim_act"]

# Monte Carlo draws
a_sim <- rnorm(n_sim, mean = a_est, sd = se_a)
b_sim <- rnorm(n_sim, mean = b_est, sd = se_b)
indirect_sim <- a_sim * b_sim

result <- mc_ci(indirect_sim, a_est, b_est, "PA love neg")

cat(sprintf("Indirect effect: %.4f, 95%% MC CI [%.4f, %.4f], p = %s\n",
            result$point, result$ci_low, result$ci_hi,
            ifelse(result$p < .001, "< .001", sprintf("%.3f", result$p))))
cat(sprintf("Total effect (c): %.4f\n", c_total))
cat(sprintf("Direct effect (c'): %.4f\n", cprime))
```

#### Monte Carlo CI – Closeness (negative interaction)

```{r h21_PA_MC_close_neg}
#| output-fold: true
#| output-summary: "Monte Carlo indirect effect — PA closeness (negative)"

b_est_c   <- fixef(h21_PA_out_close_neg)["PA_sim_perc"]
se_b_c    <- sqrt(vcov(h21_PA_out_close_neg)["PA_sim_perc", "PA_sim_perc"])
cprime_c  <- fixef(h21_PA_out_close_neg)["PA_sim_act"]
c_total_c <- fixef(h11_PA_close_neg)["PA_sim_act"]

b_sim_c <- rnorm(n_sim, mean = b_est_c, sd = se_b_c)
indirect_sim_c <- a_sim * b_sim_c

result_c <- mc_ci(indirect_sim_c, a_est, b_est_c, "PA close neg")

cat(sprintf("Indirect effect: %.4f, 95%% MC CI [%.4f, %.4f], p = %s\n",
            result_c$point, result_c$ci_low, result_c$ci_hi,
            ifelse(result_c$p < .001, "< .001", sprintf("%.3f", result_c$p))))
cat(sprintf("Total effect (c): %.4f\n", c_total_c))
cat(sprintf("Direct effect (c'): %.4f\n", cprime_c))
```

### Positive Interaction

#### a-path (mediator model)

```{r h21_PA_mediator_pos}
#| output-fold: true
#| output-summary: "Mediator model PA (positive)"

h21_PA_med_pos <- lme(
  fixed = PA_sim_perc ~ PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h21_PA_med_pos)
```

#### b- and c'-paths (outcome models)

```{r h21_PA_out_love_pos}
#| output-fold: true
#| output-summary: "Outcome model PA → Love (positive)"

h21_PA_out_love_pos <- lme(
  fixed = love ~ PA_sim_perc + PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h21_PA_out_love_pos)
```

```{r h21_PA_out_close_pos}
#| output-fold: true
#| output-summary: "Outcome model PA → Closeness (positive)"

h21_PA_out_close_pos <- lme(
  fixed = close ~ PA_sim_perc + PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h21_PA_out_close_pos)
```

#### Monte Carlo CI – Love (positive interaction)

```{r h21_PA_MC_love_pos}
#| output-fold: true
#| output-summary: "Monte Carlo indirect effect — PA love (positive)"

a_est_p   <- fixef(h21_PA_med_pos)["PA_sim_act"]
se_a_p    <- sqrt(vcov(h21_PA_med_pos)["PA_sim_act", "PA_sim_act"])
b_est_p   <- fixef(h21_PA_out_love_pos)["PA_sim_perc"]
se_b_p    <- sqrt(vcov(h21_PA_out_love_pos)["PA_sim_perc", "PA_sim_perc"])
cprime_p  <- fixef(h21_PA_out_love_pos)["PA_sim_act"]
c_total_p <- fixef(h11_PA_love_pos)["PA_sim_act"]

a_sim_p <- rnorm(n_sim, mean = a_est_p, sd = se_a_p)
b_sim_p <- rnorm(n_sim, mean = b_est_p, sd = se_b_p)
indirect_sim_p <- a_sim_p * b_sim_p

result_p <- mc_ci(indirect_sim_p, a_est_p, b_est_p, "PA love pos")

cat(sprintf("Indirect effect: %.4f, 95%% MC CI [%.4f, %.4f], p = %s\n",
            result_p$point, result_p$ci_low, result_p$ci_hi,
            ifelse(result_p$p < .001, "< .001", sprintf("%.3f", result_p$p))))
cat(sprintf("Total effect (c): %.4f\n", c_total_p))
cat(sprintf("Direct effect (c'): %.4f\n", cprime_p))
```

#### Monte Carlo CI – Closeness (positive interaction)

```{r h21_PA_MC_close_pos}
#| output-fold: true
#| output-summary: "Monte Carlo indirect effect — PA closeness (positive)"

b_est_pc   <- fixef(h21_PA_out_close_pos)["PA_sim_perc"]
se_b_pc    <- sqrt(vcov(h21_PA_out_close_pos)["PA_sim_perc", "PA_sim_perc"])
cprime_pc  <- fixef(h21_PA_out_close_pos)["PA_sim_act"]
c_total_pc <- fixef(h11_PA_close_pos)["PA_sim_act"]

b_sim_pc <- rnorm(n_sim, mean = b_est_pc, sd = se_b_pc)
indirect_sim_pc <- a_sim_p * b_sim_pc

result_pc <- mc_ci(indirect_sim_pc, a_est_p, b_est_pc, "PA close pos")

cat(sprintf("Indirect effect: %.4f, 95%% MC CI [%.4f, %.4f], p = %s\n",
            result_pc$point, result_pc$ci_low, result_pc$ci_hi,
            ifelse(result_pc$p < .001, "< .001", sprintf("%.3f", result_pc$p))))
cat(sprintf("Total effect (c): %.4f\n", c_total_pc))
cat(sprintf("Direct effect (c'): %.4f\n", cprime_pc))
```

## H2.2 – NA Mediation

### Negative Interaction

#### a-path

```{r h22_NA_mediator_neg}
#| output-fold: true
#| output-summary: "Mediator model NA (negative)"

h22_NA_med_neg <- lme(
  fixed = NA_sim_perc ~ NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h22_NA_med_neg)
```

#### b- and c'-paths

```{r h22_NA_out_love_neg}
#| output-fold: true
#| output-summary: "Outcome model NA → Love (negative)"

h22_NA_out_love_neg <- lme(
  fixed = love ~ NA_sim_perc + NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h22_NA_out_love_neg)
```

```{r h22_NA_out_close_neg}
#| output-fold: true
#| output-summary: "Outcome model NA → Closeness (negative)"

h22_NA_out_close_neg <- lme(
  fixed = close ~ NA_sim_perc + NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h22_NA_out_close_neg)
```

#### Monte Carlo CI (negative interaction)

```{r h22_NA_MC_neg}
#| output-fold: true
#| output-summary: "Monte Carlo indirect effects — NA (negative)"

# Love
a_NA_neg     <- fixef(h22_NA_med_neg)["NA_sim_act"]
se_a_NA_neg  <- sqrt(vcov(h22_NA_med_neg)["NA_sim_act", "NA_sim_act"])
b_NA_love_neg <- fixef(h22_NA_out_love_neg)["NA_sim_perc"]
se_b_NA_love_neg <- sqrt(vcov(h22_NA_out_love_neg)["NA_sim_perc", "NA_sim_perc"])

a_NA_sim_neg <- rnorm(n_sim, mean = a_NA_neg, sd = se_a_NA_neg)
b_NA_sim_love_neg <- rnorm(n_sim, mean = b_NA_love_neg, sd = se_b_NA_love_neg)
ind_NA_love_neg <- a_NA_sim_neg * b_NA_sim_love_neg

res_NA_love_neg <- mc_ci(ind_NA_love_neg, a_NA_neg, b_NA_love_neg, "NA love neg")

cat("--- Love (negative) ---\n")
cat(sprintf("Indirect: %.4f, 95%% MC CI [%.4f, %.4f], p = %s\n",
            res_NA_love_neg$point, res_NA_love_neg$ci_low, res_NA_love_neg$ci_hi,
            ifelse(res_NA_love_neg$p < .001, "< .001", sprintf("%.3f", res_NA_love_neg$p))))

# Closeness
b_NA_close_neg <- fixef(h22_NA_out_close_neg)["NA_sim_perc"]
se_b_NA_close_neg <- sqrt(vcov(h22_NA_out_close_neg)["NA_sim_perc", "NA_sim_perc"])

b_NA_sim_close_neg <- rnorm(n_sim, mean = b_NA_close_neg, sd = se_b_NA_close_neg)
ind_NA_close_neg <- a_NA_sim_neg * b_NA_sim_close_neg

res_NA_close_neg <- mc_ci(ind_NA_close_neg, a_NA_neg, b_NA_close_neg, "NA close neg")

cat("\n--- Closeness (negative) ---\n")
cat(sprintf("Indirect: %.4f, 95%% MC CI [%.4f, %.4f], p = %s\n",
            res_NA_close_neg$point, res_NA_close_neg$ci_low, res_NA_close_neg$ci_hi,
            ifelse(res_NA_close_neg$p < .001, "< .001", sprintf("%.3f", res_NA_close_neg$p))))
```

### Positive Interaction

```{r h22_NA_mediator_pos}
#| output-fold: true
#| output-summary: "Mediator model NA (positive)"

h22_NA_med_pos <- lme(
  fixed = NA_sim_perc ~ NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h22_NA_med_pos)
```

```{r h22_NA_out_love_pos}
#| output-fold: true
#| output-summary: "Outcome model NA → Love (positive)"

h22_NA_out_love_pos <- lme(
  fixed = love ~ NA_sim_perc + NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h22_NA_out_love_pos)
```

```{r h22_NA_out_close_pos}
#| output-fold: true
#| output-summary: "Outcome model NA → Closeness (positive)"

h22_NA_out_close_pos <- lme(
  fixed = close ~ NA_sim_perc + NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h22_NA_out_close_pos)
```

#### Monte Carlo CI (positive interaction)

```{r h22_NA_MC_pos}
#| output-fold: true
#| output-summary: "Monte Carlo indirect effects — NA (positive)"

# Love
a_NA_pos     <- fixef(h22_NA_med_pos)["NA_sim_act"]
se_a_NA_pos  <- sqrt(vcov(h22_NA_med_pos)["NA_sim_act", "NA_sim_act"])
b_NA_love_pos <- fixef(h22_NA_out_love_pos)["NA_sim_perc"]
se_b_NA_love_pos <- sqrt(vcov(h22_NA_out_love_pos)["NA_sim_perc", "NA_sim_perc"])

a_NA_sim_pos <- rnorm(n_sim, mean = a_NA_pos, sd = se_a_NA_pos)
b_NA_sim_love_pos <- rnorm(n_sim, mean = b_NA_love_pos, sd = se_b_NA_love_pos)
ind_NA_love_pos <- a_NA_sim_pos * b_NA_sim_love_pos

res_NA_love_pos <- mc_ci(ind_NA_love_pos, a_NA_pos, b_NA_love_pos, "NA love pos")

cat("--- Love (positive) ---\n")
cat(sprintf("Indirect: %.4f, 95%% MC CI [%.4f, %.4f], p = %s\n",
            res_NA_love_pos$point, res_NA_love_pos$ci_low, res_NA_love_pos$ci_hi,
            ifelse(res_NA_love_pos$p < .001, "< .001", sprintf("%.3f", res_NA_love_pos$p))))

# Closeness
b_NA_close_pos <- fixef(h22_NA_out_close_pos)["NA_sim_perc"]
se_b_NA_close_pos <- sqrt(vcov(h22_NA_out_close_pos)["NA_sim_perc", "NA_sim_perc"])

b_NA_sim_close_pos <- rnorm(n_sim, mean = b_NA_close_pos, sd = se_b_NA_close_pos)
ind_NA_close_pos <- a_NA_sim_pos * b_NA_sim_close_pos

res_NA_close_pos <- mc_ci(ind_NA_close_pos, a_NA_pos, b_NA_close_pos, "NA close pos")

cat("\n--- Closeness (positive) ---\n")
cat(sprintf("Indirect: %.4f, 95%% MC CI [%.4f, %.4f], p = %s\n",
            res_NA_close_pos$point, res_NA_close_pos$ci_low, res_NA_close_pos$ci_hi,
            ifelse(res_NA_close_pos$p < .001, "< .001", sprintf("%.3f", res_NA_close_pos$p))))
```

# Step 3a: Moderation by Trait We-ness (H3.1–H3.2)

We examine whether the similarity–evaluation association depends on trait we-ness ($W^T_{pd}$, grand-mean centered). Models are estimated separately within each interaction context.

For actual similarity:

$$
O_{pdc} = b_{0,d} + \beta_1 \bar{S}^{act}_{pdc} + \beta_2 W^T_{pd} + \beta_3 (\bar{S}^{act}_{pdc} \times W^T_{pd}) + \varepsilon_{pdc}
$$

For perceived similarity:

$$
O_{pdc} = b_{0,d} + \beta_1 \bar{S}^{perc}_{pdc} + \beta_2 W^T_{pd} + \beta_3 (\bar{S}^{perc}_{pdc} \times W^T_{pd}) + \varepsilon_{pdc}
$$

Both with dyad-level random intercept. The conditional similarity–outcome association is $\beta_1 + \beta_3 W^T_{pd}$.

For PA (H3.1): expected $\beta_3 > 0$ (we-ness amplifies positive effects). For NA (H3.2): expected $\beta_3 > 0$ (we-ness buffers negative effects, making the negative slope less negative).

## H3.1 – PA × We-ness

### Actual PA Similarity

#### Love

```{r h31_PA_act_love_neg}
#| output-fold: true
#| output-summary: "H3.1 actual PA × we-ness → Love (negative)"

h31_PA_act_love_neg <- lme(
  fixed = love ~ PA_sim_act + cwe_ness + PA_sim_act:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h31_PA_act_love_neg)
```

```{r h31_PA_act_love_pos}
#| output-fold: true
#| output-summary: "H3.1 actual PA × we-ness → Love (positive)"

h31_PA_act_love_pos <- lme(
  fixed = love ~ PA_sim_act + cwe_ness + PA_sim_act:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h31_PA_act_love_pos)
```

#### Probing (if interaction significant)

```{r h31_PA_act_probe_neg}
#| output-fold: true
#| output-summary: "Simple slopes — actual PA × we-ness → Love (negative)"

we_sd <- sd(sc2_neg$cwe_ness, na.rm = TRUE)

slopes_h31 <- emtrends(
  h31_PA_act_love_neg,
  specs = ~ cwe_ness,
  var   = "PA_sim_act",
  at    = list(cwe_ness = c(-we_sd, 0, we_sd))
)

slopes_h31_df <- as.data.frame(summary(slopes_h31, infer = TRUE))

level_labels <- c("Low (−1 SD)", "Average", "High (+1 SD)")

tbl_h31 <- data.frame(
  `We-ness level` = level_labels,
  `We-ness value` = sprintf("%.2f", slopes_h31_df$cwe_ness),
  `Simple slope`  = sprintf("%.4f", slopes_h31_df$PA_sim_act.trend),
  `SE`            = sprintf("%.4f", slopes_h31_df$SE),
  `95% CI`        = sprintf("[%.4f, %.4f]", slopes_h31_df$lower.CL, slopes_h31_df$upper.CL),
  `t`             = sprintf("%.3f", slopes_h31_df$t.ratio),
  `p`             = ifelse(slopes_h31_df$p.value < .001, "< .001",
                           sprintf("%.3f", slopes_h31_df$p.value)),
  check.names = FALSE
)

kable(tbl_h31,
  align    = c("l", "c", "c", "c", "c", "c", "c"),
  booktabs = TRUE,
  caption  = "Simple slopes of actual PA similarity on love at low, average, and high trait we-ness (negative interaction)"
) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    position   = "left"
  ) %>%
  footnote(
    general = "Simple slopes estimated at −1 SD, mean, and +1 SD of trait we-ness (grand-mean centered).",
    general_title = "Note.",
    footnote_as_chunk = TRUE
  )
```

#### Closeness

```{r h31_PA_act_close_neg}
#| output-fold: true
#| output-summary: "H3.1 actual PA × we-ness → Closeness (negative)"

h31_PA_act_close_neg <- lme(
  fixed = close ~ PA_sim_act + cwe_ness + PA_sim_act:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h31_PA_act_close_neg)
```

```{r h31_PA_act_close_pos}
#| output-fold: true
#| output-summary: "H3.1 actual PA × we-ness → Closeness (positive)"

h31_PA_act_close_pos <- lme(
  fixed = close ~ PA_sim_act + cwe_ness + PA_sim_act:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h31_PA_act_close_pos)
```

### Perceived PA Similarity

#### Love

```{r h31_PA_perc_love_neg}
#| output-fold: true
#| output-summary: "H3.1 perceived PA × we-ness → Love (negative)"

h31_PA_perc_love_neg <- lme(
  fixed = love ~ PA_sim_perc + cwe_ness + PA_sim_perc:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h31_PA_perc_love_neg)
```

```{r h31_PA_perc_love_pos}
#| output-fold: true
#| output-summary: "H3.1 perceived PA × we-ness → Love (positive)"

h31_PA_perc_love_pos <- lme(
  fixed = love ~ PA_sim_perc + cwe_ness + PA_sim_perc:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h31_PA_perc_love_pos)
```

#### Closeness

```{r h31_PA_perc_close_neg}
#| output-fold: true
#| output-summary: "H3.1 perceived PA × we-ness → Closeness (negative)"

h31_PA_perc_close_neg <- lme(
  fixed = close ~ PA_sim_perc + cwe_ness + PA_sim_perc:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h31_PA_perc_close_neg)
```

```{r h31_PA_perc_close_pos}
#| output-fold: true
#| output-summary: "H3.1 perceived PA × we-ness → Closeness (positive)"

h31_PA_perc_close_pos <- lme(
  fixed = close ~ PA_sim_perc + cwe_ness + PA_sim_perc:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h31_PA_perc_close_pos)
```

## H3.2 – NA × We-ness

### Actual NA Similarity

#### Love

```{r h32_NA_act_love_neg}
#| output-fold: true
#| output-summary: "H3.2 actual NA × we-ness → Love (negative)"

h32_NA_act_love_neg <- lme(
  fixed = love ~ NA_sim_act + cwe_ness + NA_sim_act:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h32_NA_act_love_neg)
```

```{r h32_NA_act_love_pos}
#| output-fold: true
#| output-summary: "H3.2 actual NA × we-ness → Love (positive)"

h32_NA_act_love_pos <- lme(
  fixed = love ~ NA_sim_act + cwe_ness + NA_sim_act:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h32_NA_act_love_pos)
```

#### Closeness

```{r h32_NA_act_close_neg}
#| output-fold: true
#| output-summary: "H3.2 actual NA × we-ness → Closeness (negative)"

h32_NA_act_close_neg <- lme(
  fixed = close ~ NA_sim_act + cwe_ness + NA_sim_act:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h32_NA_act_close_neg)
```

```{r h32_NA_act_close_pos}
#| output-fold: true
#| output-summary: "H3.2 actual NA × we-ness → Closeness (positive)"

h32_NA_act_close_pos <- lme(
  fixed = close ~ NA_sim_act + cwe_ness + NA_sim_act:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h32_NA_act_close_pos)
```

### Perceived NA Similarity

#### Love

```{r h32_NA_perc_love_neg}
#| output-fold: true
#| output-summary: "H3.2 perceived NA × we-ness → Love (negative)"

h32_NA_perc_love_neg <- lme(
  fixed = love ~ NA_sim_perc + cwe_ness + NA_sim_perc:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h32_NA_perc_love_neg)
```

```{r h32_NA_perc_love_pos}
#| output-fold: true
#| output-summary: "H3.2 perceived NA × we-ness → Love (positive)"

h32_NA_perc_love_pos <- lme(
  fixed = love ~ NA_sim_perc + cwe_ness + NA_sim_perc:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h32_NA_perc_love_pos)
```

#### Closeness

```{r h32_NA_perc_close_neg}
#| output-fold: true
#| output-summary: "H3.2 perceived NA × we-ness → Closeness (negative)"

h32_NA_perc_close_neg <- lme(
  fixed = close ~ NA_sim_perc + cwe_ness + NA_sim_perc:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h32_NA_perc_close_neg)
```

```{r h32_NA_perc_close_pos}
#| output-fold: true
#| output-summary: "H3.2 perceived NA × we-ness → Closeness (positive)"

h32_NA_perc_close_pos <- lme(
  fixed = close ~ NA_sim_perc + cwe_ness + NA_sim_perc:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h32_NA_perc_close_pos)
```

# Step 3b: Context Moderation — Negative vs. Positive Interaction (H3.3–H3.4)

We examine whether the similarity–evaluation association differs between the positive and negative interaction contexts. A dummy variable $C_{pdc} = c$ ($c = 0$ negative, $c = 1$ positive) is included in pooled models. Since these models contain two observations per person, both dyad-level and person-level random intercepts are included.

For actual similarity:

$$
O_{pdc} = b_{0,pd} + \beta_1 \bar{S}^{act}_{pdc} + \beta_2 C_{pdc} + \beta_3 (\bar{S}^{act}_{pdc} \times C_{pdc}) + \varepsilon_{pdc}
$$

$$
b_{0,pd} = \beta_0 + v_{0,d} + u_{0,pd}
$$

For perceived similarity: analogous model with $\bar{S}^{perc}_{pdc}$.

$\beta_1$ is the similarity–evaluation slope for the negative interaction ($C_{pdc} = 0$), and $\beta_1 + \beta_3$ is the slope for the positive interaction.

For PA (H3.3): expected $\beta_3 < 0$ (PA similarity more beneficial in negative interaction). For NA (H3.4): expected $\beta_3 < 0$ (NA similarity more harmful in positive interaction).

## H3.3 – PA × Context

### Actual PA Similarity

#### Love

```{r h33_PA_act_love}
#| output-fold: true
#| output-summary: "H3.3 actual PA × context → Love"

h33_PA_act_love <- lme(
  fixed = love ~ PA_sim_act + context + PA_sim_act:context,
  random = ~ 1 | dyad/person,
  data = sc2,
  method = "REML",
  na.action = na.exclude
)

model_output(h33_PA_act_love)
```

```{r h33_PA_act_love_slopes}
#| output-fold: true
#| output-summary: "Simple slopes — actual PA × context → Love"

# Simple slopes at negative (context = 0) and positive (context = 1) interaction
slopes_h33 <- emtrends(
  h33_PA_act_love,
  specs = ~ context,
  var   = "PA_sim_act",
  at    = list(context = c(0, 1))
)

slopes_h33_df <- as.data.frame(summary(slopes_h33, infer = TRUE))

tbl_h33 <- data.frame(
  `Interaction context` = c("Negative (0)", "Positive (1)"),
  `Simple slope`        = sprintf("%.4f", slopes_h33_df$PA_sim_act.trend),
  `SE`                  = sprintf("%.4f", slopes_h33_df$SE),
  `95% CI`              = sprintf("[%.4f, %.4f]", slopes_h33_df$lower.CL, slopes_h33_df$upper.CL),
  `t`                   = sprintf("%.3f", slopes_h33_df$t.ratio),
  `p`                   = ifelse(slopes_h33_df$p.value < .001, "< .001",
                                 sprintf("%.3f", slopes_h33_df$p.value)),
  check.names = FALSE
)

kable(tbl_h33,
  align    = c("l", "c", "c", "c", "c", "c"),
  booktabs = TRUE,
  caption  = "Simple slopes of actual PA similarity on love in negative and positive interactions (H3.3)"
) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    position   = "left"
  )
```

#### Closeness

```{r h33_PA_act_close}
#| output-fold: true
#| output-summary: "H3.3 actual PA × context → Closeness"

h33_PA_act_close <- lme(
  fixed = close ~ PA_sim_act + context + PA_sim_act:context,
  random = ~ 1 | dyad/person,
  data = sc2,
  method = "REML",
  na.action = na.exclude
)

model_output(h33_PA_act_close)
```

### Perceived PA Similarity

#### Love

```{r h33_PA_perc_love}
#| output-fold: true
#| output-summary: "H3.3 perceived PA × context → Love"

h33_PA_perc_love <- lme(
  fixed = love ~ PA_sim_perc + context + PA_sim_perc:context,
  random = ~ 1 | dyad/person,
  data = sc2,
  method = "REML",
  na.action = na.exclude
)

model_output(h33_PA_perc_love)
```

#### Closeness

```{r h33_PA_perc_close}
#| output-fold: true
#| output-summary: "H3.3 perceived PA × context → Closeness"

h33_PA_perc_close <- lme(
  fixed = close ~ PA_sim_perc + context + PA_sim_perc:context,
  random = ~ 1 | dyad/person,
  data = sc2,
  method = "REML",
  na.action = na.exclude
)

model_output(h33_PA_perc_close)
```

## H3.4 – NA × Context

### Actual NA Similarity

#### Love

```{r h34_NA_act_love}
#| output-fold: true
#| output-summary: "H3.4 actual NA × context → Love"

h34_NA_act_love <- lme(
  fixed = love ~ NA_sim_act + context + NA_sim_act:context,
  random = ~ 1 | dyad/person,
  data = sc2,
  method = "REML",
  na.action = na.exclude
)

model_output(h34_NA_act_love)
```

```{r h34_NA_act_love_slopes}
#| output-fold: true
#| output-summary: "Simple slopes — actual NA × context → Love"

slopes_h34 <- emtrends(
  h34_NA_act_love,
  specs = ~ context,
  var   = "NA_sim_act",
  at    = list(context = c(0, 1))
)

slopes_h34_df <- as.data.frame(summary(slopes_h34, infer = TRUE))

tbl_h34 <- data.frame(
  `Interaction context` = c("Negative (0)", "Positive (1)"),
  `Simple slope`        = sprintf("%.4f", slopes_h34_df$NA_sim_act.trend),
  `SE`                  = sprintf("%.4f", slopes_h34_df$SE),
  `95% CI`              = sprintf("[%.4f, %.4f]", slopes_h34_df$lower.CL, slopes_h34_df$upper.CL),
  `t`                   = sprintf("%.3f", slopes_h34_df$t.ratio),
  `p`                   = ifelse(slopes_h34_df$p.value < .001, "< .001",
                                 sprintf("%.3f", slopes_h34_df$p.value)),
  check.names = FALSE
)

kable(tbl_h34,
  align    = c("l", "c", "c", "c", "c", "c"),
  booktabs = TRUE,
  caption  = "Simple slopes of actual NA similarity on love in negative and positive interactions (H3.4)"
) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    position   = "left"
  )
```

#### Closeness

```{r h34_NA_act_close}
#| output-fold: true
#| output-summary: "H3.4 actual NA × context → Closeness"

h34_NA_act_close <- lme(
  fixed = close ~ NA_sim_act + context + NA_sim_act:context,
  random = ~ 1 | dyad/person,
  data = sc2,
  method = "REML",
  na.action = na.exclude
)

model_output(h34_NA_act_close)
```

### Perceived NA Similarity

#### Love

```{r h34_NA_perc_love}
#| output-fold: true
#| output-summary: "H3.4 perceived NA × context → Love"

h34_NA_perc_love <- lme(
  fixed = love ~ NA_sim_perc + context + NA_sim_perc:context,
  random = ~ 1 | dyad/person,
  data = sc2,
  method = "REML",
  na.action = na.exclude
)

model_output(h34_NA_perc_love)
```

#### Closeness

```{r h34_NA_perc_close}
#| output-fold: true
#| output-summary: "H3.4 perceived NA × context → Closeness"

h34_NA_perc_close <- lme(
  fixed = close ~ NA_sim_perc + context + NA_sim_perc:context,
  random = ~ 1 | dyad/person,
  data = sc2,
  method = "REML",
  na.action = na.exclude
)

model_output(h34_NA_perc_close)
```
