---
title: "Study Component 2: Affect Similarity in Laboratory Interactions (VMR)"
---

```{r setup}
#| include: false

# 1. Load packages, paths and data configurations
source(here::here("R", "00_setup.R"))

# 2. Set a global theme for plots
theme_set(theme_minimal() + theme(legend.position = "bottom"))

```

```{r SC2_setup}
#| include: false
# load analysis datasets
vmr_data  <- readRDS(file.path(dir_data_ana, "vmr_ana.rds"))
post_data <- readRDS(file.path(dir_data_ana, "post_ana.rds"))
bg_data   <- readRDS(file.path(dir_data_ana, "bg_ana.rds"))

# plot initialisation
plot_counter <- 1
prefix <- "SC2"

# mente carlo number of simulations
n_sim <- 20000
```

# Introduction

See also [OSF preregistration](https://doi.org/10.17605/OSF.IO/V7T9F){.uri}.

## Overview

Study Component 2 examines how **interaction-averaged affect similarity** during structured positive and negative laboratory interactions relates to **post-interaction love and closeness**, using video-mediated recall (VMR) data from romantic couples.

As in Study Component 1, we distinguish:

-   **Positive Affect (PA)** and **Negative Affect (NA)** (analyzed separately)
-   **Actual similarity** (self-reported affect compared to partner's self-reported affect)
-   **Perceived similarity** (self-reported affect compared to actor's perception of partner's affect)

Similarity is again conceptualized as **shared, co-experienced elevated affect**, operationalized as dyadic elevation (mean level) penalized by mismatch (absolute difference). For Study Component 2, we average the segment-level similarity indices across the interaction.

The main research questions are:

1.  Does interaction-averaged affect similarity predict post-interaction love and closeness? (H1)
2.  Is this association mediated by perceived similarity? (H2)
3.  Are these associations moderated by trait we-ness and interaction context? (H3)

All analyses use multilevel models with persons ($p$) nested within dyads ($d$), estimated separately for the positive and negative interaction contexts ($c$). For context moderation (H3.3–H3.4), a pooled model includes both contexts.

There are two post-interaction outcomes: love and closeness. Models are estimated for both outcomes.

## Conceptualization and Operationalization

### Notation and Multilevel Structure

Let:

-   $O_{pdc}$ = post-interaction outcome (love or closeness) for person $p$ in dyad $d$ after interaction $c$
-   $A_{mpdc}$ = actor affect at VMR segment $m$
-   $P^{act}_{mpdc}$ = partner's actual affect
-   $P^{perc}_{mpdc}$ = partner affect as perceived by actor

Interaction context: $c = 0$ (negative interaction), $c = 1$ (positive interaction).

### Person-Mean Centering

Affect variables are person-mean centered across all VMR segments and across both interactions for each person $p$ in dyad $d$, as implemented in `04_data_construct.R`:

$$
X_{mpdc} = A_{mpdc} - \frac{1}{N_{pd}} \sum_{m,c} A_{mpdc}
$$

and analogously for $Y^{act}_{mpdc}$ and $Y^{perc}_{mpdc}$.

### Similarity Indices

For each VMR segment $m$ in interaction $c$ within dyad $d$:

$$
S^{act}_{mpdc} = \frac{X_{mpdc} + Y^{act}_{mpdc}}{2} - |X_{mpdc} - Y^{act}_{mpdc}|
$$

$$
S^{perc}_{mpdc} = \frac{X_{mpdc} + Y^{perc}_{mpdc}}{2} - |X_{mpdc} - Y^{perc}_{mpdc}|
$$

These are computed at the segment level in `04_data_construct.R`. For the interaction-level analyses, we average across segments within each interaction $c$:

$$
\bar{S}^{act}_{pdc} = \frac{1}{N_{pdc}} \sum_m S^{act}_{mpdc}, \quad \bar{S}^{perc}_{pdc} = \frac{1}{N_{pdc}} \sum_m S^{perc}_{mpdc}
$$

### Random Intercepts and Estimation

For models estimated separately within each interaction context (Steps 1a–1c, 2, 3a), there is one observation per person, so only a **dyad-level random intercept** is included.

For the context moderation model (Step 3b), data from both interactions are pooled (two observations per person), and random intercepts for both **dyads** and **persons nested within dyads** are included.

Unlike Study Component 1, we do **not** specify AR(1) residual structures because the outcomes are at the interaction level.

### Trait We-ness

Trait we-ness ($W^T_{pd}$) is constructed from the Dyadic Coping Inventory (DCI) as the mean of selected items reflecting positive shared dyadic coping (see `04_data_construct.R`). It is grand-mean centered so that $0$ reflects the average level of trait we-ness in the sample.

## Hypotheses Summary

-   **H1.1** (PA): Higher average actual PA similarity → higher post-interaction love/closeness ($\beta_1 > 0$).
-   **H1.2** (NA): Higher average actual NA similarity → lower post-interaction love/closeness ($\beta_1 < 0$).
-   **H1.3**: Perceived similarity added alongside actual similarity shows comparable direction; actual similarity coefficients attenuate.
-   **H2.1** (PA mediation): Actual PA similarity → perceived PA similarity → higher love/closeness (positive indirect effect).
-   **H2.2** (NA mediation): Actual NA similarity → perceived NA similarity → lower love/closeness (negative indirect effect).
-   **H3.1** (PA × we-ness): Higher trait we-ness amplifies the positive PA similarity–evaluation association ($\beta_3 > 0$).
-   **H3.2** (NA × we-ness): Higher trait we-ness attenuates the negative NA similarity–evaluation association ($\beta_3 > 0$).
-   **H3.3** (PA × context): PA similarity more beneficial in the negative than positive interaction ($\beta_3 < 0$).
-   **H3.4** (NA × context): NA similarity more detrimental in the positive than negative interaction ($\beta_3 < 0$).

# Data Preparation

## Aggregate VMR Similarity to Interaction Level

```{r aggregate_vmr}

# Average similarity indices across VMR segments within each interaction (topic)
vmr_agg <- vmr_data %>%
  group_by(dyad, person, topic) %>%
  summarise(
    PA_sim_act  = mean(PA_similarity_act,  na.rm = TRUE),
    PA_sim_perc = mean(PA_similarity_perc, na.rm = TRUE),
    NA_sim_act  = mean(NA_similarity_act,  na.rm = TRUE),
    NA_sim_perc = mean(NA_similarity_perc, na.rm = TRUE),
    n_segments  = n(),
    .groups = "drop"
  )

cat(sprintf("VMR aggregated: %d rows (persons × interactions)\n", nrow(vmr_agg)))
cat(sprintf("Unique persons: %d\n", n_distinct(vmr_agg$person)))
cat(sprintf("Unique dyads:   %d\n", n_distinct(vmr_agg$dyad)))
```

## Reshape Post-Interaction Outcomes to Long Format

```{r reshape_post}

# Reshape post-interaction data to long format with context indicator
post_long <- post_data %>%
  pivot_longer(
    cols = c(love_neg, close_neg, love_pos, close_pos),
    names_to = c(".value", "topic"),
    names_pattern = "(love|close)_(neg|pos)"
  ) %>%
  mutate(topic = ifelse(topic == "neg", "negative", "positive"))

cat(sprintf("POST reshaped: %d rows\n", nrow(post_long)))
```

## Merge Datasets

```{r merge_sc2}

# Merge VMR aggregates with post-interaction outcomes
sc2 <- vmr_agg %>%
  inner_join(post_long, by = c("dyad", "person", "topic"))

# Merge with trait we-ness from background questionnaire
sc2 <- sc2 %>%
  left_join(bg_data %>% select(person, we_ness), by = "person")

# Create context dummy (0 = negative, 1 = positive)
sc2 <- sc2 %>%
  mutate(context = ifelse(topic == "positive", 1, 0))

# Grand-mean center trait we-ness
sc2 <- sc2 %>%
  mutate(cwe_ness = we_ness - mean(we_ness, na.rm = TRUE))

cat(sprintf("SC2 merged dataset: %d rows, %d persons, %d dyads\n",
            nrow(sc2), n_distinct(sc2$person), n_distinct(sc2$dyad)))

# Grand-mean centered similarity indices for Step 1c interaction models
# Centering is done within each context as specified in the preregistration
sc2 <- sc2 %>%
  group_by(topic) %>%
  mutate(
    PA_sim_act_gmc  = PA_sim_act  - mean(PA_sim_act,  na.rm = TRUE),
    PA_sim_perc_gmc = PA_sim_perc - mean(PA_sim_perc, na.rm = TRUE),
    NA_sim_act_gmc  = NA_sim_act  - mean(NA_sim_act,  na.rm = TRUE),
    NA_sim_perc_gmc = NA_sim_perc - mean(NA_sim_perc, na.rm = TRUE)
  ) %>%
  ungroup()

# Split by context for separate analyses
sc2_neg <- sc2 %>% filter(topic == "negative")
sc2_pos <- sc2 %>% filter(topic == "positive")

cat(sprintf("Negative context: %d obs\n", nrow(sc2_neg)))
cat(sprintf("Positive context: %d obs\n", nrow(sc2_pos)))
```

## Model Specification in R

For models estimated within a single context (Steps 1a–1c, 2, 3a), one observation per person per interaction, only a dyad-level random intercept is estimable:

> `lme(fixed = outcome ~ predictor(s), random = ~ 1 | dyad, data = ..., method = "REML", na.action = na.exclude)`

For the pooled context moderation model (Step 3b), two observations per person allow a nested random intercept structure:

> `lme(fixed = outcome ~ predictor(s), random = ~ 1 | dyad/person, data = ..., method = "REML", na.action = na.exclude)`

# Step 1a: Main Effects of Actual Similarity (H1.1–H1.2)

Primary model, estimated separately within each interaction context $c$:

$$
O_{pdc} = b_{0,d} + \beta_1 \bar{S}^{act}_{pdc} + \varepsilon_{pdc}
$$

Random intercept structure:

$$
b_{0,d} = \beta_0 + v_{0,d}
$$

-   $v_{0,d} \sim N(0, \tau_d^2)$: dyad-level random intercept

Interpretation:

-   $\beta_1$: association between interaction-averaged actual similarity and the post-interaction outcome. For PA (H1.1), expected $\beta_1 > 0$; for NA (H1.2), expected $\beta_1 < 0$.

## H1.1 – Actual PA Similarity

### Love

#### Negative interaction

```{r h11_PA_love_neg}
#| output-fold: true
#| output-summary: "H1.1 PA → Love (negative interaction)"

h11_PA_love_neg <- lme(
  fixed = love ~ PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h11_PA_love_neg)
```

**Intercept** $\beta_0 = 89.05$, 95% CI $[79.88,98.23]$, $p < .001$

This is the expected post-interaction love when $\bar{S}^{act}_{pdc} = 0$, i.e., when a person's interaction-averaged actual PA similarity is zero. As we discussed, that means the elevation and mismatch exactly cancel out across the interaction. On average, the partners were at their affect baselines or any shared elevation was exactly offset by divergence.

The value of 89 is high on a 0–100 scale, suggesting that even after a **negative** interaction (disagreement topic), couples report feeling quite loving toward their partner. This might reflect that the lab setting is relatively benign, or that love is a more stable construct that does not drop sharply from a single 8-minute disagreement.

**Slope** $\beta_1 = 0.379$, 95% CI $[0.060,0.697]$, $p=.020$

For every one-unit increase in interaction-averaged actual PA similarity, post-interaction love increases by 0.38 points. Since PA similarity is on a scale derived from 0–100 affect ratings, a 10-unit increase in average similarity corresponds to +3.8 love points. The effect is significant but modest, and the CI is wide, reflecting substantial uncertainty with only 98 dyad-level data points informing this estimate. The direction is consistent with H1.1: more shared PA during the negative interaction predicts more love afterward.

**Random effects and variance decomposition**

**Dyad intercept** $SD=8.84$, variance $= 78.16$

**Residual** $SD=18.35$, variance $= 336.86$

The total variance is $78.16+336.86=415.02$

The **ICC\*** (proportion of variance at the dyad level) is $78.16/415.02=0.188$, or about $19\%$. This means that roughly one fifth of the variability in post-interaction love is attributable to stable differences between couples, while the remaining 81% is within-dyad, differences between the two partners within the same couple. The residual is quite large ($SD=18.35$) relative to the dyad intercept ($SD=8.84$). This means partners within the same dyad can differ substantially in how much love they report after the same interaction. That makes sense: two people can experience the same disagreement quite differently. An ICC of 0.19 is modest but non-negligible, confirming that the dyad random intercept is justified: ignoring the nesting would underestimate standard errors.

\*Because the model includes only a random intercept and no random slope for PA similarity, this (partial) ICC concerns between-couple differences in average post-interaction love, not between-couple differences in the strength of the PA similarity effect.

**The model provides initial support for H1.1** in the negative interaction context: couples who share more PA on average during the disagreement interaction report more love afterward. But the effect is modest, estimated with considerable uncertainty

#### Positive interaction

```{r h11_PA_love_pos}
#| output-fold: true
#| output-summary: "H1.1 PA → Love (positive interaction)"

h11_PA_love_pos <- lme( fixed = love ~ PA_sim_act, random = ~ 1 | dyad, data = sc2_pos, method = "REML", na.action = na.exclude )

model_output(h11_PA_love_pos)
```

**Intercept** $\beta_0 = 85.97$, 95% CI $[82.79, 89.15]$, $p < .001$.

Expected post-interaction love when interaction-averaged actual PA similarity is zero. The intercept is lower than in the negative interaction model ($89.05$), which might seem counterintuitive since one might expect higher love after a positive interaction. However, this comparison is not straightforward because the intercept is evaluated at $\bar{S}^{act} = 0$, and the distribution of PA similarity likely differs between contexts (the positive interaction may have a higher mean PA similarity such that $\bar{S}^{act} = 0$ indicates a 'below average couple' , while hte negative interaction may have lower mean PA similarity so the intercept is looking at a 'higher than average couple'. The intercept being lower in the positive context ($85.97$) than the negative context ($89.05$) may simply mean: a below-average-similarity couple in a positive interaction reports less love than an above-average-similarity couple in a negative interaction.

**Slope** $\beta_1 = 0.459$, 95% CI $[0.127, 0.792]$, $p = .007$.

For every one-unit increase in interaction-averaged actual PA similarity during the positive interaction, post-interaction love increases by $0.46$ points. A 10-unit increase in average similarity corresponds to $+4.6$ love points. The effect is somewhat stronger and more precisely estimated than in the negative interaction ($\beta_1 = 0.379$, $p = .020$), and the evidence is clearer ($p = .007$ vs. $p = .020$). Like the negative interaction, **the direction is consistent with H1.1**.

**Random effects and variance decomposition**

The dyad random intercept is $SD = 10.47$ (variance $= 109.64$) and the residual is $SD = 13.28$ (variance $= 176.33$), yielding a total variance of $285.97$. The ICC is $109.64 / 285.97 = 0.383$, notably higher than in the negative interaction ($0.188$). This means that **after the positive interaction, about 38% of the variability in love is between couples rather than between partners within couples. Partners agree more about their love after a positive interaction than after a negative one, which makes sense: a positive discussion may produce a more uniform relational experience within couples, while a disagreement is more likely to be experienced asymmetrically.**

The residual $SD = 13.28$ is considerably smaller than in the negative interaction ($SD = 18.35$). **Combined with the higher ICC, this indicates that post-interaction love after a positive interaction is both more homogeneous within couples and less variable overall.**

The standardized residuals range from $-3.63$ to $+1.97$, showing the same left-skewed pattern observed across all love models (ceiling compression).

------------------------------------------------------------------------

### Closeness

#### Negative interaction

```{r h11_PA_close_neg}
#| output-fold: true
#| output-summary: "H1.1 PA → Closeness (negative interaction)"

h11_PA_close_neg <- lme(
  fixed = close ~ PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h11_PA_close_neg)
```

**Intercept** $\beta_0 = 92.07$, 95% CI $[82.10, 102.04]$, $p < .001$.

Expected post-interaction closeness when interaction-averaged actual PA similarity is zero. The value is high on the 0–100 scale and even higher than the corresponding love intercept in the same context ($89.05$), suggesting that couples report feeling very close after a negative interaction. There is a wide CI ($\approx 20$ points).

**Slope** $\beta_1 = 0.432$, 95% CI $[0.086, 0.778]$, $p = .015$.

For every one-unit increase in interaction-averaged actual PA similarity during the negative interaction, post-interaction closeness increases by $0.43$ points. A 10-unit increase corresponds to $+4.3$ closeness points. The effect is slightly larger than for love in the same context ($\beta_1 = 0.379$, $p = .020$) and has a comparable level of evidence. T**he direction is consistent with H1.1: sharing more PA during the disagreement predicts feeling closer afterward.**

**Random effects and variance decomposition**

The dyad random intercept is $SD = 12.75$ (variance $= 162.65$) and the residual is $SD = 16.06$ (variance $= 257.79$), yielding a total variance of $420.44$. The ICC is $162.65 / 420.44 = 0.387$. This is substantially higher than the ICC for love in the same context ($0.188$), meaning that partners within the same couple agree much more about their closeness after a negative interaction than about their love. Nearly 40% of the variability in post-interaction closeness is between couples, with the remaining 60% reflecting differences between partners within couples.

#### Positive interaction

```{r h11_PA_close_pos}
#| output-fold: true
#| output-summary: "H1.1 PA → Closeness (positive interaction)"

h11_PA_close_pos <- lme(
  fixed = close ~ PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h11_PA_close_pos)
```

**Intercept** $\beta_0 = 87.27$, 95% CI $[84.36, 90.17]$, $p < .001$.

Expected post-interaction closeness when interaction-averaged actual PA similarity is zero. As with love, the intercept is lower than in the negative interaction ($92.07$), likely reflecting the same issue: the distribution of PA similarity differs between contexts, so the intercept evaluates at a different relative position in the data rather than reflecting a genuine difference in closeness across contexts.

**Slope** $\beta_1 = 0.376$, 95% CI $[0.072, 0.679]$, $p = .016$.

For every one-unit increase in interaction-averaged actual PA similarity during the positive interaction, post-interaction closeness increases by $0.38$ points. A 10-unit increase corresponds to $+3.8$ closeness points. The effect is comparable to the love outcome in the positive interaction ($\beta_1 = 0.459$, $p = .007$) and to closeness in the negative interaction ($\beta_1 = 0.432$, $p = .015$), though slightly smaller in magnitude. T**he direction is again consistent with H1.1.**

**Random effects and variance decomposition**

The dyad random intercept is $SD = 9.37$ (variance $= 87.81$) and the residual is $SD = 12.42$ (variance $= 154.31$), yielding a total variance of $242.12$. The ICC is $87.81 / 242.12 = 0.363$. This is similar to the love ICC in the positive interaction ($0.383$) and to the closeness ICC in the negative interaction ($0.387$), suggesting that roughly a third to two fifths of the variability in post-interaction evaluations is between couples across both outcomes in both contexts, with the exception of love after the negative interaction, where the ICC is notably lower ($0.188$). L**ove after a disagreement is the outcome where partners diverge the most.**

The total variance ($242.12$) is the lowest of the four Step 1a models, indicating that closeness after a positive interaction is the least variable outcome overall. This is consistent with a positive discussion producing both strong agreement within couples and modest differentiation across couples.

### Summary of results

+-------------------------------------------------------------------------+---------------------------------------------------------+--------------------------------------------------------+
|                                                                         | Negative Interaction                                    | Positive Interaction                                   |
+:========================================================================+:========================================================+:=======================================================+
| **Love**                                                                | $\beta_0 = 89.05$, $p < .001$, 95% CI $[79.88,98.23]$   | $\beta_0 = 85.97$, $p < .001$, 95% CI $[82.79, 89.15]$ |
|                                                                         |                                                         |                                                        |
| *('hoe liefdevol voel je je op dit momentten opzichte van je partner')* | $\beta_1 = 0.379$, $p = .020$, 95% CI $[0.060, 0.697]$  | $\beta_1 = 0.459$, $p = .007$, 95% CI $[0.127, 0.792]$ |
|                                                                         |                                                         |                                                        |
|                                                                         | ICC $= 18.8\%$                                          | ICC $= 38.3\%$                                         |
+-------------------------------------------------------------------------+---------------------------------------------------------+--------------------------------------------------------+
| **Closeness**                                                           | $\beta_0 = 92.07$, $p < .001$, 95% CI $[82.10, 102.04]$ | $\beta_0 = 87.27$, $p < .001$, 95% CI $[84.36, 90.17]$ |
|                                                                         |                                                         |                                                        |
| *('hoe hecht voel je je op dit moment met je partner')*                 | $\beta_1 = 0.432$, $p = .015$, 95% CI $[0.086, 0.778]$  | $\beta_1 = 0.376$, $p = .016$, 95% CI $[0.072, 0.679]$ |
|                                                                         |                                                         |                                                        |
|                                                                         | ICC $= 38.7\%$                                          | ICC $= 36.3\%$                                         |
+-------------------------------------------------------------------------+---------------------------------------------------------+--------------------------------------------------------+

: *Note.* Actual PA similarity predicting post-interaction evaluations (H1.1). All models include a dyad-level random intercept. ICC reflects the proportion of variance between dyads (i.e., the expected within-dyad partner correlation).

So from these results it follows that **negative interactions uniquely reduce within-couple agreement for love, but not for closeness.** Closeness seems to be a more consistently dyadic state; love reports seems to be more context-sensitive and individually filtered, especially after disagreement.

But, looking at the separate variance evolution, we see:

-   **For love as outcome, couples converge while partners within couples diverge (scissor effect):**

|             |            |           |          |      |
|-------------|------------|-----------|----------|------|
| **Context** | σ²_between | σ²_within | σ²_total | ICC  |
| Positive    | 109.6      | 176.3     | 285.9    | .383 |
| Negative    | 78.2       | 336.9     | 415.0    | .188 |

-   For **closeness, both couples as partners within couples diverge**:

| Context  | σ²_between | σ²_within | σ²_total | ICC  |
|----------|------------|-----------|----------|------|
| Positive | 87.8       | 154.3     | 242.1    | .363 |
| Negative | 162.6      | 257.8     | 420.4    | .387 |

Thus, for both outcomes, the within dyad variance increases in the negative context model vs. the positive contetx model (partners are scoring more different), while the between dyad variance decreases for love and increased for closeness (couples becoming more similar to each other in their love ratings but less similar in their closeness ratings).

Also noteworthy is that the total variance increases from positive to negative contexts for both outcomes, which suggests that negative interactions are more broadly destabilizing.

### Diagnostics

#### Outcome variable distribution

```{r outcome_variable_distribution}

# Combine the four outcome distributions in one long data frame
plot_df <- bind_rows(
  sc2_neg %>% transmute(dataset = "Negative interaction", love, close),
  sc2_pos %>% transmute(dataset = "Positive interaction", love, close)
) %>%
  pivot_longer(
    cols = c(love, close),
    names_to = "outcome",
    values_to = "value"
  ) %>%
  mutate(
    outcome = dplyr::recode(outcome,
                     love = "Love",
                     close = "Closeness")
  )

# Histogram + density, faceted by outcome and dataset
ggplot(plot_df, aes(x = value)) +
  geom_histogram(aes(y = after_stat(density)),
                 bins = 20, boundary = 0, closed = "left") +
  geom_density(na.rm = TRUE, linewidth = 0.8) +
  facet_grid(outcome ~ dataset) +
  scale_x_continuous(limits = c(0, 100), breaks = seq(0, 100, 10)) +
  labs(
    x = "Score",
    y = "Density",
    title = "Distributions of love and closeness in SC2 datasets"
  ) +
  theme_minimal()
```

#### **Diagnostics - H1.1 Actual PA → Love (Negative Interaction)**

```{r h11_PA_love_neg_diagnostics}
#| warning: false
#| output-fold: true
#| output-summary: "Diagnostics H1.1 PA love (negative interaction)"

lme_diagnostics(h11_PA_love_neg)
```

**Residuals vs. fitted values.** The LOESS curve deviates from the zero line, dipping below zero at lower fitted values (around 55–65) and running slightly above zero for higher fitted values (75–90). This suggests some nonlinearity: the model somewhat overpredicts love for couples with low predicted values and slightly underpredicts at the high end. There is also visible heteroscedasticity: the residual spread is larger at lower fitted values and compresses toward higher fitted values, consistent with a ceiling effect where most observations cluster near the upper bound of the 0–100 scale. The bulk of the data concentrates at fitted values above 75, with relatively few observations below 70 informing the left side of the plot.

**QQ-plot.** The central range (theoretical quantiles between approximately $-1.5$ and $+1.5$) follows the reference line reasonably well, despite the clear tail deviations, indicating approximate normality in the bulk of the distribution. However, the left tail deviates substantially: observed residuals are more extreme than expected under normality, with several points falling well below the line. The right tail is compressed (points fall below the line), meaning the model produces fewer large positive residuals than a normal distribution would predict. This asymmetry is the same ceiling-driven pattern observed in the SC1 models: love ratings near 100 leave no room for positive residuals, producing a left-skewed residual distribution.

These patterns are not ideal but are expected given the bounded outcome scale and the high baseline love observed in this sample. The fixed-effect estimates remain interpretable, but inference should be interpreted with some caution.

#### Diagnostics - H1.1 Actual PA → Love (Positive Interaction)

```{r h11_PA_love_pos_diagnostics}
#| warning: false
#| output-fold: true
#| output-summary: "Diagnostics H1.1 PA love (positive interaction)"

lme_diagnostics(h11_PA_love_pos)
```

**Residuals vs. fitted values.** The LOESS curve runs below zero for lower fitted values (50–75) and approaches zero for higher fitted values (80–95). The pattern is similar to the negative interaction but more pronounced: there are a few isolated observations with very low fitted values (around 50–60) and large negative residuals, pulling the LOESS curve down on the left. Heteroscedasticity is again present: residual spread is wider at lower fitted values and compresses toward the ceiling. The data concentration is shifted further right compared to the negative interaction, with most observations above fitted value 80.

**QQ-plot.** The central range follows the reference line well. The left tail shows the familiar deviation (observed residuals more extreme than expected) with a few notable outliers around theoretical quantile $-3$. The right tail is compressed. Overall comparable to the negative interaction diagnostics.

#### Diagnostics - H1.1 Actual PA → Closeness (Negative Interaction)

```{r h11_PA_close_neg_diagnostics}
#| warning: false
#| output-fold: true
#| output-summary: "Diagnostics H1.1 PA closeness (negative interaction)"

lme_diagnostics(h11_PA_close_neg)
```

**Residuals vs. fitted values.** The LOESS curve shows a similar pattern to the love models: dipping below zero at lower fitted values (55–70) and levelling off near zero at higher fitted values (75–90). The heteroscedasticity pattern is consistent with a wider residual spread at lower fitted values and compression at higher values. The data distribution is very similar to the love model in the same context.

**QQ-plot.** Slightly better behaved than the love models. The central range tracks the reference line closely. The left tail deviates but somewhat less dramatically, with fewer extreme outliers pulling away from the line. The right tail shows moderate compression. Overall, the normality assumption is approximately met in the central range with the expected ceiling-driven departures in the tails.

#### Diagnostics - H1.1 Actual PA → Closeness (Positive Interaction)

```{r h11_PA_close_pos_diagnostics}
#| warning: false
#| output-fold: true
#| output-summary: "Diagnostics H1.1 PA closeness (positive[] interaction)"

lme_diagnostics(h11_PA_close_pos)
```

**Residuals vs. fitted values.** Very similar to the other three models: LOESS below zero at lower fitted values, converging toward zero at higher values. The residual spread again compresses toward higher fitted values. The bulk of the data sits above fitted value 75, consistent with the positive interaction generally producing high closeness ratings.

**QQ-plot.** The best-behaved of the four models. The central range follows the reference line closely, and the left-tail deviations are present but somewhat milder. There is one notable outlier below theoretical quantile $-2.5$. The right tail is compressed as in the other models.

#### Diagnostics - Summary

All four diagnostic panels show the same constellation: some nonlinearity in the LOESS curve driven by sparse low-fitted-value observations, heteroscedasticity from ceiling compression, and left-skewed residuals in the QQ-plot. These patterns are structural consequences of the bounded 0–100 outcome scale combined with high baseline love and closeness in this sample. The closeness models show slightly better residual behaviour than the love models, and the positive interaction models show slightly denser concentration at high fitted values. None of the model diagnostics suggest severe enough violations to make conclusions by default unreliable but the ceiling-driven violations warrant caution for inference.

## H1.2 – Actual NA Similarity

### Love

#### Negative interaction

```{r h12_NA_love_neg}
#| output-fold: true
#| output-summary: "H1.2 NA → Love (negative interaction)"

h12_NA_love_neg <- lme(
  fixed = love ~ NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h12_NA_love_neg)
```

**Intercept** $\beta_0 = 77.53$, 95% CI $[72.57, 82.49]$, $p < .001$

This is the expected post-interaction love when $\bar{S}^{act}_{pdc} = 0$, that is, when a person’s interaction-averaged actual NA similarity is zero. As discussed earlier, this means that shared elevation in negative affect is exactly balanced by dissimilarity across the interaction, so there is no net actual NA similarity on average.

The intercept is notably lower than in the corresponding PA model for the negative interaction ($89.05$), suggesting that when the net similarity score is zero in the NA model, predicted love is lower than when the net similarity score is zero in the PA model.

**Slope** $\beta_1 = -0.141$, 95% CI $[-0.559, 0.276]$, $p = .504$

For every one-unit increase in interaction-averaged actual NA similarity, post-interaction love decreases by about 0.14 points. A 10-unit increase in average NA similarity would thus correspond to a decrease of about 1.4 love points. However, the effect is small, the confidence interval includes both small negative and small positive effects, and the association is clearly non-significant. In other words, there is no evidence in this model that couples who share more negative affect during the disagreement interaction report either less or more love afterward.

**Random effects and variance decomposition**

Dyad intercept $SD = 9.55$, variance $= 91.25$

Residual $SD = 18.35$, variance $= 336.86$

The total variance is $91.25 + 336.86 = 428.11$

The ICC (proportion of variance at the dyad level) is $91.25/428.11 = 0.213$, or about $21\%$. This means that about one fifth of the variability in post-interaction love is attributable to stable differences between couples, while the remaining $79\%$ reflects differences between the two partners within the same couple. As in the PA model for the negative interaction, the residual variance is much larger than the dyad-level variance, indicating substantial divergence between partners in how much love they report after the same disagreement interaction.

The ICC is slightly higher than in the corresponding PA model ($0.188$), but the overall pattern is similar: post-interaction love after a negative interaction contains a substantial within-couple component, with partners often differing meaningfully in their reported love. This again supports the inclusion of a dyad random intercept.

**Overall, the model does not support H1.2 in the negative interaction context: actual NA similarity during the disagreement interaction does not reliably predict post-interaction love.**

#### Positive interaction

```{r h12_NA_love_pos}
#| output-fold: true
#| output-summary: "H1.2 NA → Love (positive interaction)"

h12_NA_love_pos <- lme(
  fixed = love ~ NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h12_NA_love_pos)
```

**Intercept** $\beta_0 = 91.58$, 95% CI $[85.21, 97.95]$, $p < .001$

This is the expected post-interaction love when $\bar{S}^{act}_{pdc} = 0$, that is, when a person’s interaction-averaged actual NA similarity is zero.

The intercept is high on the 0–100 scale, indicating that when actual NA similarity is zero, couples report high love after the positive interaction. It is also clearly higher than the corresponding intercept in the negative interaction NA model ($77.53$), suggesting that post-interaction love is generally higher in the positive interaction context.

**Slope** $\beta_1 = 0.352$, 95% CI $[0.089, 0.615]$, $p = .009$

For every one-unit increase in interaction-averaged actual NA similarity, post-interaction love increases by about 0.35 points. A 10-unit increase in average NA similarity corresponds to an increase of about 3.5 love points.

The direction is noteworthy: greater actual similarity in negative affect during the positive interaction predicts more love afterward. This is opposite to the expectation that shared negative affect would undermine post-interaction relationship evaluations. One possible interpretation is that, in the context of an otherwise positive interaction, sharing negative affect may reflect emotional attunement or mutual engagement rather than relational strain. Regardless of interpretation, the model does not support H1.2.

**Random effects and variance decomposition**

Dyad intercept $SD = 10.51$, variance $= 110.54$

Residual $SD = 13.28$, variance $= 176.33$

The total variance is $110.54 + 176.33 = 286.87$

The ICC (proportion of variance at the dyad level) is $110.54/286.87 = 0.385$, or about $39\%$. This means that roughly two fifths of the variability in post-interaction love is attributable to stable differences between couples, while the remaining $61\%$ reflects within-dyad differences between partners within the same couple. Compared with the negative interaction NA model (ICC $= 0.213$), post-interaction love is more strongly clustered at the dyad level in the positive interaction context (each couple has its own level, and the two partners within that couple ar more alike).

Overall, the model indicates that actual NA similarity during the positive interaction is positively associated with post-interaction love. This finding is statistically reliable but runs counter to the hypothesized negative association, so it does not support H1.2 as originally formulated.

### Closeness

#### Negative interaction

```{r h12_NA_close_neg}
#| output-fold: true
#| output-summary: "H1.2 NA → Closeness (negative interaction)"

h12_NA_close_neg <- lme(
  fixed = close ~ NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h12_NA_close_neg)
```

**Intercept** $\beta_0 = 78.97$, 95% CI $[73.56, 84.38]$, $p < .001$

This is the expected post-interaction closeness when $\bar{S}^{act}_{pdc} = 0$, that is, when a person's interaction-averaged actual NA similarity is zero. The value is lower than the corresponding PA closeness intercept in the same context ($92.07$), mirroring the pattern observed for love where the NA model intercept ($77.53$) was lower than the PA model intercept ($89.05$). It is also very close to the love intercept in the same NA negative-interaction model ($77.53$), suggesting that when net NA similarity is zero, predicted closeness and love after the disagreement interaction are at comparable levels.

**Slope** $\beta_1 = -0.157$, 95% CI $[-0.612, 0.299]$, $p = .497$

For every one-unit increase in interaction-averaged actual NA similarity, post-interaction closeness decreases by about 0.16 points. A 10-unit increase in average NA similarity would correspond to a decrease of about 1.6 closeness points. However, the effect is small, the confidence interval comfortably includes zero in both directions, and the association is clearly non-significant. As with the love outcome in the same context ($\beta_1 = -0.141$, $p = .504$), there is no evidence that couples who share more negative affect during the disagreement interaction report feeling either less or more close afterward. The near-identical magnitude and direction across the two outcomes is notable: actual NA similarity during the negative interaction appears to be uninformative for both post-interaction evaluations.

**Random effects and variance decomposition**

Dyad intercept $SD = 13.41$, variance $= 179.76$

Residual $SD = 16.06$, variance $= 257.79$

The total variance is $179.76 + 257.79 = 437.55$

The ICC is $179.76 / 437.55 = 0.411$, or about $41\%$. This means that roughly two fifths of the variability in post-interaction closeness is attributable to stable differences between couples, while the remaining $59\%$ reflects differences between partners within the same couple. The ICC is notably higher than for love in the same context ($0.213$), consistent with the pattern already observed in the H1.1 models: partners within the same couple agree more about their closeness than about their love after a negative interaction. The dyad-level variance ($179.76$) is also substantially larger than in the love model ($91.25$), meaning that couples differ more from each other in post-interaction closeness than in love.

**Overall, the model does not support H1.2 in the negative interaction context: actual NA similarity during the disagreement interaction does not reliably predict post-interaction closeness.** This mirrors the null finding for love in the same context.

#### Positive interaction

```{r h12_NA_close_pos}
#| output-fold: true
#| output-summary: "H1.2 NA → Closeness (positive interaction)"

h12_NA_close_pos <- lme(
  fixed = close ~ NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h12_NA_close_pos)
```

**Intercept** $\beta_0 = 93.91$, 95% CI $[88.23, 99.59]$, $p < .001$

This is the expected post-interaction closeness when $\bar{S}^{act}_{pdc} = 0$. The intercept is high on the 0–100 scale and clearly higher than the corresponding intercept in the negative interaction NA model ($78.97$), indicating that post-interaction closeness is generally higher after the positive interaction. It is also higher than the love intercept in the same positive-interaction NA model ($91.58$), suggesting that couples report feeling slightly closer than loving after a positive interaction when net NA similarity is zero.

**Slope** $\beta_1 = 0.383$, 95% CI $[0.148, 0.617]$, $p = .002$

For every one-unit increase in interaction-averaged actual NA similarity, post-interaction closeness increases by about 0.38 points. A 10-unit increase in average NA similarity corresponds to an increase of about 3.8 closeness points. The effect is significant and virtually identical in magnitude to the love outcome in the same context ($\beta_1 = 0.352$, $p = .009$), though estimated with somewhat greater precision (narrower CI, smaller $p$-value).

As with love, the direction is opposite to what H1.2 predicts: greater actual similarity in negative affect during the positive interaction predicts feeling *closer* afterward, not less close. The same interpretive logic applies: in the context of a positive interaction, sharing negative affect may signal emotional attunement or mutual engagement rather than relational strain. The convergence across both outcomes strengthens this interpretation — the positive NA similarity effect in the positive interaction is not outcome-specific but a consistent pattern across both relational evaluations.

The contrast with the negative interaction is also notable: in the negative interaction, NA similarity was unrelated to closeness ($\beta_1 = -0.157$, $p = .497$), whereas in the positive interaction the association is clearly positive and significant. This context dependence — a null effect during disagreement but a positive effect during a positive discussion — suggests that the relational meaning of shared negative affect may differ across interaction contexts.

**Random effects and variance decomposition**

Dyad intercept $SD = 9.00$, variance $= 80.91$

Residual $SD = 12.42$, variance $= 154.31$

The total variance is $80.91 + 154.31 = 235.22$

The ICC is $80.91 / 235.22 = 0.344$, or about $34\%$. This means roughly one third of the variability in post-interaction closeness is attributable to stable differences between couples, while the remaining $66\%$ reflects within-dyad differences between partners. The ICC is lower than in the negative interaction closeness model ($0.411$), and notably both variance components are smaller: the dyad-level variance drops from $179.76$ to $80.91$ and the residual from $257.79$ to $154.31$. Post-interaction closeness after a positive interaction is thus less variable overall, with couples becoming more homogeneous and partners within couples also diverging less.

**Overall, the model does not support H1.2 in the positive interaction context.** Instead, actual NA similarity during the positive interaction is significantly and positively associated with post-interaction closeness, mirroring the finding for love. Across both outcomes, the positive interaction yields a consistent pattern: sharing negative affect predicts *better*, not worse, post-interaction relational evaluations.

### Summary of results

+-------------------------------------------------------------------------+----------------------------------------------------------+--------------------------------------------------------+
|                                                                         | Negative Interaction                                     | Positive Interaction                                   |
+:========================================================================+:=========================================================+:=======================================================+
| **Love**                                                                | $\beta_0 = 77.53$, $p < .001$,\                          | $\beta_0 = 91.58$, $p < .001$,\                        |
|                                                                         | 95% CI $[72.57, 82.49]$                                  | 95% CI $[85.21, 97.95]$                                |
| *('hoe liefdevol voel je je op dit momentten opzichte van je partner')* |                                                          |                                                        |
|                                                                         | $\beta_1 = -0.141$, $p = .504$, 95% CI $[-0.559, 0.276]$ | $\beta_1 = 0.352$, $p = .009$, 95% CI $[0.089, 0.615]$ |
|                                                                         |                                                          |                                                        |
|                                                                         | ICC $= 21.3\%$                                           | ICC $= 38.5\%$                                         |
+-------------------------------------------------------------------------+----------------------------------------------------------+--------------------------------------------------------+
| **Closeness**                                                           | $\beta_0 = 78.97$, $p < .001$,\                          | $\beta_0 = 93.91$, $p < .001$,\                        |
|                                                                         | 95% CI $[73.56, 84.38]$                                  | 95% CI $[88.23, 99.59]$                                |
| *('hoe hecht voel je je op dit moment met je partner')*                 |                                                          |                                                        |
|                                                                         | $\beta_1 = -0.157$, $p = .497$,\                         | $\beta_1 = 0.383$, $p = .002$,\                        |
|                                                                         | 95% CI $[-0.612, 0.299]$                                 | 95% CI $[0.148,0.617]$                                 |
|                                                                         |                                                          |                                                        |
|                                                                         | ICC $= 41.1\%$                                           | ICC $= 34.4\%$                                         |
+-------------------------------------------------------------------------+----------------------------------------------------------+--------------------------------------------------------+

: *Note.* Actual NA similarity predicting post-interaction evaluations (H1.2). All models include a dyad-level random intercept. ICC reflects the proportion of variance between dyads (i.e., the expected within-dyad partner correlation).

The results show a clear context split. In the negative interaction, actual NA similarity is unrelated to both love and closeness (both $p > .49$), with near-zero slopes that are virtually identical across the two outcomes ($-0.141$ and $-0.157$). In the positive interaction, actual NA similarity significantly predicts both outcomes in the *positive* direction ($\beta = 0.352$ for love, $\beta = 0.383$ for closeness), counter to the hypothesized negative association.

**H1.2 is not supported in either context.** In the negative interaction, the expected negative association is absent entirely. In the positive interaction, the association is significant but in the wrong direction: sharing more negative affect during a positive discussion predicts *higher* love and closeness afterward.

The intercepts also show a clear context difference. In the negative interaction, the intercepts for both outcomes are in the high 70s ($77.53$ and $78.97$), whereas in the positive interaction they are in the low 90s ($91.58$ and $93.91$). This context gap is expected: post-interaction evaluations are simply higher after a positive discussion than after a disagreement.

The variance decomposition shows:

-   **For love:**

+------------+----------------------+---------------------+--------------------+------------+
| Context    | $\sigma^2_{between}$ | $\sigma^2_{within}$ | $\sigma^2_{total}$ | ICC        |
+============+======================+=====================+====================+============+
| Positive   | 110.5                | 176.3               | 286.9              | .385       |
+------------+----------------------+---------------------+--------------------+------------+
| Negative   | 91.3                 | 336.9               | 428.1              | .213       |
+------------+----------------------+---------------------+--------------------+------------+

-   **For closeness:**

+------------+----------------------+---------------------+--------------------+------------+
| Context    | $\sigma^2_{between}$ | $\sigma^2_{within}$ | $\sigma^2_{total}$ | ICC        |
+============+======================+=====================+====================+============+
| Positive   | 80.9                 | 154.3               | 235.2              | .344       |
+------------+----------------------+---------------------+--------------------+------------+
| Negative   | 179.8                | 257.8               | 437.6              | .411       |
+------------+----------------------+---------------------+--------------------+------------+

The variance patterns closely replicate those from H1.1. For both outcomes, total variance increases substantially from positive to negative contexts, confirming that negative interactions are more broadly destabilizing. The within-dyad variance increases from positive to negative for both outcomes (partners diverge more after disagreement), while the between-dyad variance decreases for love but increases for closeness. This replicates the same scissor effect for love and the uniform expansion for closeness observed in the PA models, which is reassuring: these variance patterns are properties of the outcome distributions and the interaction contexts, not artefacts of the specific predictor entered in the model.

# Step 1b: Adding Perceived Similarity (H1.3)

$$
O_{pdc} = b_{0,d} + \beta_1 \bar{S}^{act}_{pdc} + \beta_2 \bar{S}^{perc}_{pdc} + \varepsilon_{pdc}
$$

with the same dyad-level random intercept as in Step 1a.

Evidence consistent with H1.3: $\beta_2$ in the expected direction, and attenuation of $\beta_1$ relative to Step 1a.

::: callout-important
## Collinearity between actual and perceived similarity in Step 1b and 1c models

When actual and perceived similarity are entered as simultaneous predictors, they share substantial variance because both indices are constructed from overlapping affect ratings (actor affect enters both computations). Model-level correlations between the two predictors range from $r = -0.51$ to $r = -0.69$ across the Step 1b and 1c models. This shared variance inflates the standard errors of both coefficients, because each predictor's unique contribution must be estimated after removing the variance it shares with the other. The practical consequence is that a coefficient that was significant in Step 1a (where actual similarity was the sole predictor) can drop below significance in Step 1b or 1c not because the point estimate changed meaningfully, but because the standard error widened. As will be observed throughout the Step 1b and 1c results, the actual similarity coefficients are broadly stable in magnitude across steps, but their $p$-values weaken due to this shared variance.

This collinearity is structurally present in Study Component 1 as well, but it is less consequential there for two reasons. First, SC1 has many repeated ESM beeps per person, yielding a much larger effective sample size; standard errors are smaller to begin with, so the additional inflation from collinearity is proportionally less impactful. Second, the SC1 predictors are decomposed into within-person and between-person components, and the within-person fluctuations in actual and perceived similarity (how much they co-vary from beep to beep around a person's own mean) may be less correlated than the interaction-level averages used in SC2. Aggregating across VMR segments to a single value per interaction compresses variation and can increase the predictor correlation, because segment-level noise that might differentiate the two indices gets averaged out.
:::

## H1.3 - PA Similarity

### Love

#### Negative interaction

```{r h13_PA_love_neg}
#| output-fold: true
#| output-summary: "H1.3 PA → Love (negative interaction)"

h13_PA_love_neg <- lme(
  fixed = love ~ PA_sim_act + PA_sim_perc,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h13_PA_love_neg)
```

**Intercept** $\beta_0 = 89.10$, 95% CI $[79.90, 98.29]$, $p < .001$

Virtually identical to the Step 1a intercept ($89.05$), as expected: adding perceived similarity does not shift the expected outcome at the zero point of both predictors.

**Actual PA similarity** $\beta_1 = 0.444$, 95% CI $[0.005, 0.883]$, $p = .048$

The coefficient for actual PA similarity *increases* slightly from the Step 1a estimate ($0.379$, $p = .020$) rather than attenuating as H1.3 would predict. The wider confidence interval and weaker $p$-value reflect the added collinearity from including both predictors (the correlation between `PA_sim_act` and `PA_sim_perc` in the model is $r = -0.687$). In other words, after partialling out perceived PA similarity, the unique contribution of actual PA similarity remains significant and, if anything, slightly larger than the total effect in Step 1a.

**Perceived PA similarity** $\beta_2 = -0.081$, 95% CI $[-0.459, 0.296]$, $p = .670$

Perceived PA similarity is non-significant and close to zero when entered alongside actual similarity. The direction is slightly negative, which is opposite to what H1.3 expects (a positive coefficient mirroring the actual similarity effect). Perceived similarity adds no explanatory value beyond actual similarity in this model.

**Random effects and variance decomposition**

Dyad intercept $SD = 8.85$, variance $= 78.25$; residual $SD = 18.40$, variance $= 338.47$. These are virtually unchanged from the Step 1a model (dyad variance $= 78.16$, residual $= 336.86$), confirming that adding perceived similarity absorbs almost no additional variance.

**The model does not support H1.3 in this context.** The expected pattern was for perceived similarity to show a positive coefficient and for actual similarity to attenuate, suggesting that part of the actual similarity effect operates through perceived similarity. Instead, actual similarity retains (and slightly increases) its predictive value, while perceived similarity contributes nothing. This suggests that in the negative interaction, the association between shared PA and post-interaction love reflects the actual affective overlap itself, not the perception of that overlap.

#### Positive interaction

```{r h13_PA_love_pos}
#| output-fold: true
#| output-summary: "H1.3 PA → Love (positive interaction)"

h13_PA_love_pos <- lme(
  fixed = love ~ PA_sim_act + PA_sim_perc,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h13_PA_love_pos)

```

Here is the write-up for `h13_PA_love_pos`, to be placed after the closing ```` ``` ```` of the code chunk at line 675:

------------------------------------------------------------------------

**Intercept** $\beta_0 = 85.52$, 95% CI $[82.03, 89.01]$, $p < .001$

Close to the Step 1a intercept ($85.97$), with a minor shift from entering both predictors.

**Actual PA similarity** $\beta_1 = 0.386$, 95% CI $[-0.022, 0.793]$, $p = .063$

The coefficient decreases slightly from the Step 1a estimate ($0.459$, $p = .007$) and drops below conventional significance. This is the direction of change that H1.3 predicts (attenuation of actual similarity when perceived similarity is added), but the attenuation is modest (from $0.459$ to $0.386$), and the loss of significance is largely attributable to the increased standard error from collinearity between the two predictors ($r = -0.581$) rather than a substantive redistribution of variance to perceived similarity.

**Perceived PA similarity** $\beta_2 = 0.105$, 95% CI $[-0.232, 0.441]$, $p = .538$

Perceived PA similarity is non-significant and small. The direction is positive, which is at least consistent with H1.3's expectation, but the effect is far from reliable and the confidence interval is wide.

**Random effects and variance decomposition**

Dyad intercept $SD = 10.37$, variance $= 107.60$; residual $SD = 13.36$, variance $= 178.37$. These are nearly identical to the Step 1a values (dyad variance $= 109.64$, residual $= 176.33$), indicating that adding perceived similarity absorbs minimal additional variance.

**The model provides at best very weak and inconclusive evidence for H1.3 in the positive interaction.** The attenuation of actual similarity is in the expected direction, and perceived similarity carries the expected positive sign, but neither pattern reaches a level that would constitute meaningful support: the attenuation is small, perceived similarity is clearly non-significant, and the shift in the actual similarity $p$-value (from $.007$ to $.063$) is better attributed to collinearity inflation than to mediation. Taken together with the negative interaction result, Step 1b suggests that perceived PA similarity does not play a meaningful mediating role for post-interaction love in either context.

### Closeness

#### Negative interaction

```{r h13_PA_close_neg}
#| output-fold: true
#| output-summary: "H1.3 PA → Closeness (negative interaction)"

h13_PA_close_neg <- lme(
  fixed = close ~ PA_sim_act + PA_sim_perc,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h13_PA_close_neg)
```

**Intercept** $\beta_0 = 92.08$, 95% CI $[82.09, 102.08]$, $p < .001$

Virtually identical to the Step 1a intercept ($92.07$).

**Actual PA similarity** $\beta_1 = 0.456$, 95% CI $[0.011, 0.901]$, $p = .045$

As in the love model for this context, the coefficient for actual PA similarity increases slightly from the Step 1a estimate ($0.432$, $p = .015$) rather than attenuating. The wider confidence interval and weaker $p$-value again reflect the collinearity from including both predictors ($r = -0.626$). After partialling out perceived PA similarity, the unique contribution of actual PA similarity remains significant and marginally larger than the total effect.

**Perceived PA similarity** $\beta_2 = -0.031$, 95% CI $[-0.379, 0.318]$, $p = .863$

Perceived PA similarity is clearly non-significant and essentially zero. As with love in the same context, the direction is slightly negative rather than the positive sign expected under H1.3. Perceived similarity adds no explanatory value beyond actual similarity.

**Random effects and variance decomposition**

Dyad intercept $SD = 12.76$, variance $= 162.87$; residual $SD = 16.11$, variance $= 259.45$. These are virtually unchanged from Step 1a (dyad variance $= 162.65$, residual $= 257.79$), confirming that perceived similarity absorbs no additional variance.

**The model does not support H1.3 for closeness in the negative interaction.** The pattern is nearly identical to the love result in the same context: actual similarity retains its predictive value, perceived similarity contributes nothing, and there is no evidence of attenuation. Across both outcomes in the negative interaction, the Step 1b models consistently indicate that the actual affective overlap itself, not its perception, drives the association with post-interaction evaluations.

#### Positive interaction

```{r h13_PA_close_pos}
#| output-fold: true
#| output-summary: "H1.3 PA → Closeness (positive interaction)"

h13_PA_close_pos <- lme(
  fixed = close ~ PA_sim_act + PA_sim_perc,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h13_PA_close_pos)
```

**Intercept** $\beta_0 = 86.28$, 95% CI $[83.08, 89.47]$, $p < .001$

Close to the Step 1a intercept ($87.27$), with a minor shift from entering both predictors.

**Actual PA similarity** $\beta_1 = 0.215$, 95% CI $[-0.158, 0.588]$, $p = .256$

The coefficient decreases substantially from the Step 1a estimate ($0.376$, $p = .016$) and is no longer significant. This is the most pronounced attenuation observed across any of the four Step 1b PA models. Unlike the love model in the same context, where the attenuation was modest (from $0.459$ to $0.386$), here the coefficient drops by about $43\%$ and the confidence interval shifts well into non-significance.

**Perceived PA similarity** $\beta_2 = 0.229$, 95% CI $[-0.082, 0.540]$, $p = .147$

Perceived PA similarity is non-significant but shows the largest positive coefficient of any Step 1b PA model. The magnitude ($0.229$) is comparable to the attenuated actual similarity coefficient ($0.215$), suggesting that the Step 1a total effect is split roughly evenly between the two predictors when they are entered together. Neither reaches significance individually, which is consistent with two correlated predictors ($r = -0.585$) sharing a modest total effect.

**Random effects and variance decomposition**

Dyad intercept $SD = 9.35$, variance $= 87.44$; residual $SD = 12.38$, variance $= 153.32$. These are very close to the Step 1a values (dyad variance $= 87.81$, residual $= 154.31$), indicating that splitting the predictor into actual and perceived components does not explain additional variance overall.

**This is the only Step 1b PA model that shows a pattern partially consistent with H1.3.** Actual similarity attenuates meaningfully, and perceived similarity carries a positive coefficient of comparable magnitude. However, neither predictor is individually significant, and the evidence remains inconclusive: the pattern is suggestive of shared variance between actual and perceived similarity in predicting closeness after a positive interaction, but the data do not clearly establish perceived similarity as the operative pathway. The contrast with the negative interaction, where actual similarity retained its full effect and perceived similarity was essentially zero, suggests that whatever role perceived similarity plays may be context-dependent, emerging only during positive interactions where couples may be more attuned to (or accurate about) each other's affective states.

### Summary of results

+------------------------------------------------------------------------+---------------------------------------------------------------+---------------------------------------------------------------+
| **Love**                                                               | Negative Interaction                                          | Positive Interaction                                          |
|                                                                        |                                                               |                                                               |
| *('hoe liefdevol voel je je op dit momenten opzichte van je partner')* | $\beta_1^{act} = 0.444$, $p = .048$ (1a: $0.379$, $p = .020$) | $\beta_1^{act} = 0.386$, $p = .063$ (1a: $0.459$, $p = .007$) |
|                                                                        |                                                               |                                                               |
|                                                                        | $\beta_2^{perc} = -0.081$, $p = .670$                         | $\beta_2^{perc} = 0.105$, $p = .538$                          |
|                                                                        |                                                               |                                                               |
|                                                                        | Actual similarity: **increased**                              | Actual similarity: **attenuated (modest)**                    |
+------------------------------------------------------------------------+---------------------------------------------------------------+---------------------------------------------------------------+
| **Closeness**                                                          | $\beta_1^{act} = 0.456$, $p = .045$ (1a: $0.432$, $p = .015$) | $\beta_1^{act} = 0.215$, $p = .256$ (1a: $0.376$, $p = .016$) |
|                                                                        |                                                               |                                                               |
| *('hoe hecht voel je je op dit moment met je partner')*                | $\beta_2^{perc} = -0.031$, $p = .863$                         | $\beta_2^{perc} = 0.229$, $p = .147$                          |
|                                                                        |                                                               |                                                               |
|                                                                        | Actual similarity: **increased**                              | Actual similarity: **attenuated (substantial)**               |
+------------------------------------------------------------------------+---------------------------------------------------------------+---------------------------------------------------------------+

: *Note.* Step 1b PA models: actual and perceived PA similarity entered simultaneously predicting post-interaction evaluations (H1.3). Step 1a estimates shown in parentheses for comparison. All models include a dyad-level random intercept.

The results split clearly by interaction context. In the **negative interaction**, adding perceived PA similarity has no effect on the actual similarity coefficients, which if anything increase slightly (love: $0.379 \rightarrow 0.444$; closeness: $0.432 \rightarrow 0.456$). Perceived similarity is essentially zero and carries a slightly negative sign in both models. There is no evidence of mediation; the actual affective overlap itself is what predicts post-interaction evaluations.

In the **positive interaction**, the pattern is different. Actual similarity attenuates in both models, modestly for love ($0.459 \rightarrow 0.386$) and substantially for closeness ($0.376 \rightarrow 0.215$). Perceived similarity carries a positive sign in both cases, and for closeness the two coefficients are nearly equal in magnitude ($0.215$ vs. $0.229$), suggesting the Step 1a total effect is split roughly evenly between actual and perceived components. However, neither predictor reaches significance in any of the positive interaction models, so the evidence remains inconclusive.

**H1.3 is not supported in the negative interaction for either outcome.** In the positive interaction, the closeness model shows the pattern most consistent with H1.3 (meaningful attenuation, positive perceived similarity coefficient of comparable size), but the evidence does not reach conventional significance levels. Overall, perceived PA similarity does not emerge as a reliable predictor of post-interaction evaluations in any of the four models, and the attenuation patterns are context-dependent and outcome-dependent rather than consistent.

## H1.3 - NA Similarity

### Love

#### Negative interaction

```{r h13_NA_love_neg}
#| output-fold: true
#| output-summary: "H1.3 NA → Love (negative interaction)"

h13_NA_love_neg <- lme(
  fixed = love ~ NA_sim_act + NA_sim_perc,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h13_NA_love_neg)
```

**Intercept** $\beta_0 = 77.31$, 95% CI $[72.25, 82.37]$, $p < .001$

Virtually identical to the Step 1a intercept ($77.53$).

**Actual NA similarity** $\beta_1 = -0.217$, 95% CI $[-0.726, 0.292]$, $p = .400$

The coefficient remains non-significant and small, consistent with the Step 1a estimate ($-0.141$, $p = .504$). The slight increase in magnitude is not meaningful given the wide confidence interval. As in Step 1a, there is no evidence that actual NA similarity predicts post-interaction love after the negative interaction.

**Perceived NA similarity** $\beta_2 = 0.108$, 95% CI $[-0.302, 0.518]$, $p = .603$

Perceived NA similarity is clearly non-significant and small. The positive direction is opposite to what H1.3 would expect for NA (where the hypothesized association is negative), but the estimate is too imprecise to warrant interpretation. With actual similarity already null in Step 1a, there is no meaningful effect for perceived similarity to mediate or absorb.

**Random effects and variance decomposition**

Dyad intercept $SD = 9.77$, variance $= 95.50$; residual $SD = 18.31$, variance $= 335.12$. These are close to the Step 1a values (dyad variance $= 91.25$, residual $= 336.86$), confirming that neither predictor explains meaningful variance in this model.

**The model does not support H1.3 for NA similarity predicting love in the negative interaction.** Both predictors are non-significant, the total effect was already null in Step 1a, and the question of attenuation is moot when there is no reliable association to attenuate. This is consistent with the broader pattern from H1.2: actual NA similarity during the disagreement interaction is uninformative for post-interaction love, and adding perceived similarity does not change this picture.

#### Positive interaction

```{r h13_NA_love_pos}
#| output-fold: true
#| output-summary: "H1.3 NA → Love (positive interaction)"

h13_NA_love_pos <- lme(
  fixed = love ~ NA_sim_act + NA_sim_perc,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h13_NA_love_pos)
```

**Intercept** $\beta_0 = 91.61$, 95% CI $[85.23, 98.00]$, $p < .001$

Virtually identical to the Step 1a intercept ($91.58$).

**Actual NA similarity** $\beta_1 = 0.332$, 95% CI $[-0.017, 0.681]$, $p = .062$

The coefficient attenuates slightly from the Step 1a estimate ($0.352$, $p = .009$) and drops below conventional significance. The attenuation is modest in magnitude (from $0.352$ to $0.332$), and as in the PA love positive-interaction model, the loss of significance is largely driven by increased standard error from collinearity between the two predictors ($r = -0.656$) rather than a substantive redistribution of variance to perceived similarity.

**Perceived NA similarity** $\beta_2 = 0.025$, 95% CI $[-0.256, 0.305]$, $p = .862$

Perceived NA similarity is clearly non-significant and essentially zero. Despite the significant total effect of actual similarity in Step 1a, perceived similarity does not absorb any meaningful portion of the association when entered alongside actual similarity.

**Random effects and variance decomposition**

Dyad intercept $SD = 10.49$, variance $= 109.97$; residual $SD = 13.34$, variance $= 177.90$. These are virtually unchanged from the Step 1a values (dyad variance $= 110.54$, residual $= 176.33$), confirming that adding perceived similarity does not explain additional variance.

**The model does not support H1.3 for NA similarity predicting love in the positive interaction.** The Step 1a model showed a significant positive association between actual NA similarity and love. Adding perceived similarity produces a slight attenuation in the expected direction, but the perceived similarity coefficient is essentially zero ($0.025$), indicating that there is no meaningful mediating role. The significant total effect from Step 1a appears to be driven by the actual affective overlap itself rather than by the actor's perception of shared negative affect. This mirrors the pattern observed across the PA models: perceived similarity consistently fails to emerge as a meaningful predictor of post-interaction love when entered alongside actual similarity.

### Closeness

#### Negative interaction

```{r h13_NA_close_neg}
#| output-fold: true
#| output-summary: "H1.3 NA → Closeness (negative interaction)"

h13_NA_close_neg <- lme(
  fixed = close ~ NA_sim_act + NA_sim_perc,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h13_NA_close_neg)
```

**Intercept** $\beta_0 = 78.92$, 95% CI $[73.44, 84.40]$, $p < .001$

Virtually identical to the Step 1a intercept ($78.97$).

**Actual NA similarity** $\beta_1 = -0.175$, 95% CI $[-0.705, 0.355]$, $p = .513$

The coefficient remains non-significant and small, consistent with the Step 1a estimate ($-0.157$, $p = .497$). As with love in the same context, there is no evidence that actual NA similarity predicts post-interaction closeness after the negative interaction.

**Perceived NA similarity** $\beta_2 = 0.026$, 95% CI $[-0.357, 0.410]$, $p = .892$

Perceived NA similarity is clearly non-significant and essentially zero. The pattern is identical to the love model in the same context ($\beta_2 = 0.108$, $p = .603$), though the coefficient is even smaller here. With actual similarity already null in Step 1a, there is nothing for perceived similarity to mediate.

**Random effects and variance decomposition**

Dyad intercept $SD = 13.44$, variance $= 180.56$; residual $SD = 16.10$, variance $= 259.17$. These are virtually unchanged from Step 1a (dyad variance $= 179.76$, residual $= 257.79$), confirming that neither predictor explains meaningful variance.

**The model does not support H1.3 for NA similarity predicting closeness in the negative interaction.** Both predictors are non-significant, the total effect was already null in Step 1a, and the question of attenuation is again moot. This completes a consistent pattern across both outcomes in the negative interaction: actual NA similarity is uninformative for post-interaction evaluations, and adding perceived similarity does not change this picture.

#### Positive interaction

```{r h13_NA_close_pos}
#| output-fold: true
#| output-summary: "H1.3 NA → Closeness (positive interaction)"

h13_NA_close_pos <- lme(
  fixed = close ~ NA_sim_act + NA_sim_perc,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h13_NA_close_pos)
```

**Intercept** $\beta_0 = 93.68$, 95% CI $[87.96, 99.41]$, $p < .001$

Virtually identical to the Step 1a intercept ($93.91$).

**Actual NA similarity** $\beta_1 = 0.528$, 95% CI $[0.212, 0.843]$, $p = .001$

The coefficient *increases* from the Step 1a estimate ($0.383$, $p = .002$) rather than attenuating. This is the opposite of what H1.3 predicts and mirrors the pattern observed in the PA models for the negative interaction, where actual similarity also increased slightly when perceived similarity was partialled out. The increase here is more pronounced (from $0.383$ to $0.528$, a $38\%$ increase), suggesting that perceived NA similarity acts as a suppressor in this model: controlling for perceived similarity reveals a stronger unique contribution of actual similarity.

**Perceived NA similarity** $\beta_2 = -0.177$, 95% CI $[-0.434, 0.080]$, $p = .174$

Perceived NA similarity is non-significant but carries a negative coefficient, opposite to the positive direction of the actual similarity effect. This negative sign is what produces the suppression pattern: perceived similarity shares variance with actual similarity (the two are correlated), but its unique contribution, once actual similarity is controlled, trends negative. This means that to the extent that perceived NA similarity has any unique association with closeness beyond actual similarity, it is in the direction of *lower* closeness, not higher. The effect does not reach significance, but the sign is interpretively noteworthy because it diverges from the total effect pattern.

**Random effects and variance decomposition**

Dyad intercept $SD = 9.18$, variance $= 84.25$; residual $SD = 12.30$, variance $= 151.27$. The residual variance decreases slightly from Step 1a ($154.31$), and the dyad-level variance also decreases slightly ($80.91$ to $84.25$, though this small increase is within estimation noise). The model explains marginally more within-dyad variance than Step 1a, consistent with the stronger actual similarity coefficient.

**The model does not support H1.3 for NA similarity predicting closeness in the positive interaction.** Rather than attenuating, actual similarity strengthens when perceived similarity is added, indicating a suppression pattern rather than mediation. This is the clearest suppression effect observed across the Step 1b models and stands in contrast to the love outcome in the same context, where actual similarity attenuated modestly and perceived similarity was essentially zero. For closeness, the actual and perceived components of NA similarity appear to pull in opposite directions once their shared variance is partialled out: actual overlap predicts more closeness, while perceived overlap (uniquely) trends toward less closeness. This divergence suggests that the two similarity channels may carry different relational information during positive interactions, at least for closeness as the outcome.

### Summary of results

+------------------------------------------------------------------------+-----------------------------------------------------------------+---------------------------------------------------------------+
|                                                                        | Negative Interaction                                            | Positive Interaction                                          |
+:=======================================================================+:================================================================+:==============================================================+
| **Love**                                                               | $\beta_1^{act} = -0.217$, $p = .400$ (1a: $-0.141$, $p = .504$) | $\beta_1^{act} = 0.332$, $p = .062$ (1a: $0.352$, $p = .009$) |
|                                                                        |                                                                 |                                                               |
| *('hoe liefdevol voel je je op dit momenten opzichte van je partner')* | $\beta_2^{perc} = 0.108$, $p = .603$                            | $\beta_2^{perc} = 0.025$, $p = .862$                          |
|                                                                        |                                                                 |                                                               |
|                                                                        | Step 1a null; attenuation **moot**                              | Actual similarity: **attenuated (modest)**                    |
+------------------------------------------------------------------------+-----------------------------------------------------------------+---------------------------------------------------------------+
| **Closeness**                                                          | $\beta_1^{act} = -0.175$, $p = .513$ (1a: $-0.157$, $p = .497$  | $\beta_1^{act} = 0.528$, $p = .001$ (1a: $0.383$, $p = .002$) |
|                                                                        |                                                                 |                                                               |
| *('hoe hecht voel je je op dit moment met je partner')*                | $\beta_2^{perc} = 0.026$, $p = .892$                            | $\beta_2^{perc} = -0.177$, $p = .174$                         |
|                                                                        |                                                                 |                                                               |
|                                                                        | Step 1a null; attenuation **moot**                              | Actual similarity: **increased (suppression)**                |
+------------------------------------------------------------------------+-----------------------------------------------------------------+---------------------------------------------------------------+

: *Note.* Step 1b NA models: actual and perceived NA similarity entered simultaneously predicting post-interaction evaluations (H1.3). Step 1a estimates shown in parentheses for comparison. All models include a dyad-level random intercept.

**H1.3 is not supported for NA similarity in any of the four models.** The results fall into two distinct patterns.

In the **negative interaction**, both Step 1a effects were null, so the question of attenuation does not arise. Adding perceived similarity changes nothing: both predictors remain non-significant and small in both models. Actual NA similarity during the disagreement interaction is uninformative for post-interaction evaluations, and perceived similarity does not alter this.

In the **positive interaction**, the two outcomes diverge. For **love**, actual similarity attenuates modestly ($0.352 \rightarrow 0.332$) and drops just below significance, while perceived similarity is essentially zero ($0.025$). The pattern is similar to what was observed for PA love in the positive interaction: a slight weakening of the actual similarity coefficient that is better attributed to collinearity than to mediation.

For **closeness**, the pattern is qualitatively different. Actual similarity *increases* from $0.383$ to $0.528$ when perceived similarity is added, indicating a suppression effect. Perceived similarity carries a negative coefficient ($-0.177$, $p = .174$), opposite in sign to the actual similarity effect. This means the two similarity channels pull in opposing directions once their shared variance is partialled out: actual NA overlap predicts more closeness, while perceived NA overlap (uniquely) trends toward less closeness. This is the most pronounced suppression effect observed across all Step 1b models.

The divergence between the love and closeness outcomes in the positive interaction is noteworthy: for love, adding perceived similarity produces a conventional (if small) attenuation; for closeness, it reveals a suppression pattern where the actual similarity effect strengthens. This suggests that actual and perceived NA similarity may carry different relational information for closeness specifically, a pattern that will be relevant when interpreting the mediation models in Step 2.

# Step 1c (Exploratory): Actual × Perceived Interaction

To examine whether the association of one form of similarity with post-interaction evaluations depends on the level of the other, we include interaction terms. Grand-mean centered (within each interaction context) similarity indices are used so that main effects remain interpretable at the average level of the moderator.

$$
O_{pdc} = b_{0,d} + \beta_1 \tilde{S}^{act}_{pdc} + \beta_2 \tilde{S}^{perc}_{pdc} + \beta_3 (\tilde{S}^{act}_{pdc} \times \tilde{S}^{perc}_{pdc}) + \varepsilon_{pdc}
$$

with the same dyad-level random intercept as before. $\tilde{S}$ denotes the grand-mean centered (within each context) interaction-averaged similarity indices.

If the interaction term $\beta_3$ is significant, we probe simple slopes at low ($-1$ SD), average (0), and high ($+1$ SD) levels of the moderator.

## PA Similarity

### Love

#### Negative interaction

```{r im_PA_love_neg}
#| output-fold: true
#| output-summary: "Interaction PA → Love (negative)"

im_PA_love_neg <- lme(
  fixed = love ~ PA_sim_act_gmc + PA_sim_perc_gmc + PA_sim_act_gmc:PA_sim_perc_gmc,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(im_PA_love_neg)
```

**Intercept** $\beta_0 = 78.57$, 95% CI $[74.84, 82.29]$, $p < .001$

This is the expected post-interaction love at the grand mean of both actual and perceived PA similarity (since both predictors are grand-mean centered within context). The value is lower than the Step 1a and 1b intercepts ($\approx 89$) because the centering approach differs: in Steps 1a/1b the predictors are uncentered (zero = no net similarity), whereas here the intercept reflects the sample average level of both similarity indices.

**Actual PA similarity** $\beta_1 = 0.437$, 95% CI $[-0.006, 0.881]$, $p = .053$

The main effect of actual PA similarity at the average level of perceived similarity. The magnitude is comparable to the Step 1b estimate ($0.444$) and falls just short of conventional significance, again reflecting the collinearity between the two predictors.

**Perceived PA similarity** $\beta_2 = -0.067$, 95% CI $[-0.463, 0.329]$, $p = .738$

The main effect of perceived PA similarity at the average level of actual similarity. As in Step 1b, the coefficient is small, non-significant, and slightly negative.

**Interaction** $\beta_3 = 0.003$, 95% CI $[-0.022, 0.028]$, $p = .808$

The interaction between actual and perceived PA similarity is essentially zero and clearly non-significant. The association between actual PA similarity and post-interaction love does not depend on the level of perceived PA similarity (and vice versa). Given the null interaction, probing simple slopes is not warranted for this model.

**Random effects and variance decomposition**

Dyad intercept $SD = 8.88$, variance $= 78.84$; residual $SD = 18.44$, variance $= 339.96$. These are virtually unchanged from Steps 1a and 1b, confirming that the interaction term absorbs no additional variance.

The model confirms that perceived PA similarity neither predicts love after the negative interaction nor modulates the actual similarity effect. The actual similarity coefficient is stable across Steps 1a ($0.379$), 1b ($0.444$), and 1c ($0.437$) regardless of whether perceived similarity is entered as a main effect or allowed to interact, indicating that perceived similarity is simply irrelevant to this association.

#### Positive interaction

```{r im_PA_love_pos}
#| output-fold: true
#| output-summary: "Interaction PA → Love (positive)"

im_PA_love_pos <- lme(
  fixed = love ~ PA_sim_act_gmc + PA_sim_perc_gmc + PA_sim_act_gmc:PA_sim_perc_gmc,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(im_PA_love_pos)
```

**Intercept** $\beta_0 = 83.77$, 95% CI $[80.65, 86.89]$, $p < .001$

Expected post-interaction love at the grand mean of both actual and perceived PA similarity in the positive interaction context. As in the negative interaction model, the intercept differs from Steps 1a/1b due to the different centering approach.

**Actual PA similarity** $\beta_1 = 0.393$, 95% CI $[-0.019, 0.804]$, $p = .061$

The main effect of actual PA similarity at the average level of perceived similarity. The magnitude is very close to the Step 1b estimate ($0.386$) and again falls just below conventional significance, consistent with the collinearity pattern observed throughout.

**Perceived PA similarity** $\beta_2 = 0.099$, 95% CI $[-0.240, 0.438]$, $p = .564$

The main effect of perceived PA similarity at the average level of actual similarity. Small, non-significant, and positive, mirroring the Step 1b estimate ($0.105$) closely.

**Interaction** $\beta_3 = 0.003$, 95% CI $[-0.023, 0.029]$, $p = .798$

The interaction is essentially zero and clearly non-significant, virtually identical to the negative interaction model ($\beta_3 = 0.003$, $p = .808$). The association between actual PA similarity and love does not depend on perceived PA similarity in either context. Probing simple slopes is not warranted.

**Random effects and variance decomposition**

Dyad intercept $SD = 10.50$, variance $= 110.16$; residual $SD = 13.34$, variance $= 177.99$. Unchanged from Steps 1a and 1b, confirming that the interaction term absorbs no additional variance.

**Neither actual nor perceived PA similarity reaches significance in this model, and the interaction between them is essentially zero (**$\beta_3 = 0.003$, $p = .798$). Combined with the identical null interaction in the negative context ($\beta_3 = 0.003$, $p = .808$), the Step 1c results provide clear evidence that the two forms of similarity do not modulate each other's association with post-interaction love in either context. The non-significance of both main effects in the positive interaction reflects the collinearity-driven splitting of a modest total effect (Step 1a: $\beta = 0.459$, $p = .007$) across two correlated predictors, rather than a genuine absence of association.

### Probing (negative interaction, love)

```{r im_PA_love_neg_probe}
#| output-fold: true
#| output-summary: "Simple slopes — PA love (negative)"

# Probe points (−1 SD, 0, +1 SD)
act_sd_neg  <- sd(sc2_neg$PA_sim_act_gmc,  na.rm = TRUE)
perc_sd_neg <- sd(sc2_neg$PA_sim_perc_gmc, na.rm = TRUE)

# Simple slopes of actual PA similarity at levels of perceived PA similarity
slopes_act <- emtrends(
  im_PA_love_neg,
  specs = ~ PA_sim_perc_gmc,
  var   = "PA_sim_act_gmc",
  at    = list(PA_sim_perc_gmc = c(-perc_sd_neg, 0, perc_sd_neg))
)

summary(slopes_act, infer = TRUE)

# Simple slopes of perceived PA similarity at levels of actual PA similarity
slopes_perc <- emtrends(
  im_PA_love_neg,
  specs = ~ PA_sim_act_gmc,
  var   = "PA_sim_perc_gmc",
  at    = list(PA_sim_act_gmc = c(-act_sd_neg, 0, act_sd_neg))
)

summary(slopes_perc, infer = TRUE)
```

**Simple slopes (confirming the null interaction)**

The simple slopes of actual PA similarity on love are virtually constant across levels of perceived PA similarity: $0.404$ at $-1$ SD, $0.437$ at the mean, $0.471$ at $+1$ SD. None reaches conventional significance (all $p > .05$), which reflects the collinearity-driven widening of standard errors discussed above rather than a meaningful change in the slope. The near-identical magnitudes confirm that perceived PA similarity does not modulate the actual similarity effect.

Conversely, the simple slopes of perceived PA similarity on love are statistically zero at all levels of actual similarity ($-0.098$, $-0.067$, $-0.036$; all $p > .63$). Perceived PA similarity is uninformative for post-interaction love regardless of how much actual PA overlap couples share.

Together, the probing results corroborate the null interaction term: the actual similarity slope is constant across levels of perceived similarity, and perceived similarity is statistically zero regardless of actual similarity level. There is no synergy or compensation between the two forms of similarity. The remaining Step 1c models (positive interaction love, and both closeness models) show the same null interaction pattern, so probing results are not reported for those.

### Closeness

#### Negative interaction

```{r im_PA_close_neg}
#| output-fold: true
#| output-summary: "Interaction PA → Closeness (negative)"

im_PA_close_neg <- lme(
  fixed = close ~ PA_sim_act_gmc + PA_sim_perc_gmc + PA_sim_act_gmc:PA_sim_perc_gmc,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(im_PA_close_neg)
```

**Intercept** $\beta_0 = 80.67$, 95% CI $[76.70, 84.64]$, $p < .001$

Expected post-interaction closeness at the grand mean of both actual and perceived PA similarity in the negative interaction context.

**Actual PA similarity** $\beta_1 = 0.464$, 95% CI $[0.015, 0.914]$, $p = .043$

The main effect of actual PA similarity at the average level of perceived similarity. The magnitude is consistent with the Step 1b estimate ($0.456$) and remains just significant.

**Perceived PA similarity** $\beta_2 = -0.048$, 95% CI $[-0.416, 0.321]$, $p = .798$

Statistically zero, as in all preceding models for the negative interaction. Perceived PA similarity has no unique association with closeness at the average level of actual similarity.

**Interaction** $\beta_3 = -0.004$, 95% CI $[-0.029, 0.022]$, $p = .774$

Statistically zero. The association between actual PA similarity and closeness does not depend on the level of perceived PA similarity. This mirrors the love model in the same context ($\beta_3 = 0.003$, $p = .808$), with both interaction terms essentially zero and opposite in sign, confirming that there is no systematic synergy between the two similarity channels for either outcome after the negative interaction. Probing simple slopes is not warranted.

**Random effects and variance decomposition**

Dyad intercept $SD = 12.83$, variance $= 164.69$; residual $SD = 16.13$, variance $= 260.06$. Unchanged from Steps 1a and 1b.

**As with love, perceived PA similarity neither predicts closeness after the negative interaction nor modulates the actual similarity effect. Actual similarity retains its predictive value regardless of model specification.**

#### Positive interaction

```{r im_PA_close_pos}
#| output-fold: true
#| output-summary: "Interaction PA → Closeness (positive)"

im_PA_close_pos <- lme(
  fixed = close ~ PA_sim_act_gmc + PA_sim_perc_gmc + PA_sim_act_gmc:PA_sim_perc_gmc,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(im_PA_close_pos)
```

**Intercept** $\beta_0 = 85.45$, 95% CI $[82.60, 88.30]$, $p < .001$

Expected post-interaction closeness at the grand mean of both actual and perceived PA similarity in the positive interaction context.

**Actual PA similarity** $\beta_1 = 0.221$, 95% CI $[-0.157, 0.598]$, $p = .249$

Statistically zero. The magnitude is consistent with the Step 1b estimate ($0.215$), confirming that in the positive interaction the unique contribution of actual PA similarity to closeness is not reliably different from zero once perceived similarity is in the model.

**Perceived PA similarity** $\beta_2 = 0.224$, 95% CI $[-0.088, 0.537]$, $p = .157$

Statistically zero, though this remains the largest perceived similarity coefficient across all Step 1b and 1c PA models. As noted in Step 1b, the near-equal magnitudes of the two coefficients ($0.221$ and $0.224$) suggest that the Step 1a total effect is split roughly evenly between the two predictors, but neither has sufficient unique variance to reach significance given the collinearity and the small sample.

**Interaction** $\beta_3 = 0.003$, 95% CI $[-0.021, 0.027]$, $p = .796$

Statistically zero. As in the other three PA models, the association between actual PA similarity and the outcome does not depend on perceived PA similarity. The interaction terms across all four Step 1c PA models are remarkably consistent (ranging from $-0.004$ to $+0.003$, all $p > .77$), providing clear evidence that the two forms of similarity do not modulate each other. Probing simple slopes is not warranted.

**Random effects and variance decomposition**

Dyad intercept $SD = 9.48$, variance $= 89.88$; residual $SD = 12.36$, variance $= 152.81$. Unchanged from Steps 1a and 1b.

### Summary of results

**Across all four models (two outcomes** $\times$ two contexts), the interaction between actual and perceived PA similarity is consistently zero. **There is no evidence that the two forms of similarity synergize or compensate in predicting post-interaction evaluations.**

+-------------------------------------------------------------------------+--------------------------------------------------+-------------------------------------------------+
|                                                                         | Negative Interaction                             | Positive Interaction                            |
+:========================================================================+:=================================================+:================================================+
| **Love**                                                                | $\beta_1^{act} = 0.437$, $p = .053$              | $\beta_1^{act} = 0.393$, $p = .061$             |
|                                                                         |                                                  |                                                 |
| *('hoe liefdevol voel je je op dit momentten opzichte van je partner')* | $\beta_2^{perc} = -0.067$, $p = .738$            | $\beta_2^{perc} = 0.099$, $p = .564$            |
|                                                                         |                                                  |                                                 |
|                                                                         | $\beta_3^{act \times perc} = 0.003$, $p = .808$  | $\beta_3^{act \times perc} = 0.003$, $p = .798$ |
+-------------------------------------------------------------------------+--------------------------------------------------+-------------------------------------------------+
| **Closeness**                                                           | $\beta_1^{act} = 0.464$, $p = .043$              | $\beta_1^{act} = 0.221$, $p = .249$             |
|                                                                         |                                                  |                                                 |
| *('hoe hecht voel je je op dit moment met je partner')*                 | $\beta_2^{perc} = -0.048$, $p = .798$            | $\beta_2^{perc} = 0.224$, $p = .157$            |
|                                                                         |                                                  |                                                 |
|                                                                         | $\beta_3^{act \times perc} = -0.004$, $p = .774$ | $\beta_3^{act \times perc} = 0.003$, $p = .796$ |
+-------------------------------------------------------------------------+--------------------------------------------------+-------------------------------------------------+

: *Note.* Step 1c PA models: actual and perceived PA similarity (grand-mean centered within context) with interaction term predicting post-interaction evaluations. All models include a dyad-level random intercept.

The interaction terms are statistically zero in all four models (ranging from $-0.004$ to $+0.003$, all $p > .77$), providing clear evidence that actual and perceived PA similarity do not modulate each other's association with post-interaction evaluations.

The main effect patterns replicate Steps 1a and 1b and split by context. In the **negative interaction**, actual PA similarity retains its predictive value for closeness ($0.464$, $p = .043$) and falls just short for love ($0.437$, $p = .053$). These coefficients are stable or slightly increased relative to Step 1a (love: $0.379$; closeness: $0.432$), consistent with the pattern identified in Step 1b where actual similarity was unaffected by the inclusion of perceived similarity. Perceived similarity is statistically zero and slightly negative for both outcomes, confirming it plays no role in the negative interaction.

In the **positive interaction**, neither predictor reaches significance for love ($0.393$, $p = .061$ and $0.099$, $p = .564$), reflecting the collinearity-driven splitting of a modest total effect (Step 1a: $0.459$, $p = .007$) across two correlated predictors. For closeness, the same splitting pattern is more pronounced: both coefficients are of comparable magnitude ($0.221$ and $0.224$) but neither is significant, as the Step 1a total effect ($0.376$, $p = .016$) is divided roughly evenly between the two predictors.

## NA Similarity

### Love

#### Negative interaction

```{r im_NA_love_neg}
#| output-fold: true
#| output-summary: "Interaction NA → Love (negative)"

im_NA_love_neg <- lme(
  fixed = love ~ NA_sim_act_gmc + NA_sim_perc_gmc + NA_sim_act_gmc:NA_sim_perc_gmc,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(im_NA_love_neg)
```

**Intercept** $\beta_0 = 79.12$, 95% CI $[75.56, 82.68]$, $p < .001$

Expected post-interaction love at the grand mean of both actual and perceived NA similarity in the negative interaction context.

**Actual NA similarity** $\beta_1 = -0.200$, 95% CI $[-0.712, 0.312]$, $p = .441$

Statistically zero, consistent with the Step 1a ($-0.141$, $p = .504$) and Step 1b ($-0.217$, $p = .400$) estimates. Actual NA similarity remains uninformative for post-interaction love after the negative interaction across all model specifications.

**Perceived NA similarity** $\beta_2 = 0.096$, 95% CI $[-0.317, 0.509]$, $p = .645$

Statistically zero, consistent with the Step 1b estimate ($0.108$, $p = .603$).

**Interaction** $\beta_3 = -0.008$, 95% CI $[-0.043, 0.028]$, $p = .673$

Statistically zero. The association between actual NA similarity and love does not depend on perceived NA similarity. With both main effects already null, a null interaction is unsurprising: there is simply no signal in either form of NA similarity for the other to modulate in this context.

**Random effects and variance decomposition**

Dyad intercept $SD = 9.48$, variance $= 89.93$; residual $SD = 18.47$, variance $= 341.23$. Unchanged from Steps 1a and 1b.

**The model confirms the consistent null pattern for NA similarity in the negative interaction: neither actual nor perceived similarity predicts love, and the two do not interact. NA similarity during the disagreement interaction is uninformative for post-interaction love regardless of model specification.**

#### Positive interaction

```{r im_NA_love_pos}
#| output-fold: true
#| output-summary: "Interaction NA → Love (positive)"

im_NA_love_pos <- lme(
  fixed = love ~ NA_sim_act_gmc + NA_sim_perc_gmc + NA_sim_act_gmc:NA_sim_perc_gmc,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(im_NA_love_pos)
```

**Intercept** $\beta_0 = 83.86$, 95% CI $[80.46, 87.26]$, $p < .001$

Expected post-interaction love at the grand mean of both actual and perceived NA similarity in the positive interaction context.

**Actual NA similarity** $\beta_1 = 0.333$, 95% CI $[-0.018, 0.683]$, $p = .063$

Consistent with the Step 1b estimate ($0.332$, $p = .062$). The coefficient is just below conventional significance, as in all positive interaction models where actual NA similarity is entered alongside perceived similarity. The point estimate remains stable across Steps 1a ($0.352$), 1b ($0.332$), and 1c ($0.333$), with the marginal $p$-value reflecting collinearity-driven standard error inflation rather than a substantive change.

**Perceived NA similarity** $\beta_2 = 0.026$, 95% CI $[-0.256, 0.308]$, $p = .856$

Statistically zero, consistent with the Step 1b estimate ($0.025$, $p = .862$).

**Interaction** $\beta_3 = 0.001$, 95% CI $[-0.019, 0.021]$, $p = .928$

Statistically zero. This is the smallest interaction term observed across all eight Step 1c models. The association between actual NA similarity and love does not depend on perceived NA similarity. Probing simple slopes is not warranted.

**Random effects and variance decomposition**

Dyad intercept $SD = 10.51$, variance $= 110.36$; residual $SD = 13.38$, variance $= 178.93$. Unchanged from Steps 1a and 1b.

**The model confirms that perceived NA similarity neither predicts love after the positive interaction nor modulates the actual similarity effect. The actual similarity coefficient remains stable at** $\approx 0.33$ **across Steps 1b and 1c regardless of whether perceived similarity is entered as a main effect or allowed to interact, indicating that perceived similarity is simply irrelevant to this association.**

### Closeness

#### Negative interaction

```{r im_NA_close_neg}
#| output-fold: true
#| output-summary: "Interaction NA → Closeness (negative)"

im_NA_close_neg <- lme(
  fixed = close ~ NA_sim_act_gmc + NA_sim_perc_gmc + NA_sim_act_gmc:NA_sim_perc_gmc,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(im_NA_close_neg)
```

**Intercept** $\beta_0 = 80.46$, 95% CI $[76.62, 84.29]$, $p < .001$

Expected post-interaction closeness at the grand mean of both actual and perceived NA similarity in the negative interaction context.

**Actual NA similarity** $\beta_1 = -0.171$, 95% CI $[-0.706, 0.364]$, $p = .527$

Statistically zero, consistent with the Step 1a ($-0.157$, $p = .497$) and Step 1b ($-0.175$, $p = .513$) estimates. The coefficient is stable across all three model specifications.

**Perceived NA similarity** $\beta_2 = 0.024$, 95% CI $[-0.363, 0.410]$, $p = .903$

Statistically zero, consistent with the Step 1b estimate ($0.026$, $p = .892$).

**Interaction** $\beta_3 = -0.002$, 95% CI $[-0.037, 0.034]$, $p = .921$

Statistically zero. As with the love model in the same context ($\beta_3 = -0.008$, $p = .673$), both main effects are null and the interaction is null. There is no signal in either form of NA similarity for the other to modulate. Probing simple slopes is not warranted.

**Random effects and variance decomposition**

Dyad intercept $SD = 13.43$, variance $= 180.33$; residual $SD = 16.16$, variance $= 261.19$. Unchanged from Steps 1a and 1b.

**The model completes the negative interaction NA series across Steps 1a, 1b, and 1c. Across all model specifications and both outcomes, NA similarity during the disagreement interaction is consistently uninformative for post-interaction evaluations. Neither actual nor perceived similarity predicts love or closeness, and the two do not interact.**

#### Positive interaction

```{r im_NA_close_pos}
#| output-fold: true
#| output-summary: "Interaction NA → Closeness (positive)"

im_NA_close_pos <- lme(
  fixed = close ~ NA_sim_act_gmc + NA_sim_perc_gmc + NA_sim_act_gmc:NA_sim_perc_gmc,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(im_NA_close_pos)
```

**Intercept** $\beta_0 = 85.09$, 95% CI $[82.04, 88.14]$, $p < .001$

Expected post-interaction closeness at the grand mean of both actual and perceived NA similarity in the positive interaction context.

**Actual NA similarity** $\beta_1 = 0.531$, 95% CI $[0.215, 0.847]$, $p = .001$

The coefficient remains significant and stable across Steps 1a ($0.383$), 1b ($0.528$), and 1c ($0.531$). The increase from Step 1a to 1b/1c reflects the suppression effect noted in Step 1b: controlling for perceived similarity reveals a stronger unique contribution of actual similarity.

**Perceived NA similarity** $\beta_2 = -0.169$, 95% CI $[-0.428, 0.090]$, $p = .198$

Statistically zero, though the negative direction is consistent with the Step 1b estimate ($-0.177$, $p = .174$). The suppression pattern persists: perceived NA similarity trends in the opposite direction to actual similarity, but does not reach significance.

**Interaction** $\beta_3 = 0.005$, 95% CI $[-0.013, 0.024]$, $p = .548$

Statistically zero. The suppression pattern observed in the main effects does not depend on the level of the other predictor. Probing simple slopes is not warranted.

**Random effects and variance decomposition**

Dyad intercept $SD = 9.09$, variance $= 82.69$; residual $SD = 12.37$, variance $= 152.95$. Unchanged from Steps 1a and 1b.

**The model confirms that the suppression pattern for NA similarity predicting closeness in the positive interaction is robust to the inclusion of an interaction term. Actual similarity retains its strengthened coefficient (**$\approx 0.53$), perceived similarity continues to trend negatively but remains non-significant, and the two do not modulate each other. This is the only model across the entire Step 1c series where a main effect is clearly significant, and it is the actual (not perceived) similarity channel that carries the signal.

### Summary of results

+-------------------------------------------------------------------------+---------------------------------------------------+-------------------------------------------------+
|                                                                         | Negative Interaction                              | Positive Interaction                            |
+:========================================================================+:==================================================+:================================================+
| **Love**                                                                | $\beta_1^{act} = -0.200$, $p = .441$              | $\beta_1^{act} = 0.333$, $p = .063$             |
|                                                                         |                                                   |                                                 |
| *('hoe liefdevol voel je je op dit momentten opzichte van je partner')* | $\beta_2^{perc} = 0.096$, $p = .645$              | $\beta_2^{perc} = 0.026$, $p = .856$            |
|                                                                         |                                                   |                                                 |
|                                                                         | $\beta_3^{act \times perc} = -0.008$, $p = .673$  | $\beta_3^{act \times perc} = 0.001$, $p = .928$ |
+-------------------------------------------------------------------------+---------------------------------------------------+-------------------------------------------------+
| **Closeness**                                                           | $\beta_1^{act} = -0.171$, $p = .527$              | $\beta_1^{act} = 0.531$, $p = .001$             |
|                                                                         |                                                   |                                                 |
| *('hoe hecht voel je je op dit moment met je partner')*                 | $\beta_2^{perc} = 0.024$, $p = .903$              | $\beta_2^{perc} = -0.169$, $p = .198$           |
|                                                                         |                                                   |                                                 |
|                                                                         | $\beta_3^{act \times perc} = -0.002$, $p = .921$  | $\beta_3^{act \times perc} = 0.005$, $p = .548$ |
+-------------------------------------------------------------------------+---------------------------------------------------+-------------------------------------------------+

: *Note.* Step 1c NA models: actual and perceived NA similarity (grand-mean centered within context) with interaction term predicting post-interaction evaluations. All models include a dyad-level random intercept.

The interaction terms are statistically zero in all four models (all $p > .54$), confirming that actual and perceived NA similarity do not modulate each other's association with post-interaction evaluations. This mirrors the Step 1c PA results, where all interaction terms were also null (all $p > .77$). Across all eight Step 1c models (two affect types $\times$ two outcomes $\times$ two contexts), there is no evidence that the two forms of similarity interact.

The main effect patterns replicate what was observed in Steps 1a and 1b. In the **negative interaction**, all coefficients are statistically zero for both outcomes: NA similarity is uninformative for post-interaction evaluations regardless of model specification. In the **positive interaction**, the two outcomes diverge. For love, actual NA similarity falls just short of significance ($0.333$, $p = .063$), stable across Steps 1b ($0.332$) and 1c ($0.333$), with the marginal $p$-value reflecting collinearity rather than a substantive change. For closeness, actual NA similarity is clearly significant ($0.531$, $p = .001$), strengthened relative to Step 1a ($0.383$) by the suppression effect identified in Step 1b. This remains the only clearly significant main effect across the entire Step 1c series, and it is the actual (not perceived) similarity channel that carries the signal.

The perceived NA similarity coefficient for closeness in the positive interaction ($-0.169$, $p = .198$) is the largest non-significant perceived similarity effect across the NA models. Its negative direction, opposite to the actual similarity effect, sustains the suppression pattern noted in Step 1b but does not reach significance in any specification.

# Summary of steps 1a, 1b, and 1c

```{r step1_summary_table}
#| tbl-cap: "Summary of Steps 1a, 1b, and 1c: Fixed effects for actual and perceived affect similarity predicting post-interaction love and closeness"

# --- Build rows ---
rows <- list(
  # ---- Closeness / PA ----
  list("<b>Closeness</b>", "<b>PA Similarity</b>", "<b>Actual</b>",
       fmt_cell(h11_PA_close_pos, "PA_sim_act"),
       fmt_cell(h13_PA_close_pos, "PA_sim_act"),
       fmt_cell(im_PA_close_pos,  "PA_sim_act_gmc"),
       fmt_cell(h11_PA_close_neg, "PA_sim_act"),
       fmt_cell(h13_PA_close_neg, "PA_sim_act"),
       fmt_cell(im_PA_close_neg,  "PA_sim_act_gmc")),
  list("<b>Closeness</b>", "<b>PA Similarity</b>", "<b>Perceived</b>",
       "",
       fmt_cell(h13_PA_close_pos, "PA_sim_perc"),
       fmt_cell(im_PA_close_pos,  "PA_sim_perc_gmc"),
       "",
       fmt_cell(h13_PA_close_neg, "PA_sim_perc"),
       fmt_cell(im_PA_close_neg,  "PA_sim_perc_gmc")),
  list("<b>Closeness</b>", "<b>PA Similarity</b>", "<b>Act × Perc</b>",
       "", "",
       fmt_cell(im_PA_close_pos, "PA_sim_act_gmc:PA_sim_perc_gmc"),
       "", "",
       fmt_cell(im_PA_close_neg, "PA_sim_act_gmc:PA_sim_perc_gmc")),
  # ---- Closeness / NA ----
  list("<b>Closeness</b>", "<b>NA Similarity</b>", "<b>Actual</b>",
       fmt_cell(h12_NA_close_pos, "NA_sim_act"),
       fmt_cell(h13_NA_close_pos, "NA_sim_act"),
       fmt_cell(im_NA_close_pos,  "NA_sim_act_gmc"),
       fmt_cell(h12_NA_close_neg, "NA_sim_act"),
       fmt_cell(h13_NA_close_neg, "NA_sim_act"),
       fmt_cell(im_NA_close_neg,  "NA_sim_act_gmc")),
  list("<b>Closeness</b>", "<b>NA Similarity</b>", "<b>Perceived</b>",
       "",
       fmt_cell(h13_NA_close_pos, "NA_sim_perc"),
       fmt_cell(im_NA_close_pos,  "NA_sim_perc_gmc"),
       "",
       fmt_cell(h13_NA_close_neg, "NA_sim_perc"),
       fmt_cell(im_NA_close_neg,  "NA_sim_perc_gmc")),
  list("<b>Closeness</b>", "<b>NA Similarity</b>", "<b>Act × Perc</b>",
       "", "",
       fmt_cell(im_NA_close_pos, "NA_sim_act_gmc:NA_sim_perc_gmc"),
       "", "",
       fmt_cell(im_NA_close_neg, "NA_sim_act_gmc:NA_sim_perc_gmc")),
  # ---- Love / PA ----
  list("<b>Love</b>", "<b>PA Similarity</b>", "<b>Actual</b>",
       fmt_cell(h11_PA_love_pos, "PA_sim_act"),
       fmt_cell(h13_PA_love_pos, "PA_sim_act"),
       fmt_cell(im_PA_love_pos,  "PA_sim_act_gmc"),
       fmt_cell(h11_PA_love_neg, "PA_sim_act"),
       fmt_cell(h13_PA_love_neg, "PA_sim_act"),
       fmt_cell(im_PA_love_neg,  "PA_sim_act_gmc")),
  list("<b>Love</b>", "<b>PA Similarity</b>", "<b>Perceived</b>",
       "",
       fmt_cell(h13_PA_love_pos, "PA_sim_perc"),
       fmt_cell(im_PA_love_pos,  "PA_sim_perc_gmc"),
       "",
       fmt_cell(h13_PA_love_neg, "PA_sim_perc"),
       fmt_cell(im_PA_love_neg,  "PA_sim_perc_gmc")),
  list("<b>Love</b>", "<b>PA Similarity</b>", "<b>Act × Perc</b>",
       "", "",
       fmt_cell(im_PA_love_pos, "PA_sim_act_gmc:PA_sim_perc_gmc"),
       "", "",
       fmt_cell(im_PA_love_neg, "PA_sim_act_gmc:PA_sim_perc_gmc")),
  # ---- Love / NA ----
  list("<b>Love</b>", "<b>NA Similarity</b>", "<b>Actual</b>",
       fmt_cell(h12_NA_love_pos, "NA_sim_act"),
       fmt_cell(h13_NA_love_pos, "NA_sim_act"),
       fmt_cell(im_NA_love_pos,  "NA_sim_act_gmc"),
       fmt_cell(h12_NA_love_neg, "NA_sim_act"),
       fmt_cell(h13_NA_love_neg, "NA_sim_act"),
       fmt_cell(im_NA_love_neg,  "NA_sim_act_gmc")),
  list("<b>Love</b>", "<b>NA Similarity</b>", "<b>Perceived</b>",
       "",
       fmt_cell(h13_NA_love_pos, "NA_sim_perc"),
       fmt_cell(im_NA_love_pos,  "NA_sim_perc_gmc"),
       "",
       fmt_cell(h13_NA_love_neg, "NA_sim_perc"),
       fmt_cell(im_NA_love_neg,  "NA_sim_perc_gmc")),
  list("<b>Love</b>", "<b>NA Similarity</b>", "<b>Act × Perc</b>",
       "", "",
       fmt_cell(im_NA_love_pos, "NA_sim_act_gmc:NA_sim_perc_gmc"),
       "", "",
       fmt_cell(im_NA_love_neg, "NA_sim_act_gmc:NA_sim_perc_gmc"))
)

# Assemble into data frame
tbl_df <- do.call(rbind, lapply(rows, function(r) {
  data.frame(
    Outcome   = r[[1]],
    Affect    = r[[2]],
    Predictor = r[[3]],
    `1a`      = r[[4]],
    `1b`      = r[[5]],
    `1c`      = r[[6]],
    `1a `     = r[[7]],
    `1b `     = r[[8]],
    `1c `     = r[[9]],
    check.names = FALSE,
    stringsAsFactors = FALSE
  )
}))

kable(tbl_df,
  format    = "html",
  col.names = c("Outcome", "Affect Similarity", "Predictor",
                "1a", "1b", "1c", "1a", "1b", "1c"),
  align     = c("l", "l", "l", rep("c", 6)),
  booktabs  = TRUE,
  escape    = FALSE
) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE,
    position   = "left",
    font_size  = 12
  ) %>%
  add_header_above(c(" " = 3,
                      "Positive Interaction" = 3,
                      "Negative Interaction" = 3)) %>%
  collapse_rows(columns = 1:2, valign = "middle") %>%
  footnote(
    general = "Coefficients with * p < .05, ** p < .01, *** p < .001. Significant coefficients (p < .05) in bold. Step 1a includes actual similarity only. Step 1b adds perceived similarity. Step 1c adds the actual × perceived interaction (predictors grand-mean centered within context). All models include a dyad-level random intercept. Empty cells indicate that the predictor is not in the model at that step.",
    general_title = "Note.",
    footnote_as_chunk = TRUE,
    escape = FALSE
  )
```

# Step 2: Mediation by Perceived Similarity (H2.1–H2.2)

We test whether interaction-averaged perceived similarity mediates the association between interaction-averaged actual similarity and post-interaction evaluations. Mediation is estimated separately within each interaction context.

**Mediator Model:**

$$
\bar{S}^{perc}_{pdc} = t_{0,d} + \theta_1 \bar{S}^{act}_{pdc} + \varepsilon^M_{pdc}
$$

**Outcome Model:**

$$
O_{pdc} = f_{0,d} + \varphi_1 \bar{S}^{perc}_{pdc} + \varphi_2 \bar{S}^{act}_{pdc} + \varepsilon^O_{pdc}
$$

Both with dyad-level random intercepts.

Indirect effect: $\theta_1 \varphi_1$. Confidence intervals via Monte Carlo simulation.

## H2.1 – PA Similarity Mediation

### Negative Interaction

#### c-path (total effect)

The c-paths are the Step 1a models: `h11_PA_love_neg` and `h11_PA_close_neg`.

#### a-path (mediator model)

```{r h21_PA_mediator_neg}
#| output-fold: true
#| output-summary: "Mediator model PA (negative)"

h21_PA_med_neg <- lme(
  fixed = PA_sim_perc ~ PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h21_PA_med_neg)
```

**Intercept** $\theta_0 = 0.507$, 95% CI $[-2.546, 3.559]$, $p = .743$

Statistically zero. When actual PA similarity is zero (elevation and mismatch cancel out), perceived PA similarity is also approximately zero on average. This makes sense: if a couple has no net actual PA similarity, the actor's perception of shared PA should also center near zero.

**a-path** $\theta_1 = 0.798$, 95% CI $[0.692, 0.904]$, $p < .001$

Actual PA similarity is a strong predictor of perceived PA similarity. For every one-unit increase in interaction-averaged actual PA similarity, perceived PA similarity increases by about 0.80 units. The coefficient is large and precisely estimated, indicating that actors' perceptions of shared PA track the actual affective overlap closely during the negative interaction. This strong a-path is a necessary condition for mediation.

**Random effects**

The dyad intercept is essentially zero ($SD = 0.0005$, variance $\approx 0$), meaning there are no stable between-couple differences in perceived PA similarity once actual PA similarity is controlled. This makes sense at the interaction level: with only one observation per person per context, and with actual similarity accounting for most of the variance, there is nothing left for the dyad intercept to capture. The residual $SD = 7.39$ reflects the remaining within-dyad scatter around the regression line.

**The a-path is clearly established: actual PA similarity strongly predicts perceived PA similarity in the negative interaction.** Whether this translates into a significant indirect effect depends on the b-path (perceived similarity predicting the outcome in the presence of actual similarity), which from the Step 1b results was non-significant for both love ($-0.081$, $p = .670$) and closeness ($-0.031$, $p = .863$). The strong a-path combined with null b-paths foreshadows null indirect effects for both outcomes.

#### b- and c'-paths (outcome model — love)

```{r h21_PA_out_love_neg}
#| output-fold: true
#| output-summary: "Outcome model PA → Love (negative)"

h21_PA_out_love_neg <- lme(
  fixed = love ~ PA_sim_perc + PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h21_PA_out_love_neg)
```

This is the Step 1b model re-estimated with predictor order swapped, so the coefficients are identical to those already reported. In the mediation framework:

**b-path** $\varphi_1 = -0.081$, 95% CI $[-0.459, 0.296]$, $p = .670$

Statistically zero. Perceived PA similarity does not predict love after the negative interaction once actual similarity is controlled. This is the critical path for mediation, and it is null.

**c'-path (direct effect)** $\varphi_2 = 0.444$, 95% CI $[0.005, 0.883]$, $p = .048$

The direct effect of actual PA similarity on love, controlling for perceived similarity. As discussed in Step 1b, this is slightly larger than the total effect from Step 1a ($c = 0.379$, $p = .020$), indicating no attenuation.

#### b- and c'-paths (outcome model — closeness)

```{r h21_PA_out_close_neg}
#| output-fold: true
#| output-summary: "Outcome model PA → Closeness (negative)"

h21_PA_out_close_neg <- lme(
  fixed = close ~ PA_sim_perc + PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h21_PA_out_close_neg)
```

Again the Step 1b model with predictor order swapped. In the mediation framework:

**b-path** $\varphi_1 = -0.031$, 95% CI $[-0.379, 0.318]$, $p = .863$

Statistically zero. Perceived PA similarity does not predict closeness after the negative interaction once actual similarity is controlled.

**c'-path (direct effect)** $\varphi_2 = 0.456$, 95% CI $[0.011, 0.901]$, $p = .045$

The direct effect of actual PA similarity on closeness, controlling for perceived similarity. As in the love model, this is slightly larger than the total effect from Step 1a ($c = 0.432$, $p = .015$), indicating no attenuation.

#### Monte Carlo CI (negative interaction)

```{r h21_PA_MC_neg}
#| output-fold: true
#| output-summary: "Monte Carlo indirect effect — PA closeness (negative)"

# love
set.seed(42)

# Extract coefficients and SEs from mediator and outcome models
a_est     <- fixef(h21_PA_med_neg)["PA_sim_act"]
se_a      <- sqrt(vcov(h21_PA_med_neg)["PA_sim_act", "PA_sim_act"])
b_est     <- fixef(h21_PA_out_love_neg)["PA_sim_perc"]
se_b      <- sqrt(vcov(h21_PA_out_love_neg)["PA_sim_perc", "PA_sim_perc"])
cprime    <- fixef(h21_PA_out_love_neg)["PA_sim_act"]
c_total   <- fixef(h11_PA_love_neg)["PA_sim_act"]

# Monte Carlo draws
a_sim <- rnorm(n_sim, mean = a_est, sd = se_a)
b_sim <- rnorm(n_sim, mean = b_est, sd = se_b)
indirect_sim <- a_sim * b_sim

result <- mc_ci(indirect_sim, a_est, b_est, "PA love neg")

cat(sprintf("Indirect effect: %.4f, 95%% MC CI [%.4f, %.4f], p = %s\n",
            result$point, result$ci_low, result$ci_hi,
            ifelse(result$p < .001, "< .001", sprintf("%.3f", result$p))))
cat(sprintf("Total effect (c): %.4f\n", c_total))
cat(sprintf("Direct effect (c'): %.4f\n", cprime))

# closeness
set.seed(42)

# Extract coefficients and SEs from outcome models
b_est_c   <- fixef(h21_PA_out_close_neg)["PA_sim_perc"]
se_b_c    <- sqrt(vcov(h21_PA_out_close_neg)["PA_sim_perc", "PA_sim_perc"])
cprime_c  <- fixef(h21_PA_out_close_neg)["PA_sim_act"]
c_total_c <- fixef(h11_PA_close_neg)["PA_sim_act"]

b_sim_c <- rnorm(n_sim, mean = b_est_c, sd = se_b_c)
indirect_sim_c <- a_sim * b_sim_c

result_c <- mc_ci(indirect_sim_c, a_est, b_est_c, "PA close neg")

cat(sprintf("Indirect effect: %.4f, 95%% MC CI [%.4f, %.4f], p = %s\n",
            result_c$point, result_c$ci_low, result_c$ci_hi,
            ifelse(result_c$p < .001, "< .001", sprintf("%.3f", result_c$p))))
cat(sprintf("Total effect (c): %.4f\n", c_total_c))
cat(sprintf("Direct effect (c'): %.4f\n", cprime_c))
```

**Monte Carlo indirect effect** $\theta_1 \times \varphi_1 = -0.065$, 95% MC CI $[-0.364, 0.235]$, $p = .677$

The indirect effect is non-significant, as anticipated from the null b-path. The CI is wide and centered near zero, confirming that perceived PA similarity does not mediate the association between actual PA similarity and love in the negative interaction. The total effect ($c = 0.379$) is fully retained in the direct effect ($c' = 0.444$), with the slight increase reflecting the non-mediating role of perceived similarity rather than suppression in any meaningful sense (the b-path is too close to zero).

**Monte Carlo indirect effect** $\theta_1 \times \varphi_1 = -0.024$, 95% MC CI $[-0.263, 0.284]$, $p = .858$

The indirect effect is non-significant and even smaller than for love ($-0.065$), consistent with the weaker b-path for closeness ($-0.031$) compared to love ($-0.081$). The total effect ($c = 0.432$) is again fully retained in the direct effect ($c' = 0.456$).

#### Mediation diagramme

```{r h21_PA_mediation_diagram_neg}
#| fig-cap: "PA mediation model — Negative interaction"

# Extract coefficients and CIs
a_est  <- fixef(h21_PA_med_neg)["PA_sim_act"]
a_ci   <- intervals(h21_PA_med_neg, which = "fixed")$fixed["PA_sim_act", ]
a_p    <- summary(h21_PA_med_neg)$tTable["PA_sim_act", "p-value"]

b_love_est  <- fixef(h13_PA_love_neg)["PA_sim_perc"]
b_love_ci   <- intervals(h13_PA_love_neg, which = "fixed")$fixed["PA_sim_perc", ]
b_love_p    <- summary(h13_PA_love_neg)$tTable["PA_sim_perc", "p-value"]

cp_love_est <- fixef(h13_PA_love_neg)["PA_sim_act"]
cp_love_ci  <- intervals(h13_PA_love_neg, which = "fixed")$fixed["PA_sim_act", ]
cp_love_p   <- summary(h13_PA_love_neg)$tTable["PA_sim_act", "p-value"]

c_love_est  <- fixef(h11_PA_love_neg)["PA_sim_act"]
c_love_ci   <- intervals(h11_PA_love_neg, which = "fixed")$fixed["PA_sim_act", ]
c_love_p    <- summary(h11_PA_love_neg)$tTable["PA_sim_act", "p-value"]

b_close_est  <- fixef(h13_PA_close_neg)["PA_sim_perc"]
b_close_ci   <- intervals(h13_PA_close_neg, which = "fixed")$fixed["PA_sim_perc", ]
b_close_p    <- summary(h13_PA_close_neg)$tTable["PA_sim_perc", "p-value"]

cp_close_est <- fixef(h13_PA_close_neg)["PA_sim_act"]
cp_close_ci  <- intervals(h13_PA_close_neg, which = "fixed")$fixed["PA_sim_act", ]
cp_close_p   <- summary(h13_PA_close_neg)$tTable["PA_sim_act", "p-value"]

c_close_est  <- fixef(h11_PA_close_neg)["PA_sim_act"]
c_close_ci   <- intervals(h11_PA_close_neg, which = "fixed")$fixed["PA_sim_act", ]
c_close_p    <- summary(h11_PA_close_neg)$tTable["PA_sim_act", "p-value"]

# --- Love diagram ---
nodes_love <- data.frame(
  name  = c("X", "M", "Y"),
  label = c("Actual PA\nSimilarity", "Perceived PA\nSimilarity", "Love"),
  shape = c("rect", "rect", "rect")
)

layout_love <- get_layout(
  "X", "M",
  NA,  "Y",
  rows = 2
)

edges_love <- data.frame(
  from      = c("X", "M", "X", "X"),
  to        = c("M", "Y", "Y", "Y"),
  label     = c(paste0("a: ",  fmt(a_est, a_ci["lower"], a_ci["upper"], a_p)),
                paste0("b: ",  fmt(b_love_est, b_love_ci["lower"], b_love_ci["upper"], b_love_p)),
                paste0("c': ", fmt(cp_love_est, cp_love_ci["lower"], cp_love_ci["upper"], cp_love_p)),
                paste0("c: ",  fmt(c_love_est, c_love_ci["lower"], c_love_ci["upper"], c_love_p))),
  linetype  = c(1, 1, 1, 2),
  colour    = c(sig_col(a_p), sig_col(b_love_p), sig_col(cp_love_p), sig_col(c_love_p)),
  curvature = c(NA, NA, NA, -30)
)

p_love <- prepare_graph(
  nodes  = nodes_love,
  edges  = edges_love,
  layout = layout_love,
  spacing_x = 3, spacing_y = 2,
  rect_width = 0.8, rect_height = 0.5
)

edges(p_love)$linetype    <- edges_love$linetype
edges(p_love)$colour      <- edges_love$colour
edges(p_love)$label_colour <- edges_love$colour
edges(p_love)$label_hjust <- "center"

p1 <- plot(p_love) + ggtitle("PA Mediation — Negative Interaction — Love")

# --- Closeness diagram ---
nodes_close <- data.frame(
  name  = c("X", "M", "Y"),
  label = c("Actual PA\nSimilarity", "Perceived PA\nSimilarity", "Closeness"),
  shape = c("rect", "rect", "rect")
)

edges_close <- data.frame(
  from      = c("X", "M", "X", "X"),
  to        = c("M", "Y", "Y", "Y"),
  label     = c(paste0("a: ",  fmt(a_est, a_ci["lower"], a_ci["upper"], a_p)),
                paste0("b: ",  fmt(b_close_est, b_close_ci["lower"], b_close_ci["upper"], b_close_p)),
                paste0("c': ", fmt(cp_close_est, cp_close_ci["lower"], cp_close_ci["upper"], cp_close_p)),
                paste0("c: ",  fmt(c_close_est, c_close_ci["lower"], c_close_ci["upper"], c_close_p))),
  linetype  = c(1, 1, 1, 2),
  colour    = c(sig_col(a_p), sig_col(b_close_p), sig_col(cp_close_p), sig_col(c_close_p)),
  curvature = c(NA, NA, NA, -30)
)

p_close <- prepare_graph(
  nodes  = nodes_close,
  edges  = edges_close,
  layout = layout_love,
  spacing_x = 3, spacing_y = 2,
  rect_width = 0.8, rect_height = 0.5
)

edges(p_close)$linetype    <- edges_close$linetype
edges(p_close)$colour      <- edges_close$colour
edges(p_close)$label_colour <- edges_close$colour
edges(p_close)$label_hjust <- "center"

p2 <- plot(p_close) + ggtitle("PA Mediation — Negative Interaction — Closeness")

p1 / p2
```

#### Mediation summary for love in the negative interaction

**Mediation summary for love in the negative interaction:**

The a-path is strong ($\theta_1 = 0.798$, $p < .001$): actors accurately perceive the actual PA overlap. But the b-path is null ($\varphi_1 = -0.081$, $p = .670$): this accurate perception does not translate into higher love. The indirect effect ($\theta_1 \times \varphi_1 = 0.798 \times -0.081 = -0.065$) is expected to be non-significant, which will be confirmed by the Monte Carlo CI below. The entire association between actual PA similarity and love operates as a direct effect; perceived similarity does not mediate.

**H2.1 is not supported for love in the negative interaction.** The mediation pathway is blocked at the b-path: actors perceive actual PA overlap accurately, but this perception carries no unique information for post-interaction love.

#### Mediation summary for closeness in the negative interaction

**Mediation summary for closeness in the negative interaction:**

The pattern is identical to love. The a-path is strong ($\theta_1 = 0.798$, $p < .001$), but the b-path is null ($\varphi_1 = -0.031$, $p = .863$). The indirect effect ($\theta_1 \times \varphi_1 = 0.798 \times -0.031 = -0.025$) is expected to be non-significant. Across both outcomes in the negative interaction, the PA mediation story is the same: actors perceive the actual PA overlap accurately, but this perception has no unique downstream consequence for post-interaction evaluations. The association between actual PA similarity and both love and closeness operates entirely as a direct effect.

**H2.1 is not supported for closeness in the negative interaction.** The pattern is identical to love: the mediation pathway is blocked at the b-path. Across both outcomes, the PA mediation results for the negative interaction are clear and consistent: actual PA similarity predicts post-interaction evaluations directly, and perceived similarity plays no mediating role despite being accurately tracked (strong a-path).

### Positive Interaction

#### c-path (total effect)

The c-paths are the Step 1a models: `h11_PA_love_pos` and `h11_PA_close_pos`.

#### a-path (mediator model)

```{r h21_PA_mediator_pos}
#| output-fold: true
#| output-summary: "Mediator model PA (positive)"

h21_PA_med_pos <- lme(
  fixed = PA_sim_perc ~ PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h21_PA_med_pos)
```

**Intercept** $\theta_0 = 4.322$, 95% CI $[3.268, 5.377]$, $p < .001$

Unlike the negative interaction where the intercept was essentially zero ($0.507$, $p = .743$), here the intercept is significantly positive. This means that when actual PA similarity is zero, perceived PA similarity is on average about 4.3 units above zero. Actors in the positive interaction tend to perceive more shared PA than actually exists, even when the actual overlap is zero. This positive bias is plausible: the positive interaction context may create a general expectation of shared positivity that inflates perception.

**a-path** $\theta_1 = 0.702$, 95% CI $[0.592, 0.812]$, $p < .001$

Actual PA similarity is a strong predictor of perceived PA similarity. The coefficient is somewhat lower than in the negative interaction ($0.798$), meaning that perception tracks actuality slightly less closely during the positive interaction. The combination of a higher intercept and a lower slope suggests that in the positive interaction, actors start from a higher baseline perception of shared PA but are somewhat less sensitive to actual variation in overlap.

**Random effects**

The dyad intercept is again essentially zero ($SD = 0.001$), as in the negative interaction. The residual $SD = 6.60$ is slightly smaller than in the negative interaction ($7.39$), indicating marginally less scatter around the regression line.

**The a-path is clearly established in the positive interaction.** As in the negative interaction, whether this translates into significant indirect effects depends on the b-paths, which from Step 1b were non-significant for love ($0.105$, $p = .538$) and closeness ($0.229$, $p = .147$).

#### b- and c'-paths (outcome model - love)

```{r h21_PA_out_love_pos}
#| output-fold: true
#| output-summary: "Outcome model PA → Love (positive)"

h21_PA_out_love_pos <- lme(
  fixed = love ~ PA_sim_perc + PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h21_PA_out_love_pos)
```

**b-path** $\varphi_1 = 0.105$, 95% CI $[-0.232, 0.441]$, $p = .538$

Statistically zero. Perceived PA similarity does not predict love after the positive interaction once actual similarity is controlled.

**c'-path (direct effect)** $\varphi_2 = 0.386$, 95% CI $[-0.022, 0.793]$, $p = .063$

The direct effect of actual PA similarity on love, controlling for perceived similarity. This attenuates modestly from the total effect ($c = 0.459$, $p = .007$) and drops below conventional significance. As discussed in Step 1b, this attenuation is the most consistent with H1.3 among the PA love models, but the b-path is null, so the attenuation does not reflect a functioning mediation pathway.

#### b- and c'-paths (outcome model - closeness)

```{r h21_PA_out_close_pos}
#| output-fold: true
#| output-summary: "Outcome model PA → Closeness (positive)"

h21_PA_out_close_pos <- lme(
  fixed = close ~ PA_sim_perc + PA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h21_PA_out_close_pos)
```

**b-path** $\varphi_1 = 0.229$, 95% CI $[-0.082, 0.540]$, $p = .147$

Statistically zero, though this is the largest b-path coefficient across all PA mediation models. It does not reach significance.

**c'-path (direct effect)** $\varphi_2 = 0.215$, 95% CI $[-0.158, 0.588]$, $p = .256$

Statistically zero. The direct effect attenuates substantially from the total effect ($c = 0.376$, $p = .016$), more so than for love in the same context. As discussed in Step 1b, this is the model where the Step 1a total effect splits most evenly between the two predictors ($0.215$ vs. $0.229$), but neither reaches significance individually.

#### Monte Carlo CI (positive interaction)

```{r h21_PA_MC_pos}
#| output-fold: true
#| output-summary: "Monte Carlo indirect effect — PA love (positive)"

# love
set.seed(42)

a_est_p   <- fixef(h21_PA_med_pos)["PA_sim_act"]
se_a_p    <- sqrt(vcov(h21_PA_med_pos)["PA_sim_act", "PA_sim_act"])
b_est_p   <- fixef(h21_PA_out_love_pos)["PA_sim_perc"]
se_b_p    <- sqrt(vcov(h21_PA_out_love_pos)["PA_sim_perc", "PA_sim_perc"])
cprime_p  <- fixef(h21_PA_out_love_pos)["PA_sim_act"]
c_total_p <- fixef(h11_PA_love_pos)["PA_sim_act"]

a_sim_p <- rnorm(n_sim, mean = a_est_p, sd = se_a_p)
b_sim_p <- rnorm(n_sim, mean = b_est_p, sd = se_b_p)
indirect_sim_p <- a_sim_p * b_sim_p

result_p <- mc_ci(indirect_sim_p, a_est_p, b_est_p, "PA love pos")

cat(sprintf("Indirect effect: %.4f, 95%% MC CI [%.4f, %.4f], p = %s\n",
            result_p$point, result_p$ci_low, result_p$ci_hi,
            ifelse(result_p$p < .001, "< .001", sprintf("%.3f", result_p$p))))
cat(sprintf("Total effect (c): %.4f\n", c_total_p))
cat(sprintf("Direct effect (c'): %.4f\n", cprime_p))

# closeness
set.seed(42)

b_est_pc   <- fixef(h21_PA_out_close_pos)["PA_sim_perc"]
se_b_pc    <- sqrt(vcov(h21_PA_out_close_pos)["PA_sim_perc", "PA_sim_perc"])
cprime_pc  <- fixef(h21_PA_out_close_pos)["PA_sim_act"]
c_total_pc <- fixef(h11_PA_close_pos)["PA_sim_act"]

b_sim_pc <- rnorm(n_sim, mean = b_est_pc, sd = se_b_pc)
indirect_sim_pc <- a_sim_p * b_sim_pc

result_pc <- mc_ci(indirect_sim_pc, a_est_p, b_est_pc, "PA close pos")

cat(sprintf("Indirect effect: %.4f, 95%% MC CI [%.4f, %.4f], p = %s\n",
            result_pc$point, result_pc$ci_low, result_pc$ci_hi,
            ifelse(result_pc$p < .001, "< .001", sprintf("%.3f", result_pc$p))))
cat(sprintf("Total effect (c): %.4f\n", c_total_pc))
cat(sprintf("Direct effect (c'): %.4f\n", cprime_pc))
```

**Love - Monte Carlo indirect effect** $\theta_1 \times \varphi_1 = 0.074$, 95% MC CI $[-0.161, 0.313]$, $p = .544$

The indirect effect is non-significant, as anticipated from the null b-path. The total effect ($c = 0.459$) attenuates modestly to the direct effect ($c' = 0.386$), but this reduction is not carried by a functioning mediation pathway.

**Closeness - Monte Carlo indirect effect** $\theta_1 \times \varphi_1 = 0.161$, 95% MC CI $[-0.049, 0.435]$, $p = .151$

This is the largest PA indirect effect across all four models, and the CI comes closest to excluding zero, but it remains non-significant. The total effect ($c = 0.376$) attenuates substantially to the direct effect ($c' = 0.215$), with the indirect pathway accounting for the difference. However, without a significant b-path or a significant indirect effect, this attenuation pattern does not constitute evidence for mediation.

#### Mediation diagramme positive interaction

```{r h21_PA_mediation_diagram_pos}
#| fig-cap: "PA mediation model — Positive interaction"

# Extract coefficients and CIs
a_est  <- fixef(h21_PA_med_pos)["PA_sim_act"]
a_ci   <- intervals(h21_PA_med_pos, which = "fixed")$fixed["PA_sim_act", ]
a_p    <- summary(h21_PA_med_pos)$tTable["PA_sim_act", "p-value"]

b_love_est  <- fixef(h13_PA_love_pos)["PA_sim_perc"]
b_love_ci   <- intervals(h13_PA_love_pos, which = "fixed")$fixed["PA_sim_perc", ]
b_love_p    <- summary(h13_PA_love_pos)$tTable["PA_sim_perc", "p-value"]

cp_love_est <- fixef(h13_PA_love_pos)["PA_sim_act"]
cp_love_ci  <- intervals(h13_PA_love_pos, which = "fixed")$fixed["PA_sim_act", ]
cp_love_p   <- summary(h13_PA_love_pos)$tTable["PA_sim_act", "p-value"]

c_love_est  <- fixef(h11_PA_love_pos)["PA_sim_act"]
c_love_ci   <- intervals(h11_PA_love_pos, which = "fixed")$fixed["PA_sim_act", ]
c_love_p    <- summary(h11_PA_love_pos)$tTable["PA_sim_act", "p-value"]

b_close_est  <- fixef(h13_PA_close_pos)["PA_sim_perc"]
b_close_ci   <- intervals(h13_PA_close_pos, which = "fixed")$fixed["PA_sim_perc", ]
b_close_p    <- summary(h13_PA_close_pos)$tTable["PA_sim_perc", "p-value"]

cp_close_est <- fixef(h13_PA_close_pos)["PA_sim_act"]
cp_close_ci  <- intervals(h13_PA_close_pos, which = "fixed")$fixed["PA_sim_act", ]
cp_close_p   <- summary(h13_PA_close_pos)$tTable["PA_sim_act", "p-value"]

c_close_est  <- fixef(h11_PA_close_pos)["PA_sim_act"]
c_close_ci   <- intervals(h11_PA_close_pos, which = "fixed")$fixed["PA_sim_act", ]
c_close_p    <- summary(h11_PA_close_pos)$tTable["PA_sim_act", "p-value"]

# --- Love diagram ---
nodes_love <- data.frame(
  name  = c("X", "M", "Y"),
  label = c("Actual PA\nSimilarity", "Perceived PA\nSimilarity", "Love"),
  shape = c("rect", "rect", "rect")
)

layout_love <- get_layout(
  "X", "M",
  NA,  "Y",
  rows = 2
)

edges_love <- data.frame(
  from      = c("X", "M", "X", "X"),
  to        = c("M", "Y", "Y", "Y"),
  label     = c(paste0("a: ",  fmt(a_est, a_ci["lower"], a_ci["upper"], a_p)),
                paste0("b: ",  fmt(b_love_est, b_love_ci["lower"], b_love_ci["upper"], b_love_p)),
                paste0("c': ", fmt(cp_love_est, cp_love_ci["lower"], cp_love_ci["upper"], cp_love_p)),
                paste0("c: ",  fmt(c_love_est, c_love_ci["lower"], c_love_ci["upper"], c_love_p))),
  linetype  = c(1, 1, 1, 2),
  colour    = c(sig_col(a_p), sig_col(b_love_p), sig_col(cp_love_p), sig_col(c_love_p)),
  curvature = c(NA, NA, NA, -30)
)

p_love <- prepare_graph(
  nodes  = nodes_love,
  edges  = edges_love,
  layout = layout_love,
  spacing_x = 3, spacing_y = 2,
  rect_width = 0.8, rect_height = 0.5
)

edges(p_love)$linetype    <- edges_love$linetype
edges(p_love)$colour      <- edges_love$colour
edges(p_love)$label_colour <- edges_love$colour
edges(p_love)$label_hjust <- "center"

p1 <- plot(p_love) + ggtitle("PA Mediation — Positive Interaction — Love")

# --- Closeness diagram ---
nodes_close <- data.frame(
  name  = c("X", "M", "Y"),
  label = c("Actual PA\nSimilarity", "Perceived PA\nSimilarity", "Closeness"),
  shape = c("rect", "rect", "rect")
)

edges_close <- data.frame(
  from      = c("X", "M", "X", "X"),
  to        = c("M", "Y", "Y", "Y"),
  label     = c(paste0("a: ",  fmt(a_est, a_ci["lower"], a_ci["upper"], a_p)),
                paste0("b: ",  fmt(b_close_est, b_close_ci["lower"], b_close_ci["upper"], b_close_p)),
                paste0("c': ", fmt(cp_close_est, cp_close_ci["lower"], cp_close_ci["upper"], cp_close_p)),
                paste0("c: ",  fmt(c_close_est, c_close_ci["lower"], c_close_ci["upper"], c_close_p))),
  linetype  = c(1, 1, 1, 2),
  colour    = c(sig_col(a_p), sig_col(b_close_p), sig_col(cp_close_p), sig_col(c_close_p)),
  curvature = c(NA, NA, NA, -30)
)

p_close <- prepare_graph(
  nodes  = nodes_close,
  edges  = edges_close,
  layout = layout_love,
  spacing_x = 3, spacing_y = 2,
  rect_width = 0.8, rect_height = 0.5
)

edges(p_close)$linetype    <- edges_close$linetype
edges(p_close)$colour      <- edges_close$colour
edges(p_close)$label_colour <- edges_close$colour
edges(p_close)$label_hjust <- "center"

p2 <- plot(p_close) + ggtitle("PA Mediation — Positive Interaction — Closeness")

p1 / p2
```

#### **Mediation summary for love in the positive interaction**

The a-path is strong ($\theta_1 = 0.702$, $p < .001$), but the b-path is statistically zero ($\varphi_1 = 0.105$, $p = .538$). The indirect effect ($\theta_1 \times \varphi_1 = 0.702 \times 0.105 = 0.074$) is expected to be non-significant. The modest attenuation of the c'-path relative to the c-path is consistent with the two correlated predictors sharing some variance, but without a significant b-path this does not constitute evidence for mediation.

**H2.1 is not supported for love in the positive interaction.** As in the negative interaction, the mediation pathway is blocked at the b-path.

#### **Mediation summary for closeness in the positive interaction**

The a-path is strong ($\theta_1 = 0.702$, $p < .001$), and the b-path is the largest observed across the PA models ($0.229$), but it remains non-significant ($p = .147$). The indirect effect ($\theta_1 \times \varphi_1 = 0.702 \times 0.229 = 0.161$) will be the largest PA indirect effect tested, but given the non-significant b-path, it is unlikely to reach significance. This is the PA model closest to supporting H2.1, but the evidence does not cross the threshold: the total effect is real (Step 1a), the a-path is strong, yet the b-path fails to establish that perceived similarity is the channel through which actual similarity reaches closeness.

**H2.1 is not supported for closeness in the positive interaction**, though this is the model where the mediation pattern comes closest. Across all four PA mediation models (two outcomes $\times$ two contexts), the conclusion is consistent: the a-path is always strong (actors perceive actual PA similarity accurately), but the b-path never reaches significance (this perception does not uniquely predict post-interaction evaluations). PA mediation is not supported in SC2.

## H2.2 – NA Similarity Mediation

### Negative Interaction

#### c-path (total effect)

The c-paths are the Step 1a models: `h11_NA_love_neg` and `h11_NA_close_neg`.

#### a-path (mediator model)

```{r h22_NA_mediator_neg}
#| output-fold: true
#| output-summary: "Mediator model NA (negative)"

h22_NA_med_neg <- lme(
  fixed = NA_sim_perc ~ NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h22_NA_med_neg)
```

**Intercept** $\theta_0 = 2.067$, 95% CI $[0.555, 3.579]$, $p = .008$

Unlike the PA negative interaction model where the intercept was essentially zero, here it is small but significantly positive. When actual NA similarity is zero, perceived NA similarity is on average about 2.1 units above zero, suggesting a slight positive perceptual bias: actors tend to perceive marginally more shared NA than actually exists, even when the actual overlap is zero.

**a-path** $\theta_1 = 0.701$, 95% CI $[0.574, 0.828]$, $p < .001$

Actual NA similarity is a strong predictor of perceived NA similarity. The coefficient is lower than the PA a-path in the same context ($0.798$), indicating that perception tracks actuality somewhat less closely for NA than for PA during the negative interaction. This is consistent with negative affect being harder to read accurately in a partner than positive affect.

**Random effects**

The dyad intercept is essentially zero ($SD = 0.0006$), as in all mediator models. The residual $SD = 6.95$ is comparable to the PA mediator model ($7.39$).

**The a-path is clearly established for NA in the negative interaction.** From the Step 1b results, however, the b-paths were statistically zero for both love ($0.108$, $p = .603$) and closeness ($0.026$, $p = .892$), and the total effects (Step 1a) were also null. The mediation pathway is blocked at multiple points: there is no total effect to mediate, and no b-path to carry an indirect effect.

#### b- and c'-paths (outcome model - love)

```{r h22_NA_out_love_neg}
#| output-fold: true
#| output-summary: "Outcome model NA → Love (negative)"

h22_NA_out_love_neg <- lme(
  fixed = love ~ NA_sim_perc + NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h22_NA_out_love_neg)
```

**b-path** $\varphi_1 = 0.108$, 95% CI $[-0.302, 0.518]$, $p = .603$

Statistically zero. Perceived NA similarity does not predict love after the negative interaction once actual similarity is controlled.

**c'-path (direct effect)** $\varphi_2 = -0.217$, 95% CI $[-0.726, 0.292]$, $p = .400$

Statistically zero, consistent with the null total effect from Step 1a ($c = -0.141$, $p = .504$).

#### b- and c'-paths (outcome model - closeness)

```{r h22_NA_out_close_neg}
#| output-fold: true
#| output-summary: "Outcome model NA → Closeness (negative)"

h22_NA_out_close_neg <- lme(
  fixed = close ~ NA_sim_perc + NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h22_NA_out_close_neg)
```

**b-path** $\varphi_1 = 0.026$, 95% CI $[-0.357, 0.410]$, $p = .892$

Statistically zero. Perceived NA similarity does not predict closeness after the negative interaction once actual similarity is controlled.

**c'-path (direct effect)** $\varphi_2 = -0.175$, 95% CI $[-0.705, 0.355]$, $p = .513$

Statistically zero, consistent with the null total effect from Step 1a ($c = -0.157$, $p = .497$).

#### Monte Carlo CI (negative interaction)

```{r h22_NA_MC_neg}
#| output-fold: true
#| output-summary: "Monte Carlo indirect effects — NA (negative)"

# love
set.seed(42)

a_NA_neg     <- fixef(h22_NA_med_neg)["NA_sim_act"]
se_a_NA_neg  <- sqrt(vcov(h22_NA_med_neg)["NA_sim_act", "NA_sim_act"])
b_NA_love_neg <- fixef(h22_NA_out_love_neg)["NA_sim_perc"]
se_b_NA_love_neg <- sqrt(vcov(h22_NA_out_love_neg)["NA_sim_perc", "NA_sim_perc"])

a_NA_sim_neg <- rnorm(n_sim, mean = a_NA_neg, sd = se_a_NA_neg)
b_NA_sim_love_neg <- rnorm(n_sim, mean = b_NA_love_neg, sd = se_b_NA_love_neg)
ind_NA_love_neg <- a_NA_sim_neg * b_NA_sim_love_neg

res_NA_love_neg <- mc_ci(ind_NA_love_neg, a_NA_neg, b_NA_love_neg, "NA love neg")

cat("--- Love (negative) ---\n")
cat(sprintf("Indirect: %.4f, 95%% MC CI [%.4f, %.4f], p = %s\n",
            res_NA_love_neg$point, res_NA_love_neg$ci_low, res_NA_love_neg$ci_hi,
            ifelse(res_NA_love_neg$p < .001, "< .001", sprintf("%.3f", res_NA_love_neg$p))))

# closeness
set.seed(42)

b_NA_close_neg <- fixef(h22_NA_out_close_neg)["NA_sim_perc"]
se_b_NA_close_neg <- sqrt(vcov(h22_NA_out_close_neg)["NA_sim_perc", "NA_sim_perc"])

b_NA_sim_close_neg <- rnorm(n_sim, mean = b_NA_close_neg, sd = se_b_NA_close_neg)
ind_NA_close_neg <- a_NA_sim_neg * b_NA_sim_close_neg

res_NA_close_neg <- mc_ci(ind_NA_close_neg, a_NA_neg, b_NA_close_neg, "NA close neg")

cat("\n--- Closeness (negative) ---\n")
cat(sprintf("Indirect: %.4f, 95%% MC CI [%.4f, %.4f], p = %s\n",
            res_NA_close_neg$point, res_NA_close_neg$ci_low, res_NA_close_neg$ci_hi,
            ifelse(res_NA_close_neg$p < .001, "< .001", sprintf("%.3f", res_NA_close_neg$p))))
```

**Love - Monte Carlo indirect effect** $\theta_1 \times \varphi_1 = 0.076$, 95% MC CI $[-0.209, 0.368]$, $p = .605$

**Closeness - Monte Carlo indirect effect** $\theta_1 \times \varphi_1 = 0.019$, 95% MC CI $[-0.205, 0.336]$, $p = .891$

#### Mediation diagramme for negative interaction

```{r h22_NA_mediation_diagram_neg}
#| fig-cap: "NA mediation model — Negative interaction"

# Extract coefficients and CIs
a_est  <- fixef(h22_NA_med_neg)["NA_sim_act"]
a_ci   <- intervals(h22_NA_med_neg, which = "fixed")$fixed["NA_sim_act", ]
a_p    <- summary(h22_NA_med_neg)$tTable["NA_sim_act", "p-value"]

b_love_est  <- fixef(h13_NA_love_neg)["NA_sim_perc"]
b_love_ci   <- intervals(h13_NA_love_neg, which = "fixed")$fixed["NA_sim_perc", ]
b_love_p    <- summary(h13_NA_love_neg)$tTable["NA_sim_perc", "p-value"]

cp_love_est <- fixef(h13_NA_love_neg)["NA_sim_act"]
cp_love_ci  <- intervals(h13_NA_love_neg, which = "fixed")$fixed["NA_sim_act", ]
cp_love_p   <- summary(h13_NA_love_neg)$tTable["NA_sim_act", "p-value"]

c_love_est  <- fixef(h12_NA_love_neg)["NA_sim_act"]
c_love_ci   <- intervals(h12_NA_love_neg, which = "fixed")$fixed["NA_sim_act", ]
c_love_p    <- summary(h12_NA_love_neg)$tTable["NA_sim_act", "p-value"]

b_close_est  <- fixef(h13_NA_close_neg)["NA_sim_perc"]
b_close_ci   <- intervals(h13_NA_close_neg, which = "fixed")$fixed["NA_sim_perc", ]
b_close_p    <- summary(h13_NA_close_neg)$tTable["NA_sim_perc", "p-value"]

cp_close_est <- fixef(h13_NA_close_neg)["NA_sim_act"]
cp_close_ci  <- intervals(h13_NA_close_neg, which = "fixed")$fixed["NA_sim_act", ]
cp_close_p   <- summary(h13_NA_close_neg)$tTable["NA_sim_act", "p-value"]

c_close_est  <- fixef(h12_NA_close_neg)["NA_sim_act"]
c_close_ci   <- intervals(h12_NA_close_neg, which = "fixed")$fixed["NA_sim_act", ]
c_close_p    <- summary(h12_NA_close_neg)$tTable["NA_sim_act", "p-value"]

# --- Love diagram ---
nodes_love <- data.frame(
  name  = c("X", "M", "Y"),
  label = c("Actual NA\nSimilarity", "Perceived NA\nSimilarity", "Love"),
  shape = c("rect", "rect", "rect")
)

layout_love <- get_layout(
  "X", "M",
  NA,  "Y",
  rows = 2
)

edges_love <- data.frame(
  from      = c("X", "M", "X", "X"),
  to        = c("M", "Y", "Y", "Y"),
  label     = c(paste0("a: ",  fmt(a_est, a_ci["lower"], a_ci["upper"], a_p)),
                paste0("b: ",  fmt(b_love_est, b_love_ci["lower"], b_love_ci["upper"], b_love_p)),
                paste0("c': ", fmt(cp_love_est, cp_love_ci["lower"], cp_love_ci["upper"], cp_love_p)),
                paste0("c: ",  fmt(c_love_est, c_love_ci["lower"], c_love_ci["upper"], c_love_p))),
  linetype  = c(1, 1, 1, 2),
  colour    = c(sig_col(a_p), sig_col(b_love_p), sig_col(cp_love_p), sig_col(c_love_p)),
  curvature = c(NA, NA, NA, -30)
)

p_love <- prepare_graph(
  nodes  = nodes_love,
  edges  = edges_love,
  layout = layout_love,
  spacing_x = 3, spacing_y = 2,
  rect_width = 0.8, rect_height = 0.5
)

edges(p_love)$linetype     <- edges_love$linetype
edges(p_love)$colour       <- edges_love$colour
edges(p_love)$label_colour <- edges_love$colour
edges(p_love)$label_hjust  <- "center"

p1 <- plot(p_love) + ggtitle("NA Mediation — Negative Interaction — Love")

# --- Closeness diagram ---
nodes_close <- data.frame(
  name  = c("X", "M", "Y"),
  label = c("Actual NA\nSimilarity", "Perceived NA\nSimilarity", "Closeness"),
  shape = c("rect", "rect", "rect")
)

edges_close <- data.frame(
  from      = c("X", "M", "X", "X"),
  to        = c("M", "Y", "Y", "Y"),
  label     = c(paste0("a: ",  fmt(a_est, a_ci["lower"], a_ci["upper"], a_p)),
                paste0("b: ",  fmt(b_close_est, b_close_ci["lower"], b_close_ci["upper"], b_close_p)),
                paste0("c': ", fmt(cp_close_est, cp_close_ci["lower"], cp_close_ci["upper"], cp_close_p)),
                paste0("c: ",  fmt(c_close_est, c_close_ci["lower"], c_close_ci["upper"], c_close_p))),
  linetype  = c(1, 1, 1, 2),
  colour    = c(sig_col(a_p), sig_col(b_close_p), sig_col(cp_close_p), sig_col(c_close_p)),
  curvature = c(NA, NA, NA, -30)
)

p_close <- prepare_graph(
  nodes  = nodes_close,
  edges  = edges_close,
  layout = layout_love,
  spacing_x = 3, spacing_y = 2,
  rect_width = 0.8, rect_height = 0.5
)

edges(p_close)$linetype     <- edges_close$linetype
edges(p_close)$colour       <- edges_close$colour
edges(p_close)$label_colour <- edges_close$colour
edges(p_close)$label_hjust  <- "center"

p2 <- plot(p_close) + ggtitle("NA Mediation — Negative Interaction — Closeness")

p1 / p2
```

#### Mediation summary for love in the negative interaction

The a-path is strong ($\theta_1 = 0.701$, $p < .001$), but every other component of the mediation is null: the total effect is non-significant, the b-path is non-significant, and the direct effect is non-significant. The indirect effect ($\theta_1 \times \varphi_1 = 0.701 \times 0.108 = 0.076$) is expected to be non-significant. This is a case where mediation is not merely unsupported but structurally inapplicable: there is no association between actual NA similarity and love to be mediated in the first place.

The indirect effect is non-significant, as expected given both the null b-path and the null total effect. **H2.2 is not supported for love in the negative interaction.**

#### Mediation summary for closeness in the negative interaction

The pattern is identical to love: strong a-path ($\theta_1 = 0.701$, $p < .001$), but null total effect, null b-path, and null direct effect. The indirect effect ($\theta_1 \times \varphi_1 = 0.701 \times 0.026 = 0.018$) is expected to be non-significant. As with love, mediation is structurally inapplicable: there is no association to mediate. Across both outcomes in the negative interaction, the NA mediation picture is clear: actors perceive actual NA overlap accurately, but this perception carries no information for post-interaction evaluations, and actual NA similarity itself is unrelated to both love and closeness.

The indirect effect is non-significant and essentially zero. **H2.2 is not supported for closeness in the negative interaction.** Across both outcomes, the NA mediation results for the negative interaction are uniformly null at every stage: no total effect, no b-path, no indirect effect. The only functioning component is the a-path, which confirms that actors perceive actual NA overlap accurately but this perception leads nowhere.

### Positive Interaction

#### c-path (total effect)

The c-paths are the Step 1a models: `h12_NA_love_pos` and `h12_NA_close_pos`.

#### a-path (mediator model)

```{r h22_NA_mediator_pos}
#| output-fold: true
#| output-summary: "Mediator model NA (positive)"

h22_NA_med_pos <- lme(
  fixed = NA_sim_perc ~ NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h22_NA_med_pos)
```

**Intercept** $\theta_0 = -1.286$, 95% CI $[-3.745, 1.173]$, $p = .302$

Statistically zero. When actual NA similarity is zero, perceived NA similarity is approximately zero on average in the positive interaction. This contrasts with the negative interaction, where the intercept was small but significantly positive ($2.067$, $p = .008$), and with the PA positive interaction, where the intercept was clearly positive ($4.322$, $p < .001$). In the positive interaction, actors do not show a systematic perceptual bias for NA similarity at zero actual overlap.

**a-path** $\theta_1 = 0.817$, 95% CI $[0.716, 0.919]$, $p < .001$

Actual NA similarity is a strong predictor of perceived NA similarity. This is the strongest a-path across all four mediator models (PA negative: $0.798$; PA positive: $0.702$; NA negative: $0.701$; NA positive: $0.817$). Perception tracks actuality most closely for NA during the positive interaction, which may reflect the salience of shared negative affect in an otherwise positive context making it easier to detect.

**Random effects**

The dyad intercept is essentially zero ($SD = 0.0005$), as in all mediator models. The residual $SD = 7.70$ is comparable to the other mediator models.

**The a-path is clearly established for NA in the positive interaction.** From the Step 1b results, the b-path for love was statistically zero ($0.025$, $p = .862$), while for closeness the b-path trended negative ($-0.177$, $p = .174$) as part of the suppression pattern. The total effects (Step 1a) were significant for both outcomes ($0.352$ for love, $0.383$ for closeness), so unlike the negative interaction, there is an association that could in principle be mediated.

#### b- and c'-paths (outcome model - love)

```{r h22_NA_out_love_pos}
#| output-fold: true
#| output-summary: "Outcome model NA → Love (positive)"

h22_NA_out_love_pos <- lme(
  fixed = love ~ NA_sim_perc + NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h22_NA_out_love_pos)
```

**b-path** $\varphi_1 = 0.025$, 95% CI $[-0.256, 0.305]$, $p = .862$

Statistically zero. Perceived NA similarity does not predict love after the positive interaction once actual similarity is controlled. Despite the strong a-path ($0.817$), the perceptual channel carries no information to the outcome.

**c'-path (direct effect)** $\varphi_2 = 0.332$, 95% CI $[-0.017, 0.681]$, $p = .062$

The direct effect attenuates modestly from the total effect ($c = 0.352$, $p = .009$) and drops below conventional significance. As in the PA love positive interaction model, the loss of significance reflects collinearity rather than a functioning mediation pathway, since the b-path is essentially zero.

#### b- and c'-paths (outcome model - closeness)

```{r h22_NA_out_close_pos}
#| output-fold: true
#| output-summary: "Outcome model NA → Closeness (positive)"

h22_NA_out_close_pos <- lme(
  fixed = close ~ NA_sim_perc + NA_sim_act,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h22_NA_out_close_pos)
```

**b-path** $\varphi_1 = -0.177$, 95% CI $[-0.434, 0.080]$, $p = .174$

Statistically zero, though the negative direction is consistent with the suppression pattern identified in Step 1b. Perceived NA similarity trends toward predicting *less* closeness once actual similarity is controlled, opposite to the direction of the actual similarity effect. This is the largest b-path magnitude across all eight mediation outcome models, but it does not reach significance.

**c'-path (direct effect)** $\varphi_2 = 0.528$, 95% CI $[0.212, 0.843]$, $p = .001$

The direct effect *increases* from the total effect ($c = 0.383$, $p = .002$), confirming the suppression pattern: controlling for perceived similarity strengthens rather than attenuates the actual similarity coefficient. This is the opposite of what mediation produces, and it is the clearest suppression effect across all mediation models.

#### Monte Carlo CI (positive interaction)

```{r h22_NA_MC_pos}
#| output-fold: true
#| output-summary: "Monte Carlo indirect effects — NA (positive)"

# Love
set.seed(42)

a_NA_pos     <- fixef(h22_NA_med_pos)["NA_sim_act"]
se_a_NA_pos  <- sqrt(vcov(h22_NA_med_pos)["NA_sim_act", "NA_sim_act"])
b_NA_love_pos <- fixef(h22_NA_out_love_pos)["NA_sim_perc"]
se_b_NA_love_pos <- sqrt(vcov(h22_NA_out_love_pos)["NA_sim_perc", "NA_sim_perc"])

a_NA_sim_pos <- rnorm(n_sim, mean = a_NA_pos, sd = se_a_NA_pos)
b_NA_sim_love_pos <- rnorm(n_sim, mean = b_NA_love_pos, sd = se_b_NA_love_pos)
ind_NA_love_pos <- a_NA_sim_pos * b_NA_sim_love_pos

res_NA_love_pos <- mc_ci(ind_NA_love_pos, a_NA_pos, b_NA_love_pos, "NA love pos")

cat("--- Love (positive) ---\n")
cat(sprintf("Indirect: %.4f, 95%% MC CI [%.4f, %.4f], p = %s\n",
            res_NA_love_pos$point, res_NA_love_pos$ci_low, res_NA_love_pos$ci_hi,
            ifelse(res_NA_love_pos$p < .001, "< .001", sprintf("%.3f", res_NA_love_pos$p))))

# Closeness
set.seed(42)

b_NA_close_pos <- fixef(h22_NA_out_close_pos)["NA_sim_perc"]
se_b_NA_close_pos <- sqrt(vcov(h22_NA_out_close_pos)["NA_sim_perc", "NA_sim_perc"])

b_NA_sim_close_pos <- rnorm(n_sim, mean = b_NA_close_pos, sd = se_b_NA_close_pos)
ind_NA_close_pos <- a_NA_sim_pos * b_NA_sim_close_pos

res_NA_close_pos <- mc_ci(ind_NA_close_pos, a_NA_pos, b_NA_close_pos, "NA close pos")

cat("\n--- Closeness (positive) ---\n")
cat(sprintf("Indirect: %.4f, 95%% MC CI [%.4f, %.4f], p = %s\n",
            res_NA_close_pos$point, res_NA_close_pos$ci_low, res_NA_close_pos$ci_hi,
            ifelse(res_NA_close_pos$p < .001, "< .001", sprintf("%.3f", res_NA_close_pos$p))))
```

**Love:**

**Monte Carlo indirect effect** $\theta_1 \times \varphi_1 = 0.020$, 95% MC CI $[-0.207, 0.250]$, $p = .856$

The indirect effect is non-significant and essentially zero. **H2.2 is not supported for love in the positive interaction.** Despite the strongest a-path across all mediator models ($0.817$), the null b-path ($0.025$) ensures that the mediation pathway carries no signal.

**Closeness:**

**Monte Carlo indirect effect** $\theta_1 \times \varphi_1 = -0.145$, 95% MC CI $[-0.311, 0.070]$, $p = .173$

The indirect effect is negative and the largest in absolute magnitude across all eight mediation tests, but it does not reach significance. The CI just includes zero. This is the suppression pathway identified in the outcome model: actors who perceive more shared NA than actually exists tend to report slightly less closeness, partially counteracting the positive direct effect of actual similarity. The suppression is not statistically reliable at conventional levels, but the direction and magnitude are consistent with the pattern observed across Steps 1b and 1c.

#### Mediation diagramme for the positive interaction

```{r h22_NA_mediation_diagram_pos}
#| fig-cap: "NA mediation model — Positive interaction"
#| fig-width: 8
#| fig-height: 4

# Extract coefficients and CIs
a_est  <- fixef(h22_NA_med_pos)["NA_sim_act"]
a_ci   <- intervals(h22_NA_med_pos, which = "fixed")$fixed["NA_sim_act", ]
a_p    <- summary(h22_NA_med_pos)$tTable["NA_sim_act", "p-value"]

b_love_est  <- fixef(h22_NA_out_love_pos)["NA_sim_perc"]
b_love_ci   <- intervals(h22_NA_out_love_pos, which = "fixed")$fixed["NA_sim_perc", ]
b_love_p    <- summary(h22_NA_out_love_pos)$tTable["NA_sim_perc", "p-value"]

cp_love_est <- fixef(h22_NA_out_love_pos)["NA_sim_act"]
cp_love_ci  <- intervals(h22_NA_out_love_pos, which = "fixed")$fixed["NA_sim_act", ]
cp_love_p   <- summary(h22_NA_out_love_pos)$tTable["NA_sim_act", "p-value"]

c_love_est  <- fixef(h12_NA_love_pos)["NA_sim_act"]
c_love_ci   <- intervals(h12_NA_love_pos, which = "fixed")$fixed["NA_sim_act", ]
c_love_p    <- summary(h12_NA_love_pos)$tTable["NA_sim_act", "p-value"]

b_close_est  <- fixef(h22_NA_out_close_pos)["NA_sim_perc"]
b_close_ci   <- intervals(h22_NA_out_close_pos, which = "fixed")$fixed["NA_sim_perc", ]
b_close_p    <- summary(h22_NA_out_close_pos)$tTable["NA_sim_perc", "p-value"]

cp_close_est <- fixef(h22_NA_out_close_pos)["NA_sim_act"]
cp_close_ci  <- intervals(h22_NA_out_close_pos, which = "fixed")$fixed["NA_sim_act", ]
cp_close_p   <- summary(h22_NA_out_close_pos)$tTable["NA_sim_act", "p-value"]

c_close_est  <- fixef(h12_NA_close_pos)["NA_sim_act"]
c_close_ci   <- intervals(h12_NA_close_pos, which = "fixed")$fixed["NA_sim_act", ]
c_close_p    <- summary(h12_NA_close_pos)$tTable["NA_sim_act", "p-value"]

# --- Love diagram ---
nodes_love <- data.frame(
  name  = c("X", "M", "Y"),
  label = c("Actual NA\nSimilarity", "Perceived NA\nSimilarity", "Love"),
  shape = c("rect", "rect", "rect")
)

layout_love <- get_layout(
  "X", "M",
  NA,  "Y",
  rows = 2
)

edges_love <- data.frame(
  from      = c("X", "M", "X", "X"),
  to        = c("M", "Y", "Y", "Y"),
  label     = c(paste0("a: ",  fmt(a_est, a_ci["lower"], a_ci["upper"], a_p)),
                paste0("b: ",  fmt(b_love_est, b_love_ci["lower"], b_love_ci["upper"], b_love_p)),
                paste0("c': ", fmt(cp_love_est, cp_love_ci["lower"], cp_love_ci["upper"], cp_love_p)),
                paste0("c: ",  fmt(c_love_est, c_love_ci["lower"], c_love_ci["upper"], c_love_p))),
  linetype  = c(1, 1, 1, 2),
  colour    = c(sig_col(a_p), sig_col(b_love_p), sig_col(cp_love_p), sig_col(c_love_p)),
  curvature = c(NA, NA, NA, -30)
)

p_love <- prepare_graph(
  nodes  = nodes_love,
  edges  = edges_love,
  layout = layout_love,
  spacing_x = 3, spacing_y = 2,
  rect_width = 0.8, rect_height = 0.5
)

edges(p_love)$linetype     <- edges_love$linetype
edges(p_love)$colour       <- edges_love$colour
edges(p_love)$label_colour <- edges_love$colour
edges(p_love)$label_hjust  <- "center"

p1 <- plot(p_love) + ggtitle("NA Mediation — Positive Interaction — Love")

# --- Closeness diagram ---
nodes_close <- data.frame(
  name  = c("X", "M", "Y"),
  label = c("Actual NA\nSimilarity", "Perceived NA\nSimilarity", "Closeness"),
  shape = c("rect", "rect", "rect")
)

edges_close <- data.frame(
  from      = c("X", "M", "X", "X"),
  to        = c("M", "Y", "Y", "Y"),
  label     = c(paste0("a: ",  fmt(a_est, a_ci["lower"], a_ci["upper"], a_p)),
                paste0("b: ",  fmt(b_close_est, b_close_ci["lower"], b_close_ci["upper"], b_close_p)),
                paste0("c': ", fmt(cp_close_est, cp_close_ci["lower"], cp_close_ci["upper"], cp_close_p)),
                paste0("c: ",  fmt(c_close_est, c_close_ci["lower"], c_close_ci["upper"], c_close_p))),
  linetype  = c(1, 1, 1, 2),
  colour    = c(sig_col(a_p), sig_col(b_close_p), sig_col(cp_close_p), sig_col(c_close_p)),
  curvature = c(NA, NA, NA, -30)
)

p_close <- prepare_graph(
  nodes  = nodes_close,
  edges  = edges_close,
  layout = layout_love,
  spacing_x = 3, spacing_y = 2,
  rect_width = 0.8, rect_height = 0.5
)

edges(p_close)$linetype     <- edges_close$linetype
edges(p_close)$colour       <- edges_close$colour
edges(p_close)$label_colour <- edges_close$colour
edges(p_close)$label_hjust  <- "center"

p2 <- plot(p_close) + ggtitle("NA Mediation — Positive Interaction — Closeness")

p1 / p2
```

#### Mediation summary for love in the positive interaction

The a-path is the strongest observed across all mediator models ($\theta_1 = 0.817$, $p < .001$), but the b-path is statistically zero ($\varphi_1 = 0.025$, $p = .862$). The indirect effect ($\theta_1 \times \varphi_1 = 0.817 \times 0.025 = 0.020$) is expected to be non-significant. Despite a significant total effect in Step 1a, the mediation pathway is blocked at the b-path: actors perceive actual NA overlap very accurately during the positive interaction, but this perception does not translate into higher or lower love.

#### Mediation summary for closeness in the positive interaction

The a-path is strong ($\theta_1 = 0.817$, $p < .001$) and the b-path trends negative ($\varphi_1 = -0.177$, $p = .174$), producing an indirect effect in the negative direction ($\theta_1 \times \varphi_1 = 0.817 \times -0.177 = -0.145$). If significant, this would indicate *suppression* rather than mediation: perceived similarity would be partially masking the actual similarity effect, explaining why the total effect ($0.383$) is weaker than the direct effect ($0.528$). The Monte Carlo CI will determine whether this suppression pathway is statistically reliable.

**H2.2 is not supported for closeness in the positive interaction.** The pattern is the opposite of mediation: rather than carrying the actual similarity effect to the outcome, perceived similarity partially counteracts it. Actors who perceive more shared NA (beyond the actual overlap) tend to report slightly less closeness, while the actual overlap itself predicts more closeness. The two channels pull in opposing directions, consistent with the suppression pattern observed throughout Steps 1b and 1c for this specific model.

**H2.2 is not supported for closeness in the positive interaction**, though this model comes closest to revealing a statistically reliable indirect pathway, and the pathway that nearly emerges is suppression rather than mediation. Across all eight NA mediation tests (two outcomes $\times$ two contexts), no indirect effect reaches significance. The a-path is consistently strong, but the b-path never reliably connects perceived similarity to the outcome, blocking mediation at every turn.

# Step 3a: Moderation by Trait We-ness (H3.1–H3.2)

We examine whether the similarity–evaluation association depends on trait we-ness ($W^T_{pd}$, grand-mean centered). Models are estimated separately within each interaction context.

For actual similarity:

$$
O_{pdc} = b_{0,d} + \beta_1 \bar{S}^{act}_{pdc} + \beta_2 W^T_{pd} + \beta_3 (\bar{S}^{act}_{pdc} \times W^T_{pd}) + \varepsilon_{pdc}
$$

For perceived similarity:

$$
O_{pdc} = b_{0,d} + \beta_1 \bar{S}^{perc}_{pdc} + \beta_2 W^T_{pd} + \beta_3 (\bar{S}^{perc}_{pdc} \times W^T_{pd}) + \varepsilon_{pdc}
$$

Both with dyad-level random intercept. The conditional similarity–outcome association is $\beta_1 + \beta_3 W^T_{pd}$.

For PA (H3.1): expected $\beta_3 > 0$ (we-ness amplifies positive effects). For NA (H3.2): expected $\beta_3 > 0$ (we-ness buffers negative effects, making the negative slope less negative).

## H3.1 – PA × We-ness

### Actual PA Similarity

#### Love

```{r h31_PA_act_love_neg}
#| output-fold: true
#| output-summary: "H3.1 actual PA × we-ness → Love (negative)"

h31_PA_act_love_neg <- lme(
  fixed = love ~ PA_sim_act + cwe_ness + PA_sim_act:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h31_PA_act_love_neg)
```

```{r h31_PA_act_love_pos}
#| output-fold: true
#| output-summary: "H3.1 actual PA × we-ness → Love (positive)"

h31_PA_act_love_pos <- lme(
  fixed = love ~ PA_sim_act + cwe_ness + PA_sim_act:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h31_PA_act_love_pos)
```

#### Probing (if interaction significant)

```{r h31_PA_act_probe_neg}
#| output-fold: true
#| output-summary: "Simple slopes — actual PA × we-ness → Love (negative)"

we_sd <- sd(sc2_neg$cwe_ness, na.rm = TRUE)

slopes_h31 <- emtrends(
  h31_PA_act_love_neg,
  specs = ~ cwe_ness,
  var   = "PA_sim_act",
  at    = list(cwe_ness = c(-we_sd, 0, we_sd))
)

slopes_h31_df <- as.data.frame(summary(slopes_h31, infer = TRUE))

level_labels <- c("Low (−1 SD)", "Average", "High (+1 SD)")

tbl_h31 <- data.frame(
  `We-ness level` = level_labels,
  `We-ness value` = sprintf("%.2f", slopes_h31_df$cwe_ness),
  `Simple slope`  = sprintf("%.4f", slopes_h31_df$PA_sim_act.trend),
  `SE`            = sprintf("%.4f", slopes_h31_df$SE),
  `95% CI`        = sprintf("[%.4f, %.4f]", slopes_h31_df$lower.CL, slopes_h31_df$upper.CL),
  `t`             = sprintf("%.3f", slopes_h31_df$t.ratio),
  `p`             = ifelse(slopes_h31_df$p.value < .001, "< .001",
                           sprintf("%.3f", slopes_h31_df$p.value)),
  check.names = FALSE
)

kable(tbl_h31,
  align    = c("l", "c", "c", "c", "c", "c", "c"),
  booktabs = TRUE,
  caption  = "Simple slopes of actual PA similarity on love at low, average, and high trait we-ness (negative interaction)"
) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    position   = "left"
  ) %>%
  footnote(
    general = "Simple slopes estimated at −1 SD, mean, and +1 SD of trait we-ness (grand-mean centered).",
    general_title = "Note.",
    footnote_as_chunk = TRUE
  )
```

#### Closeness

```{r h31_PA_act_close_neg}
#| output-fold: true
#| output-summary: "H3.1 actual PA × we-ness → Closeness (negative)"

h31_PA_act_close_neg <- lme(
  fixed = close ~ PA_sim_act + cwe_ness + PA_sim_act:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h31_PA_act_close_neg)
```

```{r h31_PA_act_close_pos}
#| output-fold: true
#| output-summary: "H3.1 actual PA × we-ness → Closeness (positive)"

h31_PA_act_close_pos <- lme(
  fixed = close ~ PA_sim_act + cwe_ness + PA_sim_act:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h31_PA_act_close_pos)
```

### Perceived PA Similarity

#### Love

```{r h31_PA_perc_love_neg}
#| output-fold: true
#| output-summary: "H3.1 perceived PA × we-ness → Love (negative)"

h31_PA_perc_love_neg <- lme(
  fixed = love ~ PA_sim_perc + cwe_ness + PA_sim_perc:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h31_PA_perc_love_neg)
```

```{r h31_PA_perc_love_pos}
#| output-fold: true
#| output-summary: "H3.1 perceived PA × we-ness → Love (positive)"

h31_PA_perc_love_pos <- lme(
  fixed = love ~ PA_sim_perc + cwe_ness + PA_sim_perc:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h31_PA_perc_love_pos)
```

#### Closeness

```{r h31_PA_perc_close_neg}
#| output-fold: true
#| output-summary: "H3.1 perceived PA × we-ness → Closeness (negative)"

h31_PA_perc_close_neg <- lme(
  fixed = close ~ PA_sim_perc + cwe_ness + PA_sim_perc:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h31_PA_perc_close_neg)
```

```{r h31_PA_perc_close_pos}
#| output-fold: true
#| output-summary: "H3.1 perceived PA × we-ness → Closeness (positive)"

h31_PA_perc_close_pos <- lme(
  fixed = close ~ PA_sim_perc + cwe_ness + PA_sim_perc:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h31_PA_perc_close_pos)
```

## H3.2 – NA × We-ness

### Actual NA Similarity

#### Love

```{r h32_NA_act_love_neg}
#| output-fold: true
#| output-summary: "H3.2 actual NA × we-ness → Love (negative)"

h32_NA_act_love_neg <- lme(
  fixed = love ~ NA_sim_act + cwe_ness + NA_sim_act:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h32_NA_act_love_neg)
```

```{r h32_NA_act_love_pos}
#| output-fold: true
#| output-summary: "H3.2 actual NA × we-ness → Love (positive)"

h32_NA_act_love_pos <- lme(
  fixed = love ~ NA_sim_act + cwe_ness + NA_sim_act:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h32_NA_act_love_pos)
```

#### Closeness

```{r h32_NA_act_close_neg}
#| output-fold: true
#| output-summary: "H3.2 actual NA × we-ness → Closeness (negative)"

h32_NA_act_close_neg <- lme(
  fixed = close ~ NA_sim_act + cwe_ness + NA_sim_act:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h32_NA_act_close_neg)
```

```{r h32_NA_act_close_pos}
#| output-fold: true
#| output-summary: "H3.2 actual NA × we-ness → Closeness (positive)"

h32_NA_act_close_pos <- lme(
  fixed = close ~ NA_sim_act + cwe_ness + NA_sim_act:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h32_NA_act_close_pos)
```

### Perceived NA Similarity

#### Love

```{r h32_NA_perc_love_neg}
#| output-fold: true
#| output-summary: "H3.2 perceived NA × we-ness → Love (negative)"

h32_NA_perc_love_neg <- lme(
  fixed = love ~ NA_sim_perc + cwe_ness + NA_sim_perc:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h32_NA_perc_love_neg)
```

```{r h32_NA_perc_love_pos}
#| output-fold: true
#| output-summary: "H3.2 perceived NA × we-ness → Love (positive)"

h32_NA_perc_love_pos <- lme(
  fixed = love ~ NA_sim_perc + cwe_ness + NA_sim_perc:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h32_NA_perc_love_pos)
```

#### Closeness

```{r h32_NA_perc_close_neg}
#| output-fold: true
#| output-summary: "H3.2 perceived NA × we-ness → Closeness (negative)"

h32_NA_perc_close_neg <- lme(
  fixed = close ~ NA_sim_perc + cwe_ness + NA_sim_perc:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_neg,
  method = "REML",
  na.action = na.exclude
)

model_output(h32_NA_perc_close_neg)
```

```{r h32_NA_perc_close_pos}
#| output-fold: true
#| output-summary: "H3.2 perceived NA × we-ness → Closeness (positive)"

h32_NA_perc_close_pos <- lme(
  fixed = close ~ NA_sim_perc + cwe_ness + NA_sim_perc:cwe_ness,
  random = ~ 1 | dyad,
  data = sc2_pos,
  method = "REML",
  na.action = na.exclude
)

model_output(h32_NA_perc_close_pos)
```

# Step 3b: Context Moderation — Negative vs. Positive Interaction (H3.3–H3.4)

We examine whether the similarity–evaluation association differs between the positive and negative interaction contexts. A dummy variable $C_{pdc} = c$ ($c = 0$ negative, $c = 1$ positive) is included in pooled models. Since these models contain two observations per person, both dyad-level and person-level random intercepts are included.

For actual similarity:

$$
O_{pdc} = b_{0,pd} + \beta_1 \bar{S}^{act}_{pdc} + \beta_2 C_{pdc} + \beta_3 (\bar{S}^{act}_{pdc} \times C_{pdc}) + \varepsilon_{pdc}
$$

$$
b_{0,pd} = \beta_0 + v_{0,d} + u_{0,pd}
$$

For perceived similarity: analogous model with $\bar{S}^{perc}_{pdc}$.

$\beta_1$ is the similarity–evaluation slope for the negative interaction ($C_{pdc} = 0$), and $\beta_1 + \beta_3$ is the slope for the positive interaction.

For PA (H3.3): expected $\beta_3 < 0$ (PA similarity more beneficial in negative interaction). For NA (H3.4): expected $\beta_3 < 0$ (NA similarity more harmful in positive interaction).

## H3.3 – PA × Context

### Actual PA Similarity

#### Love

```{r h33_PA_act_love}
#| output-fold: true
#| output-summary: "H3.3 actual PA × context → Love"

h33_PA_act_love <- lme(
  fixed = love ~ PA_sim_act + context + PA_sim_act:context,
  random = ~ 1 | dyad/person,
  data = sc2,
  method = "REML",
  na.action = na.exclude
)

model_output(h33_PA_act_love)
```

```{r h33_PA_act_love_slopes}
#| output-fold: true
#| output-summary: "Simple slopes — actual PA × context → Love"

# Simple slopes at negative (context = 0) and positive (context = 1) interaction
slopes_h33 <- emtrends(
  h33_PA_act_love,
  specs = ~ context,
  var   = "PA_sim_act",
  at    = list(context = c(0, 1))
)

slopes_h33_df <- as.data.frame(summary(slopes_h33, infer = TRUE))

tbl_h33 <- data.frame(
  `Interaction context` = c("Negative (0)", "Positive (1)"),
  `Simple slope`        = sprintf("%.4f", slopes_h33_df$PA_sim_act.trend),
  `SE`                  = sprintf("%.4f", slopes_h33_df$SE),
  `95% CI`              = sprintf("[%.4f, %.4f]", slopes_h33_df$lower.CL, slopes_h33_df$upper.CL),
  `t`                   = sprintf("%.3f", slopes_h33_df$t.ratio),
  `p`                   = ifelse(slopes_h33_df$p.value < .001, "< .001",
                                 sprintf("%.3f", slopes_h33_df$p.value)),
  check.names = FALSE
)

kable(tbl_h33,
  align    = c("l", "c", "c", "c", "c", "c"),
  booktabs = TRUE,
  caption  = "Simple slopes of actual PA similarity on love in negative and positive interactions (H3.3)"
) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    position   = "left"
  )
```

#### Closeness

```{r h33_PA_act_close}
#| output-fold: true
#| output-summary: "H3.3 actual PA × context → Closeness"

h33_PA_act_close <- lme(
  fixed = close ~ PA_sim_act + context + PA_sim_act:context,
  random = ~ 1 | dyad/person,
  data = sc2,
  method = "REML",
  na.action = na.exclude
)

model_output(h33_PA_act_close)
```

### Perceived PA Similarity

#### Love

```{r h33_PA_perc_love}
#| output-fold: true
#| output-summary: "H3.3 perceived PA × context → Love"

h33_PA_perc_love <- lme(
  fixed = love ~ PA_sim_perc + context + PA_sim_perc:context,
  random = ~ 1 | dyad/person,
  data = sc2,
  method = "REML",
  na.action = na.exclude
)

model_output(h33_PA_perc_love)
```

#### Closeness

```{r h33_PA_perc_close}
#| output-fold: true
#| output-summary: "H3.3 perceived PA × context → Closeness"

h33_PA_perc_close <- lme(
  fixed = close ~ PA_sim_perc + context + PA_sim_perc:context,
  random = ~ 1 | dyad/person,
  data = sc2,
  method = "REML",
  na.action = na.exclude
)

model_output(h33_PA_perc_close)
```

## H3.4 – NA × Context

### Actual NA Similarity

#### Love

```{r h34_NA_act_love}
#| output-fold: true
#| output-summary: "H3.4 actual NA × context → Love"

h34_NA_act_love <- lme(
  fixed = love ~ NA_sim_act + context + NA_sim_act:context,
  random = ~ 1 | dyad/person,
  data = sc2,
  method = "REML",
  na.action = na.exclude
)

model_output(h34_NA_act_love)
```

```{r h34_NA_act_love_slopes}
#| output-fold: true
#| output-summary: "Simple slopes — actual NA × context → Love"

slopes_h34 <- emtrends(
  h34_NA_act_love,
  specs = ~ context,
  var   = "NA_sim_act",
  at    = list(context = c(0, 1))
)

slopes_h34_df <- as.data.frame(summary(slopes_h34, infer = TRUE))

tbl_h34 <- data.frame(
  `Interaction context` = c("Negative (0)", "Positive (1)"),
  `Simple slope`        = sprintf("%.4f", slopes_h34_df$NA_sim_act.trend),
  `SE`                  = sprintf("%.4f", slopes_h34_df$SE),
  `95% CI`              = sprintf("[%.4f, %.4f]", slopes_h34_df$lower.CL, slopes_h34_df$upper.CL),
  `t`                   = sprintf("%.3f", slopes_h34_df$t.ratio),
  `p`                   = ifelse(slopes_h34_df$p.value < .001, "< .001",
                                 sprintf("%.3f", slopes_h34_df$p.value)),
  check.names = FALSE
)

kable(tbl_h34,
  align    = c("l", "c", "c", "c", "c", "c"),
  booktabs = TRUE,
  caption  = "Simple slopes of actual NA similarity on love in negative and positive interactions (H3.4)"
) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    position   = "left"
  )
```

#### Closeness

```{r h34_NA_act_close}
#| output-fold: true
#| output-summary: "H3.4 actual NA × context → Closeness"

h34_NA_act_close <- lme(
  fixed = close ~ NA_sim_act + context + NA_sim_act:context,
  random = ~ 1 | dyad/person,
  data = sc2,
  method = "REML",
  na.action = na.exclude
)

model_output(h34_NA_act_close)
```

### Perceived NA Similarity

#### Love

```{r h34_NA_perc_love}
#| output-fold: true
#| output-summary: "H3.4 perceived NA × context → Love"

h34_NA_perc_love <- lme(
  fixed = love ~ NA_sim_perc + context + NA_sim_perc:context,
  random = ~ 1 | dyad/person,
  data = sc2,
  method = "REML",
  na.action = na.exclude
)

model_output(h34_NA_perc_love)
```

#### Closeness

```{r h34_NA_perc_close}
#| output-fold: true
#| output-summary: "H3.4 perceived NA × context → Closeness"

h34_NA_perc_close <- lme(
  fixed = close ~ NA_sim_perc + context + NA_sim_perc:context,
  random = ~ 1 | dyad/person,
  data = sc2,
  method = "REML",
  na.action = na.exclude
)

model_output(h34_NA_perc_close)
```
