---
title: "Exploratory Extension: Response Surface Analysis of Affect Similarity (ESM)"
---

```{r setup}
#| include: false

# 1. Load packages, paths and data configurations
source(here::here("R", "00_setup.R"))

# 2. Set a global theme for plots
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

```{r RSA_setup}
#| include: false
# load data
esm_data  <- readRDS(file.path(dir_data_ana, "esm_ana.rds"))
# load data configuration file 
load_config("ESM")
# plot initialisation
plot_counter <- 1
prefix <- "RSA"
```

# Introduction

This document presents an **exploratory extension** to Study Component 1 (`06_study_component_1.qmd`). It is not part of the preregistered analysis plan and should be read as a complementary perspective rather than a confirmatory test.

In our preregistered analyses, we operationalized affect similarity as a single composite index:

$$
S_{mpd} = \underbrace{\frac{X_{mpd} + Y_{mpd}}{2}}_{\text{elevation}} - \underbrace{|X_{mpd} - Y_{mpd}|}_{\text{distance}}
$$

where $X$ is the actor's person-mean centered affect and $Y$ is the partner's (actual or perceived) person-mean centered affect. This index captures the idea that similarity means shared, elevated affect, rewarding co-experienced high affect and penalizing mismatch. However, this operationalization collapses a fundamentally two-dimensional predictor space (actor affect $\times$ partner affect) into a single number. In doing so, it imposes specific assumptions about the functional form of the relationship between individual affect levels and love.

**Response surface analysis** (RSA; @edwardsAlternativesDifferenceScores2002; @nestlerResponseSurfaceAnalysis2019) offers a complementary approach. Rather than constructing a similarity score upfront, RSA fits a second-degree polynomial regression with both individual affect levels as separate predictors and then analyzes the shape of the resulting three-dimensional response surface. This allows the data to reveal whether love is maximized along the line where actor and partner affect match (a congruence pattern), whether higher shared levels additionally matter (a rising ridge), or whether some other, more complex pattern emerges, without imposing the specific constraints baked into the composite similarity index.

## Why RSA as an exploratory complement?

Our composite similarity index served the preregistered hypotheses well, but it makes two implicit constraints that RSA can relax:

1.  **Symmetry of mismatch.** The absolute difference $|X - Y|$ treats actor-exceeds-partner and partner-exceeds-actor mismatches identically. RSA can detect asymmetric effects (e.g., love decreasing more steeply when the partner's affect exceeds the actor's than vice versa).

2.  **Fixed functional form.** The composite index imposes a specific nonlinear mapping from $(X, Y)$ to a single similarity score. RSA lets the polynomial regression freely estimate how combinations of actor and partner affect relate to love, and then inspects the resulting surface.

RSA is not a replacement for our preregistered approach, it answers a related but distinct question. Where our similarity index asks "does our specific operationalization of co-experienced elevated affect predict love?", RSA asks "what does the full landscape of the actor affect $\times$ partner affect association with love look like, and is it consistent with a congruence pattern?"

# Methodology

## The polynomial regression model

RSA begins with a second-degree polynomial regression. For moment $m$, person $p$ in dyad $d$:

$$
L_{mpd} = b_0 + b_1 X_{mpd} + b_2 Y_{mpd} + b_3 X^2_{mpd} + b_4 X_{mpd} Y_{mpd} + b_5 Y^2_{mpd} + \varepsilon_{mpd}
$$

where $L$ is momentary love, $X$ is the actor's person-mean centered affect, and $Y$ is the partner's person-mean centered affect (actual or perceived). The five predictors capture linear effects of each partner's affect ($b_1$, $b_2$), curvilinear effects ($b_3$, $b_5$), and their interaction ($b_4$).

This produces a three-dimensional response surface with actor affect on the $x$-axis, partner affect on the $y$-axis, and predicted love on the $z$-axis.

## Multilevel extension

Following @nestlerResponseSurfaceAnalysis2019, we embed this polynomial regression in our three-level structure (moments within persons within dyads). We fit the model with `nlme::lme()` using the same specification conventions as in SC1:

-   Random intercepts for dyads and persons within dyads
-   Continuous-time AR(1) residual correlation structure
-   Person-mean centered predictors (the `c*` variables already constructed in `04_data_construct.R`)

The fixed effects $\hat{b}_1$ through $\hat{b}_5$ define the **average response surface** across all persons and dyads.

### A note on centering

@nestlerResponseSurfaceAnalysis2019 recommend centering both predictor variables on their combined (pooled) grand mean to ensure a shared zero point. In our data, actor and partner affect variables are already person-mean centered separately (each relative to its own person mean). Since both are measured on the same 0–100 scale, the person means are comparable and the zero points approximately aligned. For this exploratory analysis, we use the existing person-mean centered variables as-is.

## Response surface parameters

The polynomial coefficients can be transformed into five **response surface parameters** that describe interpretable features of the surface:

### Line of Congruence (LOC)

The LOC traces the surface along the diagonal where actor and partner affect are equal ($X = Y$):

$$
\hat{L} = \hat{b}_0 + a_1 X + a_2 X^2
$$

with

$$
\begin{aligned}
a_1 &= b_1 + b_2 \quad &\text{(slope of the LOC at the origin)} \\
a_2 &= b_3 + b_4 + b_5 \quad &\text{(curvature of the LOC)}
\end{aligned}
$$

The LOC tells us what happens to predicted love as both partners' affect increases together. A positive $a_1$ indicates a "rising ridge": love is higher when both partners share high affect than when both share low affect.

### Line of Incongruence (LOIC)

The LOIC is orthogonal to the LOC and passes through the origin ($X = -Y$):

$$
\hat{L} = \hat{b}_0 + a_3 X + a_4 X^2
$$

with

$$
\begin{aligned}
a_3 &= b_1 - b_2 \quad &\text{(slope of the LOIC at the origin)} \\
a_4 &= b_3 - b_4 + b_5 \quad &\text{(curvature of the LOIC)}
\end{aligned}
$$

The LOIC tells us what happens when the partners diverge. A negative $a_4$ (inverted-U/concave curvature) means love decreases as actor and partner affect diverge from each other.

### First Principal Axis (FPA) condition

$$
a_5 = b_3 - b_5
$$

When $a_5 = 0$, the ridge of the surface (the FPA) lies exactly on the LOC ($X = Y$). If $a_5 \neq 0$, the ridge is shifted, meaning the "optimal" partner affect for maximizing love is not equal to the actor's own affect.

## Interpreting surface patterns

A **strict congruence effect** requires five conditions simultaneously (Edwards, 2002; Humberg et al., 2019; Nestler et al., 2019):

| Condition | Meaning                                                        |
|---------------|---------------------------------------------------------|
| $a_1 = 0$ | LOC is flat: congruence at high vs. low levels equally good    |
| $a_2 = 0$ | LOC is linear (no concavity/convexity)                         |
| $a_3 = 0$ | LOIC is symmetric around the origin/LOC                        |
| $a_4 < 0$ | Inverted-U/concavity on LOIC: love drops when partners diverge |
| $a_5 = 0$ | FPA aligns with the LOC                                        |

In practice, a **rising ridge** pattern ($a_1 > 0$, $a_4 < 0$) in affect would mean: congruence matters, but shared high affect is better than shared low affect. This pattern would align well with the elevation component of our preregistered similarity index.

## Expectations from this exploration

Translating the preregistered hypotheses into RSA terms:

**PA similarity (H1.1/H1.3).** Our preregistered index captures both elevation and congruence. In RSA, we would expect a **rising ridge** pattern: a positive $a_1$ (love increases along the LOC, i.e., shared high PA is better than shared low PA) combined with a negative $a_4$ (love drops when partners' PA diverges). This would confirm that the similarity index's two components — shared elevation and low mismatch — both contribute independently.

**NA similarity (H1.2/H1.4).** The preregistered hypothesis predicted that higher NA similarity is associated with lower love (within dyads) — but the SC1 results showed the opposite at the within-dyad level. In RSA terms, the within-dyad finding ($\beta_1 = -0.15$, $p < .001$) suggests a pattern where shared negative affect relates to lower love. We might expect a negative $a_1$ (love decreases along the LOC as shared NA rises). Whether a congruence effect ($a_4 < 0$) emerges for NA is an open question: it would mean that NA mismatch is worse for love than matched NA levels, regardless of the direction.

# Data Preparation

## Constructing polynomial terms

The person-mean centered affect variables are already available in the data (see `04_data_construct.R`). We construct the quadratic and interaction terms from these centered predictors.

```{r rsa_data_prep}
esm_data <- esm_data %>% arrange(dyad, person, time)

# --- Actual affect: actor (X) and partner's self-report (Y) ---
esm_data <- esm_data %>%
  mutate(
    # PA actual
    PA_own_sq      = cPA_own^2,
    PA_part_act_sq = cPA_part_act^2,
    PA_own_x_part  = cPA_own * cPA_part_act,
    # NA actual
    NA_own_sq      = cNA_own^2,
    NA_part_act_sq = cNA_part_act^2,
    NA_own_x_part  = cNA_own * cNA_part_act
  )

# --- Perceived affect: actor (X) and actor's perception of partner (Y) ---
esm_data <- esm_data %>%
  mutate(
    # PA perceived
    PA_part_perc_sq  = cPA_part_perc^2,
    PA_own_x_perc    = cPA_own * cPA_part_perc,
    # NA perceived
    NA_part_perc_sq  = cNA_part_perc^2,
    NA_own_x_perc    = cNA_own * cNA_part_perc
  )
```

## Helper functions

```{r rsa_helpers}
# --- Compute RSA surface parameters from polynomial coefficients ---
rsa_params <- function(model) {
  b  <- fixef(model)
  # b[1] = intercept, b[2] = b1 (X), b[3] = b2 (Y), 
  # b[4] = b3 (X²), b[5] = b4 (XY), b[6] = b5 (Y²)
  b1 <- b[2]; b2 <- b[3]; b3 <- b[4]; b4 <- b[5]; b5 <- b[6]
  
  a1 <- b1 + b2       # LOC slope
  a2 <- b3 + b4 + b5  # LOC curvature
  a3 <- b1 - b2       # LOIC slope
  a4 <- b3 - b4 + b5  # LOIC curvature
  a5 <- b3 - b5       # FPA = LOC condition
  
  # Standard errors via delta method (variance of linear combos)
  V <- vcov(model)
  # indices 2:6 correspond to b1..b5
  se_a1 <- sqrt(V[2,2] + V[3,3] + 2*V[2,3])
  se_a2 <- sqrt(V[4,4] + V[5,5] + V[6,6] + 2*V[4,5] + 2*V[4,6] + 2*V[5,6])
  se_a3 <- sqrt(V[2,2] + V[3,3] - 2*V[2,3])
  se_a4 <- sqrt(V[4,4] + V[5,5] + V[6,6] - 2*V[4,5] + 2*V[4,6] - 2*V[5,6])
  se_a5 <- sqrt(V[4,4] + V[6,6] - 2*V[4,6])
  
  params <- data.frame(
    Parameter = c("a1 (LOC slope)", "a2 (LOC curvature)", 
                   "a3 (LOIC slope)", "a4 (LOIC curvature)", 
                   "a5 (FPA = LOC)"),
    Meaning   = c("Slope along X = Y at origin",
                   "Curvature along X = Y",
                   "Slope along X = −Y at origin",
                   "Curvature along X = −Y",
                   "Ridge displacement from LOC"),
    Estimate = c(a1, a2, a3, a4, a5),
    SE       = c(se_a1, se_a2, se_a3, se_a4, se_a5),
    stringsAsFactors = FALSE
  )
  params$z    <- params$Estimate / params$SE
  params$p    <- 2 * pnorm(-abs(params$z))
  params$sig  <- ifelse(params$p < .001, "***",
                 ifelse(params$p < .01,  "**",
                 ifelse(params$p < .05,  "*", "")))
  params
}

# --- 3D response surface plot with plotly ---
plot_rsa_surface <- function(model, data, x_var, y_var,
                              x_lab = "Actor affect (centered)",
                              y_lab = "Partner affect (centered)",
                              title = "Response Surface",
                              z_lab = "Momentary love",
                              z_lim = c(0, 100),
                              x_range = c(-95, 95),
                              y_range = c(-95, 95)) {
  
  # ---- fixed effects: intercept, x, y, x^2, x:y, y^2 ----
  b <- as.numeric(nlme::fixef(model))
  if (length(b) < 6) {
    stop("fixef(model) must return at least 6 coefficients ordered as: intercept, x, y, x^2, x:y, y^2")
  }
  b <- b[1:6]
  
  poly_fun <- function(x, y) {
    b[1] + b[2] * x + b[3] * y + b[4] * x^2 + b[5] * x * y + b[6] * y^2
  }
  
  # ---- plotting ranges ----
  x_range <- x_range
  y_range <- y_range
  
  n_grid <- 300L
  x_seq  <- seq(x_range[1], x_range[2], length.out = n_grid)
  y_seq  <- seq(y_range[1], y_range[2], length.out = n_grid)
  
  Xmat <- outer(x_seq, y_seq, function(x, y) x)
  Ymat <- outer(x_seq, y_seq, function(x, y) y)
  Zmat <- poly_fun(Xmat, Ymat)
  
  vx <- as.vector(Xmat)
  vy <- as.vector(Ymat)
  vz <- as.vector(Zmat)
  
  vid <- function(ix, iy) ix + (iy - 1L) * n_grid
  
  # ---- diverging palette (fixed to love's natural scale, not per-plot z-range) ----
  # Using z_lim as the colour anchor means all four response surfaces share a
  # comparable colour mapping — a blue patch on one plot represents the same
  # predicted love as a blue patch on another.
  ramp_lo <- grDevices::colorRamp(c("#2166ac", "#f7f7f7"))
  ramp_hi <- grDevices::colorRamp(c("#f7f7f7", "#b2182b"))
  
  z_to_col <- function(z) {
    u <- (z - z_lim[1]) / (z_lim[2] - z_lim[1])
    u <- pmin(pmax(u, 0), 1)
    
    out <- character(length(u))
    lo  <- u <= 0.5
    hi  <- !lo
    
    if (any(lo)) {
      rgb_lo <- ramp_lo(u[lo] / 0.5)
      out[lo] <- grDevices::rgb(rgb_lo[, 1], rgb_lo[, 2], rgb_lo[, 3],
                                maxColorValue = 255)
    }
    if (any(hi)) {
      rgb_hi <- ramp_hi((u[hi] - 0.5) / 0.5)
      out[hi] <- grDevices::rgb(rgb_hi[, 1], rgb_hi[, 2], rgb_hi[, 3],
                                maxColorValue = 255)
    }
    out
  }
  
  # ---- explicit triangulation ----
  n_faces <- 2L * (n_grid - 1L) * (n_grid - 1L)
  ii <- integer(n_faces); jj <- integer(n_faces); kk <- integer(n_faces)
  face_x <- numeric(n_faces); face_y <- numeric(n_faces); face_z <- numeric(n_faces)
  loc_hit  <- logical(n_faces)
  loic_hit <- logical(n_faces)
  
  pos <- 1L
  for (iy in 1:(n_grid - 1L)) {
    for (ix in 1:(n_grid - 1L)) {
      v00 <- vid(ix, iy)
      v10 <- vid(ix + 1L, iy)
      v11 <- vid(ix + 1L, iy + 1L)
      v01 <- vid(ix, iy + 1L)
      
      # triangle 1: (v00, v10, v11)
      tri1 <- c(v00, v10, v11)
      x1 <- vx[tri1]; y1 <- vy[tri1]; z1 <- vz[tri1]
      ii[pos] <- tri1[1] - 1L; jj[pos] <- tri1[2] - 1L; kk[pos] <- tri1[3] - 1L
      face_x[pos] <- mean(x1); face_y[pos] <- mean(y1); face_z[pos] <- mean(z1)
      loc_hit[pos]  <- (min(x1 - y1) <= 0) && (max(x1 - y1) >= 0)
      loic_hit[pos] <- (min(x1 + y1) <= 0) && (max(x1 + y1) >= 0)
      pos <- pos + 1L
      
      # triangle 2: (v00, v11, v01)
      tri2 <- c(v00, v11, v01)
      x2 <- vx[tri2]; y2 <- vy[tri2]; z2 <- vz[tri2]
      ii[pos] <- tri2[1] - 1L; jj[pos] <- tri2[2] - 1L; kk[pos] <- tri2[3] - 1L
      face_x[pos] <- mean(x2); face_y[pos] <- mean(y2); face_z[pos] <- mean(z2)
      loc_hit[pos]  <- (min(x2 - y2) <= 0) && (max(x2 - y2) >= 0)
      loic_hit[pos] <- (min(x2 + y2) <= 0) && (max(x2 + y2) >= 0)
      pos <- pos + 1L
    }
  }
  
  face_col <- z_to_col(face_z)
  
  # ---- LOIC dashed pattern ----
  loic_tmin <- max(x_range[1], -y_range[2])
  loic_tmax <- min(x_range[2], -y_range[1])
  loic_span <- loic_tmax - loic_tmin
  
  if (is.finite(loic_span) && loic_span > 0) {
    t_loic   <- (face_x - face_y) / 2
    dash_len <- loic_span / 30
    gap_len  <- dash_len * 0.4
    phase    <- (t_loic - loic_tmin) %% (dash_len + gap_len)
    loic_on  <- loic_hit & (phase <= dash_len)
  } else {
    loic_on <- rep(FALSE, length(face_col))
  }
  
  # paint LOC/LOIC into the face colour vector
  face_col[loic_on] <- "#000000"
  face_col[loc_hit] <- "#000000"
  
  # ---- observed data points (subsample for rendering speed) ----
  obs <- data %>%
    dplyr::select(x = dplyr::all_of(x_var),
                  y = dplyr::all_of(y_var),
                  z = love) %>%
    dplyr::filter(stats::complete.cases(.)) %>%
    dplyr::filter(x >= x_range[1], x <= x_range[2],
                  y >= y_range[1], y <= y_range[2])
  if (nrow(obs) > 2000) obs <- obs %>% dplyr::slice_sample(n = 2000)
  
  # ---- colorbar trace: invisible scatter3d carrying the colour scale ----
  # mesh3d with facecolor can't display a colorbar, so we attach one to a
  # hidden marker trace placed off-screen. The colorscale here must match the
  # z_to_col() mapping above so the legend reflects what the surface shows.
  cbar_scale <- list(
    c(0.0, "#2166ac"),
    c(0.5, "#f7f7f7"),
    c(1.0, "#b2182b")
  )
  
  plotly::plot_ly() %>%
    # main surface
    plotly::add_trace(
      type = "mesh3d",
      x = vx, y = vy, z = vz,
      i = ii, j = jj, k = kk,
      facecolor = I(face_col),
      flatshading = FALSE,
      opacity = 0.65,
      # hoverinfo = "skip",
      lighting = list(
        ambient = 0.85, diffuse = 0.7, specular = 0.03,
        roughness = 0.9, fresnel = 0.02
      ),
      lightposition = list(x = 100, y = 200, z = 0),
      name = "Surface",
      showlegend = FALSE
    ) %>%
    # observed data points
    plotly::add_markers(
      x = obs$x, y = obs$y, z = obs$z,
      marker = list(size = 1.5, color = "black", opacity = 0.25),
      name = "Observed",
      # hoverinfo = "skip",
      showlegend = FALSE
    ) %>%
    # hidden colorbar trace
    plotly::add_trace(
      type = "scatter3d",
      mode = "markers",
      x = c(x_range[1], x_range[1]),
      y = c(y_range[1], y_range[1]),
      z = z_lim,
      marker = list(
        size = 0.1,
        color = z_lim,
        colorscale = cbar_scale,
        cmin = z_lim[1],
        cmax = z_lim[2],
        showscale = TRUE,
        colorbar = list(
          title = list(text = "Predicted\nlove"),
          len = 0.7
        )
      ),
      hoverinfo = "skip",
      showlegend = FALSE
    ) %>%
    plotly::layout(
      title = list(text = title),
      scene = list(
        xaxis = list(title = x_lab),
        yaxis = list(title = y_lab),
        zaxis = list(title = z_lab, range = z_lim),
        camera = list(eye = list(x = 1.5, y = -1.5, z = 1.2))
      ),
      annotations = list(
        list(
          x = 0.01, y = 0.99, xref = "paper", yref = "paper",
          text = "<b>LOC</b>: solid black&nbsp;&nbsp;&nbsp;<b>LOIC</b>: dashed",
          showarrow = FALSE, xanchor = "left", yanchor = "top",
          bgcolor = "rgba(255,255,255,0.75)",
          bordercolor = "rgba(0,0,0,0.15)",
          borderwidth = 1
        )
      )
    )
}

# --- 2D contour plot as a static alternative ---
plot_rsa_contour <- function(model, data, x_var, y_var,
                              x_lab = "Actor affect (centered)",
                              y_lab = "Partner affect (centered)",
                              title = "Response Surface (contour)") {
  
  b <- fixef(model)
  
  x_range <- c(-50, 50)
  y_range <- c(-50, 50)
  
  grid <- expand.grid(
    X = seq(x_range[1], x_range[2], length.out = 100),
    Y = seq(y_range[1], y_range[2], length.out = 100)
  )
  
  grid$Z <- b[1] + b[2]*grid$X + b[3]*grid$Y + 
            b[4]*grid$X^2 + b[5]*grid$X*grid$Y + b[6]*grid$Y^2
  
  ggplot(grid, aes(x = X, y = Y, z = Z)) +
    geom_contour_filled(alpha = 0.8) +
    geom_abline(intercept = 0, slope = 1, linetype = "solid", colour = "black") +
    geom_abline(intercept = 0, slope = -1, linetype = "dashed", colour = "black") +
    annotate("text", x = x_range[2]*0.85, y = x_range[2]*0.85, 
             label = "LOC (X = Y)", angle = 45, size = 3, colour = "black") +
    annotate("text", x = x_range[2]*0.85, y = -x_range[2]*0.85, 
             label = "LOIC (X = −Y)", angle = -45, size = 3, colour = "black") +
    labs(x = x_lab, y = y_lab, fill = "Predicted\nlove", title = title) +
    coord_fixed() +
    theme_minimal()
}
```

# Model 1: Actual PA — Response Surface

## Model specification

$$
L_{mpd} = b_{0,pd} + b_1 \, cPA^{own}_{mpd} + b_2 \, cPA^{part,act}_{mpd} + b_3 \, (cPA^{own}_{mpd})^2 + b_4 \, cPA^{own}_{mpd} \cdot cPA^{part,act}_{mpd} + b_5 \, (cPA^{part,act}_{mpd})^2 + \varepsilon_{mpd}
$$

with random intercepts $b_{0,pd} = \beta_0 + v_{0,d} + u_{0,pd}$ and continuous-time AR(1) residuals.

**Expectation:** Given that our preregistered index found a significant within-dyad effect of actual PA similarity on love ($\beta = 0.113$, $p < .001$), we expect the response surface to show both a rising ridge ($a_1 > 0$: shared high PA predicts more love) and a congruence component ($a_4 < 0$: PA mismatch reduces love). The rising ridge would correspond to the elevation component and the inverted-U LOIC to the distance penalty in our composite index.

```{r rsa_PA_act}
#| output-fold: true
#| output-summary: "Model 1: Actual PA polynomial regression"

rsa_PA_act <- lme(
  fixed = love ~ cPA_own + cPA_part_act + PA_own_sq + PA_own_x_part + PA_part_act_sq,
  random = ~ 1 | dyad/person,
  correlation = corCAR1(form = ~ time | dyad/person),
  data = esm_data,
  method = "REML",
  na.action = na.exclude,
  control = lmeControl(opt = "nlminb")
)

model_output(rsa_PA_act)
```

## Surface parameters

```{r rsa_PA_act_params}
params_PA_act <- rsa_params(rsa_PA_act)
kable(params_PA_act, digits = 3, caption = "RSA parameters — Actual PA → Love") %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)
```

## 3D response surface

```{r rsa_PA_act_3d}
#| fig-cap: "Response surface: actual PA (actor × partner) → love. Black dots are observed data points (random subsample). Dashed line on contour plot = LOC (X = Y)."

plot_rsa_surface(
  rsa_PA_act, esm_data, "cPA_own", "cPA_part_act",
  x_lab = "Actor PA (person-centered)",
  y_lab = "Partner PA (person-centered)",
  title = "Actual PA → Love"
)

plot_rsa_surface(
  rsa_PA_act, esm_data, "cPA_own", "cPA_part_act",
  x_lab = "Actor PA (person-centered)",
  y_lab = "Partner PA (person-centered)",
  title = "Actual PA → Love",
  x_range = c(-30,30),
  y_range = c(-30,30)
)
```

```{r rsa_PA_act_contour}
#| fig-cap: "Contour projection of the actual PA response surface. Dashed = LOC; dotted = LOIC."

plot_rsa_contour(
  rsa_PA_act, esm_data, "cPA_own", "cPA_part_act",
  x_lab = "Actor PA (person-centered)",
  y_lab = "Partner PA (person-centered)",
  title = "Actual PA → Love (contour)"
)
```

# Model 2: Actual NA — Response Surface

## Model specification

$$
L_{mpd} = b_{0,pd} + b_1 \, cNA^{own}_{mpd} + b_2 \, cNA^{part,act}_{mpd} + b_3 \, (cNA^{own}_{mpd})^2 + b_4 \, cNA^{own}_{mpd} \cdot cNA^{part,act}_{mpd} + b_5 \, (cNA^{part,act}_{mpd})^2 + \varepsilon_{mpd}
$$

**Expectation:** The SC1 results for H1.2 showed a negative within-dyad association between NA similarity and love ($\beta_1 = -0.15$, $p < .001$) and a positive between-dyad association ($\beta_2 = 0.62$, $p < .01$). Since this model estimates within-person effects, we focus on the within-person pattern. The negative within-dyad finding suggests $a_1 < 0$ (love decreasing as both partners' NA rises together). Whether a congruence pattern ($a_4 < 0$) emerges is less clear: the composite index's mismatch penalty may or may not correspond to a clean inverted-U on the LOIC when modeled as a free polynomial.

```{r rsa_NA_act}
#| output-fold: true
#| output-summary: "Model 2: Actual NA polynomial regression"

rsa_NA_act <- lme(
  fixed = love ~ cNA_own + cNA_part_act + NA_own_sq + NA_own_x_part + NA_part_act_sq,
  random = ~ 1 | dyad/person,
  correlation = corCAR1(form = ~ time | dyad/person),
  data = esm_data,
  method = "REML",
  na.action = na.exclude,
  control = lmeControl(opt = "nlminb")
)

model_output(rsa_NA_act)
```

## Surface parameters

```{r rsa_NA_act_params}
params_NA_act <- rsa_params(rsa_NA_act)
kable(params_NA_act, digits = 3, caption = "RSA parameters — Actual NA → Love") %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)
```

## 3D response surface

```{r rsa_NA_act_3d}
plot_rsa_surface(
  rsa_NA_act, esm_data, "cNA_own", "cNA_part_act",
  x_lab = "Actor NA (person-centered)",
  y_lab = "Partner NA (person-centered)",
  title = "Actual NA → Love"
)
```

```{r rsa_NA_act_contour}
plot_rsa_contour(
  rsa_NA_act, esm_data, "cNA_own", "cNA_part_act",
  x_lab = "Actor NA (person-centered)",
  y_lab = "Partner NA (person-centered)",
  title = "Actual NA → Love (contour)"
)
```

# Model 3: Perceived PA — Response Surface

## Model specification

$$
L_{mpd} = b_{0,pd} + b_1 \, cPA^{own}_{mpd} + b_2 \, cPA^{part,perc}_{mpd} + b_3 \, (cPA^{own}_{mpd})^2 + b_4 \, cPA^{own}_{mpd} \cdot cPA^{part,perc}_{mpd} + b_5 \, (cPA^{part,perc}_{mpd})^2 + \varepsilon_{mpd}
$$

**Expectation:** The preregistered H1.3 tested perceived PA similarity as a mediator. The direct effect of perceived PA similarity on love was significant within persons ($\beta = 0.11$, $p < .001$) in the Step 1a models. In RSA terms, we would again expect a rising ridge ($a_1 > 0$), potentially with a congruence component ($a_4 < 0$). If the perceived similarity surface closely resembles the actual similarity surface, this would suggest that perceived and actual congruence tap into the same underlying process.

```{r rsa_PA_perc}
#| output-fold: true
#| output-summary: "Model 3: Perceived PA polynomial regression"

rsa_PA_perc <- lme(
  fixed = love ~ cPA_own + cPA_part_perc + PA_own_sq + PA_own_x_perc + PA_part_perc_sq,
  random = ~ 1 | dyad/person,
  correlation = corCAR1(form = ~ time | dyad/person),
  data = esm_data,
  method = "REML",
  na.action = na.exclude,
  control = lmeControl(opt = "nlminb")
)

model_output(rsa_PA_perc)
```

## Surface parameters

```{r rsa_PA_perc_params}
params_PA_perc <- rsa_params(rsa_PA_perc)
kable(params_PA_perc, digits = 3, caption = "RSA parameters — Perceived PA → Love") %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)
```

## 3D response surface

```{r rsa_PA_perc_3d}
plot_rsa_surface(
  rsa_PA_perc, esm_data, "cPA_own", "cPA_part_perc",
  x_lab = "Actor PA (person-centered)",
  y_lab = "Perceived partner PA (person-centered)",
  title = "Perceived PA → Love"
)
```

```{r rsa_PA_perc_contour}
plot_rsa_contour(
  rsa_PA_perc, esm_data, "cPA_own", "cPA_part_perc",
  x_lab = "Actor PA (person-centered)",
  y_lab = "Perceived partner PA (person-centered)",
  title = "Perceived PA → Love (contour)"
)
```

# Model 4: Perceived NA — Response Surface

## Model specification

$$
L_{mpd} = b_{0,pd} + b_1 \, cNA^{own}_{mpd} + b_2 \, cNA^{part,perc}_{mpd} + b_3 \, (cNA^{own}_{mpd})^2 + b_4 \, cNA^{own}_{mpd} \cdot cNA^{part,perc}_{mpd} + b_5 \, (cNA^{part,perc}_{mpd})^2 + \varepsilon_{mpd}
$$

**Expectation:** As with Model 2, the NA surface may show a descending ridge ($a_1 < 0$) reflecting that shared high NA predicts lower love. The congruence component is exploratory: it is unclear whether a clean inverted-U on the LOIC will emerge for perceived NA, particularly given that the perceived NA similarity effect in the preregistered analyses was significant within persons.

```{r rsa_NA_perc}
#| output-fold: true
#| output-summary: "Model 4: Perceived NA polynomial regression"

rsa_NA_perc <- lme(
  fixed = love ~ cNA_own + cNA_part_perc + NA_own_sq + NA_own_x_perc + NA_part_perc_sq,
  random = ~ 1 | dyad/person,
  correlation = corCAR1(form = ~ time | dyad/person),
  data = esm_data,
  method = "REML",
  na.action = na.exclude,
  control = lmeControl(opt = "nlminb")
)

model_output(rsa_NA_perc)
```

## Surface parameters

```{r rsa_NA_perc_params}
params_NA_perc <- rsa_params(rsa_NA_perc)
kable(params_NA_perc, digits = 3, caption = "RSA parameters — Perceived NA → Love") %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)
```

## 3D response surface

```{r rsa_NA_perc_3d}
plot_rsa_surface(
  rsa_NA_perc, esm_data, "cNA_own", "cNA_part_perc",
  x_lab = "Actor NA (person-centered)",
  y_lab = "Perceived partner NA (person-centered)",
  title = "Perceived NA → Love"
)
```

```{r rsa_NA_perc_contour}
plot_rsa_contour(
  rsa_NA_perc, esm_data, "cNA_own", "cNA_part_perc",
  x_lab = "Actor NA (person-centered)",
  y_lab = "Perceived partner NA (person-centered)",
  title = "Perceived NA → Love (contour)"
)
```

# Summary Comparison

## RSA parameters across models

```{r rsa_summary_table}
summary_tbl <- bind_rows(
  params_PA_act  %>% mutate(Model = "Actual PA"),
  params_NA_act  %>% mutate(Model = "Actual NA"),
  params_PA_perc %>% mutate(Model = "Perceived PA"),
  params_NA_perc %>% mutate(Model = "Perceived NA")
) %>%
  select(Model, Parameter, Estimate, SE, z, p, sig)

kable(summary_tbl, digits = 3, 
      caption = "RSA surface parameters across all four models") %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE) %>%
  pack_rows(index = c("Actual PA" = 5, "Actual NA" = 5, 
                       "Perceived PA" = 5, "Perceived NA" = 5))
```

## Connecting RSA to the preregistered similarity index

The response surfaces above provide a complementary lens on the same data analyzed in Study Component 1. A few points of connection:

**Elevation vs. congruence decomposition.** Where the preregistered similarity index collapses elevation and mismatch into a single score, RSA separates them. The LOC parameters ($a_1$, $a_2$) capture the elevation component, how love changes as both partners' affect rises together. The LOIC parameters ($a_3$, $a_4$) capture the congruence component, how love changes as partners diverge. If both the LOC and LOIC contribute to the surface shape, this supports the idea that our composite index captures two meaningfully distinct processes.

**Asymmetric mismatch.** The composite similarity index assumes symmetric mismatch penalties ($|X - Y|$). If the RSA surface is asymmetric around the LOC (visible as $a_3 \neq 0$ or an asymmetric contour pattern), this suggests that actor-exceeds-partner and partner-exceeds-actor mismatches have different implications for love.

**Nonlinearity.** The polynomial model can capture curvilinear effects ($a_2 \neq 0$) that the preregistered linear models cannot. A dome-shaped surface ($a_2 < 0$) would suggest diminishing returns of shared elevated affect — shared PA above some threshold does not further increase love.

# Limitations of This Exploration

This analysis is exploratory and should be interpreted with appropriate caution:

1.  **Not preregistered.** These models were not part of the analysis plan and inflate the number of statistical tests conducted on this dataset.

2.  **Within-person focus.** Because person-mean centering removes between-person variance, the RSA surfaces reflect within-person (momentary) associations only. The between-dyad effects examined in SC1 are not captured here. A full multilevel RSA following Nestler et al. (2019) would include group means and their polynomial terms as Level 2 predictors to separate within and between effects — a possible extension for future work.

3.  **Centering approximation.** Separate person-mean centering (rather than pooled centering) is an approximation. For commensurable scales measured on the same VAS, this is acceptable for exploratory purposes.

4.  **No random slopes.** We fit random intercepts only, consistent with SC1. Adding random slopes for the five polynomial terms (as Nestler et al. recommend for the full approach) would quickly exhaust the data, given our sample size.

5.  **Polynomial boundary.** Second-degree polynomials can produce implausible predictions (e.g., negative love or values $> 100$) in regions of the predictor space with sparse data. The 3D plots are trimmed to the central 98% of the observed data to mitigate this, but extrapolation beyond the data range should not be trusted.

# References

-   Edwards, J. R. (2002). Alternatives to difference scores: Polynomial regression analysis and response surface methodology. In F. Drasgow & N. Schmitt (Eds.), *Measuring and analyzing behavior in organizations* (pp. 350–400). Jossey-Bass.
-   Humberg, S., Nestler, S., & Back, M. D. (2019). Response surface analysis in personality and social psychology: Checklist and clarifications for the case of congruence hypotheses. *Social Psychological and Personality Science*, *10*(3), 409–419.
-   Nestler, S., Humberg, S., & Schönbrodt, F. D. (2019). Response surface analysis with multilevel data: Illustration for the case of congruence hypotheses. *Psychological Methods*, *24*(3), 291–308.
-   Shanock, L. R., Baran, B. E., Gentry, W. A., Pattison, S. C., & Heggestad, E. D. (2010). Polynomial regression with response surface analysis: A powerful approach for examining moderation and overcoming limitations of difference scores. *Journal of Business and Psychology*, *25*, 543–554.
