---
title: "Manuscript Tables"
---

```{r setup}
#| include: false

source(here::here("R", "00_setup.R"))
# dir_tables is defined and created by R/01_paths.R (sourced above)

# Change MS_FONT to match whatever font you use in Word (e.g. "Calibri")
MS_FONT      <- "Times New Roman"
MS_SIZE_BODY <- 11
MS_SIZE_NOTE <-  9

rule_thick <- fp_border(width = 1.5)
rule_thin  <- fp_border(width = 0.75)
```

# SC1 Step 1 Tables

## Main and Additive Effects (H1.1–H1.3)

```{r sc1-step1-data}
#| include: false

esm_data <- readRDS(file.path(dir_data_ana, "esm_ana.rds"))
load_config("ESM")
esm_data <- esm_data %>% arrange(dyad, person, time)

sc1_ctrl <- lmeControl(opt = "nlminb")
sc1_cor  <- corCAR1(form = ~ time | dyad/person)

# ── Step 1a: actual similarity only (H1.1 PA, H1.2 NA) ──
h11 <- lme(love ~ PA_similarity_act_w + PA_similarity_act_b,
           random = ~1|dyad/person, correlation = sc1_cor,
           data = esm_data, method = "REML", na.action = na.exclude, control = sc1_ctrl)

h12 <- lme(love ~ NA_similarity_act_w + NA_similarity_act_b,
           random = ~1|dyad/person, correlation = sc1_cor,
           data = esm_data, method = "REML", na.action = na.exclude, control = sc1_ctrl)

# ── Step 1b: actual + perceived similarity (H1.3) ──
h13_PA <- lme(love ~ PA_similarity_act_w + PA_similarity_act_b +
                     PA_similarity_perc_w + PA_similarity_perc_b,
              random = ~1|dyad/person, correlation = sc1_cor,
              data = esm_data, method = "REML", na.action = na.exclude, control = sc1_ctrl)

h13_NA <- lme(love ~ NA_similarity_act_w + NA_similarity_act_b +
                     NA_similarity_perc_w + NA_similarity_perc_b,
              random = ~1|dyad/person, correlation = sc1_cor,
              data = esm_data, method = "REML", na.action = na.exclude, control = sc1_ctrl)
```

```{r table-sc1-step1}
coef_cols_s1 <- c("PA_1a", "PA_1b", "NA_1a", "NA_1b")

tbl_s1 <- tibble(
  predictor = c("Act. similarity (within)", "Act. similarity (between)",
                "Perc. similarity (within)", "Perc. similarity (between)"),
  PA_1a = c(pull_fmt(h11, "PA_similarity_act_w"),
            pull_fmt(h11, "PA_similarity_act_b"), "—", "—"),
  PA_1b = c(pull_fmt(h13_PA, "PA_similarity_act_w"),
            pull_fmt(h13_PA, "PA_similarity_act_b"),
            pull_fmt(h13_PA, "PA_similarity_perc_w"),
            pull_fmt(h13_PA, "PA_similarity_perc_b")),
  NA_1a = c(pull_fmt(h12, "NA_similarity_act_w"),
            pull_fmt(h12, "NA_similarity_act_b"), "—", "—"),
  NA_1b = c(pull_fmt(h13_NA, "NA_similarity_act_w"),
            pull_fmt(h13_NA, "NA_similarity_act_b"),
            pull_fmt(h13_NA, "NA_similarity_perc_w"),
            pull_fmt(h13_NA, "NA_similarity_perc_b"))
)

p_s1 <- tibble(
  PA_1a = c(pull_p(h11, "PA_similarity_act_w"),
            pull_p(h11, "PA_similarity_act_b"), NA, NA),
  PA_1b = c(pull_p(h13_PA, "PA_similarity_act_w"),
            pull_p(h13_PA, "PA_similarity_act_b"),
            pull_p(h13_PA, "PA_similarity_perc_w"),
            pull_p(h13_PA, "PA_similarity_perc_b")),
  NA_1a = c(pull_p(h12, "NA_similarity_act_w"),
            pull_p(h12, "NA_similarity_act_b"), NA, NA),
  NA_1b = c(pull_p(h13_NA, "NA_similarity_act_w"),
            pull_p(h13_NA, "NA_similarity_act_b"),
            pull_p(h13_NA, "NA_similarity_perc_w"),
            pull_p(h13_NA, "NA_similarity_perc_b"))
)

note_s1 <- paste0(
  "Note. Model 1a = actual similarity only (H1.1 for PA, H1.2 for NA); Model 1b = actual + perceived ",
  "similarity (H1.3). All models include random intercepts for persons nested in dyads and a ",
  "continuous-time AR(1) residual structure. '—' = predictor not included in model. ",
  "Bold = p < .05. * p < .05, ** p < .01, *** p < .001."
)

ft_s1 <- flextable(tbl_s1) %>%
  set_header_labels(
    predictor = "Predictor",
    PA_1a = "Model 1a", PA_1b = "Model 1b",
    NA_1a = "Model 1a", NA_1b = "Model 1b"
  ) %>%
  add_header_row(
    values    = c("", "PA Similarity (H1.1 & H1.3)", "NA Similarity (H1.2 & H1.3)"),
    colwidths = c(1, 2, 2)
  ) %>%
  merge_v(j = "predictor", part = "header") %>%
  bold_sig(p_s1, coef_cols_s1) %>%
  font(fontname = MS_FONT, part = "all") %>%
  fontsize(size = MS_SIZE_BODY, part = "all") %>%
  bold(part = "header") %>%
  align(align = "center", part = "header") %>%
  align(j = "predictor", align = "left", part = "body") %>%
  align(j = coef_cols_s1, align = "right", part = "body") %>%
  border_remove() %>%
  hline_top(border = rule_thick, part = "header") %>%
  hline(i = 1, border = rule_thin,  part = "header") %>%
  hline_bottom(border = rule_thick, part = "header") %>%
  hline(i = 2, border = rule_thin,  part = "body") %>%
  hline_bottom(border = rule_thick, part = "body") %>%
  add_footer_row(values = note_s1, colwidths = 5) %>%
  font(fontname = MS_FONT, part = "footer") %>%
  fontsize(size = MS_SIZE_NOTE, part = "footer") %>%
  italic(part = "footer") %>%
  set_table_properties(layout = "autofit")

save_as_docx(ft_s1, path = file.path(dir_tables, "table_sc1_step1_main_additive.docx"))
ft_s1
```

## Summary (Models 1a–1c)

```{r sc1-step1-1c-data}
#| include: false

h1c_PA <- lme(love ~ PA_similarity_act_w + PA_similarity_act_b +
                     PA_similarity_perc_w + PA_similarity_perc_b +
                     PA_similarity_act_w:PA_similarity_perc_w +
                     PA_similarity_act_b:PA_similarity_perc_b,
              random = ~1|dyad/person, correlation = sc1_cor,
              data = esm_data, method = "REML", na.action = na.exclude, control = sc1_ctrl)

h1c_NA <- lme(love ~ NA_similarity_act_w + NA_similarity_act_b +
                     NA_similarity_perc_w + NA_similarity_perc_b +
                     NA_similarity_act_w:NA_similarity_perc_w +
                     NA_similarity_act_b:NA_similarity_perc_b,
              random = ~1|dyad/person, correlation = sc1_cor,
              data = esm_data, method = "REML", na.action = na.exclude, control = sc1_ctrl)
```

```{r table-sc1-step1-summary}
coef_cols_s1_sum <- c("PA_1a", "PA_1b", "PA_1c", "NA_1a", "NA_1b", "NA_1c")

tbl_s1_sum <- tibble(
  predictor = c(rep("Act. Similarity", 2), rep("Perc. Similarity", 2), rep("Act. × Perc. Similarity", 2)),
  level     = rep(c("within", "between"), 3),
  PA_1a = c(pull_fmt(h11, "PA_similarity_act_w"),
            pull_fmt(h11, "PA_similarity_act_b"),
            "—", "—", "—", "—"),
  PA_1b = c(pull_fmt(h13_PA, "PA_similarity_act_w"),
            pull_fmt(h13_PA, "PA_similarity_act_b"),
            pull_fmt(h13_PA, "PA_similarity_perc_w"),
            pull_fmt(h13_PA, "PA_similarity_perc_b"),
            "—", "—"),
  PA_1c = c(pull_fmt(h1c_PA, "PA_similarity_act_w"),
            pull_fmt(h1c_PA, "PA_similarity_act_b"),
            pull_fmt(h1c_PA, "PA_similarity_perc_w"),
            pull_fmt(h1c_PA, "PA_similarity_perc_b"),
            pull_fmt(h1c_PA, "PA_similarity_act_w:PA_similarity_perc_w"),
            pull_fmt(h1c_PA, "PA_similarity_act_b:PA_similarity_perc_b")),
  NA_1a = c(pull_fmt(h12, "NA_similarity_act_w"),
            pull_fmt(h12, "NA_similarity_act_b"),
            "—", "—", "—", "—"),
  NA_1b = c(pull_fmt(h13_NA, "NA_similarity_act_w"),
            pull_fmt(h13_NA, "NA_similarity_act_b"),
            pull_fmt(h13_NA, "NA_similarity_perc_w"),
            pull_fmt(h13_NA, "NA_similarity_perc_b"),
            "—", "—"),
  NA_1c = c(pull_fmt(h1c_NA, "NA_similarity_act_w"),
            pull_fmt(h1c_NA, "NA_similarity_act_b"),
            pull_fmt(h1c_NA, "NA_similarity_perc_w"),
            pull_fmt(h1c_NA, "NA_similarity_perc_b"),
            pull_fmt(h1c_NA, "NA_similarity_act_w:NA_similarity_perc_w"),
            pull_fmt(h1c_NA, "NA_similarity_act_b:NA_similarity_perc_b"))
)

p_s1_sum <- tibble(
  PA_1a = c(pull_p(h11, "PA_similarity_act_w"),
            pull_p(h11, "PA_similarity_act_b"),
            NA, NA, NA, NA),
  PA_1b = c(pull_p(h13_PA, "PA_similarity_act_w"),
            pull_p(h13_PA, "PA_similarity_act_b"),
            pull_p(h13_PA, "PA_similarity_perc_w"),
            pull_p(h13_PA, "PA_similarity_perc_b"),
            NA, NA),
  PA_1c = c(pull_p(h1c_PA, "PA_similarity_act_w"),
            pull_p(h1c_PA, "PA_similarity_act_b"),
            pull_p(h1c_PA, "PA_similarity_perc_w"),
            pull_p(h1c_PA, "PA_similarity_perc_b"),
            pull_p(h1c_PA, "PA_similarity_act_w:PA_similarity_perc_w"),
            pull_p(h1c_PA, "PA_similarity_act_b:PA_similarity_perc_b")),
  NA_1a = c(pull_p(h12, "NA_similarity_act_w"),
            pull_p(h12, "NA_similarity_act_b"),
            NA, NA, NA, NA),
  NA_1b = c(pull_p(h13_NA, "NA_similarity_act_w"),
            pull_p(h13_NA, "NA_similarity_act_b"),
            pull_p(h13_NA, "NA_similarity_perc_w"),
            pull_p(h13_NA, "NA_similarity_perc_b"),
            NA, NA),
  NA_1c = c(pull_p(h1c_NA, "NA_similarity_act_w"),
            pull_p(h1c_NA, "NA_similarity_act_b"),
            pull_p(h1c_NA, "NA_similarity_perc_w"),
            pull_p(h1c_NA, "NA_similarity_perc_b"),
            pull_p(h1c_NA, "NA_similarity_act_w:NA_similarity_perc_w"),
            pull_p(h1c_NA, "NA_similarity_act_b:NA_similarity_perc_b"))
)

note_s1_sum <- paste0(
  "Note. Model 1a = actual similarity only (H1.1 for PA, H1.2 for NA); ",
  "Model 1b = actual + perceived similarity (H1.3); ",
  "Model 1c = actual + perceived + Actual × Perceived interaction (exploratory). ",
  "All models include random intercepts for persons nested in dyads and a ",
  "continuous-time AR(1) residual structure. '—' = predictor not included in model. ",
  "Bold = p < .05. * p < .05, ** p < .01, *** p < .001."
)

ft_s1_sum <- flextable(tbl_s1_sum) %>%
  set_header_labels(
    predictor = "Predictor",
    level     = "Level",
    PA_1a = "1a", PA_1b = "1b", PA_1c = "1c",
    NA_1a = "1a", NA_1b = "1b", NA_1c = "1c"
  ) %>%
  add_header_row(
    values    = c("", "", "PA Similarity", "NA Similarity"),
    colwidths = c(1, 1, 3, 3)
  ) %>%
  merge_v(j = c("predictor", "level"), part = "header") %>%
  merge_v(j = "predictor", part = "body") %>%
  bold_sig(p_s1_sum, coef_cols_s1_sum) %>%
  font(fontname = MS_FONT, part = "all") %>%
  fontsize(size = MS_SIZE_BODY, part = "all") %>%
  bold(part = "header") %>%
  align(align = "center", part = "header") %>%
  align(j = c("predictor", "level"), align = "left", part = "body") %>%
  align(j = coef_cols_s1_sum, align = "right", part = "body") %>%
  border_remove() %>%
  hline_top(border = rule_thick, part = "header") %>%
  hline(i = 1, border = rule_thin,  part = "header") %>%
  hline_bottom(border = rule_thick, part = "header") %>%
  hline(i = 2, border = rule_thin, part = "body") %>%
  hline(i = 4, border = rule_thin, part = "body") %>%
  hline_bottom(border = rule_thick, part = "body") %>%
  add_footer_row(values = note_s1_sum, colwidths = 8) %>%
  font(fontname = MS_FONT, part = "footer") %>%
  fontsize(size = MS_SIZE_NOTE, part = "footer") %>%
  italic(part = "footer") %>%
  set_table_properties(layout = "autofit")

save_as_docx(ft_s1_sum, path = file.path(dir_tables, "table_sc1_step1_summary.docx"))
ft_s1_sum
```

# SC1 Step 2 Tables

## Mediation Summary (H2.1–H2.2)

```{r sc1-step2-data}
#| include: false

esm_data <- readRDS(file.path(dir_data_ana, "esm_ana.rds"))
load_config("ESM")
esm_data <- esm_data %>% arrange(dyad, person, time)

set.seed(42)

sc1_ctrl <- lmeControl(opt = "nlminb")
sc1_cor  <- corCAR1(form = ~ time | dyad/person)

# ── Total-effect (c-path) models ──
h11 <- lme(love ~ PA_similarity_act_w + PA_similarity_act_b,
           random = ~1|dyad/person, correlation = sc1_cor,
           data = esm_data, method = "REML", na.action = na.exclude, control = sc1_ctrl)
h12 <- lme(love ~ NA_similarity_act_w + NA_similarity_act_b,
           random = ~1|dyad/person, correlation = sc1_cor,
           data = esm_data, method = "REML", na.action = na.exclude, control = sc1_ctrl)

# ── Mediator models (a-paths) ──
h21_mediator <- lme(PA_similarity_perc ~ PA_similarity_act_w + PA_similarity_act_b,
                    random = ~1|dyad/person, correlation = sc1_cor,
                    data = esm_data, method = "REML", na.action = na.exclude, control = sc1_ctrl)
h22_mediator <- lme(NA_similarity_perc ~ NA_similarity_act_w + NA_similarity_act_b,
                    random = ~1|dyad/person, correlation = sc1_cor,
                    data = esm_data, method = "REML", na.action = na.exclude, control = sc1_ctrl)

# ── Outcome models (b- and c'-paths) ──
h21_outcome <- lme(love ~ PA_similarity_perc_w + PA_similarity_perc_b +
                          PA_similarity_act_w  + PA_similarity_act_b,
                   random = ~1|dyad/person, correlation = sc1_cor,
                   data = esm_data, method = "REML", na.action = na.exclude, control = sc1_ctrl)
h22_outcome <- lme(love ~ NA_similarity_perc_w + NA_similarity_perc_b +
                          NA_similarity_act_w  + NA_similarity_act_b,
                   random = ~1|dyad/person, correlation = sc1_cor,
                   data = esm_data, method = "REML", na.action = na.exclude, control = sc1_ctrl)

# ── Monte Carlo indirect effects (20,000 draws each) ──
mc_PA_w <- run_mc(h21_mediator, "PA_similarity_act_w", h21_outcome, "PA_similarity_perc_w")
mc_PA_b <- run_mc(h21_mediator, "PA_similarity_act_b", h21_outcome, "PA_similarity_perc_b")
mc_NA_w <- run_mc(h22_mediator, "NA_similarity_act_w", h22_outcome, "NA_similarity_perc_w")
mc_NA_b <- run_mc(h22_mediator, "NA_similarity_act_b", h22_outcome, "NA_similarity_perc_b")
```

```{r table-sc1-step2-mediation}
coef_cols_s2 <- c("PA_sim", "NA_sim")

tbl_s2 <- tibble(
  level  = c(rep("Within", 5), rep("Between", 5)),
  path   = rep(c("a-path", "b-path", "c-path (total)", "c′-path (direct)", "Indirect (MC)"), 2),
  PA_sim = c(
    pull_fmt(h21_mediator, "PA_similarity_act_w"),
    pull_fmt(h21_outcome,  "PA_similarity_perc_w"),
    pull_fmt(h11,          "PA_similarity_act_w"),
    pull_fmt(h21_outcome,  "PA_similarity_act_w"),
    fmt_b(mc_PA_w$point,   mc_PA_w$p),
    pull_fmt(h21_mediator, "PA_similarity_act_b"),
    pull_fmt(h21_outcome,  "PA_similarity_perc_b"),
    pull_fmt(h11,          "PA_similarity_act_b"),
    pull_fmt(h21_outcome,  "PA_similarity_act_b"),
    fmt_b(mc_PA_b$point,   mc_PA_b$p)
  ),
  NA_sim = c(
    pull_fmt(h22_mediator, "NA_similarity_act_w"),
    pull_fmt(h22_outcome,  "NA_similarity_perc_w"),
    pull_fmt(h12,          "NA_similarity_act_w"),
    pull_fmt(h22_outcome,  "NA_similarity_act_w"),
    fmt_b(mc_NA_w$point,   mc_NA_w$p),
    pull_fmt(h22_mediator, "NA_similarity_act_b"),
    pull_fmt(h22_outcome,  "NA_similarity_perc_b"),
    pull_fmt(h12,          "NA_similarity_act_b"),
    pull_fmt(h22_outcome,  "NA_similarity_act_b"),
    fmt_b(mc_NA_b$point,   mc_NA_b$p)
  )
)

p_s2 <- tibble(
  PA_sim = c(
    pull_p(h21_mediator, "PA_similarity_act_w"),
    pull_p(h21_outcome,  "PA_similarity_perc_w"),
    pull_p(h11,          "PA_similarity_act_w"),
    pull_p(h21_outcome,  "PA_similarity_act_w"),
    mc_PA_w$p,
    pull_p(h21_mediator, "PA_similarity_act_b"),
    pull_p(h21_outcome,  "PA_similarity_perc_b"),
    pull_p(h11,          "PA_similarity_act_b"),
    pull_p(h21_outcome,  "PA_similarity_act_b"),
    mc_PA_b$p
  ),
  NA_sim = c(
    pull_p(h22_mediator, "NA_similarity_act_w"),
    pull_p(h22_outcome,  "NA_similarity_perc_w"),
    pull_p(h12,          "NA_similarity_act_w"),
    pull_p(h22_outcome,  "NA_similarity_act_w"),
    mc_NA_w$p,
    pull_p(h22_mediator, "NA_similarity_act_b"),
    pull_p(h22_outcome,  "NA_similarity_perc_b"),
    pull_p(h12,          "NA_similarity_act_b"),
    pull_p(h22_outcome,  "NA_similarity_act_b"),
    mc_NA_b$p
  )
)

note_s2 <- paste0(
  "Note. a-path: actual similarity predicting perceived similarity. b-path: perceived similarity ",
  "predicting love, controlling for actual similarity. c-path (total): total effect of actual ",
  "similarity on love. c′-path (direct): direct effect of actual similarity, controlling for ",
  "perceived similarity. Indirect (MC): Monte Carlo indirect effect (20,000 draws; a × b product). ",
  "Within = within-person/moment component; Between = between-dyad component. All models include ",
  "random intercepts for persons nested in dyads and a continuous-time AR(1) residual structure. ",
  "Bold = p < .05. * p < .05, ** p < .01, *** p < .001."
)

ft_s2 <- flextable(tbl_s2) %>%
  set_header_labels(
    level  = "Level",
    path   = "Path",
    PA_sim = "PA Similarity (H2.1)",
    NA_sim = "NA Similarity (H2.2)"
  ) %>%
  merge_v(j = "level", part = "body") %>%
  bold_sig(p_s2, coef_cols_s2) %>%
  font(fontname = MS_FONT, part = "all") %>%
  fontsize(size = MS_SIZE_BODY, part = "all") %>%
  bold(part = "header") %>%
  align(align = "center", part = "header") %>%
  align(j = c("level", "path"), align = "left", part = "body") %>%
  align(j = coef_cols_s2, align = "right", part = "body") %>%
  border_remove() %>%
  hline_top(border = rule_thick, part = "header") %>%
  hline_bottom(border = rule_thick, part = "header") %>%
  hline(i = 5, border = rule_thin,  part = "body") %>%
  hline_bottom(border = rule_thick, part = "body") %>%
  add_footer_row(values = note_s2, colwidths = 4) %>%
  font(fontname = MS_FONT, part = "footer") %>%
  fontsize(size = MS_SIZE_NOTE, part = "footer") %>%
  italic(part = "footer") %>%
  set_table_properties(layout = "autofit")

save_as_docx(ft_s2, path = file.path(dir_tables, "table_sc1_step2_mediation.docx"))
ft_s2
```

# SC1 Step 3a Tables

## Responsiveness Moderation Summary (H3.1–H3.2)

```{r sc1-step3a-data}
#| include: false

esm_data <- readRDS(file.path(dir_data_ana, "esm_ana.rds"))
load_config("ESM")
esm_data <- esm_data %>% arrange(dyad, person, time)

sc1_ctrl <- lmeControl(opt = "nlminb")
sc1_cor  <- corCAR1(form = ~ time | dyad/person)

# ── H3.1: PA × Responsiveness ──
h31a <- lme(love ~ PA_similarity_act_w  + PA_similarity_act_b  +
                   cperc_resp + PA_similarity_act_w:cperc_resp,
            random = ~1|dyad/person, correlation = sc1_cor,
            data = esm_data, method = "REML", na.action = na.exclude, control = sc1_ctrl)
h31p <- lme(love ~ PA_similarity_perc_w + PA_similarity_perc_b +
                   cperc_resp + PA_similarity_perc_w:cperc_resp,
            random = ~1|dyad/person, correlation = sc1_cor,
            data = esm_data, method = "REML", na.action = na.exclude, control = sc1_ctrl)

# ── H3.2: NA × Responsiveness ──
h32a <- lme(love ~ NA_similarity_act_w  + NA_similarity_act_b  +
                   cperc_resp + NA_similarity_act_w:cperc_resp,
            random = ~1|dyad/person, correlation = sc1_cor,
            data = esm_data, method = "REML", na.action = na.exclude, control = sc1_ctrl)
h32p <- lme(love ~ NA_similarity_perc_w + NA_similarity_perc_b +
                   cperc_resp + NA_similarity_perc_w:cperc_resp,
            random = ~1|dyad/person, correlation = sc1_cor,
            data = esm_data, method = "REML", na.action = na.exclude, control = sc1_ctrl)
```

```{r table-sc1-step3a-responsiveness}
coef_cols_3a_sc1 <- "estimate"

tbl_3a_sc1 <- tibble(
  affect     = rep(c("PA Similarity (H3.1)", "NA Similarity (H3.2)"), each = 8),
  similarity = rep(c("Actual", "Perceived"), each = 4, times = 2),
  parameter  = rep(c("Similarity (within)", "Similarity (between)",
                     "Responsiveness", "Sim × Responsiveness"), times = 4),
  estimate   = c(
    pull_fmt(h31a, "PA_similarity_act_w"),
    pull_fmt(h31a, "PA_similarity_act_b"),
    pull_fmt(h31a, "cperc_resp"),
    pull_fmt(h31a, "PA_similarity_act_w:cperc_resp"),
    pull_fmt(h31p, "PA_similarity_perc_w"),
    pull_fmt(h31p, "PA_similarity_perc_b"),
    pull_fmt(h31p, "cperc_resp"),
    pull_fmt(h31p, "PA_similarity_perc_w:cperc_resp"),
    pull_fmt(h32a, "NA_similarity_act_w"),
    pull_fmt(h32a, "NA_similarity_act_b"),
    pull_fmt(h32a, "cperc_resp"),
    pull_fmt(h32a, "NA_similarity_act_w:cperc_resp"),
    pull_fmt(h32p, "NA_similarity_perc_w"),
    pull_fmt(h32p, "NA_similarity_perc_b"),
    pull_fmt(h32p, "cperc_resp"),
    pull_fmt(h32p, "NA_similarity_perc_w:cperc_resp")
  )
)

p_mat_3a_sc1 <- tibble(
  estimate = c(
    pull_p(h31a, "PA_similarity_act_w"),
    pull_p(h31a, "PA_similarity_act_b"),
    pull_p(h31a, "cperc_resp"),
    pull_p(h31a, "PA_similarity_act_w:cperc_resp"),
    pull_p(h31p, "PA_similarity_perc_w"),
    pull_p(h31p, "PA_similarity_perc_b"),
    pull_p(h31p, "cperc_resp"),
    pull_p(h31p, "PA_similarity_perc_w:cperc_resp"),
    pull_p(h32a, "NA_similarity_act_w"),
    pull_p(h32a, "NA_similarity_act_b"),
    pull_p(h32a, "cperc_resp"),
    pull_p(h32a, "NA_similarity_act_w:cperc_resp"),
    pull_p(h32p, "NA_similarity_perc_w"),
    pull_p(h32p, "NA_similarity_perc_b"),
    pull_p(h32p, "cperc_resp"),
    pull_p(h32p, "NA_similarity_perc_w:cperc_resp")
  )
)

note_3a_sc1 <- paste0(
  "Note. Similarity (within): main effect of within-person similarity at average responsiveness. ",
  "Similarity (between): main effect of between-dyad similarity. ",
  "Responsiveness: main effect of person-mean-centered perceived partner responsiveness. ",
  "Sim × Responsiveness: interaction term (preregistered hypothesis: positive for PA [H3.1] and ",
  "NA [H3.2]). All models include random intercepts for persons nested in dyads and a ",
  "continuous-time AR(1) residual structure. Bold = p < .05. * p < .05, ** p < .01, *** p < .001."
)

ft_3a_sc1 <- flextable(tbl_3a_sc1) %>%
  set_header_labels(
    affect     = "Affect",
    similarity = "Similarity",
    parameter  = "Parameter",
    estimate   = "Estimate"
  ) %>%
  merge_v(j = c("affect", "similarity"), part = "body") %>%
  bold_sig(p_mat_3a_sc1, coef_cols_3a_sc1) %>%
  font(fontname = MS_FONT, part = "all") %>%
  fontsize(size = MS_SIZE_BODY, part = "all") %>%
  bold(part = "header") %>%
  align(align = "center", part = "header") %>%
  align(j = c("affect", "similarity", "parameter"), align = "left", part = "body") %>%
  align(j = coef_cols_3a_sc1, align = "right", part = "body") %>%
  border_remove() %>%
  hline_top(border = rule_thick, part = "header") %>%
  hline_bottom(border = rule_thick, part = "header") %>%
  hline(i =  4, border = rule_thin,  part = "body") %>%
  hline(i =  8, border = rule_thick, part = "body") %>%
  hline(i = 12, border = rule_thin,  part = "body") %>%
  hline_bottom(border = rule_thick, part = "body") %>%
  add_footer_row(values = note_3a_sc1, colwidths = 4) %>%
  font(fontname = MS_FONT, part = "footer") %>%
  fontsize(size = MS_SIZE_NOTE, part = "footer") %>%
  italic(part = "footer") %>%
  set_table_properties(layout = "autofit")

save_as_docx(ft_3a_sc1, path = file.path(dir_tables, "table_sc1_step3a_responsiveness.docx"))
ft_3a_sc1
```

# SC1 Step 3b Tables

## Event Valence Moderation Summary (H3.3–H3.4)

```{r sc1-step3b-data}
#| include: false

esm_data <- readRDS(file.path(dir_data_ana, "esm_ana.rds"))
load_config("ESM")
esm_data <- esm_data %>% arrange(dyad, person, time)

sc1_ctrl <- lmeControl(opt = "nlminb")
sc1_cor  <- corCAR1(form = ~ time | dyad/person)

# ── H3.3: PA × Event Valence (cC) ──
h33a <- lme(love ~ PA_similarity_act_w  + PA_similarity_act_b  +
                   cC + PA_similarity_act_w:cC,
            random = ~1|dyad/person, correlation = sc1_cor,
            data = esm_data, method = "REML", na.action = na.exclude, control = sc1_ctrl)
h33p <- lme(love ~ PA_similarity_perc_w + PA_similarity_perc_b +
                   cC + PA_similarity_perc_w:cC,
            random = ~1|dyad/person, correlation = sc1_cor,
            data = esm_data, method = "REML", na.action = na.exclude, control = sc1_ctrl)

# ── H3.4: NA × Event Valence (cC) ──
h34a <- lme(love ~ NA_similarity_act_w  + NA_similarity_act_b  +
                   cC + NA_similarity_act_w:cC,
            random = ~1|dyad/person, correlation = sc1_cor,
            data = esm_data, method = "REML", na.action = na.exclude, control = sc1_ctrl)
h34p <- lme(love ~ NA_similarity_perc_w + NA_similarity_perc_b +
                   cC + NA_similarity_perc_w:cC,
            random = ~1|dyad/person, correlation = sc1_cor,
            data = esm_data, method = "REML", na.action = na.exclude, control = sc1_ctrl)
```

```{r table-sc1-step3b-valence}
coef_cols_3b_sc1 <- "estimate"

tbl_3b_sc1 <- tibble(
  affect     = rep(c("PA Similarity (H3.3)", "NA Similarity (H3.4)"), each = 8),
  similarity = rep(c("Actual", "Perceived"), each = 4, times = 2),
  parameter  = rep(c("Similarity (within)", "Similarity (between)",
                     "Event Valence", "Sim × Event Valence"), times = 4),
  estimate   = c(
    pull_fmt(h33a, "PA_similarity_act_w"),
    pull_fmt(h33a, "PA_similarity_act_b"),
    pull_fmt(h33a, "cC"),
    pull_fmt(h33a, "PA_similarity_act_w:cC"),
    pull_fmt(h33p, "PA_similarity_perc_w"),
    pull_fmt(h33p, "PA_similarity_perc_b"),
    pull_fmt(h33p, "cC"),
    pull_fmt(h33p, "PA_similarity_perc_w:cC"),
    pull_fmt(h34a, "NA_similarity_act_w"),
    pull_fmt(h34a, "NA_similarity_act_b"),
    pull_fmt(h34a, "cC"),
    pull_fmt(h34a, "NA_similarity_act_w:cC"),
    pull_fmt(h34p, "NA_similarity_perc_w"),
    pull_fmt(h34p, "NA_similarity_perc_b"),
    pull_fmt(h34p, "cC"),
    pull_fmt(h34p, "NA_similarity_perc_w:cC")
  )
)

p_mat_3b_sc1 <- tibble(
  estimate = c(
    pull_p(h33a, "PA_similarity_act_w"),
    pull_p(h33a, "PA_similarity_act_b"),
    pull_p(h33a, "cC"),
    pull_p(h33a, "PA_similarity_act_w:cC"),
    pull_p(h33p, "PA_similarity_perc_w"),
    pull_p(h33p, "PA_similarity_perc_b"),
    pull_p(h33p, "cC"),
    pull_p(h33p, "PA_similarity_perc_w:cC"),
    pull_p(h34a, "NA_similarity_act_w"),
    pull_p(h34a, "NA_similarity_act_b"),
    pull_p(h34a, "cC"),
    pull_p(h34a, "NA_similarity_act_w:cC"),
    pull_p(h34p, "NA_similarity_perc_w"),
    pull_p(h34p, "NA_similarity_perc_b"),
    pull_p(h34p, "cC"),
    pull_p(h34p, "NA_similarity_perc_w:cC")
  )
)

note_3b_sc1 <- paste0(
  "Note. Event valence (cC) = person-mean-centered net positivity of events since the previous beep ",
  "(positive − negative events; only available at beeps when the participant was not with or in recent ",
  "contact with their partner). Similarity (within): main effect at average event valence. ",
  "Similarity (between): main effect of between-dyad similarity. ",
  "Sim × Event Valence: interaction term (preregistered mismatch hypothesis: negative for PA [H3.3] ",
  "and negative for NA [H3.4]). All models include random intercepts for persons nested in dyads ",
  "and a continuous-time AR(1) residual structure. Bold = p < .05. * p < .05, ** p < .01, *** p < .001."
)

ft_3b_sc1 <- flextable(tbl_3b_sc1) %>%
  set_header_labels(
    affect     = "Affect",
    similarity = "Similarity",
    parameter  = "Parameter",
    estimate   = "Estimate"
  ) %>%
  merge_v(j = c("affect", "similarity"), part = "body") %>%
  bold_sig(p_mat_3b_sc1, coef_cols_3b_sc1) %>%
  font(fontname = MS_FONT, part = "all") %>%
  fontsize(size = MS_SIZE_BODY, part = "all") %>%
  bold(part = "header") %>%
  align(align = "center", part = "header") %>%
  align(j = c("affect", "similarity", "parameter"), align = "left", part = "body") %>%
  align(j = coef_cols_3b_sc1, align = "right", part = "body") %>%
  border_remove() %>%
  hline_top(border = rule_thick, part = "header") %>%
  hline_bottom(border = rule_thick, part = "header") %>%
  hline(i =  4, border = rule_thin,  part = "body") %>%
  hline(i =  8, border = rule_thick, part = "body") %>%
  hline(i = 12, border = rule_thin,  part = "body") %>%
  hline_bottom(border = rule_thick, part = "body") %>%
  add_footer_row(values = note_3b_sc1, colwidths = 4) %>%
  font(fontname = MS_FONT, part = "footer") %>%
  fontsize(size = MS_SIZE_NOTE, part = "footer") %>%
  italic(part = "footer") %>%
  set_table_properties(layout = "autofit")

save_as_docx(ft_3b_sc1, path = file.path(dir_tables, "table_sc1_step3b_valence.docx"))
ft_3b_sc1
```

## Extended Event Valence Moderation — Exploratory (cC_expl)

```{r sc1-step3b-expl-data}
#| include: false

esm_data <- readRDS(file.path(dir_data_ana, "esm_ana.rds"))
load_config("ESM")
esm_data <- esm_data %>% arrange(dyad, person, time)

sc1_ctrl <- lmeControl(opt = "nlminb")
sc1_cor  <- corCAR1(form = ~ time | dyad/person)

# ── PA × Extended Event Valence ──
e1_h33a <- lme(love ~ PA_similarity_act_w  + PA_similarity_act_b  +
                      cC_expl + PA_similarity_act_w:cC_expl,
               random = ~1|dyad/person, correlation = sc1_cor,
               data = esm_data, method = "REML", na.action = na.exclude, control = sc1_ctrl)
e1_h33p <- lme(love ~ PA_similarity_perc_w + PA_similarity_perc_b +
                      cC_expl + PA_similarity_perc_w:cC_expl,
               random = ~1|dyad/person, correlation = sc1_cor,
               data = esm_data, method = "REML", na.action = na.exclude, control = sc1_ctrl)

# ── NA × Extended Event Valence ──
e1_h34a <- lme(love ~ NA_similarity_act_w  + NA_similarity_act_b  +
                      cC_expl + NA_similarity_act_w:cC_expl,
               random = ~1|dyad/person, correlation = sc1_cor,
               data = esm_data, method = "REML", na.action = na.exclude, control = sc1_ctrl)
e1_h34p <- lme(love ~ NA_similarity_perc_w + NA_similarity_perc_b +
                      cC_expl + NA_similarity_perc_w:cC_expl,
               random = ~1|dyad/person, correlation = sc1_cor,
               data = esm_data, method = "REML", na.action = na.exclude, control = sc1_ctrl)
```

```{r table-sc1-step3b-expl-valence}
coef_cols_3b_e <- "estimate"

tbl_3b_e <- tibble(
  affect     = rep(c("PA Similarity", "NA Similarity"), each = 8),
  similarity = rep(c("Actual", "Perceived"), each = 4, times = 2),
  parameter  = rep(c("Similarity (within)", "Similarity (between)",
                     "Event Valence (ext.)", "Sim × Event Valence (ext.)"), times = 4),
  estimate   = c(
    pull_fmt(e1_h33a, "PA_similarity_act_w"),
    pull_fmt(e1_h33a, "PA_similarity_act_b"),
    pull_fmt(e1_h33a, "cC_expl"),
    pull_fmt(e1_h33a, "PA_similarity_act_w:cC_expl"),
    pull_fmt(e1_h33p, "PA_similarity_perc_w"),
    pull_fmt(e1_h33p, "PA_similarity_perc_b"),
    pull_fmt(e1_h33p, "cC_expl"),
    pull_fmt(e1_h33p, "PA_similarity_perc_w:cC_expl"),
    pull_fmt(e1_h34a, "NA_similarity_act_w"),
    pull_fmt(e1_h34a, "NA_similarity_act_b"),
    pull_fmt(e1_h34a, "cC_expl"),
    pull_fmt(e1_h34a, "NA_similarity_act_w:cC_expl"),
    pull_fmt(e1_h34p, "NA_similarity_perc_w"),
    pull_fmt(e1_h34p, "NA_similarity_perc_b"),
    pull_fmt(e1_h34p, "cC_expl"),
    pull_fmt(e1_h34p, "NA_similarity_perc_w:cC_expl")
  )
)

p_mat_3b_e <- tibble(
  estimate = c(
    pull_p(e1_h33a, "PA_similarity_act_w"),
    pull_p(e1_h33a, "PA_similarity_act_b"),
    pull_p(e1_h33a, "cC_expl"),
    pull_p(e1_h33a, "PA_similarity_act_w:cC_expl"),
    pull_p(e1_h33p, "PA_similarity_perc_w"),
    pull_p(e1_h33p, "PA_similarity_perc_b"),
    pull_p(e1_h33p, "cC_expl"),
    pull_p(e1_h33p, "PA_similarity_perc_w:cC_expl"),
    pull_p(e1_h34a, "NA_similarity_act_w"),
    pull_p(e1_h34a, "NA_similarity_act_b"),
    pull_p(e1_h34a, "cC_expl"),
    pull_p(e1_h34a, "NA_similarity_act_w:cC_expl"),
    pull_p(e1_h34p, "NA_similarity_perc_w"),
    pull_p(e1_h34p, "NA_similarity_perc_b"),
    pull_p(e1_h34p, "cC_expl"),
    pull_p(e1_h34p, "NA_similarity_perc_w:cC_expl")
  )
)

note_3b_e <- paste0(
  "Note. Exploratory extension of Step 3b. Extended event valence (cC_expl) = person-mean-centered ",
  "net positivity pooled across general (partner-absent beeps) and partner-related (partner-present ",
  "or partner-contact beeps) event items, covering a substantially larger portion of the data than ",
  "the preregistered cC index. Similarity (within): main effect at average event valence. ",
  "Similarity (between): main effect of between-dyad similarity. ",
  "Sim × Event Valence (ext.): interaction with extended event valence. ",
  "All models include random intercepts for persons nested in dyads and a continuous-time AR(1) ",
  "residual structure. Bold = p < .05. * p < .05, ** p < .01, *** p < .001."
)

ft_3b_e <- flextable(tbl_3b_e) %>%
  set_header_labels(
    affect     = "Affect",
    similarity = "Similarity",
    parameter  = "Parameter",
    estimate   = "Estimate"
  ) %>%
  merge_v(j = c("affect", "similarity"), part = "body") %>%
  bold_sig(p_mat_3b_e, coef_cols_3b_e) %>%
  font(fontname = MS_FONT, part = "all") %>%
  fontsize(size = MS_SIZE_BODY, part = "all") %>%
  bold(part = "header") %>%
  align(align = "center", part = "header") %>%
  align(j = c("affect", "similarity", "parameter"), align = "left", part = "body") %>%
  align(j = coef_cols_3b_e, align = "right", part = "body") %>%
  border_remove() %>%
  hline_top(border = rule_thick, part = "header") %>%
  hline_bottom(border = rule_thick, part = "header") %>%
  hline(i =  4, border = rule_thin,  part = "body") %>%
  hline(i =  8, border = rule_thick, part = "body") %>%
  hline(i = 12, border = rule_thin,  part = "body") %>%
  hline_bottom(border = rule_thick, part = "body") %>%
  add_footer_row(values = note_3b_e, colwidths = 4) %>%
  font(fontname = MS_FONT, part = "footer") %>%
  fontsize(size = MS_SIZE_NOTE, part = "footer") %>%
  italic(part = "footer") %>%
  set_table_properties(layout = "autofit")

save_as_docx(ft_3b_e, path = file.path(dir_tables, "table_sc1_step3b_expl_valence.docx"))
ft_3b_e
```

# SC2 Step 1 Tables

```{r sc2-data}
#| include: false

vmr_data  <- readRDS(file.path(dir_data_ana, "vmr_ana.rds"))
post_data <- readRDS(file.path(dir_data_ana, "post_ana.rds"))
bg_data   <- readRDS(file.path(dir_data_ana, "bg_ana.rds"))

vmr_agg <- vmr_data %>%
  group_by(dyad, person, topic) %>%
  summarise(
    PA_sim_act  = mean(PA_similarity_act,  na.rm = TRUE),
    PA_sim_perc = mean(PA_similarity_perc, na.rm = TRUE),
    NA_sim_act  = mean(NA_similarity_act,  na.rm = TRUE),
    NA_sim_perc = mean(NA_similarity_perc, na.rm = TRUE),
    .groups = "drop"
  )

post_long <- post_data %>%
  pivot_longer(
    cols = c(love_neg, close_neg, love_pos, close_pos),
    names_to  = c(".value", "topic"),
    names_pattern = "(love|close)_(neg|pos)"
  ) %>%
  mutate(topic = ifelse(topic == "neg", "negative", "positive"))

sc2 <- vmr_agg %>%
  inner_join(post_long, by = c("dyad", "person", "topic")) %>%
  left_join(bg_data %>% select(person, we_ness), by = "person") %>%
  mutate(context  = ifelse(topic == "positive", 1, 0),
         cwe_ness = we_ness - mean(we_ness, na.rm = TRUE)) %>%
  group_by(topic) %>%
  mutate(
    PA_sim_act_gmc  = PA_sim_act  - mean(PA_sim_act,  na.rm = TRUE),
    PA_sim_perc_gmc = PA_sim_perc - mean(PA_sim_perc, na.rm = TRUE),
    NA_sim_act_gmc  = NA_sim_act  - mean(NA_sim_act,  na.rm = TRUE),
    NA_sim_perc_gmc = NA_sim_perc - mean(NA_sim_perc, na.rm = TRUE)
  ) %>%
  ungroup()

sc2_neg <- sc2 %>% filter(topic == "negative")
sc2_pos <- sc2 %>% filter(topic == "positive")
```

## Negative context

```{r table-sc2-neg}
m_neg        <- fit_step1(sc2_neg)
d_neg        <- make_step1_data(m_neg)
ft_step1_neg <- make_step1_ft(d_neg, "Negative")

save_as_docx(ft_step1_neg, path = file.path(dir_tables, "table_sc2_step1_neg.docx"))
ft_step1_neg
```

## Positive context

```{r table-sc2-pos}
m_pos        <- fit_step1(sc2_pos)
d_pos        <- make_step1_data(m_pos)
ft_step1_pos <- make_step1_ft(d_pos, "Positive")

save_as_docx(ft_step1_pos, path = file.path(dir_tables, "table_sc2_step1_pos.docx"))
ft_step1_pos
```

## Summary (both contexts)

```{r table-sc2-step1-summary}
coef_cols_sum <- c("pos_1a", "pos_1b", "pos_1c", "neg_1a", "neg_1b", "neg_1c")

tbl_sum <- tibble(
  outcome   = rep(c("Closeness", "Closeness", "Love", "Love"), each = 3),
  affect    = rep(c("PA Similarity", "NA Similarity", "PA Similarity", "NA Similarity"), each = 3),
  predictor = rep(c("Actual", "Perceived", "Actual × Perceived"), times = 4),
  pos_1a = c(
    pull_fmt(m_pos$m1a_PA_close, "PA_sim_act"), "", "",
    pull_fmt(m_pos$m1a_NA_close, "NA_sim_act"), "", "",
    pull_fmt(m_pos$m1a_PA_love,  "PA_sim_act"), "", "",
    pull_fmt(m_pos$m1a_NA_love,  "NA_sim_act"), "", ""
  ),
  pos_1b = c(
    pull_fmt(m_pos$m1b_PA_close, "PA_sim_act"),  pull_fmt(m_pos$m1b_PA_close, "PA_sim_perc"), "",
    pull_fmt(m_pos$m1b_NA_close, "NA_sim_act"),  pull_fmt(m_pos$m1b_NA_close, "NA_sim_perc"), "",
    pull_fmt(m_pos$m1b_PA_love,  "PA_sim_act"),  pull_fmt(m_pos$m1b_PA_love,  "PA_sim_perc"), "",
    pull_fmt(m_pos$m1b_NA_love,  "NA_sim_act"),  pull_fmt(m_pos$m1b_NA_love,  "NA_sim_perc"), ""
  ),
  pos_1c = c(
    pull_fmt(m_pos$m1c_PA_close, "PA_sim_act_gmc"),
    pull_fmt(m_pos$m1c_PA_close, "PA_sim_perc_gmc"),
    pull_fmt(m_pos$m1c_PA_close, "PA_sim_act_gmc:PA_sim_perc_gmc"),
    pull_fmt(m_pos$m1c_NA_close, "NA_sim_act_gmc"),
    pull_fmt(m_pos$m1c_NA_close, "NA_sim_perc_gmc"),
    pull_fmt(m_pos$m1c_NA_close, "NA_sim_act_gmc:NA_sim_perc_gmc"),
    pull_fmt(m_pos$m1c_PA_love,  "PA_sim_act_gmc"),
    pull_fmt(m_pos$m1c_PA_love,  "PA_sim_perc_gmc"),
    pull_fmt(m_pos$m1c_PA_love,  "PA_sim_act_gmc:PA_sim_perc_gmc"),
    pull_fmt(m_pos$m1c_NA_love,  "NA_sim_act_gmc"),
    pull_fmt(m_pos$m1c_NA_love,  "NA_sim_perc_gmc"),
    pull_fmt(m_pos$m1c_NA_love,  "NA_sim_act_gmc:NA_sim_perc_gmc")
  ),
  neg_1a = c(
    pull_fmt(m_neg$m1a_PA_close, "PA_sim_act"), "", "",
    pull_fmt(m_neg$m1a_NA_close, "NA_sim_act"), "", "",
    pull_fmt(m_neg$m1a_PA_love,  "PA_sim_act"), "", "",
    pull_fmt(m_neg$m1a_NA_love,  "NA_sim_act"), "", ""
  ),
  neg_1b = c(
    pull_fmt(m_neg$m1b_PA_close, "PA_sim_act"),  pull_fmt(m_neg$m1b_PA_close, "PA_sim_perc"), "",
    pull_fmt(m_neg$m1b_NA_close, "NA_sim_act"),  pull_fmt(m_neg$m1b_NA_close, "NA_sim_perc"), "",
    pull_fmt(m_neg$m1b_PA_love,  "PA_sim_act"),  pull_fmt(m_neg$m1b_PA_love,  "PA_sim_perc"), "",
    pull_fmt(m_neg$m1b_NA_love,  "NA_sim_act"),  pull_fmt(m_neg$m1b_NA_love,  "NA_sim_perc"), ""
  ),
  neg_1c = c(
    pull_fmt(m_neg$m1c_PA_close, "PA_sim_act_gmc"),
    pull_fmt(m_neg$m1c_PA_close, "PA_sim_perc_gmc"),
    pull_fmt(m_neg$m1c_PA_close, "PA_sim_act_gmc:PA_sim_perc_gmc"),
    pull_fmt(m_neg$m1c_NA_close, "NA_sim_act_gmc"),
    pull_fmt(m_neg$m1c_NA_close, "NA_sim_perc_gmc"),
    pull_fmt(m_neg$m1c_NA_close, "NA_sim_act_gmc:NA_sim_perc_gmc"),
    pull_fmt(m_neg$m1c_PA_love,  "PA_sim_act_gmc"),
    pull_fmt(m_neg$m1c_PA_love,  "PA_sim_perc_gmc"),
    pull_fmt(m_neg$m1c_PA_love,  "PA_sim_act_gmc:PA_sim_perc_gmc"),
    pull_fmt(m_neg$m1c_NA_love,  "NA_sim_act_gmc"),
    pull_fmt(m_neg$m1c_NA_love,  "NA_sim_perc_gmc"),
    pull_fmt(m_neg$m1c_NA_love,  "NA_sim_act_gmc:NA_sim_perc_gmc")
  )
)

p_sum <- tibble(
  pos_1a = c(
    pull_p(m_pos$m1a_PA_close, "PA_sim_act"), NA, NA,
    pull_p(m_pos$m1a_NA_close, "NA_sim_act"), NA, NA,
    pull_p(m_pos$m1a_PA_love,  "PA_sim_act"), NA, NA,
    pull_p(m_pos$m1a_NA_love,  "NA_sim_act"), NA, NA
  ),
  pos_1b = c(
    pull_p(m_pos$m1b_PA_close, "PA_sim_act"),  pull_p(m_pos$m1b_PA_close, "PA_sim_perc"), NA,
    pull_p(m_pos$m1b_NA_close, "NA_sim_act"),  pull_p(m_pos$m1b_NA_close, "NA_sim_perc"), NA,
    pull_p(m_pos$m1b_PA_love,  "PA_sim_act"),  pull_p(m_pos$m1b_PA_love,  "PA_sim_perc"), NA,
    pull_p(m_pos$m1b_NA_love,  "NA_sim_act"),  pull_p(m_pos$m1b_NA_love,  "NA_sim_perc"), NA
  ),
  pos_1c = c(
    pull_p(m_pos$m1c_PA_close, "PA_sim_act_gmc"),
    pull_p(m_pos$m1c_PA_close, "PA_sim_perc_gmc"),
    pull_p(m_pos$m1c_PA_close, "PA_sim_act_gmc:PA_sim_perc_gmc"),
    pull_p(m_pos$m1c_NA_close, "NA_sim_act_gmc"),
    pull_p(m_pos$m1c_NA_close, "NA_sim_perc_gmc"),
    pull_p(m_pos$m1c_NA_close, "NA_sim_act_gmc:NA_sim_perc_gmc"),
    pull_p(m_pos$m1c_PA_love,  "PA_sim_act_gmc"),
    pull_p(m_pos$m1c_PA_love,  "PA_sim_perc_gmc"),
    pull_p(m_pos$m1c_PA_love,  "PA_sim_act_gmc:PA_sim_perc_gmc"),
    pull_p(m_pos$m1c_NA_love,  "NA_sim_act_gmc"),
    pull_p(m_pos$m1c_NA_love,  "NA_sim_perc_gmc"),
    pull_p(m_pos$m1c_NA_love,  "NA_sim_act_gmc:NA_sim_perc_gmc")
  ),
  neg_1a = c(
    pull_p(m_neg$m1a_PA_close, "PA_sim_act"), NA, NA,
    pull_p(m_neg$m1a_NA_close, "NA_sim_act"), NA, NA,
    pull_p(m_neg$m1a_PA_love,  "PA_sim_act"), NA, NA,
    pull_p(m_neg$m1a_NA_love,  "NA_sim_act"), NA, NA
  ),
  neg_1b = c(
    pull_p(m_neg$m1b_PA_close, "PA_sim_act"),  pull_p(m_neg$m1b_PA_close, "PA_sim_perc"), NA,
    pull_p(m_neg$m1b_NA_close, "NA_sim_act"),  pull_p(m_neg$m1b_NA_close, "NA_sim_perc"), NA,
    pull_p(m_neg$m1b_PA_love,  "PA_sim_act"),  pull_p(m_neg$m1b_PA_love,  "PA_sim_perc"), NA,
    pull_p(m_neg$m1b_NA_love,  "NA_sim_act"),  pull_p(m_neg$m1b_NA_love,  "NA_sim_perc"), NA
  ),
  neg_1c = c(
    pull_p(m_neg$m1c_PA_close, "PA_sim_act_gmc"),
    pull_p(m_neg$m1c_PA_close, "PA_sim_perc_gmc"),
    pull_p(m_neg$m1c_PA_close, "PA_sim_act_gmc:PA_sim_perc_gmc"),
    pull_p(m_neg$m1c_NA_close, "NA_sim_act_gmc"),
    pull_p(m_neg$m1c_NA_close, "NA_sim_perc_gmc"),
    pull_p(m_neg$m1c_NA_close, "NA_sim_act_gmc:NA_sim_perc_gmc"),
    pull_p(m_neg$m1c_PA_love,  "PA_sim_act_gmc"),
    pull_p(m_neg$m1c_PA_love,  "PA_sim_perc_gmc"),
    pull_p(m_neg$m1c_PA_love,  "PA_sim_act_gmc:PA_sim_perc_gmc"),
    pull_p(m_neg$m1c_NA_love,  "NA_sim_act_gmc"),
    pull_p(m_neg$m1c_NA_love,  "NA_sim_perc_gmc"),
    pull_p(m_neg$m1c_NA_love,  "NA_sim_act_gmc:NA_sim_perc_gmc")
  )
)

note_sum <- paste0(
  "Note. Model 1a = actual similarity only; 1b = actual + perceived similarity; ",
  "1c = actual + perceived + Actual × Perceived interaction (predictors grand-mean ",
  "centered within context). All models include a dyad-level random intercept. ",
  "Bold = p < .05. * p < .05, ** p < .01, *** p < .001."
)

ft_sum <- flextable(tbl_sum) %>%
  set_header_labels(
    outcome   = "Outcome",
    affect    = "Affect\nSimilarity",
    predictor = "Predictor",
    pos_1a = "1a", pos_1b = "1b", pos_1c = "1c",
    neg_1a = "1a", neg_1b = "1b", neg_1c = "1c"
  ) %>%
  add_header_row(
    values    = c("", "", "", "Positive Interaction", "Negative Interaction"),
    colwidths = c(1, 1, 1, 3, 3)
  ) %>%
  merge_v(j = c("outcome", "affect"), part = "header") %>%
  merge_v(j = c("outcome", "affect"), part = "body") %>%
  bold_sig(p_sum, coef_cols_sum) %>%
  font(fontname = MS_FONT, part = "all") %>%
  fontsize(size = MS_SIZE_BODY, part = "all") %>%
  fontsize(size = MS_SIZE_NOTE, part = "footer") %>%
  bold(part = "header") %>%
  align(align = "center", part = "header") %>%
  align(j = c("outcome", "affect", "predictor"), align = "left", part = "body") %>%
  align(j = coef_cols_sum, align = "right", part = "body") %>%
  border_remove() %>%
  hline_top(border = rule_thick, part = "header") %>%
  hline(i = 1, border = rule_thin, part = "header") %>%
  hline_bottom(border = rule_thick, part = "header") %>%
  hline(i = c(3, 6, 9), border = rule_thin, part = "body") %>%
  hline_bottom(border = rule_thick, part = "body") %>%
  add_footer_row(values = note_sum, colwidths = 9) %>%
  font(fontname = MS_FONT, part = "footer") %>%
  fontsize(size = MS_SIZE_NOTE, part = "footer") %>%
  italic(part = "footer") %>%
  set_table_properties(layout = "autofit")

save_as_docx(ft_sum, path = file.path(dir_tables, "table_sc2_step1_summary.docx"))
ft_sum
```

# SC2 Step 2 Tables

## Mediation Summary (H2.1–H2.2)

```{r sc2-step2-data}
#| include: false

set.seed(42)

# ── Mediator models ──
h21_PA_med_neg <- lme(PA_sim_perc ~ PA_sim_act, random = ~1|dyad, data = sc2_neg, method = "REML", na.action = na.exclude)
h21_PA_med_pos <- lme(PA_sim_perc ~ PA_sim_act, random = ~1|dyad, data = sc2_pos, method = "REML", na.action = na.exclude)
h22_NA_med_neg <- lme(NA_sim_perc ~ NA_sim_act, random = ~1|dyad, data = sc2_neg, method = "REML", na.action = na.exclude)
h22_NA_med_pos <- lme(NA_sim_perc ~ NA_sim_act, random = ~1|dyad, data = sc2_pos, method = "REML", na.action = na.exclude)

# ── Outcome models – love ──
h21_PA_out_love_neg  <- lme(love  ~ PA_sim_perc + PA_sim_act, random = ~1|dyad, data = sc2_neg, method = "REML", na.action = na.exclude)
h21_PA_out_love_pos  <- lme(love  ~ PA_sim_perc + PA_sim_act, random = ~1|dyad, data = sc2_pos, method = "REML", na.action = na.exclude)
h22_NA_out_love_neg  <- lme(love  ~ NA_sim_perc + NA_sim_act, random = ~1|dyad, data = sc2_neg, method = "REML", na.action = na.exclude)
h22_NA_out_love_pos  <- lme(love  ~ NA_sim_perc + NA_sim_act, random = ~1|dyad, data = sc2_pos, method = "REML", na.action = na.exclude)

# ── Outcome models – closeness ──
h21_PA_out_close_neg <- lme(close ~ PA_sim_perc + PA_sim_act, random = ~1|dyad, data = sc2_neg, method = "REML", na.action = na.exclude)
h21_PA_out_close_pos <- lme(close ~ PA_sim_perc + PA_sim_act, random = ~1|dyad, data = sc2_pos, method = "REML", na.action = na.exclude)
h22_NA_out_close_neg <- lme(close ~ NA_sim_perc + NA_sim_act, random = ~1|dyad, data = sc2_neg, method = "REML", na.action = na.exclude)
h22_NA_out_close_pos <- lme(close ~ NA_sim_perc + NA_sim_act, random = ~1|dyad, data = sc2_pos, method = "REML", na.action = na.exclude)

# ── Monte Carlo indirect effects (20,000 draws each) ──
mc_PA_love_neg  <- run_mc(h21_PA_med_neg, "PA_sim_act", h21_PA_out_love_neg,  "PA_sim_perc")
mc_PA_love_pos  <- run_mc(h21_PA_med_pos, "PA_sim_act", h21_PA_out_love_pos,  "PA_sim_perc")
mc_PA_close_neg <- run_mc(h21_PA_med_neg, "PA_sim_act", h21_PA_out_close_neg, "PA_sim_perc")
mc_PA_close_pos <- run_mc(h21_PA_med_pos, "PA_sim_act", h21_PA_out_close_pos, "PA_sim_perc")
mc_NA_love_neg  <- run_mc(h22_NA_med_neg, "NA_sim_act", h22_NA_out_love_neg,  "NA_sim_perc")
mc_NA_love_pos  <- run_mc(h22_NA_med_pos, "NA_sim_act", h22_NA_out_love_pos,  "NA_sim_perc")
mc_NA_close_neg <- run_mc(h22_NA_med_neg, "NA_sim_act", h22_NA_out_close_neg, "NA_sim_perc")
mc_NA_close_pos <- run_mc(h22_NA_med_pos, "NA_sim_act", h22_NA_out_close_pos, "NA_sim_perc")
```

```{r table-sc2-step2-mediation}
est_cols_med <- c("PA_neg", "PA_pos", "NA_neg", "NA_pos")

tbl_med2 <- tibble(
  outcome = c("Love/Closeness", "Love", "Love", "Love",
              "Closeness", "Closeness", "Closeness"),
  path    = c("a-path",
              "b-path", "c-path (indirect)", "c′-path (direct)",
              "b-path", "c-path (indirect)", "c′-path (direct)"),
  PA_neg  = c(
    pull_fmt(h21_PA_med_neg,       "PA_sim_act"),
    pull_fmt(h21_PA_out_love_neg,  "PA_sim_perc"),
    fmt_b(mc_PA_love_neg$point,    mc_PA_love_neg$p),
    pull_fmt(h21_PA_out_love_neg,  "PA_sim_act"),
    pull_fmt(h21_PA_out_close_neg, "PA_sim_perc"),
    fmt_b(mc_PA_close_neg$point,   mc_PA_close_neg$p),
    pull_fmt(h21_PA_out_close_neg, "PA_sim_act")
  ),
  PA_pos  = c(
    pull_fmt(h21_PA_med_pos,       "PA_sim_act"),
    pull_fmt(h21_PA_out_love_pos,  "PA_sim_perc"),
    fmt_b(mc_PA_love_pos$point,    mc_PA_love_pos$p),
    pull_fmt(h21_PA_out_love_pos,  "PA_sim_act"),
    pull_fmt(h21_PA_out_close_pos, "PA_sim_perc"),
    fmt_b(mc_PA_close_pos$point,   mc_PA_close_pos$p),
    pull_fmt(h21_PA_out_close_pos, "PA_sim_act")
  ),
  NA_neg  = c(
    pull_fmt(h22_NA_med_neg,       "NA_sim_act"),
    pull_fmt(h22_NA_out_love_neg,  "NA_sim_perc"),
    fmt_b(mc_NA_love_neg$point,    mc_NA_love_neg$p),
    pull_fmt(h22_NA_out_love_neg,  "NA_sim_act"),
    pull_fmt(h22_NA_out_close_neg, "NA_sim_perc"),
    fmt_b(mc_NA_close_neg$point,   mc_NA_close_neg$p),
    pull_fmt(h22_NA_out_close_neg, "NA_sim_act")
  ),
  NA_pos  = c(
    pull_fmt(h22_NA_med_pos,       "NA_sim_act"),
    pull_fmt(h22_NA_out_love_pos,  "NA_sim_perc"),
    fmt_b(mc_NA_love_pos$point,    mc_NA_love_pos$p),
    pull_fmt(h22_NA_out_love_pos,  "NA_sim_act"),
    pull_fmt(h22_NA_out_close_pos, "NA_sim_perc"),
    fmt_b(mc_NA_close_pos$point,   mc_NA_close_pos$p),
    pull_fmt(h22_NA_out_close_pos, "NA_sim_act")
  )
)

p_med2 <- tibble(
  PA_neg = c(
    pull_p(h21_PA_med_neg,       "PA_sim_act"),
    pull_p(h21_PA_out_love_neg,  "PA_sim_perc"),
    mc_PA_love_neg$p,
    pull_p(h21_PA_out_love_neg,  "PA_sim_act"),
    pull_p(h21_PA_out_close_neg, "PA_sim_perc"),
    mc_PA_close_neg$p,
    pull_p(h21_PA_out_close_neg, "PA_sim_act")
  ),
  PA_pos = c(
    pull_p(h21_PA_med_pos,       "PA_sim_act"),
    pull_p(h21_PA_out_love_pos,  "PA_sim_perc"),
    mc_PA_love_pos$p,
    pull_p(h21_PA_out_love_pos,  "PA_sim_act"),
    pull_p(h21_PA_out_close_pos, "PA_sim_perc"),
    mc_PA_close_pos$p,
    pull_p(h21_PA_out_close_pos, "PA_sim_act")
  ),
  NA_neg = c(
    pull_p(h22_NA_med_neg,       "NA_sim_act"),
    pull_p(h22_NA_out_love_neg,  "NA_sim_perc"),
    mc_NA_love_neg$p,
    pull_p(h22_NA_out_love_neg,  "NA_sim_act"),
    pull_p(h22_NA_out_close_neg, "NA_sim_perc"),
    mc_NA_close_neg$p,
    pull_p(h22_NA_out_close_neg, "NA_sim_act")
  ),
  NA_pos = c(
    pull_p(h22_NA_med_pos,       "NA_sim_act"),
    pull_p(h22_NA_out_love_pos,  "NA_sim_perc"),
    mc_NA_love_pos$p,
    pull_p(h22_NA_out_love_pos,  "NA_sim_act"),
    pull_p(h22_NA_out_close_pos, "NA_sim_perc"),
    mc_NA_close_pos$p,
    pull_p(h22_NA_out_close_pos, "NA_sim_act")
  )
)

note_med <- paste0(
  "Note. a-path: actual similarity predicting perceived similarity ",
  "(shared across outcomes within each context). b-path: perceived similarity ",
  "predicting outcome, controlling for actual similarity. c-path (indirect): ",
  "Monte Carlo indirect effect (20,000 draws). c′-path (direct): direct ",
  "effect of actual similarity, controlling for perceived similarity. ",
  "All models include a dyad-level random intercept. ",
  "Bold = p < .05. * p < .05, ** p < .01, *** p < .001."
)

ft_med2 <- flextable(tbl_med2) %>%
  set_header_labels(
    outcome = "Outcome",
    path    = "Path",
    PA_neg  = "Negative",
    PA_pos  = "Positive",
    NA_neg  = "Negative",
    NA_pos  = "Positive"
  ) %>%
  add_header_row(
    values    = c("", "PA Similarity (H2.1)", "NA Similarity (H2.2)"),
    colwidths = c(2, 2, 2)
  ) %>%
  merge_v(j = c("outcome", "path"), part = "header") %>%
  merge_v(j = "outcome", part = "body") %>%
  bold_sig(p_med2, est_cols_med) %>%
  font(fontname = MS_FONT, part = "all") %>%
  fontsize(size = MS_SIZE_BODY, part = "all") %>%
  bold(part = "header") %>%
  align(align = "center", part = "header") %>%
  align(j = c("outcome", "path"), align = "left", part = "body") %>%
  align(j = est_cols_med, align = "right", part = "body") %>%
  border_remove() %>%
  hline_top(border = rule_thick, part = "header") %>%
  hline(i = 1, border = rule_thin,  part = "header") %>%
  hline_bottom(border = rule_thick, part = "header") %>%
  hline(i = 1, border = rule_thin,  part = "body") %>%
  hline(i = 4, border = rule_thin,  part = "body") %>%
  hline_bottom(border = rule_thick, part = "body") %>%
  add_footer_row(values = note_med, colwidths = 6) %>%
  font(fontname = MS_FONT, part = "footer") %>%
  fontsize(size = MS_SIZE_NOTE, part = "footer") %>%
  italic(part = "footer") %>%
  set_table_properties(layout = "autofit")

save_as_docx(ft_med2, path = file.path(dir_tables, "table_sc2_step2_mediation_summary.docx"))
ft_med2
```

# SC2 Step 3a Tables

## We-ness Moderation Summary (H3.1–H3.2)

```{r sc2-step3a-data}
#| include: false

# ── H3.1: PA Similarity × We-ness (actual PA) ──
h31_PA_act_love_neg  <- lme(love  ~ PA_sim_act + cwe_ness + PA_sim_act:cwe_ness,  random = ~1|dyad, data = sc2_neg, method = "REML", na.action = na.exclude)
h31_PA_act_love_pos  <- lme(love  ~ PA_sim_act + cwe_ness + PA_sim_act:cwe_ness,  random = ~1|dyad, data = sc2_pos, method = "REML", na.action = na.exclude)
h31_PA_act_close_neg <- lme(close ~ PA_sim_act + cwe_ness + PA_sim_act:cwe_ness,  random = ~1|dyad, data = sc2_neg, method = "REML", na.action = na.exclude)
h31_PA_act_close_pos <- lme(close ~ PA_sim_act + cwe_ness + PA_sim_act:cwe_ness,  random = ~1|dyad, data = sc2_pos, method = "REML", na.action = na.exclude)

# ── H3.1: PA Similarity × We-ness (perceived PA) ──
h31_PA_perc_love_neg  <- lme(love  ~ PA_sim_perc + cwe_ness + PA_sim_perc:cwe_ness, random = ~1|dyad, data = sc2_neg, method = "REML", na.action = na.exclude)
h31_PA_perc_love_pos  <- lme(love  ~ PA_sim_perc + cwe_ness + PA_sim_perc:cwe_ness, random = ~1|dyad, data = sc2_pos, method = "REML", na.action = na.exclude)
h31_PA_perc_close_neg <- lme(close ~ PA_sim_perc + cwe_ness + PA_sim_perc:cwe_ness, random = ~1|dyad, data = sc2_neg, method = "REML", na.action = na.exclude)
h31_PA_perc_close_pos <- lme(close ~ PA_sim_perc + cwe_ness + PA_sim_perc:cwe_ness, random = ~1|dyad, data = sc2_pos, method = "REML", na.action = na.exclude)

# ── H3.2: NA Similarity × We-ness (actual NA) ──
h32_NA_act_love_neg  <- lme(love  ~ NA_sim_act + cwe_ness + NA_sim_act:cwe_ness,  random = ~1|dyad, data = sc2_neg, method = "REML", na.action = na.exclude)
h32_NA_act_love_pos  <- lme(love  ~ NA_sim_act + cwe_ness + NA_sim_act:cwe_ness,  random = ~1|dyad, data = sc2_pos, method = "REML", na.action = na.exclude)
h32_NA_act_close_neg <- lme(close ~ NA_sim_act + cwe_ness + NA_sim_act:cwe_ness,  random = ~1|dyad, data = sc2_neg, method = "REML", na.action = na.exclude)
h32_NA_act_close_pos <- lme(close ~ NA_sim_act + cwe_ness + NA_sim_act:cwe_ness,  random = ~1|dyad, data = sc2_pos, method = "REML", na.action = na.exclude)

# ── H3.2: NA Similarity × We-ness (perceived NA) ──
h32_NA_perc_love_neg  <- lme(love  ~ NA_sim_perc + cwe_ness + NA_sim_perc:cwe_ness, random = ~1|dyad, data = sc2_neg, method = "REML", na.action = na.exclude)
h32_NA_perc_love_pos  <- lme(love  ~ NA_sim_perc + cwe_ness + NA_sim_perc:cwe_ness, random = ~1|dyad, data = sc2_pos, method = "REML", na.action = na.exclude)
h32_NA_perc_close_neg <- lme(close ~ NA_sim_perc + cwe_ness + NA_sim_perc:cwe_ness, random = ~1|dyad, data = sc2_neg, method = "REML", na.action = na.exclude)
h32_NA_perc_close_pos <- lme(close ~ NA_sim_perc + cwe_ness + NA_sim_perc:cwe_ness, random = ~1|dyad, data = sc2_pos, method = "REML", na.action = na.exclude)
```

```{r table-sc2-step3a-weness}
est_cols_3a <- c("pos", "neg")

tbl_3a2 <- tibble(
  outcome    = rep(c("Closeness", "Love"), each = 12),
  affect     = rep(c("PA Similarity", "NA Similarity"), each = 6, times = 2),
  similarity = rep(c("Actual", "Perceived"), each = 3, times = 4),
  parameter  = rep(c("Similarity", "We-ness", "Sim × We-ness"), times = 8),
  pos = c(
    pull_fmt(h31_PA_act_close_pos,  "PA_sim_act"),
    pull_fmt(h31_PA_act_close_pos,  "cwe_ness"),
    pull_fmt(h31_PA_act_close_pos,  "PA_sim_act:cwe_ness"),
    pull_fmt(h31_PA_perc_close_pos, "PA_sim_perc"),
    pull_fmt(h31_PA_perc_close_pos, "cwe_ness"),
    pull_fmt(h31_PA_perc_close_pos, "PA_sim_perc:cwe_ness"),
    pull_fmt(h32_NA_act_close_pos,  "NA_sim_act"),
    pull_fmt(h32_NA_act_close_pos,  "cwe_ness"),
    pull_fmt(h32_NA_act_close_pos,  "NA_sim_act:cwe_ness"),
    pull_fmt(h32_NA_perc_close_pos, "NA_sim_perc"),
    pull_fmt(h32_NA_perc_close_pos, "cwe_ness"),
    pull_fmt(h32_NA_perc_close_pos, "NA_sim_perc:cwe_ness"),
    pull_fmt(h31_PA_act_love_pos,   "PA_sim_act"),
    pull_fmt(h31_PA_act_love_pos,   "cwe_ness"),
    pull_fmt(h31_PA_act_love_pos,   "PA_sim_act:cwe_ness"),
    pull_fmt(h31_PA_perc_love_pos,  "PA_sim_perc"),
    pull_fmt(h31_PA_perc_love_pos,  "cwe_ness"),
    pull_fmt(h31_PA_perc_love_pos,  "PA_sim_perc:cwe_ness"),
    pull_fmt(h32_NA_act_love_pos,   "NA_sim_act"),
    pull_fmt(h32_NA_act_love_pos,   "cwe_ness"),
    pull_fmt(h32_NA_act_love_pos,   "NA_sim_act:cwe_ness"),
    pull_fmt(h32_NA_perc_love_pos,  "NA_sim_perc"),
    pull_fmt(h32_NA_perc_love_pos,  "cwe_ness"),
    pull_fmt(h32_NA_perc_love_pos,  "NA_sim_perc:cwe_ness")
  ),
  neg = c(
    pull_fmt(h31_PA_act_close_neg,  "PA_sim_act"),
    pull_fmt(h31_PA_act_close_neg,  "cwe_ness"),
    pull_fmt(h31_PA_act_close_neg,  "PA_sim_act:cwe_ness"),
    pull_fmt(h31_PA_perc_close_neg, "PA_sim_perc"),
    pull_fmt(h31_PA_perc_close_neg, "cwe_ness"),
    pull_fmt(h31_PA_perc_close_neg, "PA_sim_perc:cwe_ness"),
    pull_fmt(h32_NA_act_close_neg,  "NA_sim_act"),
    pull_fmt(h32_NA_act_close_neg,  "cwe_ness"),
    pull_fmt(h32_NA_act_close_neg,  "NA_sim_act:cwe_ness"),
    pull_fmt(h32_NA_perc_close_neg, "NA_sim_perc"),
    pull_fmt(h32_NA_perc_close_neg, "cwe_ness"),
    pull_fmt(h32_NA_perc_close_neg, "NA_sim_perc:cwe_ness"),
    pull_fmt(h31_PA_act_love_neg,   "PA_sim_act"),
    pull_fmt(h31_PA_act_love_neg,   "cwe_ness"),
    pull_fmt(h31_PA_act_love_neg,   "PA_sim_act:cwe_ness"),
    pull_fmt(h31_PA_perc_love_neg,  "PA_sim_perc"),
    pull_fmt(h31_PA_perc_love_neg,  "cwe_ness"),
    pull_fmt(h31_PA_perc_love_neg,  "PA_sim_perc:cwe_ness"),
    pull_fmt(h32_NA_act_love_neg,   "NA_sim_act"),
    pull_fmt(h32_NA_act_love_neg,   "cwe_ness"),
    pull_fmt(h32_NA_act_love_neg,   "NA_sim_act:cwe_ness"),
    pull_fmt(h32_NA_perc_love_neg,  "NA_sim_perc"),
    pull_fmt(h32_NA_perc_love_neg,  "cwe_ness"),
    pull_fmt(h32_NA_perc_love_neg,  "NA_sim_perc:cwe_ness")
  )
)

p_mat_3a <- tibble(
  pos = c(
    pull_p(h31_PA_act_close_pos,  "PA_sim_act"),
    pull_p(h31_PA_act_close_pos,  "cwe_ness"),
    pull_p(h31_PA_act_close_pos,  "PA_sim_act:cwe_ness"),
    pull_p(h31_PA_perc_close_pos, "PA_sim_perc"),
    pull_p(h31_PA_perc_close_pos, "cwe_ness"),
    pull_p(h31_PA_perc_close_pos, "PA_sim_perc:cwe_ness"),
    pull_p(h32_NA_act_close_pos,  "NA_sim_act"),
    pull_p(h32_NA_act_close_pos,  "cwe_ness"),
    pull_p(h32_NA_act_close_pos,  "NA_sim_act:cwe_ness"),
    pull_p(h32_NA_perc_close_pos, "NA_sim_perc"),
    pull_p(h32_NA_perc_close_pos, "cwe_ness"),
    pull_p(h32_NA_perc_close_pos, "NA_sim_perc:cwe_ness"),
    pull_p(h31_PA_act_love_pos,   "PA_sim_act"),
    pull_p(h31_PA_act_love_pos,   "cwe_ness"),
    pull_p(h31_PA_act_love_pos,   "PA_sim_act:cwe_ness"),
    pull_p(h31_PA_perc_love_pos,  "PA_sim_perc"),
    pull_p(h31_PA_perc_love_pos,  "cwe_ness"),
    pull_p(h31_PA_perc_love_pos,  "PA_sim_perc:cwe_ness"),
    pull_p(h32_NA_act_love_pos,   "NA_sim_act"),
    pull_p(h32_NA_act_love_pos,   "cwe_ness"),
    pull_p(h32_NA_act_love_pos,   "NA_sim_act:cwe_ness"),
    pull_p(h32_NA_perc_love_pos,  "NA_sim_perc"),
    pull_p(h32_NA_perc_love_pos,  "cwe_ness"),
    pull_p(h32_NA_perc_love_pos,  "NA_sim_perc:cwe_ness")
  ),
  neg = c(
    pull_p(h31_PA_act_close_neg,  "PA_sim_act"),
    pull_p(h31_PA_act_close_neg,  "cwe_ness"),
    pull_p(h31_PA_act_close_neg,  "PA_sim_act:cwe_ness"),
    pull_p(h31_PA_perc_close_neg, "PA_sim_perc"),
    pull_p(h31_PA_perc_close_neg, "cwe_ness"),
    pull_p(h31_PA_perc_close_neg, "PA_sim_perc:cwe_ness"),
    pull_p(h32_NA_act_close_neg,  "NA_sim_act"),
    pull_p(h32_NA_act_close_neg,  "cwe_ness"),
    pull_p(h32_NA_act_close_neg,  "NA_sim_act:cwe_ness"),
    pull_p(h32_NA_perc_close_neg, "NA_sim_perc"),
    pull_p(h32_NA_perc_close_neg, "cwe_ness"),
    pull_p(h32_NA_perc_close_neg, "NA_sim_perc:cwe_ness"),
    pull_p(h31_PA_act_love_neg,   "PA_sim_act"),
    pull_p(h31_PA_act_love_neg,   "cwe_ness"),
    pull_p(h31_PA_act_love_neg,   "PA_sim_act:cwe_ness"),
    pull_p(h31_PA_perc_love_neg,  "PA_sim_perc"),
    pull_p(h31_PA_perc_love_neg,  "cwe_ness"),
    pull_p(h31_PA_perc_love_neg,  "PA_sim_perc:cwe_ness"),
    pull_p(h32_NA_act_love_neg,   "NA_sim_act"),
    pull_p(h32_NA_act_love_neg,   "cwe_ness"),
    pull_p(h32_NA_act_love_neg,   "NA_sim_act:cwe_ness"),
    pull_p(h32_NA_perc_love_neg,  "NA_sim_perc"),
    pull_p(h32_NA_perc_love_neg,  "cwe_ness"),
    pull_p(h32_NA_perc_love_neg,  "NA_sim_perc:cwe_ness")
  )
)

note_3a <- paste0(
  "Note. Similarity: main effect of similarity at average we-ness. ",
  "We-ness: main effect of trait we-ness at average similarity. ",
  "Sim × We-ness: interaction term (H3.1 for PA similarity, H3.2 for NA ",
  "similarity; preregistered hypothesis: positive). All models include a ",
  "dyad-level random intercept. Bold = p < .05. * p < .05, ** p < .01, *** p < .001."
)

ft_3a <- flextable(tbl_3a2) %>%
  set_header_labels(
    outcome    = "Outcome",
    affect     = "Affect",
    similarity = "Similarity",
    parameter  = "Parameter",
    pos        = "Positive",
    neg        = "Negative"
  ) %>%
  add_header_row(
    values    = c("", "Interaction Valence"),
    colwidths = c(4, 2)
  ) %>%
  merge_v(j = c("outcome", "affect", "similarity", "parameter"), part = "header") %>%
  merge_v(j = c("outcome", "affect", "similarity"), part = "body") %>%
  bold_sig(p_mat_3a, est_cols_3a) %>%
  font(fontname = MS_FONT, part = "all") %>%
  fontsize(size = MS_SIZE_BODY, part = "all") %>%
  bold(part = "header") %>%
  align(align = "center", part = "header") %>%
  align(j = c("outcome", "affect", "similarity", "parameter"), align = "left", part = "body") %>%
  align(j = est_cols_3a, align = "right", part = "body") %>%
  border_remove() %>%
  hline_top(border = rule_thick, part = "header") %>%
  hline(i = 1, border = rule_thin,  part = "header") %>%
  hline_bottom(border = rule_thick, part = "header") %>%
  hline(i =  3, border = rule_thin,  part = "body") %>%
  hline(i =  6, border = rule_thin,  part = "body") %>%
  hline(i =  9, border = rule_thin,  part = "body") %>%
  hline(i = 12, border = rule_thick, part = "body") %>%
  hline(i = 15, border = rule_thin,  part = "body") %>%
  hline(i = 18, border = rule_thin,  part = "body") %>%
  hline(i = 21, border = rule_thin,  part = "body") %>%
  hline_bottom(border = rule_thick, part = "body") %>%
  add_footer_row(values = note_3a, colwidths = 6) %>%
  font(fontname = MS_FONT, part = "footer") %>%
  fontsize(size = MS_SIZE_NOTE, part = "footer") %>%
  italic(part = "footer") %>%
  set_table_properties(layout = "autofit")

save_as_docx(ft_3a, path = file.path(dir_tables, "table_sc2_step3a_weness_moderation_summary.docx"))
ft_3a
```

# SC2 Step 3b Tables

## Context Moderation Summary (H3.3–H3.4)

```{r sc2-step3b-data}
#| include: false

# ── H3.3: PA × Context ──
h33_PA_act_love   <- lme(love  ~ PA_sim_act  + context + PA_sim_act:context,  random = ~1|dyad/person, data = sc2, method = "REML", na.action = na.exclude)
h33_PA_act_close  <- lme(close ~ PA_sim_act  + context + PA_sim_act:context,  random = ~1|dyad/person, data = sc2, method = "REML", na.action = na.exclude)
h33_PA_perc_love  <- lme(love  ~ PA_sim_perc + context + PA_sim_perc:context, random = ~1|dyad/person, data = sc2, method = "REML", na.action = na.exclude)
h33_PA_perc_close <- lme(close ~ PA_sim_perc + context + PA_sim_perc:context, random = ~1|dyad/person, data = sc2, method = "REML", na.action = na.exclude)

# ── H3.4: NA × Context ──
h34_NA_act_love   <- lme(love  ~ NA_sim_act  + context + NA_sim_act:context,  random = ~1|dyad/person, data = sc2, method = "REML", na.action = na.exclude)
h34_NA_act_close  <- lme(close ~ NA_sim_act  + context + NA_sim_act:context,  random = ~1|dyad/person, data = sc2, method = "REML", na.action = na.exclude)
h34_NA_perc_love  <- lme(love  ~ NA_sim_perc + context + NA_sim_perc:context, random = ~1|dyad/person, data = sc2, method = "REML", na.action = na.exclude)
h34_NA_perc_close <- lme(close ~ NA_sim_perc + context + NA_sim_perc:context, random = ~1|dyad/person, data = sc2, method = "REML", na.action = na.exclude)
```

```{r table-sc2-step3b-context}
est_cols_3b <- "estimate"

tbl_3b2 <- tibble(
  outcome    = rep(c("Closeness", "Love"), each = 12),
  affect     = rep(c("PA Similarity", "NA Similarity"), each = 6, times = 2),
  similarity = rep(c("Actual", "Perceived"), each = 3, times = 4),
  parameter  = rep(c("Similarity", "Context", "Sim × Context"), times = 8),
  estimate = c(
    pull_fmt(h33_PA_act_close,  "PA_sim_act"),
    pull_fmt(h33_PA_act_close,  "context"),
    pull_fmt(h33_PA_act_close,  "PA_sim_act:context"),
    pull_fmt(h33_PA_perc_close, "PA_sim_perc"),
    pull_fmt(h33_PA_perc_close, "context"),
    pull_fmt(h33_PA_perc_close, "PA_sim_perc:context"),
    pull_fmt(h34_NA_act_close,  "NA_sim_act"),
    pull_fmt(h34_NA_act_close,  "context"),
    pull_fmt(h34_NA_act_close,  "NA_sim_act:context"),
    pull_fmt(h34_NA_perc_close, "NA_sim_perc"),
    pull_fmt(h34_NA_perc_close, "context"),
    pull_fmt(h34_NA_perc_close, "NA_sim_perc:context"),
    pull_fmt(h33_PA_act_love,   "PA_sim_act"),
    pull_fmt(h33_PA_act_love,   "context"),
    pull_fmt(h33_PA_act_love,   "PA_sim_act:context"),
    pull_fmt(h33_PA_perc_love,  "PA_sim_perc"),
    pull_fmt(h33_PA_perc_love,  "context"),
    pull_fmt(h33_PA_perc_love,  "PA_sim_perc:context"),
    pull_fmt(h34_NA_act_love,   "NA_sim_act"),
    pull_fmt(h34_NA_act_love,   "context"),
    pull_fmt(h34_NA_act_love,   "NA_sim_act:context"),
    pull_fmt(h34_NA_perc_love,  "NA_sim_perc"),
    pull_fmt(h34_NA_perc_love,  "context"),
    pull_fmt(h34_NA_perc_love,  "NA_sim_perc:context")
  )
)

p_mat_3b <- tibble(
  estimate = c(
    pull_p(h33_PA_act_close,  "PA_sim_act"),
    pull_p(h33_PA_act_close,  "context"),
    pull_p(h33_PA_act_close,  "PA_sim_act:context"),
    pull_p(h33_PA_perc_close, "PA_sim_perc"),
    pull_p(h33_PA_perc_close, "context"),
    pull_p(h33_PA_perc_close, "PA_sim_perc:context"),
    pull_p(h34_NA_act_close,  "NA_sim_act"),
    pull_p(h34_NA_act_close,  "context"),
    pull_p(h34_NA_act_close,  "NA_sim_act:context"),
    pull_p(h34_NA_perc_close, "NA_sim_perc"),
    pull_p(h34_NA_perc_close, "context"),
    pull_p(h34_NA_perc_close, "NA_sim_perc:context"),
    pull_p(h33_PA_act_love,   "PA_sim_act"),
    pull_p(h33_PA_act_love,   "context"),
    pull_p(h33_PA_act_love,   "PA_sim_act:context"),
    pull_p(h33_PA_perc_love,  "PA_sim_perc"),
    pull_p(h33_PA_perc_love,  "context"),
    pull_p(h33_PA_perc_love,  "PA_sim_perc:context"),
    pull_p(h34_NA_act_love,   "NA_sim_act"),
    pull_p(h34_NA_act_love,   "context"),
    pull_p(h34_NA_act_love,   "NA_sim_act:context"),
    pull_p(h34_NA_perc_love,  "NA_sim_perc"),
    pull_p(h34_NA_perc_love,  "context"),
    pull_p(h34_NA_perc_love,  "NA_sim_perc:context")
  )
)

note_3b <- paste0(
  "Note. Similarity: slope in the negative interaction context (context = 0). ",
  "Context: difference in intercept between positive and negative interaction. ",
  "Sim × Context: difference in similarity slope between contexts ",
  "(H3.3 for PA similarity, H3.4 for NA similarity; preregistered hypothesis: ",
  "negative). All models include dyad-level and person-level random intercepts. ",
  "Bold = p < .05. * p < .05, ** p < .01, *** p < .001."
)

ft_3b <- flextable(tbl_3b2) %>%
  set_header_labels(
    outcome    = "Outcome",
    affect     = "Affect",
    similarity = "Similarity",
    parameter  = "Parameter",
    estimate   = "Estimate"
  ) %>%
  merge_v(j = c("outcome", "affect", "similarity"), part = "body") %>%
  bold_sig(p_mat_3b, est_cols_3b) %>%
  font(fontname = MS_FONT, part = "all") %>%
  fontsize(size = MS_SIZE_BODY, part = "all") %>%
  bold(part = "header") %>%
  align(align = "center", part = "header") %>%
  align(j = c("outcome", "affect", "similarity", "parameter"), align = "left", part = "body") %>%
  align(j = est_cols_3b, align = "right", part = "body") %>%
  border_remove() %>%
  hline_top(border = rule_thick, part = "header") %>%
  hline_bottom(border = rule_thick, part = "header") %>%
  hline(i =  3, border = rule_thin,  part = "body") %>%
  hline(i =  6, border = rule_thin,  part = "body") %>%
  hline(i =  9, border = rule_thin,  part = "body") %>%
  hline(i = 12, border = rule_thick, part = "body") %>%
  hline(i = 15, border = rule_thin,  part = "body") %>%
  hline(i = 18, border = rule_thin,  part = "body") %>%
  hline(i = 21, border = rule_thin,  part = "body") %>%
  hline_bottom(border = rule_thick, part = "body") %>%
  add_footer_row(values = note_3b, colwidths = 5) %>%
  font(fontname = MS_FONT, part = "footer") %>%
  fontsize(size = MS_SIZE_NOTE, part = "footer") %>%
  italic(part = "footer") %>%
  set_table_properties(layout = "autofit")

save_as_docx(ft_3b, path = file.path(dir_tables, "table_sc2_step3b_context_moderation_summary.docx"))
ft_3b
```

# RSA Parameter Table

```{r esm-data}
#| include: false

esm_data <- readRDS(file.path(dir_data_ana, "esm_ana.rds"))
load_config("ESM")

esm_data <- esm_data %>%
  arrange(dyad, person, time) %>%
  mutate(
    PA_own_sq        = cPA_own^2,
    PA_part_act_sq   = cPA_part_act^2,
    PA_own_x_part    = cPA_own * cPA_part_act,
    NA_own_sq        = cNA_own^2,
    NA_part_act_sq   = cNA_part_act^2,
    NA_own_x_part    = cNA_own * cNA_part_act,
    PA_part_perc_sq  = cPA_part_perc^2,
    PA_own_x_perc    = cPA_own * cPA_part_perc,
    NA_part_perc_sq  = cNA_part_perc^2,
    NA_own_x_perc    = cNA_own * cNA_part_perc
  )
```

```{r rsa-models}

rsa_ctrl <- lmeControl(opt = "nlminb")

rsa_PA_act <- lme(
  fixed  = love ~ cPA_own + cPA_part_act + PA_own_sq + PA_own_x_part + PA_part_act_sq,
  random = ~1 | dyad/person,
  correlation = corCAR1(form = ~time | dyad/person),
  data = esm_data, method = "REML", na.action = na.exclude, control = rsa_ctrl
)

rsa_NA_act <- lme(
  fixed  = love ~ cNA_own + cNA_part_act + NA_own_sq + NA_own_x_part + NA_part_act_sq,
  random = ~1 | dyad/person,
  correlation = corCAR1(form = ~time | dyad/person),
  data = esm_data, method = "REML", na.action = na.exclude, control = rsa_ctrl
)

rsa_PA_perc <- lme(
  fixed  = love ~ cPA_own + cPA_part_perc + PA_own_sq + PA_own_x_perc + PA_part_perc_sq,
  random = ~1 | dyad/person,
  correlation = corCAR1(form = ~time | dyad/person),
  data = esm_data, method = "REML", na.action = na.exclude, control = rsa_ctrl
)

rsa_NA_perc <- lme(
  fixed  = love ~ cNA_own + cNA_part_perc + NA_own_sq + NA_own_x_perc + NA_part_perc_sq,
  random = ~1 | dyad/person,
  correlation = corCAR1(form = ~time | dyad/person),
  data = esm_data, method = "REML", na.action = na.exclude, control = rsa_ctrl
)
```

```{r table-rsa}
params_PA_act  <- rsa_params(rsa_PA_act)
params_NA_act  <- rsa_params(rsa_NA_act)
params_PA_perc <- rsa_params(rsa_PA_perc)
params_NA_perc <- rsa_params(rsa_NA_perc)

rsa_tbl <- bind_rows(
  pivot_rsa(params_PA_act)  %>% mutate(Actor = "Actual PA",  Partner = "Actual PA"),
  pivot_rsa(params_PA_perc) %>% mutate(Actor = "Actual PA",  Partner = "Perceived PA"),
  pivot_rsa(params_NA_act)  %>% mutate(Actor = "Actual NA",  Partner = "Actual NA"),
  pivot_rsa(params_NA_perc) %>% mutate(Actor = "Actual NA",  Partner = "Perceived NA")
) %>%
  select(Actor, Partner, a1, a2, a3, a4, a5)

rsa_p <- bind_rows(
  pivot_rsa_p(params_PA_act)  %>% mutate(Actor = "Actual PA",  Partner = "Actual PA"),
  pivot_rsa_p(params_PA_perc) %>% mutate(Actor = "Actual PA",  Partner = "Perceived PA"),
  pivot_rsa_p(params_NA_act)  %>% mutate(Actor = "Actual NA",  Partner = "Actual NA"),
  pivot_rsa_p(params_NA_perc) %>% mutate(Actor = "Actual NA",  Partner = "Perceived NA")
) %>%
  select(a1, a2, a3, a4, a5)

rsa_note <- paste0(
  "Note. LOC = line of congruence (X = Y); LOIC = line of incongruence (X = −Y); ",
  "FPA = first principal axis. Parameters are derived from second-degree polynomial ",
  "multilevel models with random intercepts for dyads and persons, and continuous-time ",
  "AR(1) residuals. All predictors are person-mean centered. *** p < .001."
)

rsa_param_cols <- c("a1", "a2", "a3", "a4", "a5")

ft_rsa <- flextable(rsa_tbl) %>%
  set_header_labels(
    Actor   = "Actor",
    Partner = "Partner",
    a1 = "a₁\n(LOC slope)",
    a2 = "a₂\n(LOC curv.)",
    a3 = "a₃\n(LOIC slope)",
    a4 = "a₄\n(LOIC curv.)",
    a5 = "a₅\n(FPA = LOC)"
  ) %>%
  add_header_row(
    values    = c("Model", ""),
    colwidths = c(2, 5)
  ) %>%
  merge_v(j = c("Actor", "Partner"), part = "header") %>%
  merge_v(j = "Actor", part = "body") %>%
  bold_sig(rsa_p, rsa_param_cols) %>%
  font(fontname = MS_FONT, part = "all") %>%
  fontsize(size = MS_SIZE_BODY, part = "all") %>%
  fontsize(size = MS_SIZE_NOTE, part = "footer") %>%
  bold(part = "header") %>%
  align(align = "center", part = "header") %>%
  align(j = c("Actor", "Partner"), align = "left", part = "body") %>%
  align(j = rsa_param_cols, align = "right", part = "body") %>%
  border_remove() %>%
  hline_top(border = rule_thick, part = "header") %>%
  hline(i = 1, border = rule_thin,  part = "header") %>%
  hline_bottom(border = rule_thick, part = "header") %>%
  hline(i = 2, border = rule_thin, part = "body") %>%
  hline_bottom(border = rule_thick, part = "body") %>%
  add_footer_row(values = rsa_note, colwidths = 7) %>%
  font(fontname = MS_FONT, part = "footer") %>%
  fontsize(size = MS_SIZE_NOTE, part = "footer") %>%
  italic(part = "footer") %>%
  set_table_properties(layout = "autofit")

save_as_docx(ft_rsa, path = file.path(dir_tables, "table_rsa_params.docx"))
ft_rsa
```